### 2020-10-22 doc/temp-sentiment.tex
begin manuscript

\documentclass{article}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[a4paper, margin=1.25in]{geometry}
\usepackage{breakcites}
\usepackage{subcaption}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{textgreek}
\usepackage{authblk}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{lineno}
\usepackage[
  style=numeric,
  citestyle=numeric-comp,
  backend=biber,
  doi=true,
  natbib=true,
  sorting=none
]{biblatex}

\addbibresource{library.bib}

\begin{document}

\title{Assessing Vulnerability to Mental Health Impacts of Weather}

\author[1, *]{Matthew Cooper}

\affil[1]{T.H. Chan School of Public Health, Harvard University}
\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}

\maketitle
\begin{abstract}
  This is the Abstract!
\end{abstract}

\section{Introduction}
This work builds on existing work by \citep{baylis_weather_2018}

\section{Data}

\section{Methods}

\section{Results}

\section{Discussion}

\section{Conclusion}

\printbibliography

\section*{Supplemental Info}
\setcounter{table}{0}
\setcounter{figure}{0}
\setcounter{section}{0}
\renewcommand{\thetable}{S\arabic{table}}
\renewcommand{\thefigure}{S\arabic{figure}}
\renewcommand{\thesection}{S\arabic{section}}

\end{document}

### 2020-10-22 extract/Extract_NLCD.py
new earth engine extraction scripts

@@ -1,0 +1,105 @@
+#source $HOME/miniconda3/bin/activate
+#conda activate ee
+
+import json
+import ee
+import os
+from datetime import datetime, timedelta, time
+import pandas as pd
+import numpy as np
+import time as sleeptime #otherwise interferes with datetime.time
+import progressbar
+
+ee.Initialize()
+
+os.chdir('/home/ubuntu/tweets/tweets')
+
+fs = sorted(os.listdir())
+fs = [f for f in fs if 'json' in f]
+
+def parse_res(res, val, to_round):
+    if to_round:
+        return [round(x['properties'][val], 2) if x['properties']['landcover'] is not None else np.nan for x in res['features']]
+    else:
+        return [x['properties'][val] if x['properties']['landcover'] is not None else np.nan for x in res['features'] ]
+
+for f in fs:
+    print(datetime.now(), f)
+    #From the docs: The dataset uses a day-ending naming convention, e.g., a day ending at 1200 UTC on 1 January is labeled 1 January. 
+    today = datetime.strptime(f[:10], '%Y-%m-%d')
+    name = today + timedelta(days=1)
+    
+    #Separate climate data for whether tweet was before or after 12 UTC
+    nlcd2011 = ee.Image("USGS/NLCD/NLCD2011")
+    nlcd2016 = ee.Image("USGS/NLCD/NLCD2016")
+    #after12 = ee.Image("OREGONSTATE/PRISM/AN81d/" + name.strftime("%Y%m%d"))
+    #before12 = ee.Image("OREGONSTATE/PRISM/AN81d/" + today.strftime("%Y%m%d"))
+    with open(f, 'r') as c:
+        dat = json.loads(c.read())
+    
+    if len(dat) == 0:
+        print('passing')
+        continue
+    
+    #Get collections of points, separated by year before 2014 and year after 2014 
+    before_points = []
+    after_points = []
+    for d in dat:
+        lon = d['geo']['coordinates'][0]
+        lat = d['geo']['coordinates'][1]
+        created_at = datetime.strptime(d['tweet_created_at'], '%Y-%m-%d %H:%M:%S')
+        is_early = (created_at.year < 2014)
+        pt = ee.Geometry.Point(lon, lat)
+        feat = ee.Feature(pt, {'tweet_created_at': d['tweet_created_at'], 'id': str(d['id'])})
+        if is_early:
+            before_points.append(feat)
+        else:
+            after_points.append(feat)
+     
[…]

### 2020-10-22 extract/Extract_NLDAS.py

@@ -1,0 +1,134 @@
+#source $HOME/miniconda3/bin/activate
+#conda activate ee
+
+import json
+import ee
+import os
+from datetime import datetime, timedelta, time
+import pandas as pd
+import numpy as np
+import time as sleeptime #otherwise interferes with datetime.time
+from functools import reduce
+import math
+
+ee.Initialize()
+
+os.chdir('/home/ubuntu/tweets/tweets')
+
+fs = sorted(os.listdir())
+fs = [f for f in fs if 'json' in f]
+
+def parse_res(res, var):
+    df = pd.DataFrame([x['properties'] for x in res['features']])
+    df = df.rename(columns={'first': var})
+    return(df)
+
+def sh2rh(sh, temp, press):
+    #Adapted from https://earthscience.stackexchange.com/questions/2360/how-do-i-convert-specific-humidity-to-relative-humidity
+    press = press * 0.01 #convert milibars to pascals
+    es = 6.112 * math.exp((17.67*temp)/(temp + 243.5))
+    e = sh*press / (0.378 * sh + 0.622)
+    rh = e/es
+    if rh > 1:
+        return(1)
+    if rh < 0:
+        return(0)
+    return(rh)
+
+
+def heatindex(t, rh):
+    #Translated from R weathermetrics::heat.index.algorithm
+    #For more info: https://doi.org/10.1016/j.envres.2018.09.032
+    
+    #Convert to fahrenheit
+    t = t*(9/5) + 32
+    
+    #Convert rh to percent
+    rh = rh*100
+    
+    if t <= 40:
+        hi = t
+    else:
+        alpha = 61 + ((t - 68) * 1.2) + (rh * 0.094)
+        hi = 0.5 * (alpha + t)
+        if hi > 79:
+            hi = -42.379 + 2.04901523 * t + 10.14333127 * rh - 0.22475541 * t * rh - 6.83783 * 10**-3 * t**2 - 5.481717 * 10**-2 * rh**2 + 1.22874 * 10**-3 * t**2 * rh + 8.5282 * 10**-4 * t * rh**2 - 1.99 * 10**-6 * t**2 * rh**2
+            if (rh <= 13) & (t >= 80) & (t <= 112):
+                adjustment1 = (13 - rh)/4
+                adjustment2 = math.sqrt((17 - abs(t - 95))/17)
+                totaladjustment = adjustment1 * adjustment2
+                hi = hi - totaladjustment
+            elif (rh > 85) & (t >= 80) & (t <= 87):
[…]

### 2020-10-22 extract/Extract_PRISM.py

@@ -23,7 +23,7 @@ def parse_res(res, val, to_round):
     else:
         return [x['properties'][val] if x['properties']['ppt'] is not None else np.nan for x in res['features'] ]
 
-for f in fs[2801:]:
+for f in fs:
     print(datetime.now(), f)
     #From the docs: The dataset uses a day-ending naming convention, e.g., a day ending at 1200 UTC on 1 January is labeled 1 January. 
     today = datetime.strptime(f[:10], '%Y-%m-%d')

### 2020-10-22 extract/Combine.R
use Vader and write sample dfs

@@ -4,7 +4,7 @@ library(lubridate)
 
 setwd('/home/ubuntu/tweets')
 
-setDTthreads(64)
+setDTthreads(32)
 
 #To combine the daily csvs, I ran
 # cat sentiment/*.csv | grep -v tweet_created_at > sentiment_all.csv
@@ -13,9 +13,9 @@ setDTthreads(64)
 sen <- fread('sentiment_all.csv',
              col.names=c('id', 'tweet_created_at', 'weather_term', 
                         # 'afinn', 'textblob', 
-                        'hedono'),
-                        # 'vader', 'swn', 'wkwsci'),
-             drop=c(4, 5, 7, 8, 9))
+                        # 'hedono'),
+                         'vader'), # 'swn', 'wkwsci'),
+             drop=c(4, 5, 6, 8, 9))
 sen <- unique(sen, by=c('id', 'tweet_created_at'))
 
 #Get Climate Data
@@ -59,4 +59,12 @@ all$income_percap_q <- class[as.numeric(Hmisc::cut2(all$income_percap, g=5))]
 fwrite(all, 'all.csv', row.names=F)
 system('~/telegram.sh "Donezo!"')
 
+#Write samples after deleting tweet-specific data
+all[ , id:=NULL]
+all[ , tweet_created_at:=NULL]
+all[ , tweet_created_at_local:=NULL]
+all[ , lat:=NULL]
+all[ , lon:=NULL]
 
+fwrite(all[sample(.N, .N*0.01)], 'all_samp_1pct.csv', row.names=F)
+fwrite(all[sample(.N, .N*0.1)], 'all_samp_10pct.csv', row.names=F)

### 2020-12-03 doc/temp-sentiment.tex

@@
 
-\addbibresource{library.bib}
+\addbibresource{TextDataClimateShocks.bib}
 
@@
 
-\author[1, *]{Matthew Cooper}
+%\author[1, *]{Matthew Cooper}
 
-\affil[1]{T.H. Chan School of Public Health, Harvard University}
-\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
+%\affil[1]{T.H. Chan School of Public Health, Harvard University}
+%\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
 
 \maketitle
+
 \begin{abstract}
-  This is the Abstract!
+Weather can affect people’s mood and well-being.
+Traditional studies exploring the effects of weather on mental health require significant amounts of time to plan, fund and execute.
+Continuous in situ data streams from social media, newspapers, and other big data sources offer an opportunity to examine how people perceive and respond to environmental conditions.
+Recent work demonstrates that prevailing weather conditions can affect people’s sentiment as expressed on Twitter and other social media.
+Studies to date however, have modeled the weather as having the same effect on sentiment across all locations and individuals (i.e., in aggregate).
+In this study, we explore how weather affects individual expressed sentiment (via Twitter) across different weather gradients and locations.
+We analyzed the sentiment from a quarter of a billion geolocated Tweets from 2009 to 2019, using multiple measures of sentiment, overlaid with data about prevailing weather conditions, as well as local land cover, income levels, and demographic characteristics in the vicinity of Twitter users.
+We find that weather affects people differently across a range of incomes, with people in higher income brackets generally happier with increasing temperature, while people in lower income brackets are less happy with increasing temperature.
+This project extends existing research about the relationship between temperature and mood (sentiment) by examining how varying climatic conditions across different geographies and socio-economic groups are correlated with expressed sentiment.
+This project brings new insights into how humans perceive and respond to environmental shocks, with important implications for economics, policy and adaptation and resilience planning.
+
 \end{abstract}
 
+Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation
+
 \section{Introduction}
-This work builds on existing work by \citep{baylis_weather_2018}
+This work builds on existing work by \citep{baylis_weather_2018}.
+
+Human moods and mental states are important aspects of overall well-being and can be influenced by ambient and persistent environmental conditions.
+Climate change is accelerating the rate and variability around meteorological norms like the timing, magnitude, intensity and duration of precipitation events and air temperature minima and maxima \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
+Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meterological conditions.
+Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to meteorological conditions.
+Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.
+
+Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes .
+Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
+(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
+
+[from our Jan/Feb 2020 draft working manuscript in Google Docs]
+Climate change is a universal issue as it affects every country in the world.
+Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
+Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
+(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
+
+Meteorological conditions can impact human physical (cite) and emotional states (cite).
+Emotional states and well-being are associated with physiological functioning and mental acuity which can affect social relationships, workplace productivity (cite) and health risks (cite).
+People that are more reliant on livelihoods which require them to be outdoors (e.g., farming, agriculture, logging, etc.), living and working in places where they are exposed to environmental minima and maxima, or with prolonged exposure to ambient meteorological extremes are particularly vulnerable to changes in environmental exposure \citep{frimpong_heat_2017} and associated health risks (cite), with implications for workplace productivity and livelihoods \cite{kjellstrom_impact_2016}.
+For example, Nigerian maize farmers experience significant declines in productivity (2-8percent per degree above 17degC) and substantive increases in health impacts including dehydration, muscle cramps, headaches and dizziness, heat exhaustion, sun stroke, and even death \cite{sadiq_impact_2019}.
+Other research indicates farm health and labor productivity were compromised under extreme heat and cold events in the Nepali Food Bowl region \cite{budhathoki_socio-economic_2019}.
+Similarly, Oregon farmers reported health impacts of working in both hot and cold conditions but the negative impacts were attenuated relative to the high heat situations \cite{bethel_heat-related_2014}.
+Other research found occupational injuries in Thailand increased with increasing heat exposure and occupational heat stress \cite{tawatsupa_association_2013}.
+Heat exposure risks are also a key factor in urban areas with particular attention needed for how the impacts of changing climate play out in health inequality \cite{friel_urban_2011}.
+
+Current estimates place annual workplace productivity losses due to heat exposure at 15-20percent with that rate potentially doubling by 2050 \cite{kjellstrom_heat_2016}.
+Projected increases in future ambient temperature suggest serious short- and long-term potential health consequences from exposure to heat stress (cite) with some estimates putting potential productivity losses at several percentage points by 2030 \cite{kjellstrom_heat_2016}, with middle- and low-income countries particularly vulnerable as they are often more reliant on physical work for their livelihoods.
+
+The level of changes in weather influences people’s mental health and patterns of emotions expressed.
+A study by Sun et al.
+2018 \href{file:///Users/portia/Downloads/ijerph-16-00086-v2 20(1).pdf.}{(Portia has this PDF?)} demonstrates varied relationships between haze and negative emotions of the public under different seasons of the year.
+Being sad or happy influences one's style and value of reasoning.
+Time of day, season, location, and climate allow aggregate prediction of sentiments \cite{hannak_tweetin_2012} \href{https://www.ccs.neu.edu/~amislove/publications/Weather-ICWSM.pdf}{(link to PDF)} while positive affect from an evaluative statement enhances potential responses \cite{clore_how_2007} \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2483304/pdf/nihms40349.pdf}{(link to PDF)}.
+For instance, changes in temperature affects people’s response about weather.
+Depending on a reference temperature for an  area and time of year, people are more likely to comment on unusual weather for a particular place and time than on the same weather considered typical in another place \cite{moore_rapidly_2019}.
+
+Although climate change is an international phenomenon experienced in every country, its discussions vary from country to country.
+According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions.
+For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link} demonstrated that many people living in a state with higher mean temperature do express more positive emotions on Twitter than those living in colder states in the US.
+Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage global warming receives in the media in countries like USA, Australia and Germany.
+Studies comparing how different countries portray climate change have been widely conducted in USA, UK, France and the Netherlands \cite{vu_nationalizing_2019}.
+Such comparisons are mainly between developed countries with developing countries outside the focus.
+Schafer and O’Neil \cite{schafer_what_2013} advance the need for academic scholars to investigate transnational contexts within climate communication.
+Wealthier countries with comparatively higher GDP are likely to frame climate change as a political issue as financial resources exist for exploring research on climate change.
+Conversely, coverage of climate change news from poorer developing countries focus mainly on international relations.
+Thus, the effects of national macro economic variables that affects a country’s sociopolitical and economic development such as GDP reflects a country’s governance system in media coverage \cite{vu_what_2018}.
+Cultural theorists assert that, understanding such differential variations depicts how different entities interpret danger and respond to risk \cite{tansey_cultural_1999}.
+Inclusion of developing countries in scientific research provides an avenue for international support to such countries in responding to the effects of climate change.Yet studies attempting to explain cross-national variation in climate change public opinion is limited \cite{knight_public_2016}.
+
+Public debates on the current consequences of climate change are discussed in both scientific and non-scientific mediums.
+People can express their emotional states and feelings - sentiment - through physical, vocal and written expressions on mediums including newspapers, scientific articles, blogs and other online social media platforms.
+Twitter is one of the more popular social media platforms where people express their sentiments about any number of topics.
+Public posts on Twitter - called tweets - are widely available for public consumption and research, and offer a glimpse into collective social state.
+For example, text posted to these platforms can be used to gauge the public mood, assess opinions, measure brand affinity and for emergency planning and disaster response (cites).
+Sentiment is frequently employed as a correlate of emotional state or mood (cites).
+Weather events focus public attention in different ways of expression (cites).
+This offers a unique opportunity to analyze how events influence public discourse on their expressions and responses.
+
+Here, we report on associations between meteorological conditions and expressed sentiment of public posts on Twitter from the United States of America between 2009 and 2019.
+This work builds on research from Baylis and others \cite{baylis_weather_2018} on weather impacts on expressed sentiment in the USA.
+In particular, we extend this work by including the temperature effects on sentiment in multiple regions across the USA while drawing in various climatic variables and socio-demographic data.
+In this research, we explore two primary questions.
+First, to what extent are weather conditions correlated with changes in expressed sentiment?
+Second, is this effect moderated by wealth and local land cover conditions?
 
@@
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{../res/Boston_Map.png}
+  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
+  \label{fig:timeseries}
+\end{figure}
+
 \section{Methods}
+\subsection{Text Data Collection}
+Text data used for this project come from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose - which provides public access to a random 10\% of all public messages on Twitter.
+The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
+The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
+We used all geolocated Tweets from the Decahose for the period 2012-2016 with a bounding box polygon filter for India, Western Europe and the United States of America.
+We only considered users’ original content, thus did not include retweets in our analysis.
+
+\subsection{Analytic Approach}
+To assess whether temperature affects sentiment, we ran a linear regression model that included meteorological measures and sentiment.
+Independent variables included maximum daily temperature, precipitation and heat index.
+We also included fixed effects indicator variables at the year-month and decimal degree levels to account for potential biases in unobserved characteristics associated with seasonality and date.
+We examined the relationships between meteorological variables and sentiment using linear regression models.
+
+\subsection{Data Processing}
+We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 20193) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
+Sentiment data was analyzed using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
+We extracted meteorological data using the geolocated coordinates from the Twitter posts (latitude and longitude).
+We calculated mean climatological scores (z-scores)using baseline meteorological data from 1990 to present, calculated at a quarter degree resolution, and based on the long-term averages and standard deviation.
+We calculated z-scores for each day and the overall annual average z-score.
+For text data cleaning we used the python package tweet-preprocessor \href{https://pypi.org/project/tweet-preprocessor/}{(link)} to remove all text and characters not part of the main text phrase (e.g., URLs, mentions, GIFs, etc.).
+We converted all dates and times to UTC format.
+
+\subsection{Weather}
+We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington \href{https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2011JD016048}{link}.
+This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution \cite{xia_continental-scale_2012}.
+We extracted the data for the exact hour and location of each tweet using Google Earth Engine.
+We extracted several variables from the NLDAS-2 dataset, including temperature, specific humidity, air pressure, total precipitation, and shortwave radiation.
+Using temperature, specific humidity, and pressure, we derived relative humidity using methods described by  \cite{bolton_computation_1980}.
+Then, based on the Heat Index algorithm developed by Steadman \cite{steadman_assessment_1979} and improved upon by Rothfusz and the National Weather service \cite{rothfusz_heat_1990}, we calculated the Heat Index using temperature and relative humidity \cite{heo_comparison_2019}.
+
+ISSUES: deciduous trees lose leaves every year and may not provide much shade cover and heat relief during particular periods of the year when they are leafless.
+
+\subsection{Income \& Race}
+We used data from the US Census at the Census Block level to estimate the
+
+\subsection{Land Cover}
+To estimate land cover categories and the percentage of tree cover and impervious surfaces, we used data from the National Land Cover Database (NLCD) from the USGS' 2011  \cite{homer_completion_2015} and 2016 \cite{yang_new_2018} landcover datasets.
+These datasets are national land cover dataset prepared by the Multi-Resolution Land Characteristics (MRLC) Consortium.
+These data are at the thirty meter resolution, and provide estimates of the fraction of tree cover and impervious surfaces.
+We then associated each tweet with the land cover fraction for the nearest available year and pixel using the geolocation associated with each tweet, extracting landcover values from whichever land cover dataset year was closest in time to the date timestamp of each tweet.
+We used tree cover as a proxy for amount of potential shade and heat relief in a particular area and can be broadly indicative of generalized wealth.
+We use impervious surface cover as a proxy for relative proximity to potential natural shade and heat relief, and because individuals in urban environments may be exposed to the urban heat island effect.
+[NEED SOME CITATIONS HERE]
+
+\subsection{Sentiment}
+We assessed the sentiment expressed in text using several of the most commonly used measures of sentiment.
+Each of the measures uses validated dictionaries of words that have been assigned positive and negative scores and compares these words with words in the text corpora of interest.
+For measures of sentiment, we used the Hedonometer \href{https://hedonometer.org/timeseries/en_all/}{(link)}, Afinn \href{http://corpustext.com/reference/sentiment_afinn.html}{(link)}, Textblob \href{https://textblob.readthedocs.io/en/dev/}{(link)}, Vader , SentiwordNet and LIWC tools.
+These tools assign positive and negative scores to words, then average the scores for each text string.
+Our text data was multilingual, which can present challenges with analyzing text (cite).
+Given our interest only in the collection of words used (in aggregate) and not the sequence or nuance of the phrases, we translated all languages into English using Google Translate API and the Microsoft Azure API cognitive services (translator text API) before conducting the sentiment analyses.
+
+Some papers we might cite in this section (only for multilingual work): Vilares et al. 2017 , Lo et al. 2017, Dashtipour et al. 2016.
+
+Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! #fail]), etc.
+
+Issues with sample population (non-representative)
+
+Issues with what people say on social media being attenuated or mismatched from actual emotional state
+
+\subsubsection{Hedonometer}
+The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
+The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2)using a naive algorithm for scaling up from individual words to texts.
+For the English corpus, Dodds and others \cite{dodds_temporal_2011} collect 10,222 unique words and used crowd-sourcing platform Amazon Mechanical Turk to get human evaluation of happiness degree for each word in an integer scale from 1 to 9, representing a sad to happy spectrum.
+Score 5 represents neutral words.
+Each word will be calculated average score and then the word and happiness score are compiled into a dataset (labMT 1.0).
+Some illustrative example of words are:
+
+\[h_{avg} (\text{laughter}) = 8.50 \]
+\[h_{avg} (\text{the}) = 4.98\]
+\[h_{avg} (\text{hate}) = 2.34\]
+
+The sentiment score for single text will be the mean happiness score for all words in the text.
+
+\subsubsection{VADER}
+The VADER sentiment corpus was proposed by Gilbert and Hutto \cite{gilbert_vader_2014} and stands for 'Valence Aware Dictionary for sentiment Reasoning'.
+It is is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media by incorporating lexical features common to sentiment expression in microblogs
 
 \section{Results}
+(sub-section) DESCRIPTIVE STATISTICS
+
+Figure of heatmap of tweet density across USA (pick a year? entire dataset of georeferenced tweets?)
+
+(sub-section) TEMPERATURE and HUMIDITY
+Figure of distribution of max and min temperatures
+
+(sub-section) INCOME
+
+(sub-section?) SENTIMENT
+Figure of distribution of observed sentiment scores (by Hedonometer, Vader, etc.)
 
@@
 \section{Conclusion}
+
+\section{Acknowledgements}
+For funding, advising, SESYNC grant number?, etc.
 
@@
 
+\section*{Text Snippets Not Used Here but We Might Want to Keep}
+Gaps on above we intend to fill in:
+Our research aim and objectives
+Its contribution
+
+Dehghan et al.
+2012  The evaluation of heat stress through monitoring environmental factors and physiological responses in melting and casting industries workers.
+High prevalence of industrial workers experiencing heat stress.
+Relatively low humidity conditions made it necessary to alter their work/rest cycles.
+\href{http://www.ijehe.org/article.asp?issn=2277-9183;year=2012;volume=1;issue=1;spage=21;epage=21;aulast=Dehghan}{link}
+
+?.
+(from intro) Increased variability around environmental norms (e.g., lower/higher temperatures, changes in timing, magnitude, duration of precipitation or dry spells, etc.) can amplify the effects on systems.
+Increased uncertainty in planning for these events and ability to respond to them (adaptive capacity).
+- draw in cultural risk and social practice theories, here?
+
+NOTES: Scaling back out → we are exploring temperature and sentiment (in Tweets) across languages.
+This is a clear innovation that builds on the existing work from the Baylis et al.
+team (in their plosONE paper) and in fact is something they recommended be done - recommended to us as a team, too.
+
+We know that temps affect sentiment, but (1) Is this effect observable across languages? (2) Is this effect moderated by factors like wealth?
+
+We could ask similar questions substituting in the patent filings, too (sub-project 3 in our previous work-plan).
+These questions - and likely the methods used to answer these questions - are quite portable and compelling and may take the form for the patent work like this: Given we know that temperature affects worker productivity (increased temp increases productivity up to 13C, sharp negative decline thereafter), can we observe this pattern in patent filings?
+Across various geographies?
+Are there lags in filings after unusually cold or hot temperature events?
+
+For the Twitter data, using text translated into a base language is fully appropriate and a frequently used method if looking for latent text structures in topic modeling (see Proksch et al.
+2018 for an alternative).
+For sentiment analysis on multilingual datasets, we'll need to decide on an approach (see Ling Lo et al.
+2016 review) but there are several that are good.
+If we wanted to use another open-source option, we could follow that laid out by Lucas and others (2015).
+
+Proskch et al. 2018 paper --> \href{https://onlinelibrary.wiley.com/doi/full/10.1111/lsq.12218}{link}
+
+Ling Lo et al. 2016 paper --> \href{https://link.springer.com/article/10.1007/s10462-016-9508-4}{link}
+
+Lucas et al. 2015 paper --> \href{https://www.cambridge.org/core/journals/political-analysis/article/computerassisted-text-analysis-for-comparative-politics/CC8B2CF63A8CC36FE00A13F9839F92BB}{link}
+
 \end{document}

### 2020-12-03 doc/temp-sentiment.tex
Now includes all comments from the original working manuscript on Google Docs.

@@
 \subsection{Data Processing}
-We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 20193) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
+We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 2019) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
 Sentiment data was analyzed using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.

### 2020-12-03 model/model_sklearn_interact_continuous.py
more improvements to sparse models

@@ -3,22 +3,26 @@
 import pandas as pd
 import numpy as np
 from scipy import sparse
+from scipy.sparse.linalg import lsqr
 from sklearn.linear_model import LinearRegression
+from sklearn.linear_model import Lasso
 from sklearn.preprocessing import OneHotEncoder
 import math
 import getpass
 import os
 
 #Model run name
-MOD_RUN = 'racecat_moreknots'
+MOD_RUN = 'impervcont_moreknots'
 
-INTERACT_VAR = 'majority'
+INTERACT_VAR = 'impervious'
 
 FIXED_EFFECTS = ['dow', 'doy', 'tod', 'fips', 'year', 'statemonth']
 
 FILTER_WEATHER_TERM = True
 
 SENTIMENT_VAR = 'vader'
+
+LOG_TRANSFORM = False
 
 ### Local ###
 if getpass.getuser() == 'mattcoop':
@@ -36,6 +40,9 @@ if getpass.getuser() == 'ubuntu':
 if FILTER_WEATHER_TERM:
     data = data[data['weather_term'] != 1] 
 
+if LOG_TRANSFORM:
+    data[INTERACT_VAR] = data[INTERACT_VAR].apply(lambda x: math.log(x))
+
 #Remove columns we are not using right now
 cols = data.columns.tolist()
 keepcols = ['precip', 'srad', 'temp.hi'] + FIXED_EFFECTS + [INTERACT_VAR, SENTIMENT_VAR]
@@ -47,21 +54,6 @@ data = data.drop(columns=dropcols)
 y = data[[SENTIMENT_VAR]]
 data = data.drop(columns=[SENTIMENT_VAR])
 
-#Get categories of interact var
-cat = np.sort(data[INTERACT_VAR].unique()).tolist()
-
-# #Set knots
-# # Can inspect quantiles with
-# # cut = pd.qcut(dat['temp.hi'], q=10)
-# # cut.cat.categories
-# # Use 20 categories for temp.hi
-# # Use 5 for precip (since 85% are 0 anyway)
-# # Use 10 for srad (since 50% are 0)
-# knots = {'temp.hi': np.linspace(-36, 56, num=15).tolist(),
-#         'precip': list(map(lambda x: math.exp(x) - 1, 
-#                            np.linspace(math.log(0 + 1), math.log(73 + 1), num=5))),
-#         'srad': list(map(lambda x: math.exp(x) - 1,
-#                          np.linspace(math.log(0 + 1), math.log(1354 + 1), num=5)))}
[…]

### 2020-12-03 model/model_sklearn_interact_continuous_alt.py

@@ -3,22 +3,26 @@
 import pandas as pd
 import numpy as np
 from scipy import sparse
+from scipy.sparse.linalg import lsqr
 from sklearn.linear_model import LinearRegression
+from sklearn.linear_model import Lasso
 from sklearn.preprocessing import OneHotEncoder
 import math
 import getpass
 import os
 
 #Model run name
-MOD_RUN = 'racecat_moreknots'
+MOD_RUN = 'treecont_moreknots'
 
-INTERACT_VAR = 'majority'
+INTERACT_VAR = 'tree'
 
 FIXED_EFFECTS = ['dow', 'doy', 'tod', 'fips', 'year', 'statemonth']
 
 FILTER_WEATHER_TERM = True
 
 SENTIMENT_VAR = 'vader'
+
+LOG_TRANSFORM = False
 
 ### Local ###
 if getpass.getuser() == 'mattcoop':
@@ -36,6 +40,9 @@ if getpass.getuser() == 'ubuntu':
 if FILTER_WEATHER_TERM:
     data = data[data['weather_term'] != 1] 
 
+if LOG_TRANSFORM:
+    data[INTERACT_VAR] = data[INTERACT_VAR].apply(lambda x: math.log(x))
+
 #Remove columns we are not using right now
 cols = data.columns.tolist()
 keepcols = ['precip', 'srad', 'temp.hi'] + FIXED_EFFECTS + [INTERACT_VAR, SENTIMENT_VAR]
@@ -47,21 +54,6 @@ data = data.drop(columns=dropcols)
 y = data[[SENTIMENT_VAR]]
 data = data.drop(columns=[SENTIMENT_VAR])
 
-#Get categories of interact var
-cat = np.sort(data[INTERACT_VAR].unique()).tolist()
-
-# #Set knots
-# # Can inspect quantiles with
-# # cut = pd.qcut(dat['temp.hi'], q=10)
-# # cut.cat.categories
-# # Use 20 categories for temp.hi
-# # Use 5 for precip (since 85% are 0 anyway)
-# # Use 10 for srad (since 50% are 0)
-# knots = {'temp.hi': np.linspace(-36, 56, num=15).tolist(),
-#         'precip': list(map(lambda x: math.exp(x) - 1, 
-#                            np.linspace(math.log(0 + 1), math.log(73 + 1), num=5))),
-#         'srad': list(map(lambda x: math.exp(x) - 1,
-#                          np.linspace(math.log(0 + 1), math.log(1354 + 1), num=5)))}
[…]

### 2020-12-03 doc/temp-sentiment.tex
New text, comments from P. Baylis, additional comments embedded 12/3/20 SESYNC co-working session end.

@@
 This work builds on existing work by \citep{baylis_weather_2018}. 
+Consider going through our paper and using the ABT method for ensuring we are building up a logic string, arguments, why we're doing this, what it answers, etc.
 
@@
 
+Building on previous findings from Baylis and others, we hypothesize that local ambient weather conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1].
+Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas [H2].
+Third, we hypothesize that people living in areas with less impervious surfaces and more vegetative tree cover will exhibit higher sentiment scores [H3].
+
 \section{Data}
@@
 
-Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! #fail]), etc.
+Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! \#fail]), etc.
 
@@
 Figure of distribution of max and min temperatures
+Comment from Patrick Baylis on our "preliminary results". 2nd comment from Patrick
 
@@
 Figure of distribution of observed sentiment scores (by Hedonometer, Vader, etc.)
+
+Comment from Patrick Baylis on our "preliminary results" doc
 

### 2020-12-04 model/fixest.R
try model with fixest estimation

@@ -1,0 +1,34 @@
+## This looks like it will work, but is too large for an E16
+
+library(data.table)
+library(fixest)
+
+setwd('~/tweets/')
+
+data <- fread('all.csv')
+
+#Tempknots
+knots = list("temp.hi"= c(min(data$temp.hi), -10, 0, 5, 10, 15, 20, 25, 30, 35, 
+                          max(data$temp.hi)),
+             "precip"= c(min(data$precip), 0.000001, 0.05, 0.5, max(data$precip)),
+             "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
+
+#Define Segmenting Function
+piece.formula <- function(var.name, knots) {
+  formula.sign <- rep(" - ", length(knots))
+  formula.sign[knots < 0] <- " + "
+  paste(var.name, "+",
+        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
+}
+
+form <- paste0("vader ~ ", piece.formula("temp.hi", knots[['temp.hi']]),
+                           #piece.formula("precip", knots[['precip']]),
+                           #piece.formula("srad", knots[['srad']]),
+                           " | dow + doy + tod + fips + year + statemonth")
+
+start <- Sys.time()
+print(start)
+mod <- feols(as.formula(form), data)
+end <- Sys.time()
+end - start

### 2020-12-04 model/model_local.R
more visualizing results

@@ -1,0 +1,17 @@
+library(data.table)
+library(glmnet)
+library(Matrix)
+library(ggplot2)
+library(MatrixModels)
+
+all <- fread('~/tweets/all_samp_1pct.csv', stringsAsFactors=T)
+all <- data.frame(all)
+
+
+form <- paste0("vader ~ income_percap*temp.hi + income_percap*I(pmax(temp.hi + 0, 0)) + income_percap*I(pmax(temp.hi +  15, 0)) + income_percap*I(pmax(temp.hi + 30, 0)) + statemonth + dow + doy + tod + year + fips")
+
+mm <- sparse.model.matrix(as.formula(form), data=all)
+
+mod <- MatrixModels:::lm.fit.sparse(mm, y=all$vader)
+
+

### 2020-12-04 viz/sklearn_interact_res_ses_manual.R

@@ -4,12 +4,22 @@ setwd('~/tweets/mod-res/')
 
 options(stringsAsFactors=F)
 
-MOD_RUN <- 'weather_income_full'
-catname <- 'income_percap'
-
-pred <- read.csv(paste0(MOD_RUN, '_preds.csv'))
+MOD_RUN <- 'weather_race_full'
+catname <- 'majority'
 
 coefs <- read.csv(paste0(MOD_RUN, '_coefs.csv'))
+
+pred <- read.csv(paste0(MOD_RUN, '_preds.csv'))
+# #In case pred creation effed up
+pred <- bind_rows(list(data.frame(temp.hi=unique(pred$temp.hi), precip=0, srad=0),
+                           data.frame(precip=unique(pred$precip), srad=0, temp.hi=0),
+                           data.frame(srad=unique(pred$srad), precip=0, temp.hi=0)))
+pred <- bind_rows(pred %>% mutate(majority = 'race_black'),
+                  pred %>% mutate(majority = 'race_white'),
+                  pred %>% mutate(majority = 'race_hisp'),
+                  pred %>% mutate(majority = 'race_other'))
+
+
 
 #####################################
 # Determine model parameters/ set up
@@ -76,15 +86,15 @@ for (ct in cat[2:length(cat)]){
 #Make Predictions
 #############################
 pred$prediction <- NA
-pred$std.err <- NA
+#pred$std.err <- NA
 
 for (i in 1:nrow(pred)){
-  pred$prediction[i] <- sum(predX[i, ]*coefs$coefs[1:nrow(pred)])
-  pred$std.err[i] <- sum(predX[i, ]*coefs$standarderror[1:nrow(pred)])
+  pred$predicted[i] <- sum(predX[i, ]*coefs$coefs[1:nrow(pred)])
+  #pred$std.err[i] <- sum(predX[i, ]*coefs$standarderror[1:nrow(pred)])
 }
 
-pred$ymin <- pred$prediction - pred$std.err*qnorm(0.975)
-pred$ymax <- pred$prediction + pred$std.err*qnorm(0.975)
+#pred$ymin <- pred$prediction - pred$std.err*qnorm(0.975)
+#pred$ymax <- pred$prediction + pred$std.err*qnorm(0.975)
 
 ###########################
 # Make Plots
[…]

### 2020-12-05 model/fixest.R
get predictions from fixest

@@ -1,17 +1,18 @@
-## This looks like it will work, but is too large for an E16
+##E32
 
 library(data.table)
 library(fixest)
+library(dplyr)
+library(ggplot2)
 
 setwd('~/tweets/')
 
 data <- fread('all.csv')
 
 #Tempknots
-knots = list("temp.hi"= c(min(data$temp.hi), -10, 0, 5, 10, 15, 20, 25, 30, 35, 
-                          max(data$temp.hi)),
-             "precip"= c(min(data$precip), 0.000001, 0.05, 0.5, max(data$precip)),
-             "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
+knots = list("temp.hi"= c(-10, 0, 5, 10, 15, 20, 25, 30, 35),
+             "precip"= c(0.000001, 0.05, 0.5),
+             "srad"= c(0.000001, 250, 500, 1000))
 
 #Define Segmenting Function
 piece.formula <- function(var.name, knots) {
@@ -22,13 +23,39 @@ piece.formula <- function(var.name, knots) {
               collapse = " + ", sep=""))
 }
 
-form <- paste0("vader ~ ", piece.formula("temp.hi", knots[['temp.hi']]),
-                           #piece.formula("precip", knots[['precip']]),
-                           #piece.formula("srad", knots[['srad']]),
-                           " | dow + doy + tod + fips + year + statemonth")
 
-start <- Sys.time()
-print(start)
+system('~/telegram.sh "starting"')
+form <- paste0("vader ~ ", piece.formula("temp.hi", knots[['temp.hi']]), ' + ',
+                           piece.formula("precip", knots[['precip']]), ' + ',
+                           piece.formula("srad", knots[['srad']]),
+                           " | dow + doy + tod + fips + year + statemonth")
 mod <- feols(as.formula(form), data)
-end <- Sys.time()
-end - start
+system('~/telegram.sh "donezo!"')
+
+pred <- bind_rows(data.frame(srad=0, precip=0, temp.hi=seq(-40, 50)),
+                  data.frame(temp.hi=0, precip=0, srad=c(seq(0, 0.05, length.out=10),
+                                                         seq(0.05, 0.5, length.out=10),
+                                                         seq(0.5, 50, length.out=10))),
+                  data.frame(temp.hi=0, srad=0, precip=c(seq(0, 250, length.out=10),
[…]

### 2020-12-05 model/fixest.R
run model with income_percap; plot results

@@ -1,11 +1,15 @@
-## This looks like it will work, but is too large for an E16
+##Use E32
 
 library(data.table)
 library(fixest)
+library(survey)
+library(tidyverse)
 
 setwd('~/tweets/')
 
-data <- fread('all.csv')
+#data <- fread('all.csv')
+data <- fread('all_samp_1pct.csv')
+data$income_percap <- log(data$income_percap)
 
 #Tempknots
 knots = list("temp.hi"= c(min(data$temp.hi), -10, 0, 5, 10, 15, 20, 25, 30, 35, 
@@ -14,21 +18,126 @@ knots = list("temp.hi"= c(min(data$temp.hi), -10, 0, 5, 10, 15, 20, 25, 30, 35,
              "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
 
 #Define Segmenting Function
-piece.formula <- function(var.name, knots) {
+piece.formula <- function(var.name, knots, interact_var='') {
+  knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
   formula.sign[knots < 0] <- " + "
+  if (interact_var != ''){
+    var.name <- paste0(interact_var, "*", var.name)
+  }
   paste(var.name, "+",
         paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
               collapse = " + ", sep=""))
 }
 
-form <- paste0("vader ~ ", piece.formula("temp.hi", knots[['temp.hi']]),
-                           #piece.formula("precip", knots[['precip']]),
-                           #piece.formula("srad", knots[['srad']]),
-                           " | dow + doy + tod + fips + year + statemonth")
+formula <- paste0("vader ~ ", 
+                     piece.formula("temp.hi", knots[['temp.hi']], "income_percap"), ' + ',
+                     piece.formula("precip", knots[['precip']], "income_percap"), ' + ',
+                     piece.formula("srad", knots[['srad']], "income_percap"),
+                     " | dow + doy + tod + fips + year + statemonth")
 
-start <- Sys.time()
-print(start)
-mod <- feols(as.formula(form), data)
-end <- Sys.time()
-end - start
+mod <- feols(as.formula(formula), data)
+
[…]

### 2020-12-17 model/fixest_income_hypothesis.R
test no interaction and quadratic models

@@ -1,0 +1,191 @@
+#Use E32
+
+library(data.table)
+library(fixest)
+library(survey)
+library(tidyverse)
+library(cowplot)
+
+options(scipen=100)
+
+setwd('~/tweets/')
+
+data <- fread('all.csv')
+data$income_percap <- log(data$income_percap)
+
+data <- data[weather_term == 0, ]
+
+qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+
+formula <- paste0("vader ~ income_percap*temp + income_percap*I(temp^2) + 
+                  income_percap*precip + income_percap*I(precip^2) + 
+                  income_percap*srad + income_percap*I(srad^2)", 
+                  " | dow + doy + tod + fips + year + statemonth")
+
+start <- Sys.time()
+mod <- feols(as.formula(formula), data)
+end <- Sys.time()
+
+make_groups <- function(df, label, values){
+  final <- data.frame()
+  for (v in values){
+    df[ , label] <- v
+    final <- bind_rows(df, final)
+  }
+  final
+}
+
+######################################
+#temp
+
+preddf <- data.frame(temp=seq(min(data$temp), max(data$temp), length.out=250))
+preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
+
+preddf$vader <- 1
+
+mm <- model.matrix(vader ~ income_percap*temp + income_percap*I(temp^2),
+                   data=preddf)
+
+mm <- mm[ , colnames(mm) %in% names(coef(mod))]
+
+mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
+
+res <- svycontrast(mod, mmp)
+
+preddf <- cbind(preddf, as.data.frame(res))
+preddf$ymin <- preddf$contrast - 1.96*preddf$SE
+preddf$ymax <- preddf$contrast + 1.96*preddf$SE
+
+preddf$income_percap <- factor(preddf$income_percap)
+levels(preddf$income_percap) <- preddf$income_percap %>%
+                                   levels %>%
+                                   as.numeric %>%
+                                   exp %>%
+                                   round(0) %>%
+                                   format(big.mark = ',') %>%
+                                   paste0('$', .)
+
+levels(preddf$income_percap) <- paste0(levels(preddf$income_percap), ' (', 
[…]

### 2020-12-17 model/fixest_nointeract_quad.R

@@ -1,0 +1,133 @@
+#Use E32
+
+library(data.table)
+library(fixest)
+library(survey)
+library(tidyverse)
+library(cowplot)
+
+options(scipen=100)
+
+setwd('~/tweets/')
+
+data <- fread('all.csv')
+
+data <- data[weather_term == 0, ]
+
+formula <- paste0("vader ~ temp + temp^2 + precip + precip^2 + srad + srad^2", 
+                     " | dow + doy + tod + fips + year + statemonth")
+
+start <- Sys.time()
+mod <- feols(as.formula(formula), data)
+end <- Sys.time()
+
+######################################
+#temp
+
+preddf <- data.frame(temp=seq(min(data$temp), max(data$temp), length.out=250))
+
+preddf$vader <- 1
+
+mm <- model.matrix(vader ~ temp + I(temp^2),
+                   data=preddf)
+
+mm <- mm[ , colnames(mm) %in% names(coef(mod))]
+
+mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
+
+res <- svycontrast(mod, mmp)
+
+preddf <- cbind(preddf, as.data.frame(res))
+preddf$ymin <- preddf$contrast - 1.96*preddf$SE
+preddf$ymax <- preddf$contrast + 1.96*preddf$SE
+
+curve <- ggplot(preddf) + 
+  geom_line(aes(x=temp, y=contrast)) + 
+  geom_ribbon(aes(x=temp, ymin=ymin, ymax=ymax), alpha=0.5) + 
+  scale_x_continuous(expand=c(0, 0), limits=c(-21.4, 43.1)) + 
+  labs(x='', y='Predicted Mood of Tweet') +
+  theme(legend.position=c(0.2, 0.2))
+
+hist <- ggplot(data[sample(1:nrow(data), nrow(data)*0.01), 'temp']) +
+  geom_histogram(aes(x=temp), binwidth=1) + 
+  scale_x_continuous(expand=c(0, 0), limits=c(-21.38222, 43.02777)) + 
+  scale_y_continuous(labels=function(x){format(x*100, big.mark=',')}) + 
+  labs(y="Tweet Count", x="Temperature")
+
+plot_grid(curve, hist, align='v', axis='rl', ncol=1, rel_heights=c(0.8, 0.2))
+ggsave('~/temp-sentiment/res/quad-temp-only.png', width=6, height=7)
+
[…]

### 2020-12-17 doc/temp-sentiment.tex
Latest manuscript edits, comments, text, etc from co-working session Thursday 17Dec2020. Reflects new ABT (And But Therefore) framing and paper re-organization.

@@
 
-\title{Assessing Vulnerability to Mental Health Impacts of Weather}
+\title{Heatwaves Exacerbate Inequalities in Mental Health}
 
@@
 Weather can affect people’s mood and well-being.
-Traditional studies exploring the effects of weather on mental health require significant amounts of time to plan, fund and execute.
+Traditional studies exploring the effects of weather on mental health require significant resource investment (e.g., time, funding, execution).
 Continuous in situ data streams from social media, newspapers, and other big data sources offer an opportunity to examine how people perceive and respond to environmental conditions.
@@
 
+%Cite this new paper: https://doi.org/10.1016/S2542-5196(20)30251-5
+
 \section{Introduction}
+
+% Organize as follows:
+
+% Heatwaves are known to worsen mental health
+% AND mental health is strongly associated with income
+% AND wealthier better able to cope with heatwaves 
+% BUT we know little about the interaction between heatwaves, mental health and income
+% THEREFORE: we did this really novel study.
+
 This work builds on existing work by \citep{baylis_weather_2018}. 
 Consider going through our paper and using the ABT method for ensuring we are building up a logic string, arguments, why we're doing this, what it answers, etc.
+
+Research has demonstrated strong linkages between ambient environmental conditions and mental health, with extreme temperatures more frequently associated with poorer overall mental health.
+Mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less capable of finding help for coping with these stresses.
+Additionally, results from research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves.
+Furthermore, previous studies examining relationships between environmental extremes and health sometimes aggregate data at scales too coarse to examine whether income may be an ameliorating factor.
+Yet we know relatively little about the interaction between temperature, mental health and income.
+This research examines the interactions between these three using relatively fine-scale, publicly available data.
+This study expands on existing temperature and sentiment - one measure of mental health - research by examining finer-scale data and incorporating income data to examine correlations between extreme temperatures and sentiment.
+
+\section{Background}
 
@@
 Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.
-
-Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes .
-Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
-(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
-
-[from our Jan/Feb 2020 draft working manuscript in Google Docs]
-Climate change is a universal issue as it affects every country in the world.
-Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
-Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
-(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
 
@@
 According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions.
-For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link} demonstrated that many people living in a state with higher mean temperature do express more positive emotions on Twitter than those living in colder states in the US.
+For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature do express more positive emotions on Twitter than those living in colder states in the US.
 Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage global warming receives in the media in countries like USA, Australia and Germany.
@@
 Building on previous findings from Baylis and others, we hypothesize that local ambient weather conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1].
-Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas [H2].
+Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas.
+That is to say, people in lower income areas will be have lower sentiment scores than those in wealthier areas.
+[H2].
 Third, we hypothesize that people living in areas with less impervious surfaces and more vegetative tree cover will exhibit higher sentiment scores [H3].
@@
 \section{Data}
-
+A proposed list of tables and figures:
+Figures for location of USA Tweets (to show aggregation and coverage).
+Figures of income blocks and temperature extremes.
+Table 1. Variable descriptions for what's included in our models.
+Table 2. Descriptive statistics for datasets/variables we include.
+Table 3. Model results with variables included in the models coefficients, SEs, p-values, sample size.
+Appendix table. Sentiment scores across all types we considered.
 \begin{figure}[H]
@@
 \section{Methods}
-\subsection{Text Data Collection}
+\subsection{Text Data}
 Text data used for this project come from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose - which provides public access to a random 10\% of all public messages on Twitter.
@@
 We only considered users’ original content, thus did not include retweets in our analysis.
-
-\subsection{Analytic Approach}
-To assess whether temperature affects sentiment, we ran a linear regression model that included meteorological measures and sentiment.
-Independent variables included maximum daily temperature, precipitation and heat index.
-We also included fixed effects indicator variables at the year-month and decimal degree levels to account for potential biases in unobserved characteristics associated with seasonality and date.
-We examined the relationships between meteorological variables and sentiment using linear regression models.
-
-\subsection{Data Processing}
-We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 2019) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
-Sentiment data was analyzed using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
-We extracted meteorological data using the geolocated coordinates from the Twitter posts (latitude and longitude).
-We calculated mean climatological scores (z-scores)using baseline meteorological data from 1990 to present, calculated at a quarter degree resolution, and based on the long-term averages and standard deviation.
-We calculated z-scores for each day and the overall annual average z-score.
-For text data cleaning we used the python package tweet-preprocessor \href{https://pypi.org/project/tweet-preprocessor/}{(link)} to remove all text and characters not part of the main text phrase (e.g., URLs, mentions, GIFs, etc.).
-We converted all dates and times to UTC format.
 
@@
 
-\subsection{Income \& Race}
+\subsection{Income}
 We used data from the US Census at the Census Block level to estimate the
@@
 
+\subsection{Analytic Approach}
+To assess whether temperature affects sentiment, we ran a linear regression model that included meteorological measures and sentiment.
+Independent variables included maximum daily temperature, precipitation and heat index.
+We also included fixed effects indicator variables at the year-month and decimal degree levels to account for potential biases in unobserved characteristics associated with seasonality and date.
+We examined the relationships between meteorological variables and sentiment using linear regression models.
+
+\subsection{Data Processing}
+We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 2019) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
+Sentiment data was analyzed using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
+We extracted meteorological data using the geolocated coordinates from the Twitter posts (latitude and longitude).
+We calculated mean climatological scores (z-scores)using baseline meteorological data from 1990 to present, calculated at a quarter degree resolution, and based on the long-term averages and standard deviation.
+We calculated z-scores for each day and the overall annual average z-score.
+For text data cleaning we used the python package tweet-preprocessor \href{https://pypi.org/project/tweet-preprocessor/}{(link)} to remove all text and characters not part of the main text phrase (e.g., URLs, mentions, GIFs, etc.).
+We converted all dates and times to UTC format.
+
 \section{Results}
@@
 
+From Intro/Background
+Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes .
+Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
+(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
+
+[from our Jan/Feb 2020 draft working manuscript in Google Docs]
+Climate change is a universal issue as it affects every country in the world.
+Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
+Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
+(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
+
 Dehghan et al.

### 2020-12-17 model/fixest_income_cont.R
play with different models

@@ -46,9 +46,9 @@ formula <- paste0("vader ~ ",
                      piece.formula("temp.hi", knots[['temp.hi']], "income_percap"), ' + ',
                      piece.formula("precip", knots[['precip']], "income_percap"), ' + ',
                      piece.formula("srad", knots[['srad']], "income_percap"),
-                      ' + ', piece.formula("temp.hi", knots[['temp.hi']], ""), ' + ',
-                     piece.formula("precip", knots[['precip']], ""), ' + ',
-                     piece.formula("srad", knots[['srad']], ""), 
+                     # ' + ', piece.formula("temp.hi", knots[['temp.hi']], ""), ' + ',
+                     #piece.formula("precip", knots[['precip']], ""), ' + ',
+                     #piece.formula("srad", knots[['srad']], ""), 
                      " | dow + doy + tod + fips + year + statemonth")
 
 start <- Sys.time()
@@ -87,11 +87,14 @@ preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
 preddf$vader <- 1
 
 mm <- model.matrix(as.formula(paste0("vader ~ ",
-                   piece.formula("temp.hi", knots[['temp.hi']], ""), ' + ',
+                   #piece.formula("temp.hi", knots[['temp.hi']], ""), ' + ',
                    piece.formula("temp.hi", knots[['temp.hi']], "income_percap"))),
                    data=preddf)
 
 mm <- mm[ , colnames(mm) %in% names(coef(mod))]
+
+mm <- mm[ , grepl('temp.hi', colnames(mm)) & grepl('income_percap', colnames(mm))]
+mm <- mm[ , colnames(mm) != 'income_percap']
 
 mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
 

### 2020-12-17 model/fixest_income_hypothesis.R

@@ -11,16 +11,18 @@ options(scipen=100)
 setwd('~/tweets/')
 
 data <- fread('all.csv')
-data$income_percap <- log(data$income_percap)
-
 data <- data[weather_term == 0, ]
 
+data$income_percap <- log(data$income_percap)
 qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
 
-formula <- paste0("vader ~ income_percap*temp + income_percap*I(temp^2) + 
-                  income_percap*precip + income_percap*I(precip^2) + 
-                  income_percap*srad + income_percap*I(srad^2)", 
-                  " | dow + doy + tod + fips + year + statemonth")
+data$temp.bin <- cut(data$temp, breaks=c(-100, 5, 25, 100), labels=c("cold", "mild", "hot"))
+data$precip.bin <- cut(data$precip, breaks=c(-100, 0.000001, 1000), labels=c("norain", "rain"))
+data$srad.bin <- cut(data$srad, breaks=c(-100, 0.000001, 750, 1500), labels=c("dark", "cloudy", "sunny"))
+
+formula <- paste0("vader ~ income_percap*temp.bin + 
+                  income_percap*precip.bin + 
+                  income_percap*srad.bin | dow + doy + tod + fips + year + statemonth")
 
 start <- Sys.time()
 mod <- feols(as.formula(formula), data)
@@ -38,15 +40,16 @@ make_groups <- function(df, label, values){
 ######################################
 #temp
 
-preddf <- data.frame(temp=seq(min(data$temp), max(data$temp), length.out=250))
+preddf <- data.frame(temp.bin=c('cold', 'mild', 'hot'))
 preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
 
 preddf$vader <- 1
 
-mm <- model.matrix(vader ~ income_percap*temp + income_percap*I(temp^2),
+mm <- model.matrix(vader ~ income_percap*temp.bin,
                    data=preddf)
 
 mm <- mm[ , colnames(mm) %in% names(coef(mod))]
+mm <- mm[ , !colnames(mm) %in% 'income_percap']
 
 mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
 
@@ -57,135 +60,8 @@ preddf$ymin <- preddf$contrast - 1.96*preddf$SE
 preddf$ymax <- preddf$contrast + 1.96*preddf$SE
 
 preddf$income_percap <- factor(preddf$income_percap)
[…]

### 2021-01-06 viz/viz_bootstrap.R
new figure for manuscript

@@ -20,16 +20,8 @@ knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25,
              "prcp"= c(min(data$prcp), 0.000001, 0.05, 0.5, max(data$prcp)),
              "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
 
-#Tempknots
-knots = list("wbgt"= c(-30, -10, 0, 5, 10, 15, 20, 25, 
-                          30),
-             "prcp"= c(0, 0.000001, 0.05, 0.5, 10),
-             "srad"= c(0, 0.000001, 250, 500, 1000, 1800))
-
-
 ######################################
 #wbgt
-
 piece.formula <- function(var.name, knots, interact_var='') {
   knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
@@ -72,7 +64,21 @@ make_groups <- function(df, label, values){
   final
 }
 
-preddf <- data.frame(wbgt=fill_knots(knots$wbgt))
+#######################################
+# Get plot accessories
+#######################################
+
+ref_temps <- NULL
+for (t in seq(-20, 30, by=10)){
+  ref_temps <- c(ref_temps, median(data$temp[data$wbgt > (t - 0.25) & data$wbgt < (t + 0.25)]))
+}
+ref_temps <- round(ref_temps)
+
+
+########################################
+# WBGT
+#########################################
+preddf <- data.frame(wbgt=seq(-22, 33))
 preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
 
 preddf$vader <- 1
@@ -84,14 +90,13 @@ mm <- model.matrix(as.formula(paste0("vader ~ ",
 
 colnames(mm)[colnames(mm) == 'wbgt:income_percap'] <- 'income_percap:wbgt'
 mm <- mm[ , colnames(mm) != '(Intercept)']
-mm <- mm[ , colnames(mm) != 'income_percap']
+#mm <- mm[ , colnames(mm) != 'income_percap']
 
 mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
 
 
 #Iterate over models
 mods <- list.files()
-
 for (modf in mods){
 
   mod <- readRDS(modf)
@@ -116,21 +121,52 @@ levels(preddf$income_percap) <- paste0(levels(preddf$income_percap), ' (',
 
[…]

### 2021-01-06 doc/temp-sentiment.tex

@@
 %\author[1, *]{Matthew Cooper}
+%\author[]{Jeremiah Osborne-Gowey}
+%\author[]{Zheng Liu}
+%\author[]{Jie Liu}
+%\author[]{Portia Adade Williams}
+%\author[]{Aaron Schwartz}
+%\author[]{Patrick Baylis}
 
@@
 This work builds on existing work by \citep{baylis_weather_2018}.
-Consider going through our paper and using the ABT method for ensuring we are building up a logic string, arguments, why we're doing this, what it answers, etc.
 
+%Intro Paragraph
 Research has demonstrated strong linkages between ambient environmental conditions and mental health, with extreme temperatures more frequently associated with poorer overall mental health.
@@
 
-\section{Background}
-
 Human moods and mental states are important aspects of overall well-being and can be influenced by ambient and persistent environmental conditions.
@@
 Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.
+
+\section{Background}
+% Heatwaves are known to worsen mental health
+https://doi.org/10.1016/j.puhe.2018.06.008 (review study - will have more to cite)
+https://www.nature.com/articles/s41558-018-0222-x
+https://ehp.niehs.nih.gov/doi/full/10.1289/ehp.11339
+
+% AND mental health is strongly associated with income
+https://jamanetwork.com/journals/jamapsychiatry/fullarticle/211213
+https://doi.org/10.1002/9781118410868.wbehibs570
+https://www.sciencedirect.com/science/article/abs/pii/S027795362030527X
+https://www.sciencedirect.com/science/article/pii/S2468266717300117
+https://doi.org/10.1016/S2468-2667(17)30011-7
+https://link.springer.com/article/10.1007/s00127-017-1370-4
+
+% AND wealthier better able to cope with heatwaves 
+https://www.nature.com/articles/nclimate3253
+https://www.sciencedirect.com/science/article/abs/pii/S0378778818321327
+
+% BUT we know little about the interaction between heatwaves, mental health and income
+
+% THEREFORE: we did this really novel study.
+
+%OTHER LEFTOVER TEXT TO MAYBE BRING IN LATER
 
@@
 [H2].
-Third, we hypothesize that people living in areas with less impervious surfaces and more vegetative tree cover will exhibit higher sentiment scores [H3].
 

### 2021-01-06 doc/temp-sentiment.tex
Updates from Overleaf

@@
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
+  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
+For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
+For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
+  \label{fig:wbgt-income}
+\end{figure}
+
 Figure of heatmap of tweet density across USA (pick a year? entire dataset of georeferenced tweets?)

### 2021-01-06 model/fixest_race_cont.R
check role of race

@@ -11,11 +11,10 @@ options(scipen=100)
 setwd('~/tweets/')
 
 data <- fread('all.csv')
-data$income_percap <- log(data$income_percap)
 
 data <- data[weather_term == 0, ]
 
-qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+qs <- quantile(data$race_white, seq(0, 1, by=0.05))
 #weekdaymeans <- data[ , .(Mean=mean(vader)), .(dow)]
 #max(weekdaymeans$Mean) - min(weekdaymeans$Mean)
 
@@ -43,9 +42,9 @@ piece.formula <- function(var.name, knots, interact_var='') {
 }
 
 formula <- paste0("vader ~ ", 
-                     piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
-                     piece.formula("prcp", knots[['prcp']], "income_percap"), ' + ',
-                     piece.formula("srad", knots[['srad']], "income_percap"), ' + ',
+                     piece.formula("wbgt", knots[['wbgt']], "race_white"), ' + ',
+                     piece.formula("prcp", knots[['prcp']], "race_white"), ' + ',
+                     piece.formula("srad", knots[['srad']], "race_white"), ' + ',
                      piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
                      piece.formula("prcp", knots[['prcp']], ""), ' + ',
                      piece.formula("srad", knots[['srad']], ""), 
@@ -54,6 +53,8 @@ formula <- paste0("vader ~ ",
 start <- Sys.time()
 mod <- feols(as.formula(formula), data)
 end <- Sys.time()
+
+system('~/telegram.sh "Donezo"')
 
 fill_knots <- function(x, filln=10){
   vals <- NULL
@@ -82,20 +83,20 @@ make_groups <- function(df, label, values){
 #wbgt
 
 preddf <- data.frame(wbgt=fill_knots(knots$wbgt))
-preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
+preddf <- make_groups(preddf, 'race_white', qs[c(2, 11, 20)])
 
 preddf$vader <- 1
 
 mm <- model.matrix(as.formula(paste0("vader ~ ",
                    piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                   piece.formula("wbgt", knots[['wbgt']], "income_percap"))),
[…]

### 2021-01-06 viz/viz_bootstrap.R

@@ -90,7 +90,7 @@ mm <- model.matrix(as.formula(paste0("vader ~ ",
 
 colnames(mm)[colnames(mm) == 'wbgt:income_percap'] <- 'income_percap:wbgt'
 mm <- mm[ , colnames(mm) != '(Intercept)']
-#mm <- mm[ , colnames(mm) != 'income_percap']
+mm <- mm[ , colnames(mm) != 'income_percap']
 
 mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
 

### 2021-01-06 viz/viz_bootstrap.R
start to visualize precip and hrad from the bootstrap

@@ -170,3 +170,202 @@ xl <- get_plot_component(x2, "xlab-b")
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
 ggsave('~/temp-sentiment/res/wbgt-income.png', width=6, height=5)
+
+########################################
+# Precipitation
+#########################################
+preddf <- data.frame(prcp=seq(0, 0.1, 0.001))
+preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
+
+preddf$vader <- 1
+
+mm <- model.matrix(as.formula(paste0("vader ~ ",
+                   piece.formula("prcp", knots[['prcp']], ""), ' + ',
+                   piece.formula("prcp", knots[['prcp']], "income_percap"))),
+                   data=preddf)
+
+mm <- mm[ , colnames(mm) != '(Intercept)']
+mm <- mm[ , colnames(mm) != 'income_percap']
+
+#Iterate over models
+for (modf in mods){
+
+  mod <- readRDS(modf)
+
+  #It looks like some precip columns were dropped bc of collinearity
+  #AND the columns that were dropped varied depending on the bootstrap iteration
+  #So, we must subset the columns differently for each model
+  mms <- mm[ , colnames(mm) %in% names(coef(mod))]
+  mmp <- lapply(seq_len(nrow(mms)), function(i) mms[i,])
+
+  res <- as.data.frame(svycontrast(mod, mmp))[ , 'contrast', drop=F]
+  names(res) <- modf
+  
+  preddf <- cbind(preddf, res)
+}
+
+preddf$income_percap <- factor(preddf$income_percap)
+levels(preddf$income_percap) <- preddf$income_percap %>%
+                                   levels %>%
+                                   as.numeric %>%
+                                   exp %>%
+                                   round(0) %>%
+                                   format(big.mark = ',') %>%
+                                   paste0('$', .)
+
+levels(preddf$income_percap) <- paste0(levels(preddf$income_percap), ' (', 
+                                       names(qs[c(2, 11, 20)]), ')')
+
+preddf <- preddf %>%
+  gather(key, value, -prcp, -income_percap, -vader) %>%
[…]

### 2021-01-06 viz/viz_bootstrap.R
update bootstrap model visualizations

@@ -21,7 +21,8 @@ knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25,
              "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
 
 ######################################
-#wbgt
+# Define Functions
+#####################################
 piece.formula <- function(var.name, knots, interact_var='') {
   knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
@@ -174,7 +175,7 @@ ggsave('~/temp-sentiment/res/wbgt-income.png', width=6, height=5)
 ########################################
 # Precipitation
 #########################################
-preddf <- data.frame(prcp=seq(0, 0.1, 0.001))
+preddf <- data.frame(prcp=seq(0, 3.2, 0.01))
 preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
 
 preddf$vader <- 1
@@ -187,6 +188,7 @@ mm <- model.matrix(as.formula(paste0("vader ~ ",
 mm <- mm[ , colnames(mm) != '(Intercept)']
 mm <- mm[ , colnames(mm) != 'income_percap']
 
+mods <- list.files()
 #Iterate over models
 for (modf in mods){
 
@@ -216,6 +218,7 @@ levels(preddf$income_percap) <- preddf$income_percap %>%
 levels(preddf$income_percap) <- paste0(levels(preddf$income_percap), ' (', 
                                        names(qs[c(2, 11, 20)]), ')')
 
+
 preddf <- preddf %>%
   gather(key, value, -prcp, -income_percap, -vader) %>%
   #Normalize so that graph shows difference from prcp = X
@@ -227,53 +230,48 @@ preddf <- preddf %>%
             ymax = quantile(value, probs=0.975),
             pred = quantile(value, probs=0.5))
 
-curve <- ggplot(preddf) + 
+preddf$prcp <- log(1 + preddf$prcp*1000)
+data$prcp_l <- log(1 + data$prcp*1000)
+preddf$ymin[preddf$ymin < -0.41] <- -0.41
+
+(curve <- ggplot(preddf) + 
   geom_line(aes(x=prcp, y=pred, color=income_percap)) + 
-  geom_ribbon(aes(x=prcp, ymin=ymin, ymax=ymax, fill=income_percap), alpha=0.5) + 
[…]

### 2021-02-03 model/fixest_income_race_cont_bootstrap.R
new res

@@ -1,8 +1,14 @@
+#Use E32
+
 library(data.table)
 library(fixest)
+library(survey)
 library(tidyverse)
+library(cowplot)
 
-MOD_RUN <- 'income_percap'
+options(scipen=100)
+
+MOD_RUN <- 'race_income'
 
 setwd('~/tweets/')
 
@@ -11,12 +17,11 @@ data$income_percap <- log(data$income_percap)
 
 data <- data[weather_term == 0, ]
 
-data$raining <- data$prcp > 0
-
 #Tempknots
 knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
                           max(data$wbgt)))
 
+data$raining <- data$prcp > 0
 
 #Define Segmenting Function
 piece.formula <- function(var.name, knots, interact_var='') {
@@ -37,10 +42,13 @@ piece.formula <- function(var.name, knots, interact_var='') {
 formula <- paste0("vader ~ ", 
                      piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
                      piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     'income_percap*raining + income_percap*srad',
+                     piece.formula("wbgt", knots[['wbgt']], "race_white"), ' + ',
+                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
+                     'race_white*income_percap*raining + race_white*srad*income_percap',
                      " | dow + doy + tod + fips + year + statemonth")
 
-for (i in 2:80){
+
+for (i in 1:80){
   mod <- feols(as.formula(formula), data[sample(1:nrow(data), nrow(data), replace=T), ])
   cf <- coef(mod)
   vc <- vcov(mod)
@@ -54,8 +62,4 @@ for (i in 2:80){
   rm(list=c('mod', 'cf', 'vc', 'myobj'))
   gc()
 }
-
-
-system('sudo poweroff')
-
 

### 2021-02-03 model/fixest_race_bin_bootstrap.R

@@ -1,13 +1,13 @@
 library(data.table)
 library(fixest)
 library(tidyverse)
+library(Hmisc)
 
-MOD_RUN <- 'income_percap'
+MOD_RUN <- 'race_q'
 
 setwd('~/tweets/')
 
 data <- fread('all.csv')
-data$income_percap <- log(data$income_percap)
 
 data <- data[weather_term == 0, ]
 
@@ -17,30 +17,38 @@ data$raining <- data$prcp > 0
 knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
                           max(data$wbgt)))
 
-
 #Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='') {
+# DIFFERENT FOR GROUPING VARS
+piece.formula <- function(var.name, knots, interact_var='', interact_type=c('factor', 'continuous', 'none')) {
   knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
   formula.sign[knots < 0] <- " + "
-  if (interact_var == ''){
-    paste(var.name, "+",
+  if (interact_type == 'none'){
+    formula <- paste(var.name, "+",
         paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
               collapse = " + ", sep=""))
-  } else{
-    paste(interact_var, '*', var.name, "+",
-        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
   }
+  if (interact_type == 'continuous'){
+    formula <- paste(interact_var, '*', var.name, "+",
+          paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
+                collapse = " + ", sep=""))
+  } 
+  if (interact_type == 'factor'){
+    formula <- paste(interact_var, '*', var.name, "+",
+          paste(interact_var, "*I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+                collapse = " + ", sep=""))
+  }
+  formula
 }
 
+
 formula <- paste0("vader ~ ", 
-                     piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     'income_percap*raining + income_percap*srad',
[…]

### 2021-02-03 viz/wbgt_race_bootstrap.R

@@ -1,0 +1,174 @@
+library(tidyverse)
+library(cowplot)
+library(survey)
+library(data.table)
+
+MOD_RUN <- 'race_cont'
+
+options(scipen=100)
+
+setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
+
+data <- fread('~/tweets/all.csv')
+
+data <- data[weather_term == 0, ]
+
+data$raining <- data$prcp > 0
+
+qs <- quantile(data$race_white, seq(0, 1, by=0.05))
+
+#Tempknots
+knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
+                          max(data$wbgt)))
+
+######################################
+# Define Functions
+#####################################
+piece.formula <- function(var.name, knots, interact_var='') {
+  knots <- knots[c(-1, -length(knots))]
+  formula.sign <- rep(" - ", length(knots))
+  formula.sign[knots < 0] <- " + "
+  if (interact_var == ''){
+    paste(var.name, "+",
+        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
+  } else{
+    paste(interact_var, '*', var.name, "+",
+        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
+  }
+}
+
+vcov.bootmod <- function(object){
+  return(object$vcov)
+}
+coef.bootmod <- function(object){
+  return(object$coef)
+}
+
+
+fill_knots <- function(x, filln=10){
+  vals <- NULL
+  for (i in 2:length(x)){
+    vals <- c(vals, seq(x[i-1], x[i], length.out=filln))
+  }
+  vals <- unique(vals)
+  vals <- vals[order(vals)]
+  vals
+}
+
+make_groups <- function(df, label, values){
+  final <- data.frame()
+  for (v in values){
+    df[ , label] <- v
+    final <- bind_rows(df, final)
+  }
+  final
+}
+
+#######################################
+# Get plot accessories
+#######################################
+
+ref_temps <- NULL
+for (t in seq(-20, 30, by=10)){
+  ref_temps <- c(ref_temps, median(data$temp[data$wbgt > (t - 0.25) & data$wbgt < (t + 0.25)]))
+}
+ref_temps <- round(ref_temps)
+
+
[…]

### 2021-02-03 viz/wbgt_race_q_bootstrap.R

@@ -1,0 +1,171 @@
+library(tidyverse)
+library(cowplot)
+library(survey)
+library(data.table)
+
+MOD_RUN <- 'race_q'
+
+options(scipen=100)
+
+setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
+
+data <- fread('~/tweets/all.csv')
+
+data <- data[weather_term == 0, ]
+
+data$raining <- data$prcp > 0
+
+#Tempknots
+knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
+                          max(data$wbgt)))
+
+######################################
+# Define Functions
+#####################################
+# DIFFERENT FOR GROUPING VARS
+piece.formula <- function(var.name, knots, interact_var='', interact_type=c('factor', 'continuous', 'none')) {
+  knots <- knots[c(-1, -length(knots))]
+  formula.sign <- rep(" - ", length(knots))
+  formula.sign[knots < 0] <- " + "
+  if (interact_type == 'none'){
+    formula <- paste(var.name, "+",
+        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
+  }
+  if (interact_type == 'continuous'){
+    formula <- paste(interact_var, '*', var.name, "+",
+          paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
+                collapse = " + ", sep=""))
+  } 
+  if (interact_type == 'factor'){
+    formula <- paste(interact_var, '*', var.name, "+",
+          paste(interact_var, "*I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+                collapse = " + ", sep=""))
+  }
+  formula
+}
+
+vcov.bootmod <- function(object){
+  return(object$vcov)
+}
+coef.bootmod <- function(object){
+  return(object$coef)
+}
+
+
+fill_knots <- function(x, filln=10){
+  vals <- NULL
+  for (i in 2:length(x)){
+    vals <- c(vals, seq(x[i-1], x[i], length.out=filln))
+  }
+  vals <- unique(vals)
+  vals <- vals[order(vals)]
+  vals
+}
+
+make_groups <- function(df, label, values){
+  final <- data.frame()
+  for (v in values){
+    df[ , label] <- v
[…]

### 2021-02-03 model/fixest_income_race_bin_bootstrap.R
run income-race_q bootstrap; clean results

@@ -1,0 +1,76 @@
+#Use E32
+
+library(data.table)
+library(fixest)
+library(survey)
+library(tidyverse)
+library(cowplot)
+
+options(scipen=100)
+
+MOD_RUN <- 'race_income_q'
+
+setwd(paste0('~/tweets/bootstrap/', MOD_RUN))
+
+data <- fread('~/tweets/all.csv')
+
+data$income_q <- Hmisc::cut2(data$income_percap, g=3)
+
+levels(data$income_q) <- c('inc_low', 'inc_mid', 'inc_hi')
+
+data$income_race_q <- paste0(data$income_q, '-', data$race_majority)
+
+data <- data[weather_term == 0, ]
+
+#Tempknots
+knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
+                          max(data$wbgt)))
+
+data$raining <- data$prcp > 0
+
+#Define Segmenting Function
+# DIFFERENT FOR GROUPING VARS
+piece.formula <- function(var.name, knots, interact_var='', interact_type=c('factor', 'continuous', 'none')) {
+  knots <- knots[c(-1, -length(knots))]
+  formula.sign <- rep(" - ", length(knots))
+  formula.sign[knots < 0] <- " + "
+  if (interact_type == 'none'){
+    formula <- paste(var.name, "+",
+        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
+  }
+  if (interact_type == 'continuous'){
+    formula <- paste(interact_var, '*', var.name, "+",
+          paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
+                collapse = " + ", sep=""))
+  } 
+  if (interact_type == 'factor'){
+    formula <- paste(interact_var, '*', var.name, "+",
+          paste(interact_var, "*I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+                collapse = " + ", sep=""))
+  }
+  formula
+}
+
+
+formula <- paste0("vader ~ ", 
+                     piece.formula("wbgt", knots[['wbgt']], "income_race_q", 'factor'), ' + ',
+                     piece.formula("wbgt", knots[['wbgt']], "", 'none'), ' + ',
+                     'income_race_q*raining + income_race_q*srad',
[…]

### 2021-02-03 model/fixest_income_race_cont.R

@@ -1,224 +1,0 @@
-#Use E32
-
-library(data.table)
-library(fixest)
-library(survey)
-library(tidyverse)
-library(cowplot)
-
-options(scipen=100)
-
-setwd('~/tweets/')
-
-data <- fread('all.csv')
-data$income_percap <- log(data$income_percap)
-
-data <- data[weather_term == 0, ]
-
-qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
-rqs <- quantile(data$race_white, seq(0, 1, by=0.05))
-#weekdaymeans <- data[ , .(Mean=mean(vader)), .(dow)]
-#max(weekdaymeans$Mean) - min(weekdaymeans$Mean)
-
-#Tempknots
-knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
-                          max(data$wbgt)))
-
-data$raining <- data$prcp > 0
-
-#Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='') {
-  knots <- knots[c(-1, -length(knots))]
-  formula.sign <- rep(" - ", length(knots))
-  formula.sign[knots < 0] <- " + "
-  if (interact_var == ''){
-    paste(var.name, "+",
-        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
-  } else{
-    paste(interact_var, '*', var.name, "+",
-        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
-  }
-}
-
-formula <- paste0("vader ~ ", 
-                     piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], "race_white"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     'income_percap*raining + srad*income_percap',
-                     " | dow + doy + tod + fips + year + statemonth")
-
-start <- Sys.time()
-mod <- feols(as.formula(formula), data)
-system('~/telegram.sh "Done with Model"')
-end <- Sys.time()
-
-fill_knots <- function(x, filln=10){
-  vals <- NULL
-  for (i in 2:length(x)){
[…]

### 2021-02-03 model/fixest_nointeract_quad.R

@@ -1,133 +1,0 @@
-#Use E32
-
-library(data.table)
-library(fixest)
-library(survey)
-library(tidyverse)
-library(cowplot)
-
-options(scipen=100)
-
-setwd('~/tweets/')
-
-data <- fread('all.csv')
-
-data <- data[weather_term == 0, ]
-
-formula <- paste0("vader ~ temp + temp^2 + precip + precip^2 + srad + srad^2", 
-                     " | dow + doy + tod + fips + year + statemonth")
-
-start <- Sys.time()
-mod <- feols(as.formula(formula), data)
-end <- Sys.time()
-
-######################################
-#temp
-
-preddf <- data.frame(temp=seq(min(data$temp), max(data$temp), length.out=250))
-
-preddf$vader <- 1
-
-mm <- model.matrix(vader ~ temp + I(temp^2),
-                   data=preddf)
-
-mm <- mm[ , colnames(mm) %in% names(coef(mod))]
-
-mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
-
-res <- svycontrast(mod, mmp)
-
-preddf <- cbind(preddf, as.data.frame(res))
-preddf$ymin <- preddf$contrast - 1.96*preddf$SE
-preddf$ymax <- preddf$contrast + 1.96*preddf$SE
-
-curve <- ggplot(preddf) + 
-  geom_line(aes(x=temp, y=contrast)) + 
-  geom_ribbon(aes(x=temp, ymin=ymin, ymax=ymax), alpha=0.5) + 
-  scale_x_continuous(expand=c(0, 0), limits=c(-21.4, 43.1)) + 
-  labs(x='', y='Predicted Mood of Tweet') +
-  theme(legend.position=c(0.2, 0.2))
-
-hist <- ggplot(data[sample(1:nrow(data), nrow(data)*0.01), 'temp']) +
-  geom_histogram(aes(x=temp), binwidth=1) + 
-  scale_x_continuous(expand=c(0, 0), limits=c(-21.38222, 43.02777)) + 
-  scale_y_continuous(labels=function(x){format(x*100, big.mark=',')}) + 
-  labs(y="Tweet Count", x="Temperature")
-
-plot_grid(curve, hist, align='v', axis='rl', ncol=1, rel_heights=c(0.8, 0.2))
-ggsave('~/temp-sentiment/res/quad-temp-only.png', width=6, height=7)
-
[…]

### 2021-02-03 model/fixest_race_cont.R

@@ -1,134 +1,0 @@
-#Use E32
-
-library(data.table)
-library(fixest)
-library(survey)
-library(tidyverse)
-library(cowplot)
-
-options(scipen=100)
-
-setwd('~/tweets/')
-
-data <- fread('all.csv')
-
-data <- data[weather_term == 0, ]
-
-qs <- quantile(data$race_white, seq(0, 1, by=0.05))
-#weekdaymeans <- data[ , .(Mean=mean(vader)), .(dow)]
-#max(weekdaymeans$Mean) - min(weekdaymeans$Mean)
-
-#Tempknots
-knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
-                          max(data$wbgt)),
-             "prcp"= c(min(data$prcp), 0.000001, 0.05, 0.5, max(data$prcp)),
-             "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
-
-
-#Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='') {
-  knots <- knots[c(-1, -length(knots))]
-  formula.sign <- rep(" - ", length(knots))
-  formula.sign[knots < 0] <- " + "
-  if (interact_var == ''){
-    paste(var.name, "+",
-        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
-  } else{
-    paste(interact_var, '*', var.name, "+",
-        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
-  }
-}
-
-formula <- paste0("vader ~ ", 
-                     piece.formula("wbgt", knots[['wbgt']], "race_white"), ' + ',
-                     piece.formula("prcp", knots[['prcp']], "race_white"), ' + ',
-                     piece.formula("srad", knots[['srad']], "race_white"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     piece.formula("prcp", knots[['prcp']], ""), ' + ',
-                     piece.formula("srad", knots[['srad']], ""), 
-                     " | dow + doy + tod + fips + year + statemonth")
-
-start <- Sys.time()
-mod <- feols(as.formula(formula), data)
-end <- Sys.time()
-
[…]

### 2021-02-03 model/fixest_race_group.R

@@ -1,198 +1,0 @@
-##Use E64
-
-library(data.table)
-library(fixest)
-library(survey)
-library(tidyverse)
-library(cowplot)
-library(Hmisc)
-
-options(scipen=100)
-
-setwd('~/tweets/')
-
-data <- fread('all.csv')
-
-data$race <- factor(ifelse(data$majority == 'race_white', 'white', 'minority'))
-qs <- levels(data$race)
-#weekdaymeans <- data[ , .(Mean=mean(vader)), .(dow)]
-
-#Tempknots
-#Tempknots
-knots = list("temp"= c(min(data$temp), -10, 0, 5, 10, 15, 20, 25, 30, 35, 
-                          max(data$temp)),
-             "precip"= c(min(data$precip), 0.000001, 0.05, 0.5, max(data$precip)),
-             "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
-
-#Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='', interact_type=c('factor', 'continuous')) {
-  knots <- knots[c(-1, -length(knots))]
-  formula.sign <- rep(" - ", length(knots))
-  formula.sign[knots < 0] <- " + "
-  if (interact_type == 'continuous'){
-    formula <- paste(interact_var, '*', var.name, "+",
-          paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-                collapse = " + ", sep=""))
-  } else{
-    formula <- paste(interact_var, '*', var.name, "+",
-          paste(interact_var, "*I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
-                collapse = " + ", sep=""))
-  }
-  formula
-}
-
-formula <- paste0("vader ~ ", 
-                     piece.formula("temp", knots[['temp']], "race", 'factor'), ' + ',
-                     piece.formula("precip", knots[['precip']], "race", 'factor'), ' + ',
-                     piece.formula("srad", knots[['srad']], "race", 'factor'),
-                     " | dow + doy + tod + fips + year + statemonth")
-
-mod <- feols(as.formula(formula), data)
-
-fill_knots <- function(x, filln=10){
[…]

### 2021-02-03 model/fixest_tree_cont.R

@@ -1,188 +1,0 @@
-##Use E48
-
-library(data.table)
-library(fixest)
-library(survey)
-library(tidyverse)
-library(cowplot)
-
-options(scipen=100)
-
-setwd('~/tweets/')
-
-data <- fread('all.csv')
-qs <- quantile(data$tree, seq(0, 1, by=0.05))
-weekdaymeans <- data[ , .(Mean=mean(vader)), .(dow)]
-
-#Tempknots
-#Tempknots
-knots = list("temp.hi"= c(min(data$temp.hi), -10, 0, 5, 10, 15, 20, 25, 30, 35, 
-                          max(data$temp.hi)),
-             "precip"= c(min(data$precip), 0.000001, 0.05, 0.5, max(data$precip)),
-             "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
-
-#Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='') {
-  knots <- knots[c(-1, -length(knots))]
-  formula.sign <- rep(" - ", length(knots))
-  formula.sign[knots < 0] <- " + "
-  paste(interact_var, '*', var.name, "+",
-        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
-}
-
-formula <- paste0("vader ~ ", 
-                     piece.formula("temp.hi", knots[['temp.hi']], "tree"), ' + ',
-                     piece.formula("precip", knots[['precip']], "tree"), ' + ',
-                     piece.formula("srad", knots[['srad']], "tree"),
-                     " | dow + doy + tod + fips + year + statemonth")
-
-mod <- feols(as.formula(formula), data)
-
-fill_knots <- function(x, filln=10){
-  vals <- NULL
-  for (i in 2:length(x)){
-    vals <- c(vals, seq(x[i-1], x[i], length.out=filln))
-  }
-  vals <- unique(vals)
-  vals <- vals[order(vals)]
-  vals
-}
-
-make_groups <- function(df, label, values){
-  final <- data.frame()
-  for (v in values){
-    df[ , label] <- v
-    final <- bind_rows(df, final)
-  }
-  final
-}
-
[…]

### 2021-02-03 model/fixest_tree_group.R

@@ -1,201 +1,0 @@
-##Use E64
-
-library(data.table)
-library(fixest)
-library(survey)
-library(tidyverse)
-library(cowplot)
-library(Hmisc)
-
-options(scipen=100)
-
-setwd('~/tweets/')
-
-data <- fread('all.csv')
-data$tree_q <- cut2(data$tree, g=2)
-qs <- levels(data$tree_q)
-weekdaymeans <- data[ , .(Mean=mean(vader)), .(dow)]
-
-#Tempknots
-#Tempknots
-knots = list("temp"= c(min(data$temp), -10, 0, 5, 10, 15, 20, 25, 30, 35, 
-                          max(data$temp)),
-             "precip"= c(min(data$precip), 0.000001, 0.05, 0.5, max(data$precip)),
-             "srad"= c(min(data$srad), 0.000001, 250, 500, 1000, max(data$srad)))
-
-#Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='', interact_type=c('factor', 'continuous')) {
-  knots <- knots[c(-1, -length(knots))]
-  formula.sign <- rep(" - ", length(knots))
-  formula.sign[knots < 0] <- " + "
-  if (interact_type == 'continuous'){
-    formula <- paste(interact_var, '*', var.name, "+",
-          paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-                collapse = " + ", sep=""))
-  } else{
-    formula <- paste(interact_var, '*', var.name, "+",
-          paste(interact_var, "*I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
-                collapse = " + ", sep=""))
-  }
-  formula
-}
-
-formula <- paste0("vader ~ ", 
-                     piece.formula("temp", knots[['temp']], "tree_q", 'factor'), ' + ',
-                     piece.formula("precip", knots[['precip']], "tree_q", 'factor'), ' + ',
-                     piece.formula("srad", knots[['srad']], "tree_q", 'factor'),
-                     " | dow + doy + tod + fips + year + statemonth")
-
-mod <- feols(as.formula(formula), data)
-
-fill_knots <- function(x, filln=10){
[…]

### 2021-02-03 viz/wbgt_income_bootstrap.R

@@ -12,6 +12,8 @@ setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
 data <- fread('~/tweets/all.csv')
 data$income_percap <- log(data$income_percap)
 
+qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+
 data <- data[weather_term == 0, ]
 
 data$raining <- data$prcp > 0

### 2021-02-03 viz/wbgt_income_race_bootstrap.R

@@ -1,48 +1,43 @@
+library(data.table)
+library(fixest)
+library(survey)
 library(tidyverse)
 library(cowplot)
-library(survey)
-library(data.table)
-
-MOD_RUN <- 'race_q'
 
 options(scipen=100)
+
+MOD_RUN <- 'race_income'
 
 setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
 
 data <- fread('~/tweets/all.csv')
+data$income_percap <- log(data$income_percap)
 
 data <- data[weather_term == 0, ]
 
-data$raining <- data$prcp > 0
+inc_qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
 
 #Tempknots
 knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
                           max(data$wbgt)))
 
-######################################
-# Define Functions
-#####################################
-# DIFFERENT FOR GROUPING VARS
-piece.formula <- function(var.name, knots, interact_var='', interact_type=c('factor', 'continuous', 'none')) {
+data$raining <- data$prcp > 0
+
+#Define Segmenting Function
+piece.formula <- function(var.name, knots, interact_var='') {
   knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
   formula.sign[knots < 0] <- " + "
-  if (interact_type == 'none'){
-    formula <- paste(var.name, "+",
+  if (interact_var == ''){
+    paste(var.name, "+",
         paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
               collapse = " + ", sep=""))
+  } else{
+    paste(interact_var, '*', var.name, "+",
+        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
   }
-  if (interact_type == 'continuous'){
-    formula <- paste(interact_var, '*', var.name, "+",
-          paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-                collapse = " + ", sep=""))
-  } 
-  if (interact_type == 'factor'){
-    formula <- paste(interact_var, '*', var.name, "+",
-          paste(interact_var, "*I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
[…]

### 2021-02-03 model/fixest_income_race_bin_bootstrap.R
update failed model

@@ -1,5 +1,6 @@
-#Use E32
-
+######################################
+# DOESNT WORK EVEN ON E64
+#####################################
 library(data.table)
 library(fixest)
 library(survey)

### 2021-02-03 doc/temp-sentiment.tex
Updates from Overleaf

@@
 \begin{document}
-
-\title{Heatwaves Exacerbate Inequalities in Mental Health}
+https://www.overleaf.com/project/5f918edd1afc380001feca0d
+\title{Fine-scale Twitter Data Reveals Neighborhood Inequalities in Heat Wave Vulnerability}
 
@@
 
-Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation
+Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation, Suicide, Mortality, Morbidity, Self-harm,
 
@@
 %Intro Paragraph
-Research has demonstrated strong linkages between ambient environmental conditions and mental health, with extreme temperatures more frequently associated with poorer overall mental health.
-Mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less capable of finding help for coping with these stresses.
-Additionally, results from research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves.
-Furthermore, previous studies examining relationships between environmental extremes and health sometimes aggregate data at scales too coarse to examine whether income may be an ameliorating factor.
+Research has shown strong linkages between ambient environmental conditions and mental health, with extreme temperatures  frequently associated with poorer overall mental health.
+Individual's mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less able to find help for coping with these stresses.
+Other research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves, whether from traveling to cooler climes, accessing air conditioning, or other temperature ameliorating strategies that may require access to financial capital.
 Yet we know relatively little about the interaction between temperature, mental health and income.
-This research examines the interactions between these three using relatively fine-scale, publicly available data.
-This study expands on existing temperature and sentiment - one measure of mental health - research by examining finer-scale data and incorporating income data to examine correlations between extreme temperatures and sentiment.
+Furthermore, previous studies that did attempt to examine relationships between environmental extremes and health sometimes aggregated data at scales too coarse to examine whether income may be an ameliorating factor.
+This study examines the interactions between these three using relatively fine-scale, publicly available data.
+Our research expands on existing studies of temperature and sentiment - one measure of mental health - by examining finer-scale data and incorporating income data to examine correlations between temperature extremes and sentiment.
 
-Human moods and mental states are important aspects of overall well-being and can be influenced by ambient and persistent environmental conditions.
+Human moods and mental states are important aspects of overall well-being and can be influenced by ambient, persistent and fluctuating environmental conditions.
 Climate change is accelerating the rate and variability around meteorological norms like the timing, magnitude, intensity and duration of precipitation events and air temperature minima and maxima \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
 Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meterological conditions.
-Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to meteorological conditions.
+Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to temperature extremes.
 Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.
@@
 \section{Background}
+\subsection{Temperature and Mental Health}
 % Heatwaves are known to worsen mental health
@@
 https://ehp.niehs.nih.gov/doi/full/10.1289/ehp.11339
+https://ijmhs.biomedcentral.com/articles/10.1186/s13033-018-0210-6
+https://www.nature.com/articles/s41558-018-0102-4/
+https://doi.org/10.1016/j.jhealeco.2019.102240
 
+[ADD SUICIDE DATA TO HEALTH SECTION?]
+https://www.psychologytoday.com/us/blog/greening-the-media/201809/global-warming-and-suicide
+
+\subsection{Mental Health and Income}
 % AND mental health is strongly associated with income
@@
 https://link.springer.com/article/10.1007/s00127-017-1370-4
+Maybe this: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0116820
+https://www.sciencedirect.com/science/article/abs/pii/S0143622814002537
 
+\subsection{Income and Temperature}
 % AND wealthier better able to cope with heatwaves 
@@
 https://www.sciencedirect.com/science/article/abs/pii/S0378778818321327
+https://www.mdpi.com/2225-1154/8/1/12/htm
+
+NOT CLEAR: https://doi.org/10.1016/j.jhealeco.2019.102240 and https://www.nature.com/articles/s41558-018-0222-x find little evidence of adaptation
 
@@
 % THEREFORE: we did this really novel study.
+https://www.nature.com/articles/s41598-017-12961-9
 
@@
 Emotional states and well-being are associated with physiological functioning and mental acuity which can affect social relationships, workplace productivity (cite) and health risks (cite).
-People that are more reliant on livelihoods which require them to be outdoors (e.g., farming, agriculture, logging, etc.), living and working in places where they are exposed to environmental minima and maxima, or with prolonged exposure to ambient meteorological extremes are particularly vulnerable to changes in environmental exposure \citep{frimpong_heat_2017} and associated health risks (cite), with implications for workplace productivity and livelihoods \cite{kjellstrom_impact_2016}.
-For example, Nigerian maize farmers experience significant declines in productivity (2-8percent per degree above 17degC) and substantive increases in health impacts including dehydration, muscle cramps, headaches and dizziness, heat exhaustion, sun stroke, and even death \cite{sadiq_impact_2019}.
+People that are more reliant on 1) livelihoods which require them to be outdoors (e.g., farming, agriculture, logging, etc.), 2) living and working in places where they are exposed to environmental minima and maxima, or 3) with prolonged exposure to ambient meteorological extremes are particularly vulnerable to changes in environmental exposure \citep{frimpong_heat_2017} and associated health risks (cite).
+Exposure to extreme environmental conditions can also have implications for workplace productivity and livelihoods \cite{kjellstrom_impact_2016}.
+For example, Nigerian maize farmers experience significant declines in productivity (2-8percent per degree above 17degC) along with substantive increases in health impacts including dehydration, muscle cramps, headaches and dizziness, heat exhaustion, sun stroke, and even death \cite{sadiq_impact_2019}.
 Other research indicates farm health and labor productivity were compromised under extreme heat and cold events in the Nepali Food Bowl region \cite{budhathoki_socio-economic_2019}.
@@
 According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions.
-For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature do express more positive emotions on Twitter than those living in colder states in the US.
-Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage global warming receives in the media in countries like USA, Australia and Germany.
+For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature express more positive emotions on Twitter than those living in colder states in the US.
+Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage climate change receives in the media in countries like USA, Australia and Germany.
 Studies comparing how different countries portray climate change have been widely conducted in USA, UK, France and the Netherlands \cite{vu_nationalizing_2019}.
@@
 Weather events focus public attention in different ways of expression (cites).
-This offers a unique opportunity to analyze how events influence public discourse on their expressions and responses.
+These data offer a unique opportunity to analyze how people express and respond to events in people's daily lives, including their exposure to ambient environmental conditions.
 
-Here, we report on associations between meteorological conditions and expressed sentiment of public posts on Twitter from the United States of America between 2009 and 2019.
-This work builds on research from Baylis and others \cite{baylis_weather_2018} on weather impacts on expressed sentiment in the USA.
+Here, we report on associations between meteorological conditions and expressed sentiment of public posts on Twitter from the United States of America (USA) between 2009 and 2019.
+This work builds on research from Baylis and others \cite{baylis_weather_2018} on weather impacts on expressed sentiment.
 In particular, we extend this work by including the temperature effects on sentiment in multiple regions across the USA while drawing in various climatic variables and socio-demographic data.
 In this research, we explore two primary questions.
-First, to what extent are weather conditions correlated with changes in expressed sentiment?
-Second, is this effect moderated by wealth and local land cover conditions?
+First, to what extent are amibent environmental conditions correlated with changes in expressed sentiment?
+Second, is this effect moderated by wealth?
+and local land cover conditions?
 
-Building on previous findings from Baylis and others, we hypothesize that local ambient weather conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1].
+Building on previous findings from Baylis and others, we hypothesize that local ambient environmental conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1].
 Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas.
@@
 \section{Methods}
-\subsection{Text Data}
-Text data used for this project come from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose - which provides public access to a random 10\% of all public messages on Twitter.
+\subsection{Tweets}
+We used data from 243 million geolocated tweets from the lower 48 states from the years 2009 to 2019.
+
+Text data used for this project come from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
+The "decahose" provides public access to a random 10\% of all public messages on Twitter.
 The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
 The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
-We used all geolocated Tweets from the Decahose for the period 2012-2016 with a bounding box polygon filter for India, Western Europe and the United States of America.
+We used all geolocated Tweets from the Decahose for the period 2009-2019.
 We only considered users’ original content, thus did not include retweets in our analysis.
@@
 \subsection{Weather}
-We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington \href{https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2011JD016048}{link}.
-This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution \cite{xia_continental-scale_2012}.
-We extracted the data for the exact hour and location of each tweet using Google Earth Engine.
-We extracted several variables from the NLDAS-2 dataset, including temperature, specific humidity, air pressure, total precipitation, and shortwave radiation.
+We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington.
+This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution, and integrates a large quantity of observation-based and modeled data  \cite{xia_continental-scale_2012}.
+For the exact hour and location of each tweet, we extracted temperature, specific humidity, air pressure, total precipitation, shortwave radiation, and wind speed.
+
+Because metrics of apparent temperature that take into account humidity and other factors can better account for the impacts of heat stress on human health and wellbeing, we calculate the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet.
+WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature would show due to evaporative cooling, and can be estimated given normal temperature, relative humidity, solar radiation, and wind speed.
+Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing.
+Metrics like WBGT that account for the effects of humidity and other factors on heat stress have been associated with diminished economic output \cite{rao2020projections}, increased crime \cite{hu2017impact}, increased mortality \cite{chien2016spatiotemporal, armstrong2019role}, and worsened mental health outcomes \cite{vida2012relationship, ding2016importance}.
+
 Using temperature, specific humidity, and pressure, we derived relative humidity using methods described by  \cite{bolton_computation_1980}.
-Then, based on the Heat Index algorithm developed by Steadman \cite{steadman_assessment_1979} and improved upon by Rothfusz and the National Weather service \cite{rothfusz_heat_1990}, we calculated the Heat Index using temperature and relative humidity \cite{heo_comparison_2019}.
+We then calculated the WBGT using the formula described by \cite{heo2019comparison}.
 
-ISSUES: deciduous trees lose leaves every year and may not provide much shade cover and heat relief during particular periods of the year when they are leafless.
-
-\subsection{Income}
+\subsection{Socio-Econoimc Data}
 We used data from the US Census at the Census Block level to estimate the
-
-\subsection{Land Cover}
-To estimate land cover categories and the percentage of tree cover and impervious surfaces, we used data from the National Land Cover Database (NLCD) from the USGS' 2011  \cite{homer_completion_2015} and 2016 \cite{yang_new_2018} landcover datasets.
-These datasets are national land cover dataset prepared by the Multi-Resolution Land Characteristics (MRLC) Consortium.
-These data are at the thirty meter resolution, and provide estimates of the fraction of tree cover and impervious surfaces.
-We then associated each tweet with the land cover fraction for the nearest available year and pixel using the geolocation associated with each tweet, extracting landcover values from whichever land cover dataset year was closest in time to the date timestamp of each tweet.
-We used tree cover as a proxy for amount of potential shade and heat relief in a particular area and can be broadly indicative of generalized wealth.
-We use impervious surface cover as a proxy for relative proximity to potential natural shade and heat relief, and because individuals in urban environments may be exposed to the urban heat island effect.
-[NEED SOME CITATIONS HERE]
 
@@
 
-\subsection{Analytic Approach}
-To assess whether temperature affects sentiment, we ran a linear regression model that included meteorological measures and sentiment.
-Independent variables included maximum daily temperature, precipitation and heat index.
-We also included fixed effects indicator variables at the year-month and decimal degree levels to account for potential biases in unobserved characteristics associated with seasonality and date.
-We examined the relationships between meteorological variables and sentiment using linear regression models.
+\subsection{Modeling}
+We assessed how expressed sentiment is affect by three weather variables: wet bulb globe temperature, precipitation, and sunshine.
+Moreover, we modeled how neighborhood income affects the relationship between these weather variables and sentiment.
+
+Our model takes the following form:
+\begin{equation}
+    y = \beta_0 + f_t(t) + f_{mt}(m t) + \beta_p p + \beta_{mp} m s + \beta_s s + \beta_{ms} m s + \Phi + \epsilon
+\end{equation}
+
+Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, and $m$ is the log-transformed average income in the census block where the tweet originated.
+We estimate the
 
@@
 We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 2019) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
-Sentiment data was analyzed using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
+We analyized sentiment data using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
 We extracted meteorological data using the geolocated coordinates from the Twitter posts (latitude and longitude).
@@
 We converted all dates and times to UTC format.
+
+[NEED SOME TEXT ABOUT GOOGLE EARTH ENGINE AND MICROSOFT AI FOR EARTH] We conducted all data analyses on Microsoft AI for Earth's online cloud-computing environment using a combination of existing data analysis packages and scripts we developed.
+We used GitHub as a repository for all scripts we used in this project (link to the Git repo).
+To run ABC analysis, we used Microsoft's CBA package.
+To run XYZ analysis, we used Microsoft's ZYX package.
 
@@
 
+%%KEY POINTS TO DISCUSS
+
+% Link between twitter mood and other very bad outcomes (crime, suicide)
+
+% Baselines: this could also be interpreted as: poor are less affected by the cold\textbf{:}
+
+% Caveats:
+    % Uncertainties from census blocks (low income, young areas).  This is why we used a linaer effect instead of a binned effect.
+    % We cant speak to types of mental heath that are worsening - depression, anxiety, and post-traumatic stress, etc.  Just a general overview
+
 \section{Conclusion}
@@
 For funding, advising, SESYNC grant number?, etc.
+This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
 
@@
 \renewcommand{\thesection}{S\arabic{section}}
+
+\subsection{Land Cover}
+To estimate land cover categories and the percentage of tree cover and impervious surfaces, we used data from the National Land Cover Database (NLCD) from the USGS' 2011  \cite{homer_completion_2015} and 2016 \cite{yang_new_2018} landcover datasets.
+These datasets are national land cover dataset prepared by the Multi-Resolution Land Characteristics (MRLC) Consortium.
+These data are at the thirty meter resolution, and provide estimates of the fraction of tree cover and impervious surfaces.
+We then associated each tweet with the land cover fraction for the nearest available year and pixel using the geolocation associated with each tweet, extracting landcover values from whichever land cover dataset year was closest in time to the date timestamp of each tweet.
+We used tree cover as a proxy for amount of potential shade and heat relief in a particular area and can be broadly indicative of generalized wealth.
+We use impervious surface cover as a proxy for relative proximity to potential natural shade and heat relief, and because individuals in urban environments may be exposed to the urban heat island effect.
+[NEED SOME CITATIONS HERE]
 

### 2021-02-03 viz/wbgt_income_bootstrap.R
write income-race figure, standardize all to 5 degree baseline

@@ -126,7 +126,7 @@ preddf <- preddf %>%
   gather(key, value, -wbgt, -income_percap, -vader) %>%
   #Normalize so that graph shows difference from wbgt = X
   group_by(income_percap, key) %>%
-  mutate(value = value - value[wbgt == 0]) %>%
+  mutate(value = value - value[wbgt == 5]) %>%
   #Get ymin, ymax, and mean
   group_by(wbgt, income_percap) %>%
   summarize(ymin = quantile(value, probs=0.025),

### 2021-02-03 viz/wbgt_race_bootstrap.R

@@ -125,7 +125,7 @@ preddf <- preddf %>%
   gather(key, value, -wbgt, -race_white, -vader) %>%
   #Normalize so that graph shows difference from wbgt = X
   group_by(race_white, key) %>%
-  mutate(value = value - value[wbgt == 0]) %>%
+  mutate(value = value - value[wbgt == 5]) %>%
   #Get ymin, ymax, and mean
   group_by(wbgt, race_white) %>%
   summarize(ymin = quantile(value, probs=0.025),
@@ -172,3 +172,4 @@ xl <- get_plot_component(x2, "xlab-b")
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
 ggsave('~/temp-sentiment/res/wbgt-race.png', width=6, height=5)
+

### 2021-02-04 model/state_model.R
look at state-level heterogeneities

@@ -1,0 +1,297 @@
+library(data.table)
+library(tidyverse)
+library(fixest)
+library(broom)
+library(sf)
+library(spData)
+
+data <- fread('~/tweets/all.csv')
+
+data$state <- substr(data$statemonth, 1, 2)
+data$month <- substr(data$statemonth, 4, 5)
+
+qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
+
+data$income_percap <- log(data$income_percap)
+
+data <- data[weather_term == 0, ]
+
+data$raining <- data$prcp > 0
+
+states <- unique(data$state)
+
+##################################
+# Just effects of Temperature
+################################
+alldat <- data.frame()
+for (s in states){
+  print(s)
+  print(which(states == s))
+
+  sel <- data[data$state == s & data$wbgt > 5, ]
+
+  mod <- feols(vader ~ wbgt + raining + srad | dow + doy + tod + fips + year + month, data=sel)
+
+  alldat <- bind_rows(alldat, 
+                      tidy(mod) %>%
+                        mutate(state = s))
+}
+
+
+alldat$GEOID <- alldat$state  
+
+m <- merge(us_states, alldat %>% filter(term == 'wbgt'))
+
+ggplot(m) + 
+  geom_sf(aes(fill=estimate, color=p.value < 0.01)) + 
+  scale_fill_gradient2(low='#ca0020', mid='#f7f7f7', high='#92c5de', midpoint=0) + 
+  scale_color_manual(values=c('TRUE'='#000000', 'FALSE'=NA)) +
+  theme_void() + 
+  labs(main='Effect of Temperature on Twitter Sentiment', 
+       fill='Estimate', color='Significant (p < 0.01)')
+ggsave('~/temp-sentiment/res/map_wbgt.png')
+
+
+#####################################
+# Now effects of temp*income
+#######################################
+
+alldat2 <- data.frame()
+for (s in states){
+  print(s)
+  print(which(states == s))
+
+  sel <- data[data$state == s & data$wbgt > 5, ]
+
[…]

### 2021-04-07 doc/temp-sentiment.tex
Updates from Overleaf

@@
 \usepackage{pdflscape}
+\usepackage{subcaption}
 \usepackage{lineno}
@@
 %Cite this new paper: https://doi.org/10.1016/S2542-5196(20)30251-5
-
+% And: https://eos.org/articles/long-term-drought-harms-mental-health-in-rural-communities
 \section{Introduction}
@@
 % Heatwaves are known to worsen mental health
+
 https://doi.org/10.1016/j.puhe.2018.06.008 (review study - will have more to cite)
@@
 % AND mental health is strongly associated with income
+% 1）Childhood 2）adulthood 3）gender
+% Perceived income level
+
+The relationship between income, socioeconomic status (SES), well-being and mental health is one of the strongest established patterns in psycho-social literature (cite: Easterlin 1973, Holzer et al.
+1986, Perry 1996, Ng et al.
+2014).
+Much of the research into these patterns focuses on the inequalities and distribution of the effects on individual well-being.
+Subsequent work on mental health outcomes, one measure of well-being, have established strong associations between SES, income and mental health with the most and least privileged most commonly associated with the best and worse mental health experiences and outcomes, respectively (cites: Stevenson and Wolfers 2008, Ng et al.
+2014).
+
+For example, results from a cross-sectional comparative study of socioeconomic factors and use of mental health services by people living in Ontario, Canada and the United States, indicate clear disparities in use of mental health services between those with the highest income levels and lowest mental health morbidity and those with the lowest income levels and highest mental health morbidity - those with higher incomes tended to have lower mental health morbidity issues while utilizing mental health services more than those with lower incomes and higher mental health issues (cite Katz et al.
+1997).
+
+Results from a systematic review of the literature on associations between social inequalities and mental health disorders indicate a clear and prevailing link between  one or more indicators of less social privilege and higher prevalence of mental health disorders (cite Fryers et al.
+2003), with low income one of the most consistent markers of and associations with increases in common mental health disorders.
+Collectively, results from the income and mental health literature indicate that socially disadvantaged populations experience significantly more frequent common mental health disorders.
+Scholars use mental health disparities to indicate the disproportionate amount of mental health disorders among persons of low SES\citep{RN1292}
+
+The established association between poverty and mental health revealed that that income distribution may have a significant influence upon mental health over and above the effect of poverty \citep{HANANDITA201459}.Poverty has been identified as one of the major risk factors in mental health.
+For instance, some systematic reviews have shown that socioeconomically disadvantaged children and adolescents were 2-3 times more likely to have mental health problems than others\citep{REISS201324}, while the household increases in financial resources are generally associated with an overall reduction of mental health problems\citep{2015Does}.
+Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress (cite Kawachi and Berkman 2001).
+Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease menta the
+
+ Several reason could explain why poverty may effect mental health.
+1) the “social causation hypothesis” which suggests that stress or deprivation or decreasing the likelihood of people getting treatment may lead to poor mental health\citep{mills2015}, while social capital could reduce the likelihood of living in poverty and reduces the risk of mental disorders such as depressive symptoms and suicidal tendencies\citep{RN1291}.
+2) negative life events in a person’s life such as job loss are also associated with the risks of poverty\citep{RN1293} and mental health\citep{TAMPUBOLON201420}, and the causal inferences on the effect of poverty and mental health had also been made in several studies.
+For example, Tampubolon and Hanandita\citep{TAMPUBOLON201420} find that individual social capital is positively associated with mental health while adverse events were negatively associated; Chang.et.al \citep{RN1291} find that subjective and objective poverty is significantly associated with a higher risk of adverse life events, less social support and mental distress.
+Negative life events and social support in serial mediate the relationship between subjective poverty and mental health.
+
+For children, poverty and poor maternal mental health often co-exist and are two of the risk factors for child development\citep{LUND20111502}, and maternal mental health is significantly associated with child general psychopathology\citep{ RN1289}.
+Researchers also find that transition into poverty increased children’s socioemotional behavior problems and maternal psychological distress \citep{WICKHAM2017e141}.
+
+In general, the relationship between mental health and income level is well established.
+In the study, we explore whether income level would influence the link between weather and the sentiment in Twitter-posed text.
+
 https://jamanetwork.com/journals/jamapsychiatry/fullarticle/211213
@@
 https://www.sciencedirect.com/science/article/abs/pii/S0143622814002537
+also this: https://www.sciencedirect.com/science/article/abs/pii/S0140673614614604
 
@@
 
+\section{Results}
+We examined the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an indicator of heat stress that accounts for temperature, humidity, wind speed, and solar radiation.
+We then explored how the racial and income characteristics of neighborhoods moderate the relationship between WBGT and sentiment.
+Finally, we show how twitter sentiment is an important overall indicator of mental health by assessing the relationship between daily sentiment at the county level and the prevalence of crime and suicide.
+
+and sentiment, first examining the overall relationship, and then exploring heterogeneities based on neighborhood income per capita and race.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
+  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment in tweets.  As temperatures increase, sentiment declines.}
+  \label{fig:wbgt-income}
+\end{figure}
+
+Sentiment is highest at a WBGT of 5 degrees celsius (a dry bulb temperature typically around 12C/54F).
+The greatest declines in sentiment are observed between 20-25 degrees WBGT (dry bulb 29-36C/84-97F).
+Sentiment also declines with colder temperatures, but only slightly.
+
+\begin{figure}[H]
+\centering
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
+  %\caption{WBGT-sentiment relationship as moderated by neighborhood income.}
+  \label{fig:sub1}
+\end{subfigure}%
+\begin{subfigure}{.5\textwidth}
+  \centering
+  \includegraphics[width=\linewidth]{../res/wbgt-race_q.png}
+  %\caption{WBGT-sentiment relationship as moderated by neighborhood majority race.}
+  \label{fig:sub2}
+\end{subfigure}
+\caption{How the relationship between WBGT and sentiment is moderated }
+\label{fig:test}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
+  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
+For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
+For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
+  \label{fig:wbgt-income}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{../res/wbgt-race.png}
+  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
+For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
+For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
+  \label{fig:wbgt-income}
+\end{figure}
+
+Figure of heatmap of tweet density across USA (pick a year? entire dataset of georeferenced tweets?)
+
+(sub-section) TEMPERATURE and HUMIDITY
+Figure of distribution of max and min temperatures
+Comment from Patrick Baylis on our "preliminary results". 2nd comment from Patrick
+
+(sub-section) INCOME
+
+(sub-section?) SENTIMENT
+Figure of distribution of observed sentiment scores (by Hedonometer, Vader, etc.)
+
+Comment from Patrick Baylis on our "preliminary results" doc
+
+\section{Discussion}
+
 \section{Data}
@@
 \subsection{Socio-Econoimc Data}
-We used data from the US Census at the Census Block level to estimate the
+We used data from the American Communit   Survey administered by the US Census to income levels and the racial composition of neighborhoods where tweets were located.
+Data was at the level of the census block group, the smallest unit for which the census releases public data.
+
+ACS data is released to cover five-year periods.
+We therefore matched each tweet with census block group data from the year at the middle year of each survey's five-year range.
+For example, tweets from 2014 were matched to data from the 2012-2016 ACS.
+Because the most recent available data was from 2014-2018, all tweets from year years greater than 2016 were matched to this dataset.
+Data was downloaded from the IPUMS NHGIS service provided by the University of Minnesota.
+
+Mean annual income per capita is provided by the ACS, and we standardized this variable so that the values for each year were in 2019 dollars.
+For racial categories, we combined the various categories provided by the ACS into four racial groups: non-Hispanic white, non-Hispanic black, Hispanic of any race, and an "other" category for non-Hispanic people who were neither black nor white, such as Asian, Native American, or mixed-race people.
 
 \subsection{Sentiment}
+
+WE EXCLUDED WEATHER TERMS
+
 We assessed the sentiment expressed in text using several of the most commonly used measures of sentiment.
@@
 
+STANDARD ERRORS
+
 \subsection{Data Processing}
@@
 To run XYZ analysis, we used Microsoft's ZYX package.
-
-\section{Results}
-(sub-section) DESCRIPTIVE STATISTICS
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
-For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
-For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
-  \label{fig:wbgt-income}
-\end{figure}
-
-Figure of heatmap of tweet density across USA (pick a year? entire dataset of georeferenced tweets?)
-
-(sub-section) TEMPERATURE and HUMIDITY
-Figure of distribution of max and min temperatures
-Comment from Patrick Baylis on our "preliminary results". 2nd comment from Patrick
-
-(sub-section) INCOME
-
-(sub-section?) SENTIMENT
-Figure of distribution of observed sentiment scores (by Hedonometer, Vader, etc.)
-
-Comment from Patrick Baylis on our "preliminary results" doc
-
-\section{Discussion}
 
@@
 \renewcommand{\thesection}{S\arabic{section}}
+
+\subsection{Other Sentiment Metrics}
+
+\subsection{Precipitation}
+
+%Supplement will have:
+
+%Equations used for RH and WBGT
+
+% Robustness checks:
+
+    % Other sentiment metrics
+
+    %
+
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Paste Bin
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 

### 2021-04-07 viz/vader/wbgt_income_bootstrap.R
explore effects of precip and srad

@@ -137,7 +137,8 @@ curve <- ggplot(preddf) +
   geom_line(aes(x=wbgt, y=pred, color=income_percap)) + 
   geom_ribbon(aes(x=wbgt, ymin=ymin, ymax=ymax, fill=income_percap), alpha=0.5) + 
   scale_x_continuous(expand=c(0, 0), limits=c(-22, 33)) + 
-  #scale_y_continuous(breaks=seq(0, -0.02, -0.005)) + 
+  scale_fill_manual(values=c("#e31a1c", "#ff7f00", "#33a02c")) + 
+  scale_color_manual(values=c("#e31a1c", "#ff7f00", "#33a02c")) + 
   labs(x='', y='Change in Mood of Tweet',
        fill = 'Census Block\nIncome Per-Capita\n(Percentile)',
        color = 'Census Block\nIncome Per-Capita\n(Percentile)') +
@@ -173,3 +174,153 @@ xl <- get_plot_component(x2, "xlab-b")
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
 ggsave('~/temp-sentiment/res/wbgt-income.png', width=6, height=5)
+
+########################################
+# Precip
+#########################################
+preddf <- data.frame(raining=c(TRUE, FALSE))
+preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
+
+preddf$vader <- 1
+
+mm <- model.matrix(vader ~ income_percap*raining,
+                   data=preddf)
+
+colnames(mm)[colnames(mm) == 'wbgt:income_percap'] <- 'income_percap:wbgt'
+mm <- mm[ , colnames(mm) != '(Intercept)']
+mm <- mm[ , colnames(mm) != 'income_percap']
+
+mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
+
+#Iterate over models
+mods <- list.files()
+for (modf in mods){
+
+  mod <- readRDS(modf)
+
+  res <- as.data.frame(svycontrast(mod, mmp))[ , 'contrast', drop=F]
+  names(res) <- modf
+  
+  preddf <- cbind(preddf, res)
+}
+
+preddf$income_percap <- factor(preddf$income_percap)
+levels(preddf$income_percap) <- preddf$income_percap %>%
+                                   levels %>%
+                                   as.numeric %>%
+                                   exp %>%
+                                   round(0) %>%
[…]

### 2021-04-07 viz/vader/wbgt_income_bootstrap.R
update figures for race

@@ -232,6 +232,7 @@ ggplot(preddf %>% filter(raining)) +
   geom_errorbar(aes(x=income_percap, ymax=ymax, ymin=ymin, y=pred)) + 
   scale_fill_manual(values=c("#e31a1c", "#ff7f00", "#33a02c")) + 
   scale_color_manual(values=c("#e31a1c", "#ff7f00", "#33a02c")) + 
+  scale_y_continuous(expand=expand_scale(mult=c(0.05, 0))) + 
   guides(fill=FALSE) + 
   theme_classic() + 
   labs(x="Census Block Income Per-Capita (Percentile)", 

### 2021-04-07 viz/vader/wbgt_race_q_bootstrap.R

@@ -133,7 +133,8 @@ curve <- ggplot(preddf) +
   geom_line(aes(x=wbgt, y=pred, color=race_majority)) + 
   geom_ribbon(aes(x=wbgt, ymin=ymin, ymax=ymax, fill=race_majority), alpha=0.5) + 
   scale_x_continuous(expand=c(0, 0), limits=c(-22, 33)) + 
-  #scale_y_continuous(breaks=seq(0, -0.02, -0.005)) + 
+  scale_fill_manual(values=c("#e31a1c", "#33a02c", "#1f78b4", "#6a3d9a")) +
+  scale_color_manual(values=c("#e31a1c", "#33a02c", "#1f78b4", "#6a3d9a")) +
   labs(x='', y='Change in Mood of Tweet',
        fill = 'Census Block\nMajority Race',
        color = 'Census Block\nMajority Race') +
@@ -169,3 +170,136 @@ xl <- get_plot_component(x2, "xlab-b")
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
 ggsave('~/temp-sentiment/res/wbgt-race_q.png', width=6, height=5)
+
+########################################
+# Precip
+#########################################
+preddf <- data.frame(raining=c(TRUE, FALSE))
+preddf <- make_groups(preddf, 'race_majority', unique(data$race_majority))
+
+preddf$vader <- 1
+
+mm <- model.matrix(vader ~ race_majority*raining,
+                   data=preddf)
+
+mm <- mm[ , colnames(mm) != '(Intercept)']
+mm <- mm[ , colnames(mm) != 'race_majority']
+
+mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
+
+#Iterate over models
+mods <- list.files()
+for (modf in mods){
+
+  mod <- readRDS(modf)
+
+  res <- as.data.frame(svycontrast(mod, mmp))[ , 'contrast', drop=F]
+  names(res) <- modf
+  
+  preddf <- cbind(preddf, res)
+}
+
+preddf$race_majority <- factor(preddf$race_majority)
+levels(preddf$race_majority) <- c("Black", "Hispanic", "Other", "White")
+
+preddf <- preddf %>%
+  gather(key, value, -raining, -race_majority, -vader) %>%
+  #Normalize so that graph shows difference from wbgt = X
+  group_by(race_majority, key) %>%
[…]

### 2021-04-07 doc/pastebin.tex
work on document locally

--- 2021-01-06 doc/temp-sentiment.tex
@@
-\documentclass{article}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage[a4paper, margin=1.25in]{geometry}
-\usepackage{breakcites}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
-\usepackage{textgreek}
-\usepackage{authblk}
-\usepackage{rotating}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{pdflscape}
-\usepackage{lineno}
-\usepackage[
-  style=numeric,
-  citestyle=numeric-comp,
-  backend=biber,
-  doi=true,
-  natbib=true,
-  sorting=none
-]{biblatex}
-
-\addbibresource{TextDataClimateShocks.bib}
-
-\begin{document}
-
-\title{Heatwaves Exacerbate Inequalities in Mental Health}
-
-%\author[1, *]{Matthew Cooper}
-%\author[]{Jeremiah Osborne-Gowey}
-%\author[]{Zheng Liu}
-%\author[]{Jie Liu}
-%\author[]{Portia Adade Williams}
-%\author[]{Aaron Schwartz}
-%\author[]{Patrick Baylis}
-
-%\affil[1]{T.H. Chan School of Public Health, Harvard University}
-%\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
-
-\maketitle
-
-\begin{abstract}
-Weather can affect people’s mood and well-being.
-Traditional studies exploring the effects of weather on mental health require significant resource investment (e.g., time, funding, execution).
-Continuous in situ data streams from social media, newspapers, and other big data sources offer an opportunity to examine how people perceive and respond to environmental conditions.
-Recent work demonstrates that prevailing weather conditions can affect people’s sentiment as expressed on Twitter and other social media.
-Studies to date however, have modeled the weather as having the same effect on sentiment across all locations and individuals (i.e., in aggregate).
-In this study, we explore how weather affects individual expressed sentiment (via Twitter) across different weather gradients and locations.
-We analyzed the sentiment from a quarter of a billion geolocated Tweets from 2009 to 2019, using multiple measures of sentiment, overlaid with data about prevailing weather conditions, as well as local land cover, income levels, and demographic characteristics in the vicinity of Twitter users.
-We find that weather affects people differently across a range of incomes, with people in higher income brackets generally happier with increasing temperature, while people in lower income brackets are less happy with increasing temperature.
-This project extends existing research about the relationship between temperature and mood (sentiment) by examining how varying climatic conditions across different geographies and socio-economic groups are correlated with expressed sentiment.
-This project brings new insights into how humans perceive and respond to environmental shocks, with important implications for economics, policy and adaptation and resilience planning.
-
-\end{abstract}
-
-Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation
-
-%Cite this new paper: https://doi.org/10.1016/S2542-5196(20)30251-5
-
-\section{Introduction}
-
-% Organize as follows:
-
-% Heatwaves are known to worsen mental health
-% AND mental health is strongly associated with income
-% AND wealthier better able to cope with heatwaves 
-% BUT we know little about the interaction between heatwaves, mental health and income
-% THEREFORE: we did this really novel study.
-
-This work builds on existing work by \citep{baylis_weather_2018}.
-
-%Intro Paragraph
-Research has demonstrated strong linkages between ambient environmental conditions and mental health, with extreme temperatures more frequently associated with poorer overall mental health.
-Mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less capable of finding help for coping with these stresses.
-Additionally, results from research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves.
-Furthermore, previous studies examining relationships between environmental extremes and health sometimes aggregate data at scales too coarse to examine whether income may be an ameliorating factor.
-Yet we know relatively little about the interaction between temperature, mental health and income.
-This research examines the interactions between these three using relatively fine-scale, publicly available data.
-This study expands on existing temperature and sentiment - one measure of mental health - research by examining finer-scale data and incorporating income data to examine correlations between extreme temperatures and sentiment.
-
-Human moods and mental states are important aspects of overall well-being and can be influenced by ambient and persistent environmental conditions.
-Climate change is accelerating the rate and variability around meteorological norms like the timing, magnitude, intensity and duration of precipitation events and air temperature minima and maxima \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
-Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meterological conditions.
-Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to meteorological conditions.
-Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.
-
-\section{Background}
-% Heatwaves are known to worsen mental health
-https://doi.org/10.1016/j.puhe.2018.06.008 (review study - will have more to cite)
-https://www.nature.com/articles/s41558-018-0222-x
-https://ehp.niehs.nih.gov/doi/full/10.1289/ehp.11339
-
-% AND mental health is strongly associated with income
-https://jamanetwork.com/journals/jamapsychiatry/fullarticle/211213
-https://doi.org/10.1002/9781118410868.wbehibs570
-https://www.sciencedirect.com/science/article/abs/pii/S027795362030527X
-https://www.sciencedirect.com/science/article/pii/S2468266717300117
-https://doi.org/10.1016/S2468-2667(17)30011-7
-https://link.springer.com/article/10.1007/s00127-017-1370-4
-
-% AND wealthier better able to cope with heatwaves 
-https://www.nature.com/articles/nclimate3253
-https://www.sciencedirect.com/science/article/abs/pii/S0378778818321327
-
-% BUT we know little about the interaction between heatwaves, mental health and income
-
-% THEREFORE: we did this really novel study.
-
-%OTHER LEFTOVER TEXT TO MAYBE BRING IN LATER
-
-Meteorological conditions can impact human physical (cite) and emotional states (cite).
-Emotional states and well-being are associated with physiological functioning and mental acuity which can affect social relationships, workplace productivity (cite) and health risks (cite).
-People that are more reliant on livelihoods which require them to be outdoors (e.g., farming, agriculture, logging, etc.), living and working in places where they are exposed to environmental minima and maxima, or with prolonged exposure to ambient meteorological extremes are particularly vulnerable to changes in environmental exposure \citep{frimpong_heat_2017} and associated health risks (cite), with implications for workplace productivity and livelihoods \cite{kjellstrom_impact_2016}.
-For example, Nigerian maize farmers experience significant declines in productivity (2-8percent per degree above 17degC) and substantive increases in health impacts including dehydration, muscle cramps, headaches and dizziness, heat exhaustion, sun stroke, and even death \cite{sadiq_impact_2019}.
-Other research indicates farm health and labor productivity were compromised under extreme heat and cold events in the Nepali Food Bowl region \cite{budhathoki_socio-economic_2019}.
-Similarly, Oregon farmers reported health impacts of working in both hot and cold conditions but the negative impacts were attenuated relative to the high heat situations \cite{bethel_heat-related_2014}.
-Other research found occupational injuries in Thailand increased with increasing heat exposure and occupational heat stress \cite{tawatsupa_association_2013}.
-Heat exposure risks are also a key factor in urban areas with particular attention needed for how the impacts of changing climate play out in health inequality \cite{friel_urban_2011}.
-
-Current estimates place annual workplace productivity losses due to heat exposure at 15-20percent with that rate potentially doubling by 2050 \cite{kjellstrom_heat_2016}.
-Projected increases in future ambient temperature suggest serious short- and long-term potential health consequences from exposure to heat stress (cite) with some estimates putting potential productivity losses at several percentage points by 2030 \cite{kjellstrom_heat_2016}, with middle- and low-income countries particularly vulnerable as they are often more reliant on physical work for their livelihoods.
-
-The level of changes in weather influences people’s mental health and patterns of emotions expressed.
-A study by Sun et al.
-2018 \href{file:///Users/portia/Downloads/ijerph-16-00086-v2 20(1).pdf.}{(Portia has this PDF?)} demonstrates varied relationships between haze and negative emotions of the public under different seasons of the year.
-Being sad or happy influences one's style and value of reasoning.
-Time of day, season, location, and climate allow aggregate prediction of sentiments \cite{hannak_tweetin_2012} \href{https://www.ccs.neu.edu/~amislove/publications/Weather-ICWSM.pdf}{(link to PDF)} while positive affect from an evaluative statement enhances potential responses \cite{clore_how_2007} \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2483304/pdf/nihms40349.pdf}{(link to PDF)}.
-For instance, changes in temperature affects people’s response about weather.
-Depending on a reference temperature for an  area and time of year, people are more likely to comment on unusual weather for a particular place and time than on the same weather considered typical in another place \cite{moore_rapidly_2019}.
-
-Although climate change is an international phenomenon experienced in every country, its discussions vary from country to country.
-According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions.
-For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature do express more positive emotions on Twitter than those living in colder states in the US.
-Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage global warming receives in the media in countries like USA, Australia and Germany.
-Studies comparing how different countries portray climate change have been widely conducted in USA, UK, France and the Netherlands \cite{vu_nationalizing_2019}.
-Such comparisons are mainly between developed countries with developing countries outside the focus.
-Schafer and O’Neil \cite{schafer_what_2013} advance the need for academic scholars to investigate transnational contexts within climate communication.
-Wealthier countries with comparatively higher GDP are likely to frame climate change as a political issue as financial resources exist for exploring research on climate change.
-Conversely, coverage of climate change news from poorer developing countries focus mainly on international relations.
-Thus, the effects of national macro economic variables that affects a country’s sociopolitical and economic development such as GDP reflects a country’s governance system in media coverage \cite{vu_what_2018}.
-Cultural theorists assert that, understanding such differential variations depicts how different entities interpret danger and respond to risk \cite{tansey_cultural_1999}.
-Inclusion of developing countries in scientific research provides an avenue for international support to such countries in responding to the effects of climate change.Yet studies attempting to explain cross-national variation in climate change public opinion is limited \cite{knight_public_2016}.
-
-Public debates on the current consequences of climate change are discussed in both scientific and non-scientific mediums.
-People can express their emotional states and feelings - sentiment - through physical, vocal and written expressions on mediums including newspapers, scientific articles, blogs and other online social media platforms.
-Twitter is one of the more popular social media platforms where people express their sentiments about any number of topics.
-Public posts on Twitter - called tweets - are widely available for public consumption and research, and offer a glimpse into collective social state.
-For example, text posted to these platforms can be used to gauge the public mood, assess opinions, measure brand affinity and for emergency planning and disaster response (cites).
-Sentiment is frequently employed as a correlate of emotional state or mood (cites).
-Weather events focus public attention in different ways of expression (cites).
-This offers a unique opportunity to analyze how events influence public discourse on their expressions and responses.
-
-Here, we report on associations between meteorological conditions and expressed sentiment of public posts on Twitter from the United States of America between 2009 and 2019.
-This work builds on research from Baylis and others \cite{baylis_weather_2018} on weather impacts on expressed sentiment in the USA.
-In particular, we extend this work by including the temperature effects on sentiment in multiple regions across the USA while drawing in various climatic variables and socio-demographic data.
-In this research, we explore two primary questions.
-First, to what extent are weather conditions correlated with changes in expressed sentiment?
-Second, is this effect moderated by wealth and local land cover conditions?
-
-Building on previous findings from Baylis and others, we hypothesize that local ambient weather conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1].
-Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas.
-That is to say, people in lower income areas will be have lower sentiment scores than those in wealthier areas.
-[H2].
-
-\section{Data}
-A proposed list of tables and figures:
-Figures for location of USA Tweets (to show aggregation and coverage).
-Figures of income blocks and temperature extremes.
-Table 1. Variable descriptions for what's included in our models.
-Table 2. Descriptive statistics for datasets/variables we include.
-Table 3. Model results with variables included in the models coefficients, SEs, p-values, sample size.
-Appendix table. Sentiment scores across all types we considered.
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/Boston_Map.png}
-  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
-  \label{fig:timeseries}
-\end{figure}
-
-\section{Methods}
-\subsection{Text Data}
-Text data used for this project come from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose - which provides public access to a random 10\% of all public messages on Twitter.
-The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
-The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
-We used all geolocated Tweets from the Decahose for the period 2012-2016 with a bounding box polygon filter for India, Western Europe and the United States of America.
-We only considered users’ original content, thus did not include retweets in our analysis.
-
-\subsection{Weather}
-We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington \href{https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2011JD016048}{link}.
-This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution \cite{xia_continental-scale_2012}.
-We extracted the data for the exact hour and location of each tweet using Google Earth Engine.
-We extracted several variables from the NLDAS-2 dataset, including temperature, specific humidity, air pressure, total precipitation, and shortwave radiation.
-Using temperature, specific humidity, and pressure, we derived relative humidity using methods described by  \cite{bolton_computation_1980}.
-Then, based on the Heat Index algorithm developed by Steadman \cite{steadman_assessment_1979} and improved upon by Rothfusz and the National Weather service \cite{rothfusz_heat_1990}, we calculated the Heat Index using temperature and relative humidity \cite{heo_comparison_2019}.
-
-ISSUES: deciduous trees lose leaves every year and may not provide much shade cover and heat relief during particular periods of the year when they are leafless.
-
-\subsection{Income}
-We used data from the US Census at the Census Block level to estimate the
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
+% Paste Bin
+%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 [NEED SOME CITATIONS HERE]
-
-\subsection{Sentiment}
-We assessed the sentiment expressed in text using several of the most commonly used measures of sentiment.
-Each of the measures uses validated dictionaries of words that have been assigned positive and negative scores and compares these words with words in the text corpora of interest.
-For measures of sentiment, we used the Hedonometer \href{https://hedonometer.org/timeseries/en_all/}{(link)}, Afinn \href{http://corpustext.com/reference/sentiment_afinn.html}{(link)}, Textblob \href{https://textblob.readthedocs.io/en/dev/}{(link)}, Vader , SentiwordNet and LIWC tools.
-These tools assign positive and negative scores to words, then average the scores for each text string.
-Our text data was multilingual, which can present challenges with analyzing text (cite).
-Given our interest only in the collection of words used (in aggregate) and not the sequence or nuance of the phrases, we translated all languages into English using Google Translate API and the Microsoft Azure API cognitive services (translator text API) before conducting the sentiment analyses.
-
-Some papers we might cite in this section (only for multilingual work): Vilares et al. 2017 , Lo et al. 2017, Dashtipour et al. 2016.
-
-Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! \#fail]), etc.
-
-Issues with sample population (non-representative)
-
-Issues with what people say on social media being attenuated or mismatched from actual emotional state
-
-\subsubsection{Hedonometer}
-The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
-The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2)using a naive algorithm for scaling up from individual words to texts.
-For the English corpus, Dodds and others \cite{dodds_temporal_2011} collect 10,222 unique words and used crowd-sourcing platform Amazon Mechanical Turk to get human evaluation of happiness degree for each word in an integer scale from 1 to 9, representing a sad to happy spectrum.
-Score 5 represents neutral words.
-Each word will be calculated average score and then the word and happiness score are compiled into a dataset (labMT 1.0).
-Some illustrative example of words are:
-
-\[h_{avg} (\text{laughter}) = 8.50 \]
-\[h_{avg} (\text{the}) = 4.98\]
-\[h_{avg} (\text{hate}) = 2.34\]
-
-The sentiment score for single text will be the mean happiness score for all words in the text.
-
-\subsubsection{VADER}
-The VADER sentiment corpus was proposed by Gilbert and Hutto \cite{gilbert_vader_2014} and stands for 'Valence Aware Dictionary for sentiment Reasoning'.
-It is is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media by incorporating lexical features common to sentiment expression in microblogs
-
-\subsection{Analytic Approach}
-To assess whether temperature affects sentiment, we ran a linear regression model that included meteorological measures and sentiment.
-Independent variables included maximum daily temperature, precipitation and heat index.
-We also included fixed effects indicator variables at the year-month and decimal degree levels to account for potential biases in unobserved characteristics associated with seasonality and date.
-We examined the relationships between meteorological variables and sentiment using linear regression models.
-
-\subsection{Data Processing}
-We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 2019) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
-Sentiment data was analyzed using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
-We extracted meteorological data using the geolocated coordinates from the Twitter posts (latitude and longitude).
-We calculated mean climatological scores (z-scores)using baseline meteorological data from 1990 to present, calculated at a quarter degree resolution, and based on the long-term averages and standard deviation.
-We calculated z-scores for each day and the overall annual average z-score.
-For text data cleaning we used the python package tweet-preprocessor \href{https://pypi.org/project/tweet-preprocessor/}{(link)} to remove all text and characters not part of the main text phrase (e.g., URLs, mentions, GIFs, etc.).
-We converted all dates and times to UTC format.
-
-\section{Results}
-(sub-section) DESCRIPTIVE STATISTICS
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
-For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
-For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
-  \label{fig:wbgt-income}
-\end{figure}
-
-Figure of heatmap of tweet density across USA (pick a year? entire dataset of georeferenced tweets?)
-
-(sub-section) TEMPERATURE and HUMIDITY
-Figure of distribution of max and min temperatures
-Comment from Patrick Baylis on our "preliminary results". 2nd comment from Patrick
-
-(sub-section) INCOME
-
-(sub-section?) SENTIMENT
-Figure of distribution of observed sentiment scores (by Hedonometer, Vader, etc.)
-
-Comment from Patrick Baylis on our "preliminary results" doc
-
-\section{Discussion}
-
-\section{Conclusion}
-
-\section{Acknowledgements}
-For funding, advising, SESYNC grant number?, etc.
-
-\printbibliography
-
-\section*{Supplemental Info}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thesection}{S\arabic{section}}
 
@@
 Lucas et al. 2015 paper --> \href{https://www.cambridge.org/core/journals/political-analysis/article/computerassisted-text-analysis-for-comparative-politics/CC8B2CF63A8CC36FE00A13F9839F92BB}{link}
-
-\end{document}

### 2021-04-07 doc/temp-sentiment-supplement.tex
work on document locally

--- 2021-01-06 doc/temp-sentiment.tex
@@
 \documentclass{article}
+
 \usepackage{graphicx}
@@
 
+\pagestyle{fancy}
+\fancyhf{}
+\lfoot{Supplemental Materials for \textit{Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures}}
+\rfoot{\thepage}
+
+\renewcommand{\footrulewidth}{0.4pt}
+\renewcommand{\headrulewidth}{0pt}
+
 \addbibresource{TextDataClimateShocks.bib}
@@
 \begin{document}
+\begin{center}
+\section*{Supplemental Materials for \textit{Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures}}
+\end{center}
 
-\title{Heatwaves Exacerbate Inequalities in Mental Health}
+%Supplement will have:
 
-%\author[1, *]{Matthew Cooper}
-%\author[]{Jeremiah Osborne-Gowey}
-%\author[]{Zheng Liu}
-%\author[]{Jie Liu}
-%\author[]{Portia Adade Williams}
-%\author[]{Aaron Schwartz}
-%\author[]{Patrick Baylis}
+%Equations used for RH and WBGT
 
-%\affil[1]{T.H. Chan School of Public Health, Harvard University}
-%\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
+\section{Alternative Parameterizations}
 
-\maketitle
+\section{Effects of Precipitation and Sunshine}
 
-\begin{abstract}
-Weather can affect people’s mood and well-being.
-Traditional studies exploring the effects of weather on mental health require significant resource investment (e.g., time, funding, execution).
-Continuous in situ data streams from social media, newspapers, and other big data sources offer an opportunity to examine how people perceive and respond to environmental conditions.
-Recent work demonstrates that prevailing weather conditions can affect people’s sentiment as expressed on Twitter and other social media.
-Studies to date however, have modeled the weather as having the same effect on sentiment across all locations and individuals (i.e., in aggregate).
-In this study, we explore how weather affects individual expressed sentiment (via Twitter) across different weather gradients and locations.
-We analyzed the sentiment from a quarter of a billion geolocated Tweets from 2009 to 2019, using multiple measures of sentiment, overlaid with data about prevailing weather conditions, as well as local land cover, income levels, and demographic characteristics in the vicinity of Twitter users.
-We find that weather affects people differently across a range of incomes, with people in higher income brackets generally happier with increasing temperature, while people in lower income brackets are less happy with increasing temperature.
-This project extends existing research about the relationship between temperature and mood (sentiment) by examining how varying climatic conditions across different geographies and socio-economic groups are correlated with expressed sentiment.
-This project brings new insights into how humans perceive and respond to environmental shocks, with important implications for economics, policy and adaptation and resilience planning.
+\section{Alternative Sentiment Algorithm}
 
-\end{abstract}
-
-Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation
-
-%Cite this new paper: https://doi.org/10.1016/S2542-5196(20)30251-5
-
-\section{Introduction}
-
-% Organize as follows:
-
-% Heatwaves are known to worsen mental health
-% AND mental health is strongly associated with income
-% AND wealthier better able to cope with heatwaves 
-% BUT we know little about the interaction between heatwaves, mental health and income
-% THEREFORE: we did this really novel study.
-
-This work builds on existing work by \citep{baylis_weather_2018}.
-
-%Intro Paragraph
-Research has demonstrated strong linkages between ambient environmental conditions and mental health, with extreme temperatures more frequently associated with poorer overall mental health.
-Mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less capable of finding help for coping with these stresses.
-Additionally, results from research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves.
-Furthermore, previous studies examining relationships between environmental extremes and health sometimes aggregate data at scales too coarse to examine whether income may be an ameliorating factor.
-Yet we know relatively little about the interaction between temperature, mental health and income.
-This research examines the interactions between these three using relatively fine-scale, publicly available data.
-This study expands on existing temperature and sentiment - one measure of mental health - research by examining finer-scale data and incorporating income data to examine correlations between extreme temperatures and sentiment.
-
-Human moods and mental states are important aspects of overall well-being and can be influenced by ambient and persistent environmental conditions.
-Climate change is accelerating the rate and variability around meteorological norms like the timing, magnitude, intensity and duration of precipitation events and air temperature minima and maxima \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
-Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meterological conditions.
-Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to meteorological conditions.
-Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.
-
-\section{Background}
-% Heatwaves are known to worsen mental health
-https://doi.org/10.1016/j.puhe.2018.06.008 (review study - will have more to cite)
-https://www.nature.com/articles/s41558-018-0222-x
-https://ehp.niehs.nih.gov/doi/full/10.1289/ehp.11339
-
-% AND mental health is strongly associated with income
-https://jamanetwork.com/journals/jamapsychiatry/fullarticle/211213
-https://doi.org/10.1002/9781118410868.wbehibs570
-https://www.sciencedirect.com/science/article/abs/pii/S027795362030527X
-https://www.sciencedirect.com/science/article/pii/S2468266717300117
-https://doi.org/10.1016/S2468-2667(17)30011-7
-https://link.springer.com/article/10.1007/s00127-017-1370-4
-
-% AND wealthier better able to cope with heatwaves 
-https://www.nature.com/articles/nclimate3253
-https://www.sciencedirect.com/science/article/abs/pii/S0378778818321327
-
-% BUT we know little about the interaction between heatwaves, mental health and income
-
-% THEREFORE: we did this really novel study.
-
-%OTHER LEFTOVER TEXT TO MAYBE BRING IN LATER
-
-Meteorological conditions can impact human physical (cite) and emotional states (cite).
-Emotional states and well-being are associated with physiological functioning and mental acuity which can affect social relationships, workplace productivity (cite) and health risks (cite).
-People that are more reliant on livelihoods which require them to be outdoors (e.g., farming, agriculture, logging, etc.), living and working in places where they are exposed to environmental minima and maxima, or with prolonged exposure to ambient meteorological extremes are particularly vulnerable to changes in environmental exposure \citep{frimpong_heat_2017} and associated health risks (cite), with implications for workplace productivity and livelihoods \cite{kjellstrom_impact_2016}.
-For example, Nigerian maize farmers experience significant declines in productivity (2-8percent per degree above 17degC) and substantive increases in health impacts including dehydration, muscle cramps, headaches and dizziness, heat exhaustion, sun stroke, and even death \cite{sadiq_impact_2019}.
-Other research indicates farm health and labor productivity were compromised under extreme heat and cold events in the Nepali Food Bowl region \cite{budhathoki_socio-economic_2019}.
-Similarly, Oregon farmers reported health impacts of working in both hot and cold conditions but the negative impacts were attenuated relative to the high heat situations \cite{bethel_heat-related_2014}.
-Other research found occupational injuries in Thailand increased with increasing heat exposure and occupational heat stress \cite{tawatsupa_association_2013}.
-Heat exposure risks are also a key factor in urban areas with particular attention needed for how the impacts of changing climate play out in health inequality \cite{friel_urban_2011}.
-
-Current estimates place annual workplace productivity losses due to heat exposure at 15-20percent with that rate potentially doubling by 2050 \cite{kjellstrom_heat_2016}.
-Projected increases in future ambient temperature suggest serious short- and long-term potential health consequences from exposure to heat stress (cite) with some estimates putting potential productivity losses at several percentage points by 2030 \cite{kjellstrom_heat_2016}, with middle- and low-income countries particularly vulnerable as they are often more reliant on physical work for their livelihoods.
-
-The level of changes in weather influences people’s mental health and patterns of emotions expressed.
-A study by Sun et al.
-2018 \href{file:///Users/portia/Downloads/ijerph-16-00086-v2 20(1).pdf.}{(Portia has this PDF?)} demonstrates varied relationships between haze and negative emotions of the public under different seasons of the year.
-Being sad or happy influences one's style and value of reasoning.
-Time of day, season, location, and climate allow aggregate prediction of sentiments \cite{hannak_tweetin_2012} \href{https://www.ccs.neu.edu/~amislove/publications/Weather-ICWSM.pdf}{(link to PDF)} while positive affect from an evaluative statement enhances potential responses \cite{clore_how_2007} \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2483304/pdf/nihms40349.pdf}{(link to PDF)}.
-For instance, changes in temperature affects people’s response about weather.
-Depending on a reference temperature for an  area and time of year, people are more likely to comment on unusual weather for a particular place and time than on the same weather considered typical in another place \cite{moore_rapidly_2019}.
-
-Although climate change is an international phenomenon experienced in every country, its discussions vary from country to country.
-According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions.
-For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature do express more positive emotions on Twitter than those living in colder states in the US.
-Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage global warming receives in the media in countries like USA, Australia and Germany.
-Studies comparing how different countries portray climate change have been widely conducted in USA, UK, France and the Netherlands \cite{vu_nationalizing_2019}.
-Such comparisons are mainly between developed countries with developing countries outside the focus.
-Schafer and O’Neil \cite{schafer_what_2013} advance the need for academic scholars to investigate transnational contexts within climate communication.
-Wealthier countries with comparatively higher GDP are likely to frame climate change as a political issue as financial resources exist for exploring research on climate change.
-Conversely, coverage of climate change news from poorer developing countries focus mainly on international relations.
-Thus, the effects of national macro economic variables that affects a country’s sociopolitical and economic development such as GDP reflects a country’s governance system in media coverage \cite{vu_what_2018}.
-Cultural theorists assert that, understanding such differential variations depicts how different entities interpret danger and respond to risk \cite{tansey_cultural_1999}.
-Inclusion of developing countries in scientific research provides an avenue for international support to such countries in responding to the effects of climate change.Yet studies attempting to explain cross-national variation in climate change public opinion is limited \cite{knight_public_2016}.
-
-Public debates on the current consequences of climate change are discussed in both scientific and non-scientific mediums.
-People can express their emotional states and feelings - sentiment - through physical, vocal and written expressions on mediums including newspapers, scientific articles, blogs and other online social media platforms.
-Twitter is one of the more popular social media platforms where people express their sentiments about any number of topics.
-Public posts on Twitter - called tweets - are widely available for public consumption and research, and offer a glimpse into collective social state.
-For example, text posted to these platforms can be used to gauge the public mood, assess opinions, measure brand affinity and for emergency planning and disaster response (cites).
-Sentiment is frequently employed as a correlate of emotional state or mood (cites).
-Weather events focus public attention in different ways of expression (cites).
-This offers a unique opportunity to analyze how events influence public discourse on their expressions and responses.
-
-Here, we report on associations between meteorological conditions and expressed sentiment of public posts on Twitter from the United States of America between 2009 and 2019.
-This work builds on research from Baylis and others \cite{baylis_weather_2018} on weather impacts on expressed sentiment in the USA.
-In particular, we extend this work by including the temperature effects on sentiment in multiple regions across the USA while drawing in various climatic variables and socio-demographic data.
-In this research, we explore two primary questions.
-First, to what extent are weather conditions correlated with changes in expressed sentiment?
-Second, is this effect moderated by wealth and local land cover conditions?
-
-Building on previous findings from Baylis and others, we hypothesize that local ambient weather conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1].
-Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas.
-That is to say, people in lower income areas will be have lower sentiment scores than those in wealthier areas.
-[H2].
-
-\section{Data}
-A proposed list of tables and figures:
-Figures for location of USA Tweets (to show aggregation and coverage).
-Figures of income blocks and temperature extremes.
-Table 1. Variable descriptions for what's included in our models.
-Table 2. Descriptive statistics for datasets/variables we include.
-Table 3. Model results with variables included in the models coefficients, SEs, p-values, sample size.
-Appendix table. Sentiment scores across all types we considered.
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/Boston_Map.png}
-  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
-  \label{fig:timeseries}
-\end{figure}
-
-\section{Methods}
-\subsection{Text Data}
-Text data used for this project come from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose - which provides public access to a random 10\% of all public messages on Twitter.
-The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
-The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
-We used all geolocated Tweets from the Decahose for the period 2012-2016 with a bounding box polygon filter for India, Western Europe and the United States of America.
-We only considered users’ original content, thus did not include retweets in our analysis.
-
-\subsection{Weather}
-We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington \href{https://agupubs.onlinelibrary.wiley.com/doi/full/10.1029/2011JD016048}{link}.
-This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution \cite{xia_continental-scale_2012}.
-We extracted the data for the exact hour and location of each tweet using Google Earth Engine.
-We extracted several variables from the NLDAS-2 dataset, including temperature, specific humidity, air pressure, total precipitation, and shortwave radiation.
-Using temperature, specific humidity, and pressure, we derived relative humidity using methods described by  \cite{bolton_computation_1980}.
-Then, based on the Heat Index algorithm developed by Steadman \cite{steadman_assessment_1979} and improved upon by Rothfusz and the National Weather service \cite{rothfusz_heat_1990}, we calculated the Heat Index using temperature and relative humidity \cite{heo_comparison_2019}.
-
-ISSUES: deciduous trees lose leaves every year and may not provide much shade cover and heat relief during particular periods of the year when they are leafless.
-
-\subsection{Income}
-We used data from the US Census at the Census Block level to estimate the
-
-\subsection{Land Cover}
-To estimate land cover categories and the percentage of tree cover and impervious surfaces, we used data from the National Land Cover Database (NLCD) from the USGS' 2011  \cite{homer_completion_2015} and 2016 \cite{yang_new_2018} landcover datasets.
-These datasets are national land cover dataset prepared by the Multi-Resolution Land Characteristics (MRLC) Consortium.
-These data are at the thirty meter resolution, and provide estimates of the fraction of tree cover and impervious surfaces.
-We then associated each tweet with the land cover fraction for the nearest available year and pixel using the geolocation associated with each tweet, extracting landcover values from whichever land cover dataset year was closest in time to the date timestamp of each tweet.
-We used tree cover as a proxy for amount of potential shade and heat relief in a particular area and can be broadly indicative of generalized wealth.
-We use impervious surface cover as a proxy for relative proximity to potential natural shade and heat relief, and because individuals in urban environments may be exposed to the urban heat island effect.
-[NEED SOME CITATIONS HERE]
-
-\subsection{Sentiment}
-We assessed the sentiment expressed in text using several of the most commonly used measures of sentiment.
-Each of the measures uses validated dictionaries of words that have been assigned positive and negative scores and compares these words with words in the text corpora of interest.
-For measures of sentiment, we used the Hedonometer \href{https://hedonometer.org/timeseries/en_all/}{(link)}, Afinn \href{http://corpustext.com/reference/sentiment_afinn.html}{(link)}, Textblob \href{https://textblob.readthedocs.io/en/dev/}{(link)}, Vader , SentiwordNet and LIWC tools.
-These tools assign positive and negative scores to words, then average the scores for each text string.
-Our text data was multilingual, which can present challenges with analyzing text (cite).
-Given our interest only in the collection of words used (in aggregate) and not the sequence or nuance of the phrases, we translated all languages into English using Google Translate API and the Microsoft Azure API cognitive services (translator text API) before conducting the sentiment analyses.
-
-Some papers we might cite in this section (only for multilingual work): Vilares et al. 2017 , Lo et al. 2017, Dashtipour et al. 2016.
-
-Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! \#fail]), etc.
-
-Issues with sample population (non-representative)
-
-Issues with what people say on social media being attenuated or mismatched from actual emotional state
-
-\subsubsection{Hedonometer}
-The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
-The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2)using a naive algorithm for scaling up from individual words to texts.
-For the English corpus, Dodds and others \cite{dodds_temporal_2011} collect 10,222 unique words and used crowd-sourcing platform Amazon Mechanical Turk to get human evaluation of happiness degree for each word in an integer scale from 1 to 9, representing a sad to happy spectrum.
-Score 5 represents neutral words.
-Each word will be calculated average score and then the word and happiness score are compiled into a dataset (labMT 1.0).
-Some illustrative example of words are:
-
-\[h_{avg} (\text{laughter}) = 8.50 \]
-\[h_{avg} (\text{the}) = 4.98\]
-\[h_{avg} (\text{hate}) = 2.34\]
-
-The sentiment score for single text will be the mean happiness score for all words in the text.
-
-\subsubsection{VADER}
-The VADER sentiment corpus was proposed by Gilbert and Hutto \cite{gilbert_vader_2014} and stands for 'Valence Aware Dictionary for sentiment Reasoning'.
-It is is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media by incorporating lexical features common to sentiment expression in microblogs
-
-\subsection{Analytic Approach}
-To assess whether temperature affects sentiment, we ran a linear regression model that included meteorological measures and sentiment.
-Independent variables included maximum daily temperature, precipitation and heat index.
-We also included fixed effects indicator variables at the year-month and decimal degree levels to account for potential biases in unobserved characteristics associated with seasonality and date.
-We examined the relationships between meteorological variables and sentiment using linear regression models.
-
-\subsection{Data Processing}
-We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 2019) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
-Sentiment data was analyzed using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
-We extracted meteorological data using the geolocated coordinates from the Twitter posts (latitude and longitude).
-We calculated mean climatological scores (z-scores)using baseline meteorological data from 1990 to present, calculated at a quarter degree resolution, and based on the long-term averages and standard deviation.
-We calculated z-scores for each day and the overall annual average z-score.
-For text data cleaning we used the python package tweet-preprocessor \href{https://pypi.org/project/tweet-preprocessor/}{(link)} to remove all text and characters not part of the main text phrase (e.g., URLs, mentions, GIFs, etc.).
-We converted all dates and times to UTC format.
-
-\section{Results}
-(sub-section) DESCRIPTIVE STATISTICS
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
-For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
-For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
-  \label{fig:wbgt-income}
-\end{figure}
-
-Figure of heatmap of tweet density across USA (pick a year? entire dataset of georeferenced tweets?)
-
-(sub-section) TEMPERATURE and HUMIDITY
-Figure of distribution of max and min temperatures
-Comment from Patrick Baylis on our "preliminary results". 2nd comment from Patrick
-
-(sub-section) INCOME
-
-(sub-section?) SENTIMENT
-Figure of distribution of observed sentiment scores (by Hedonometer, Vader, etc.)
-
-Comment from Patrick Baylis on our "preliminary results" doc
-
-\section{Discussion}
-
-\section{Conclusion}
-
-\section{Acknowledgements}
-For funding, advising, SESYNC grant number?, etc.
+\setcounter{table}{0}
+\setcounter{figure}{0}
+\setcounter{section}{0}
+\renewcommand{\thetable}{S\arabic{table}}
+\renewcommand{\thefigure}{S\arabic{figure}}
+\renewcommand{\thesection}{S\arabic{section}}
 
 \printbibliography
+
+\end{document}
 
@@
 
-\section*{Text Snippets Not Used Here but We Might Want to Keep}
-Gaps on above we intend to fill in:
-Our research aim and objectives
-Its contribution
-
-From Intro/Background
-Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes .
-Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
-(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
-
-[from our Jan/Feb 2020 draft working manuscript in Google Docs]
-Climate change is a universal issue as it affects every country in the world.
-Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
-Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
-(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
-
-Dehghan et al.
-2012  The evaluation of heat stress through monitoring environmental factors and physiological responses in melting and casting industries workers.
-High prevalence of industrial workers experiencing heat stress.
-Relatively low humidity conditions made it necessary to alter their work/rest cycles.
-\href{http://www.ijehe.org/article.asp?issn=2277-9183;year=2012;volume=1;issue=1;spage=21;epage=21;aulast=Dehghan}{link}
-
-?.
-(from intro) Increased variability around environmental norms (e.g., lower/higher temperatures, changes in timing, magnitude, duration of precipitation or dry spells, etc.) can amplify the effects on systems.
-Increased uncertainty in planning for these events and ability to respond to them (adaptive capacity).
-- draw in cultural risk and social practice theories, here?
-
-NOTES: Scaling back out → we are exploring temperature and sentiment (in Tweets) across languages.
-This is a clear innovation that builds on the existing work from the Baylis et al.
-team (in their plosONE paper) and in fact is something they recommended be done - recommended to us as a team, too.
-
-We know that temps affect sentiment, but (1) Is this effect observable across languages? (2) Is this effect moderated by factors like wealth?
-
-We could ask similar questions substituting in the patent filings, too (sub-project 3 in our previous work-plan).
-These questions - and likely the methods used to answer these questions - are quite portable and compelling and may take the form for the patent work like this: Given we know that temperature affects worker productivity (increased temp increases productivity up to 13C, sharp negative decline thereafter), can we observe this pattern in patent filings?
-Across various geographies?
-Are there lags in filings after unusually cold or hot temperature events?
-
-For the Twitter data, using text translated into a base language is fully appropriate and a frequently used method if looking for latent text structures in topic modeling (see Proksch et al.
-2018 for an alternative).
-For sentiment analysis on multilingual datasets, we'll need to decide on an approach (see Ling Lo et al.
-2016 review) but there are several that are good.
-If we wanted to use another open-source option, we could follow that laid out by Lucas and others (2015).
-
-Proskch et al. 2018 paper --> \href{https://onlinelibrary.wiley.com/doi/full/10.1111/lsq.12218}{link}
-
-Ling Lo et al. 2016 paper --> \href{https://link.springer.com/article/10.1007/s10462-016-9508-4}{link}
-
-Lucas et al. 2015 paper --> \href{https://www.cambridge.org/core/journals/political-analysis/article/computerassisted-text-analysis-for-comparative-politics/CC8B2CF63A8CC36FE00A13F9839F92BB}{link}
-
 \end{document}

### 2021-04-07 doc/temp-sentiment.tex
work on document locally

@@
 \begin{document}
-https://www.overleaf.com/project/5f918edd1afc380001feca0d
-\title{Fine-scale Twitter Data Reveals Neighborhood Inequalities in Heat Wave Vulnerability}
+%https://www.overleaf.com/project/5f918edd1afc380001feca0d
+\title{Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures}
 
-%\author[1, *]{Matthew Cooper}
-%\author[]{Jeremiah Osborne-Gowey}
-%\author[]{Zheng Liu}
-%\author[]{Jie Liu}
-%\author[]{Portia Adade Williams}
-%\author[]{Aaron Schwartz}
-%\author[]{Patrick Baylis}
+\author[1, *]{Matthew Cooper}
+\author[2]{Jeremiah Osborne-Gowey}
+\author[3]{Zheng Liu}
+\author[4]{Jie Liu}
+\author[5]{Portia Adade Williams}
+\author[6]{Aaron Schwartz}
+\author[7]{Patrick Baylis}
 
-%\affil[1]{T.H. Chan School of Public Health, Harvard University}
-%\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
+\affil[1]{T.H. Chan School of Public Health, Harvard University}
+\affil[2]{University of Colorado Boulder}
+\affil[3]{Department of Geographical Sciences, University of Maryland College Park}
+\affil[4]{East China University of Science and Technology}
+\affil[5]{University of Cape Town}
+\affil[6]{University of Colorado Boulder}
+\affil[7]{University of British Columbia}
+\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
 
@@
 \begin{abstract}
-Weather can affect people’s mood and well-being.
-Traditional studies exploring the effects of weather on mental health require significant resource investment (e.g., time, funding, execution).
-Continuous in situ data streams from social media, newspapers, and other big data sources offer an opportunity to examine how people perceive and respond to environmental conditions.
-Recent work demonstrates that prevailing weather conditions can affect people’s sentiment as expressed on Twitter and other social media.
-Studies to date however, have modeled the weather as having the same effect on sentiment across all locations and individuals (i.e., in aggregate).
-In this study, we explore how weather affects individual expressed sentiment (via Twitter) across different weather gradients and locations.
-We analyzed the sentiment from a quarter of a billion geolocated Tweets from 2009 to 2019, using multiple measures of sentiment, overlaid with data about prevailing weather conditions, as well as local land cover, income levels, and demographic characteristics in the vicinity of Twitter users.
-We find that weather affects people differently across a range of incomes, with people in higher income brackets generally happier with increasing temperature, while people in lower income brackets are less happy with increasing temperature.
-This project extends existing research about the relationship between temperature and mood (sentiment) by examining how varying climatic conditions across different geographies and socio-economic groups are correlated with expressed sentiment.
-This project brings new insights into how humans perceive and respond to environmental shocks, with important implications for economics, policy and adaptation and resilience planning.
+%Limit 150 words: https://www.nature.com/nclimate/about/content
+%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
+Higher temperatures associated with climate change are expected to have major impacts on human mental health.
+Indeed, traditional analyses of heat and mental health outcomes using data collected by public agencies have found associations between elevated temperatures and increases in outcomes such as suicides and mental-health related hospitalizations.
+However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income.
+We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather and neighborhood socioeconomic conditions.
+We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.
 
@@
 %Intro Paragraph
+% I think this should be more about climate change and mental health broadly, as well as a bit about how income has been theorized to affect vulnerability, but there is no evidence yet.
+This can be shorter as well, as we have a lot of detail in the Background section -Matt
 Research has shown strong linkages between ambient environmental conditions and mental health, with extreme temperatures  frequently associated with poorer overall mental health.
@@
   \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  %\caption{WBGT-sentiment relationship as moderated by neighborhood income.}
+  \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income.}
   \label{fig:sub1}
@@
   \includegraphics[width=\linewidth]{../res/wbgt-race_q.png}
-  %\caption{WBGT-sentiment relationship as moderated by neighborhood majority race.}
+  \caption{WBGT-sentiment relationship as moderated by neighborhood majority race.}
   \label{fig:sub2}
@@
 
-\section*{Supplemental Info}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thesection}{S\arabic{section}}
-
-\subsection{Other Sentiment Metrics}
-
-\subsection{Precipitation}
-
-%Supplement will have:
-
-%Equations used for RH and WBGT
-
-% Robustness checks:
-
-    % Other sentiment metrics
-
-    %
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-% Paste Bin
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\subsection{Land Cover}
-To estimate land cover categories and the percentage of tree cover and impervious surfaces, we used data from the National Land Cover Database (NLCD) from the USGS' 2011  \cite{homer_completion_2015} and 2016 \cite{yang_new_2018} landcover datasets.
-These datasets are national land cover dataset prepared by the Multi-Resolution Land Characteristics (MRLC) Consortium.
-These data are at the thirty meter resolution, and provide estimates of the fraction of tree cover and impervious surfaces.
-We then associated each tweet with the land cover fraction for the nearest available year and pixel using the geolocation associated with each tweet, extracting landcover values from whichever land cover dataset year was closest in time to the date timestamp of each tweet.
-We used tree cover as a proxy for amount of potential shade and heat relief in a particular area and can be broadly indicative of generalized wealth.
-We use impervious surface cover as a proxy for relative proximity to potential natural shade and heat relief, and because individuals in urban environments may be exposed to the urban heat island effect.
-[NEED SOME CITATIONS HERE]
-
-\section*{Text Snippets Not Used Here but We Might Want to Keep}
-Gaps on above we intend to fill in:
-Our research aim and objectives
-Its contribution
-
-From Intro/Background
-Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes .
-Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
-(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
-
-[from our Jan/Feb 2020 draft working manuscript in Google Docs]
-Climate change is a universal issue as it affects every country in the world.
-Changes in climate result in fluctuations in the frequency, intensity, spatial extent, duration, and timing of weather resulting in unprecedented climate extremes \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
-Changing climate is putting pressures on environmental and social system health and necessitating changes in how we respond to environmental conditions and anticipate (and plan for) future conditions.
-(cite IPCC report on health and environment, also cite the Lancet report on climate and health).
-
-Dehghan et al.
-2012  The evaluation of heat stress through monitoring environmental factors and physiological responses in melting and casting industries workers.
-High prevalence of industrial workers experiencing heat stress.
-Relatively low humidity conditions made it necessary to alter their work/rest cycles.
-\href{http://www.ijehe.org/article.asp?issn=2277-9183;year=2012;volume=1;issue=1;spage=21;epage=21;aulast=Dehghan}{link}
-
-?.
-(from intro) Increased variability around environmental norms (e.g., lower/higher temperatures, changes in timing, magnitude, duration of precipitation or dry spells, etc.) can amplify the effects on systems.
-Increased uncertainty in planning for these events and ability to respond to them (adaptive capacity).
-- draw in cultural risk and social practice theories, here?
-
-NOTES: Scaling back out → we are exploring temperature and sentiment (in Tweets) across languages.
-This is a clear innovation that builds on the existing work from the Baylis et al.
-team (in their plosONE paper) and in fact is something they recommended be done - recommended to us as a team, too.
-
-We know that temps affect sentiment, but (1) Is this effect observable across languages? (2) Is this effect moderated by factors like wealth?
-
-We could ask similar questions substituting in the patent filings, too (sub-project 3 in our previous work-plan).
-These questions - and likely the methods used to answer these questions - are quite portable and compelling and may take the form for the patent work like this: Given we know that temperature affects worker productivity (increased temp increases productivity up to 13C, sharp negative decline thereafter), can we observe this pattern in patent filings?
-Across various geographies?
-Are there lags in filings after unusually cold or hot temperature events?
-
-For the Twitter data, using text translated into a base language is fully appropriate and a frequently used method if looking for latent text structures in topic modeling (see Proksch et al.
-2018 for an alternative).
-For sentiment analysis on multilingual datasets, we'll need to decide on an approach (see Ling Lo et al.
-2016 review) but there are several that are good.
-If we wanted to use another open-source option, we could follow that laid out by Lucas and others (2015).
-
-Proskch et al. 2018 paper --> \href{https://onlinelibrary.wiley.com/doi/full/10.1111/lsq.12218}{link}
-
-Ling Lo et al. 2016 paper --> \href{https://link.springer.com/article/10.1007/s10462-016-9508-4}{link}
-
-Lucas et al. 2015 paper --> \href{https://www.cambridge.org/core/journals/political-analysis/article/computerassisted-text-analysis-for-comparative-politics/CC8B2CF63A8CC36FE00A13F9839F92BB}{link}
-
 \end{document}

### 2021-04-08 doc/temp-sentiment.tex
more work on introduction paragraphs

@@
 % I think this should be more about climate change and mental health broadly, as well as a bit about how income has been theorized to affect vulnerability, but there is no evidence yet.
-This can be shorter as well, as we have a lot of detail in the Background section -Matt
+This can be shorter as well, as we have a lot of detail in the Background section
+% Cite this high-level study: https://www.nature.com/articles/s41558-018-0102-4/
+% Also this review article saying more research is needed https://www.sciencedirect.com/science/article/pii/S0033350618302130
+% -Matt
+
 Research has shown strong linkages between ambient environmental conditions and mental health, with extreme temperatures  frequently associated with poorer overall mental health.
@@
 \subsection{Temperature and Mental Health}
+Higher temperatures are associated with many indicators of worsened mental health.
+Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2006Aug, Likhvar2011Jan}.
+In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mulllins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
+
+Temperature has been hypothesized to impact mental health through a number of pathways.
+Work on biological mechanisms emphasize that there may be mental health effects from maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.
+Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}
+
 % Heatwaves are known to worsen mental health
@@
 \subsection{Mental Health and Income}
+% This is good, but probably should be shortened.  Maybe one para showing that income matters, one para discussing pathways, also maybe focus/change phrasing to be more on "vulnerability" -Matt
+
 % AND mental health is strongly associated with income
@@
 
-The relationship between income, socioeconomic status (SES), well-being and mental health is one of the strongest established patterns in psycho-social literature (cite: Easterlin 1973, Holzer et al.
-1986, Perry 1996, Ng et al.
+The relationship between income, socioeconomic status (SES), well-being and mental health is one of the strongest established patterns in psycho-social literature \citep{Easterlin1974Jan, holzer1986increased, Perry1996Sep} (I cant find Ng et al.
 2014).
 Much of the research into these patterns focuses on the inequalities and distribution of the effects on individual well-being.
-Subsequent work on mental health outcomes, one measure of well-being, have established strong associations between SES, income and mental health with the most and least privileged most commonly associated with the best and worse mental health experiences and outcomes, respectively (cites: Stevenson and Wolfers 2008, Ng et al.
+Subsequent work on mental health outcomes, one measure of well-being, have established strong associations between SES, income and mental health with the most and least privileged most commonly associated with the best and worse mental health experiences and outcomes, respectively \citep{Sevenson2008Aug}(cites: Ng et al.
 2014).
 
-For example, results from a cross-sectional comparative study of socioeconomic factors and use of mental health services by people living in Ontario, Canada and the United States, indicate clear disparities in use of mental health services between those with the highest income levels and lowest mental health morbidity and those with the lowest income levels and highest mental health morbidity - those with higher incomes tended to have lower mental health morbidity issues while utilizing mental health services more than those with lower incomes and higher mental health issues (cite Katz et al.
-1997).
+For example, results from a cross-sectional comparative study of socioeconomic factors and use of mental health services by people living in Ontario, Canada and the United States, indicate clear disparities in use of mental health services between those with the highest income levels and lowest mental health morbidity and those with the lowest income levels and highest mental health morbidity - those with higher incomes tended to have lower mental health morbidity issues while utilizing mental health services more than those with lower incomes and higher mental health issues \citep{Katz1997}.
 
-Results from a systematic review of the literature on associations between social inequalities and mental health disorders indicate a clear and prevailing link between  one or more indicators of less social privilege and higher prevalence of mental health disorders (cite Fryers et al.
-2003), with low income one of the most consistent markers of and associations with increases in common mental health disorders.
+Results from a systematic review of the literature on associations between social inequalities and mental health disorders indicate a clear and prevailing link between  one or more indicators of less social privilege and higher prevalence of mental health disorders \citep{Fryers2003}, with low income one of the most consistent markers of and associations with increases in common mental health disorders.
 Collectively, results from the income and mental health literature indicate that socially disadvantaged populations experience significantly more frequent common mental health disorders.
-Scholars use mental health disparities to indicate the disproportionate amount of mental health disorders among persons of low SES\citep{RN1292}
+Scholars use mental health disparities to indicate the disproportionate amount of mental health disorders among persons of low SES \citep{RN1292}
 
@@
 For instance, some systematic reviews have shown that socioeconomically disadvantaged children and adolescents were 2-3 times more likely to have mental health problems than others\citep{REISS201324}, while the household increases in financial resources are generally associated with an overall reduction of mental health problems\citep{2015Does}.
-Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress (cite Kawachi and Berkman 2001).
+Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress \citep{Kawachi2001Sep}.
 Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease menta the
@@
 \label{fig:test}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
-For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
-For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
-  \label{fig:wbgt-income}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-race.png}
-  \caption{Change in sentiment across census block income groups as the Wet Bulb Globe Temperature increases.
-For census blocks at the 5th percentile for income, even mild temperatures are associated with decreases in mood, while at higher temperatures, sentiment continues to decline.
-For high-income census blocks at the 95th percentile, on the other hand, temperature must be quite high before a decrease in sentiment is observed.}
-  \label{fig:wbgt-income}
 \end{figure}

### 2021-04-12 doc/temp-sentiment.tex
acknowledgements

@@
 
+% More nuanced interpretation: optimal temperature for poor & minority neighborhoods are much lower.
+
 \section{Conclusion}
@@
 \section{Acknowledgements}
-For funding, advising, SESYNC grant number?, etc.
 This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
+Additionally, cloud resources for this project were provided for the Microsoft Azure cloud by Microsoft's AI for Earth program, grant number .
 

### 2021-04-13 crime/combine_nibrs_cod_day.R
add notes on violent crimes

@@ -1,5 +1,7 @@
 library(data.table)
 
+# Definition of violent crimes: https://www.law.cornell.edu/uscode/text/18/16
+
 ###############################
 # read nibrs data
 ###############################

### 2021-04-13 crime/model_day.R

@@ -42,6 +42,7 @@ comb$doy <- substr(comb$date, 6, 10)
 gd <- expand.grid(list(out=c('violent', 'assault', 'homicides'),
                        src=c('nibrs', 'cod'),
                        ind=c('vader', 'afinn', 'hedono')))
+
 for (i in 1:nrow(gd)){
   form <- as.formula(paste0(gd$out[i], '_', gd$src[i], ' ~ ', gd$ind[i], ' + log(population) | dow + doy + year + statemonth'))
   mod <- feglm(form, data=comb, family=poisson(link=log), weight=comb$n)

### 2021-04-14 csa/make_fips_csa_table.R
preprocess fips and csa data

@@ -1,0 +1,18 @@
+library(tidyverse)
+library(sf)
+
+fips <- read_sf('~/tweets/csa/', 'USA_Counties') %>%
+  st_centroid %>%
+  select(FIPS, POPULATION)
+
+csa <- read_sf('~/tweets/csa/', 'Combined_Statistical_Areas') %>%
+  select(CSA)
+
+j <- st_join(fips, csa, join=st_within) %>%
+  filter(!is.na(CSA)) %>%
+  group_by(CSA) %>%
+  mutate(CSA_POPULATION = sum(POPULATION)) %>%
+  select(-POPULATION) %>%
+  st_drop_geometry
+
+write.csv(j, '~/tweets/csa/fips-csa.csv', row.names=F)

### 2021-04-14 csa/make_fips_csa_table.R
update csa-msa processing

@@ -1,18 +1,52 @@
 library(tidyverse)
 library(sf)
+library(USAboundaries)
 
 fips <- read_sf('~/tweets/csa/', 'USA_Counties') %>%
+  filter(!STATE_NAME %in% c('Alaska', 'Hawaii', 'Puerto Rico')) %>%
   st_centroid %>%
   select(FIPS, POPULATION)
 
+msa <- read_sf('~/tweets/csa', 'fb19722c-442d-4b5e-8649-268605c4817b2020313-1-17yu7nk.hr3gl') %>%
+  filter(CBSA_TYPE == 'Metropolitan') %>%
+  select(MSA_ID = CBSA_ID, MSA_NAME = NAME)
+
 csa <- read_sf('~/tweets/csa/', 'Combined_Statistical_Areas') %>%
-  select(CSA)
+  select(CSA_ID = CSA, CSA_NAME = NAME)
 
 j <- st_join(fips, csa, join=st_within) %>%
-  filter(!is.na(CSA)) %>%
-  group_by(CSA) %>%
-  mutate(CSA_POPULATION = sum(POPULATION)) %>%
-  select(-POPULATION) %>%
-  st_drop_geometry
+  st_join(msa, join=st_within)
 
-write.csv(j, '~/tweets/csa/fips-csa.csv', row.names=F)
+j2 <- j %>%
+  filter(!(is.na(CSA_ID) & is.na(MSA_ID))) %>%
+  group_by(CSA_ID) %>%
+  mutate(CSA_POP = sum(POPULATION)) %>%
+  group_by(MSA_ID) %>%
+  mutate(MSA_POP = sum(POPULATION)) %>%
+  st_drop_geometry %>%
+  arrange(CSA_POP, MSA_POP)
+j2$MSA_POP[is.na(j2$MSA_ID)] <- NA
+j2$CSA_POP[is.na(j2$CSA_ID)] <- NA
+
+#filter to CSA_POP > 1 mill OR MSA_POP > 1 mill
+j2 <- j2 %>%
+  filter(CSA_POP > 1000000 |  MSA_POP > 1000000) %>%
+  select(-matches('NAME')) %>%
+  mutate(id = ifelse(is.na(CSA_ID), MSA_ID, CSA_ID),
+         pop = ifelse(is.na(CSA_POP), MSA_POP, CSA_POP)) %>%
+  ungroup %>%
+  select(fips=FIPS, id, pop)
+
[…]

### 2021-04-15 csa/model_space.R
CSA and hourly analyses

@@ -1,0 +1,142 @@
+library(data.table)
+library(tidyverse)
+library(fixest)
+library(broom)
+library(sf)
+library(USAboundaries)
+
+data <- fread('~/tweets/all.csv')
+csadat <- fread('~/tweets/csa/fips-csa.csv')
+
+data <- merge(data, csadat, on='fips', all=F)
+
+data$month <- substr(data$statemonth, 4, 5)
+data$income_percap <- log(data$income_percap)
+data <- data[weather_term == 0, ]
+data$raining <- data$prcp > 0
+
+qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
+
+csas <- tab(data$id)
+csas <- as.numeric(names(csas))[order(csas, decreasing=T)]
+
+#############################################
+# Now effects of temp*income and temp*race
+#############################################
+
+allres <- data.frame()
+for (csasel in csas){
+  print(which(csas == csasel)/length(csas))
+
+  sel <- data[data$id == csasel & data$wbgt > 5, ]
+
+  mod <- feols(vader ~ wbgt + raining + srad | dow + doy + tod + fips + year + month, data=sel)
+  base_res <- tidy(mod) %>%
+    filter(term =='wbgt') %>%
+    mutate(id = csasel)
+
+  mod <- feols(vader ~ wbgt*income_percap + raining*income_percap + srad*income_percap | dow + doy + tod + fips + year + month, data=sel)
+  inc_res <- tidy(mod) %>%
+    filter(term =='wbgt:income_percap') %>%
+    mutate(id = csasel)
+
+  mod <- feols(vader ~ wbgt*race_white + raining*race_white + srad*race_white | dow + doy + tod + fips + year + month, data=sel)
+  race_res <- tidy(mod) %>%
+    filter(term == 'wbgt:race_white') %>%
+    mutate(id = csasel)
+
+  allres <- bind_rows(allres, base_res, inc_res, race_res)
+}
+
+#########################
+# Visualize
+############################
+csa_sf <- read_sf('~/tweets/csa', 'msa-csa')
+
+states <- us_states(resolution='low') %>%
+  filter(!state_name %in% c("Virgin Islands", "Alaska", "Hawaii", 
[…]

### 2021-04-15 csa/model_ts.R

@@ -1,0 +1,171 @@
+library(data.table)
+library(tidyverse)
+library(fixest)
+library(broom)
+library(hms)
+library(cowplot)
+
+options(scipen=100)
+
+data <- fread('~/tweets/all.csv')
+
+data$month <- substr(data$statemonth, 4, 5)
+data$income_percap <- log(data$income_percap)
+data <- data[weather_term == 0, ]
+data$raining <- data$prcp > 0
+data$hour <- substr(data$tod, 1, 2)
+
+qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
+
+tods <- tab(data$hour[data$wbgt > 5])
+tods <- data.frame(tod=names(tods), count=as.numeric(tods))
+
+#############################################
+# Now effects of temp*income and temp*race
+#############################################
+
+allres <- data.frame()
+for (todsel in tods$tod){
+  print(which(tods == todsel)/nrow(tods))
+
+  sel <- data[data$hour == todsel & data$wbgt > 5, ]
+
+  mod <- feols(vader ~ wbgt + raining + srad | dow + doy + tod + fips + year + month, data=sel)
+  base_res <- tidy(mod) %>%
+    filter(term =='wbgt') %>%
+    mutate(tod = todsel)
+
+   mod <- feols(vader ~ wbgt*income_percap + raining*income_percap + srad*income_percap | dow + doy + tod + fips + year + month, data=sel)
+   inc_res <- tidy(mod) %>%
+     filter(term =='wbgt:income_percap') %>%
+     mutate(tod = todsel)
+
+   mod <- feols(vader ~ wbgt*race_white + raining*race_white + srad*race_white | dow + doy + tod + fips + year + month, data=sel)
+   race_res <- tidy(mod) %>%
+     filter(term == 'wbgt:race_white') %>%
+     mutate(tod = todsel)
+
+  allres <- bind_rows(allres, base_res, inc_res, race_res)
+}
+
+#########################
+# Visualize
+############################
+
+allres$tod <- hms::as_hms(paste0(allres$tod, ':00:00'))
+tods$tods <-  hms::as_hms(paste0(tods$tod, ':00:00'))
+
+############################
+# Baseline effect
[…]

### 2021-04-15 doc/temp-sentiment.tex
Updates from Overleaf

--- 2021-04-08 doc/temp-sentiment.tex
@@
 \affil[3]{Department of Geographical Sciences, University of Maryland College Park}
-\affil[4]{East China University of Science and Technology}
+\affil[4]{School of Business, East China University of Science and Technology}
 \affil[5]{University of Cape Town}
@@
 %Cite this new paper: https://doi.org/10.1016/S2542-5196(20)30251-5
-% And: https://eos.org/articles/long-term-drought-harms-mental-health-in-rural-communities
+% And: https://eos.org/articles/long-term-drought-harms-mental-health-in-rural-communities and this is the conf. paper citation https://agu.confex.com/agu/fm20/meetingapp.cgi/Paper/771750
 \section{Introduction}
@@
 Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2006Aug, Likhvar2011Jan}.
-In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mulllins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
+In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mullins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
 
@@
 Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}
+
+% The pathways
+Higher temperature associated with climate change could influence mental health directly by exposing people to trauma.
+It could also indirectly influence mental health by affecting (1) physical health and (2)community wellbeing\citep{RN1314}.
+For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions(Anderson and Anderson 1998,Cheatwood 1995; Cohn et al.
+2004,Bouchama et al.2007.).
+Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration((Bridger 2003; Kerslake 1972; Kjellstrom et al.2009d,Leithead and Lind 1964,Schrier et al.
+1970).
+The loss of work capacity would in turn result in loss of income, which could also cause mental health problems(Kjellstrom 2009b).
+
+As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems(Miller et al.
+2009; Prince et al.
+2007), heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health((Berry 2007,World Health Organization 2008b).
 
@@
 
-\subsection{Mental Health and Income}
+\subsection{Mental Health and Vulnerability}
 % This is good, but probably should be shortened.  Maybe one para showing that income matters, one para discussing pathways, also maybe focus/change phrasing to be more on "vulnerability" -Matt
@@
 % Perceived income level
+
+% Talk more about race
 
@@
 Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress \citep{Kawachi2001Sep}.
-Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease menta the
+Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease mental health problems.
 
-Several reason could explain why poverty may effect mental health.
+Several reasons could explain why poverty may effect mental health.
 1) the “social causation hypothesis” which suggests that stress or deprivation or decreasing the likelihood of people getting treatment may lead to poor mental health\citep{mills2015}, while social capital could reduce the likelihood of living in poverty and reduces the risk of mental disorders such as depressive symptoms and suicidal tendencies\citep{RN1291}.
@@
   \includegraphics[width=\linewidth]{../res/wbgt-race_q.png}
-  \caption{WBGT-sentiment relationship as moderated by neighborhood majority race.}
+  \caption{Effect of changes in wet-bulb globe temperature on expressed entiment relationship as moderated by neighborhood majority race.}
   \label{fig:sub2}
@@
 We only considered users’ original content, thus did not include retweets in our analysis.
+
+We excluded all tweets that contained weather-related terms.
 

### 2021-04-15 extract/Export_for_Jeremiah.R
export for jeremiah

@@ -1,0 +1,27 @@
+library(data.table)
+
+dat <- fread('~/tweets/all.csv')
+
+cty <- read.csv('~/tweets/csa/fips-csa.csv')
+
+cincinnati_fips <- cty$fips[cty$id == 178]
+
+sel <- dat[dat$fips %in% cincinnati_fips, ]
+
+sel$month <- substr(sel$statemonth, 4, 5)
+
+sel2 <- sel[sel$month == '05', ]
+
+sel3 <- sel2[sel2$weather_term == 0 , 
+             .(temp = mean(temp), 
+               prcp = mean(prcp), 
+               srad = mean(srad), 
+               temp.hi = mean(temp.hi),
+               afinn = mean(afinn),
+               hedono = mean(hedono),
+               vader = mean(vader),
+               income_percap = mean(income_percap),
+               race_white = mean(race_white)),
+             .(fips, dow, year, doy)]
+
+fwrite(sel3, '~/tweets/cincinnati_sel.csv', row.names=F)

### 2021-04-28 doc/temp-sentiment.tex

@@
 \subsection{Temperature and Mental Health}
+%https://www.pnas.org/content/115/43/10953
 Higher temperatures are associated with many indicators of worsened mental health.
@@
 1970).
-The loss of work capacity would in turn result in loss of income, which could also cause mental health problems(Kjellstrom 2009b).
+The loss of work capacity would in turn result in loss of income, which could also cause mental health problems(Kjellstrom 2009b).%% more physiological terms
 
@@
 2007), heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health((Berry 2007,World Health Organization 2008b).
+Second，disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities.
+For example, extreme heat reduces the work capacity of laborers in farm fields(Kjellstrom et al.
+2009d), which further destroys agricultural-supported industries in local area(Berry et al.
+2008).
+The following economic pressures would undermine social capital and then lead to mental health problems.
 
@@
 % Talk more about race
+% This is interesting: https://www.pnas.org/content/118/17/e2019624118.abstract?etoc
 
@@
 \section{Results}
+
+\subsection{Overall Effect}
 We examined the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an indicator of heat stress that accounts for temperature, humidity, wind speed, and solar radiation.
@@
   \includegraphics[width=\linewidth]{../res/wbgt-race_q.png}
-  \caption{Effect of changes in wet-bulb globe temperature on expressed entiment relationship as moderated by neighborhood majority race.}
+  \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood majority race.}
   \label{fig:sub2}
@@
 \label{fig:test}
+\end{figure}
+
+\subsection{Effects by Combined Statistical Area}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\linewidth]{../res/map_wbgt.png}
+  \caption{Effects of rising temperatures on sentiment by Combined Statistical Areas (CSAs) with over 1 million people.
+The value shown is the predicted change in vader score for a 25-degree increase in temperatures.
+CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.}
+  \label{fig:map-wbgt}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\linewidth]{../res/map_wbgt_income.png}
+  \caption{Expected gap in sentiment scores between poor and rich neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
+The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 90th percentile for income.
+CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.}
+  \label{fig:map-wbgt-income}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\linewidth]{../res/map_wbgt_race.png}
+  \caption{Expected gap in sentiment scores between largely minority and largely white neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
+The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 95th percentile for whiteness.
+CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.}
+  \label{fig:map-wbgt-race}
+\end{figure}
+
+\subsection{Effects by Time of Day}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\linewidth]{../res/ts.png}
+  \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted change in vader score for a 25-degree increase in temperatures.}
+  \label{fig:map-wbgt}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\linewidth]{../res/ts_income.png}
+  \caption{Expected gap in sentiment scores between largely minority and largely white neighborhoods, by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 95th percentile for income.}
+  \label{fig:map-wbgt-income}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.75\linewidth]{../res/ts_race.png}
+  \caption{Expected gap in sentiment scores between largely minority and largely white neighborhoods, by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 95th percentile for whiteness.}
+  \label{fig:map-wbgt-race}
 \end{figure}
@@
 Appendix table. Sentiment scores across all types we considered.
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/Boston_Map.png}
-  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
-  \label{fig:timeseries}
-\end{figure}
 

### 2021-04-28 doc/temp-sentiment.tex
.

@@
 However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income.
-We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather and neighborhood socioeconomic conditions.
+We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
 We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.

### 2021-04-28 viz/temporal_dynamics.R
examine temporal trends

@@ -1,0 +1,40 @@
+library(data.table)
+
+setwd('~/tweets/')
+
+data <- fread('all.csv')
+
+# Day of Week, Time of Day
+dow_tod <- data[ , .(vader = mean(vader)), .(dow, tod)]
+
+dow_tod[dow_tod$vader %in% c(max(dow_tod$vader), min(dow_tod$vader)), ]
+
+        dow  tod      vader
+1: Thursday 04:0 0.07818882
+2:   Sunday 08:0 0.21021752
+
+# Day of week
+dow_df <- data[ , .(vader = mean(vader)), .(dow)]
+
+dow_df[dow_df$vader %in% c(max(dow_df$vader), min(dow_df$vader)), ]
+
+         dow     vader
+1:    Sunday 0.1366292
+2:  Saturday 0.1371815
+3:    Friday 0.1338290
+4:  Thursday 0.1305909
+5: Wednesday 0.1294679
+6:   Tuesday 0.1267888
+7:    Monday 0.1269388
+
+# Day of Year
+doy_df <- data[ , .(vader = mean(vader)), .(doy)]
+doy_df <- doy_df[doy_df$doy != '02-29', ]
+
+doy_df[doy_df$vader %in% c(max(doy_df$vader), min(doy_df$vader)), ]
+
+     doy     vader
+1: 12-25 0.2080047
+2: 12-14 0.1201150
+
+

### 2021-04-28 csa/model_space.R
filter CSA/MSAs with < 5% black

@@ -151,6 +151,8 @@ ggsave('~/temp-sentiment/res/map_wbgt_race.png', width=12, height=8)
 ###########################
 m <- merge(csa_sf, allres %>% filter(term == 'wbgt:race_black'))
 m$estimate <- 25*(m$estimate/(qs_white[20] - qs_white[2]))
+m <- m %>%
+  filter(id %in% csadat$id[csadat$black > 0.05])
 ggplot() + 
   geom_sf(data = states) + 
   geom_sf(data = m, aes(fill=estimate)) + 

### 2021-04-28 viz/temporal_dynamics.R

@@ -1,40 +1,80 @@
 library(data.table)
+library(tidyverse)
 
 setwd('~/tweets/')
 
 data <- fread('all.csv')
+data$month <- substr(data$statemonth, 4, 5)
+data$state <- substr(100000 + data$fips, 2, 3)
 
-# Day of Week, Time of Day
-dow_tod <- data[ , .(vader = mean(vader)), .(dow, tod)]
+# Export from https://wonder.cdc.gov/ucd-icd10.html
+# 1. Group Results by Month and Weekday
+# 2. Lower 48 states at all categories of urbanization
+# 3. All ages, genders, origins, races
+# 4. Years 2009-2019
+# 5. All weekdays, all values, all places
+# 6. Injry Intent and Mechanism: Injury Intent: Suicide, Mechanism: All Causes of Death
+sui <- read.delim('suicide/Suicides by Day of Week-Month.txt', nrows=1056) %>%
+  filter(Weekday != 'Unknown') %>%
+  mutate(month = substr(Month, 1, 3),
+         year = as.numeric(substr(Month.Code, 1, 4)),
+         suicides = as.numeric(Deaths)) %>%
+  select(month, year, dow=Weekday, suicides)
 
-dow_tod[dow_tod$vader %in% c(max(dow_tod$vader), min(dow_tod$vader)), ]
+rescale <- function(x){
+  x <- x - min(x)
+  x <- x/max(x)
+  x
+}
 
-        dow  tod      vader
-1: Thursday 04:0 0.07818882
-2:   Sunday 08:0 0.21021752
-
-# Day of week
+# Day of Week
 dow_df <- data[ , .(vader = mean(vader)), .(dow)]
 
-dow_df[dow_df$vader %in% c(max(dow_df$vader), min(dow_df$vader)), ]
+m <- merge(dow_df, sui %>%
+                     group_by(dow) %>%
+                     summarize(suicides=sum(suicides))) %>%
+  mutate(suicides = rescale(suicides),
+         vader = rescale(vader),
+         vader = 1 - vader) %>%
+  gather(key, val, -dow)
+m$dow <- factor(m$dow, levels=c('Sunday', 'Monday', 'Tuesday', 'Wednesday',
+                                'Thursday', 'Friday', 'Saturday'))
 
-         dow     vader
-1:    Sunday 0.1366292
-2:  Saturday 0.1371815
-3:    Friday 0.1338290
[…]

### 2021-04-28 doc/temp-sentiment.tex

@@
 \usepackage{pdflscape}
-\usepackage{subcaption}
 \usepackage{lineno}
@@
 
+% Jeremiah Osborne-Gowey: Randy Olsen (of "Don't be such a scientist" fame) and his ABT - And But Therefor framework for communicating - https://www.huffpost.com/entry/and-but-therefore-randy-o_b_8813330
+% Jeremiah Osborne-Gowey: https://blogs.scientificamerican.com/observations/how-the-word-but-could-save-the-world/
+
 This work builds on existing work by \citep{baylis_weather_2018}.
@@
 \subsection{Temperature and Mental Health}
+%Jeremiah Osborne-Gowey: We report on physiological and mental health in text here. Maybe just a heading of "health"?
+%I think physical health is beyond the scope of this paper -Matt
+
 %https://www.pnas.org/content/115/43/10953
@@
 
+% jieliu.cnah: Could delete
 For children, poverty and poor maternal mental health often co-exist and are two of the risk factors for child development\citep{LUND20111502}, and maternal mental health is significantly associated with child general psychopathology\citep{ RN1289}.
@@
 https://link.springer.com/article/10.1007/s00127-017-1370-4
+ % Jeremiah Osborne-Gowey: Not as relevant.
+Saved the citation but didn't find enough relevant text to include reference to here in our manuscript.
 Maybe this: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0116820
@@
 
+% Jeremiah Osborne-Gowey: Remove parts of this paragraph that relate to call for international research on these linkages as we now are only focusing on USA for Tweets.
 Although climate change is an international phenomenon experienced in every country, its discussions vary from country to country.
 According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions.
+% Jeremiah Osborne-Gowey: I think we can remove this entire paragraph now as it's outside the focus of this paper. Might be relevant for our 2nd paper!
+When removing the paragraph, cut from here and copy to the end of the doc where we have extra text saved. [DO NOT DELETE, just move from here]
+
+% Jeremiah Osborne-Gowey: This is perhaps the only piece of this paragraph still relevant for this paper...
 For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature express more positive emotions on Twitter than those living in colder states in the US.
@@
 Public debates on the current consequences of climate change are discussed in both scientific and non-scientific mediums. 
+% Jeremiah Osborne-Gowey: Need a better transition here between meteorological conditions and effect on mood (expressed as sentiment).
+Feels like we're missing a discussion (new paragraph?) just above here that discusses this.
+This might be where we plug in the text PAW is working on?
 People can express their emotional states and feelings - sentiment - through physical, vocal and written expressions on mediums including newspapers, scientific articles, blogs and other online social media platforms.
@@
 For example, text posted to these platforms can be used to gauge the public mood, assess opinions, measure brand affinity and for emergency planning and disaster response (cites).
+% Jeremiah Osborne-Gowey: Good to cite the top-cited public opinion papers, the Twitter flood study from CU, others from data science field including another Baylis paper
 Sentiment is frequently employed as a correlate of emotional state or mood (cites).
@@
 We examined the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an indicator of heat stress that accounts for temperature, humidity, wind speed, and solar radiation.
-We then explored how the racial and income characteristics of neighborhoods moderate the relationship between WBGT and sentiment.
-Finally, we show how twitter sentiment is an important overall indicator of mental health by assessing the relationship between daily sentiment at the county level and the prevalence of crime and suicide.
-
-and sentiment, first examining the overall relationship, and then exploring heterogeneities based on neighborhood income per capita and race.
+Controlling for a variety of fixed effects, we found that, across all tweets, higher temperatures are associated with lower sentiment (See Fig.
+\ref{fig:wbgt}).
+Sentiment is highest at a WBGT of 5 degrees Celsius (a dry bulb temperature typically around 12C/54F).
+The greatest declines in sentiment are observed between 20-25 degrees WBGT (dry bulb 29-36C/84-97F).
+Sentiment also declines with colder temperatures, but only slightly.
 
@@
   \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment in tweets.  As temperatures increase, sentiment declines.}
-  \label{fig:wbgt-income}
+  \label{fig:wbgt}
 \end{figure}
 
-Sentiment is highest at a WBGT of 5 degrees celsius (a dry bulb temperature typically around 12C/54F).
-The greatest declines in sentiment are observed between 20-25 degrees WBGT (dry bulb 29-36C/84-97F).
-Sentiment also declines with colder temperatures, but only slightly.
+We then explored how the racial and income characteristics of neighborhoods moderate the relationship between WBGT and sentiment (See Fig \ref{fig:sub1}).
+We found that neighborhood income strongly moderates the relationship between temperature and sentiment.
+As temperatures increase to moderate heat of 20 WBGT (29C/84F), sentiment decreases for most neighborhoods, but increases in neighborhoods in the 95th income percentile.
+These rich do not see decreases in sentiment until temperatures are above 20 WBGT (29C/84F), at which point sentiment decreases evenly across income ranges, still leaving a large gap in expressed sentiment from low to high-income neighborhoods.
+
+We examined patterns of temperature and sentiment for neighborhoods that had a majority population of four broad racial categories - non-hispanic black, non-hispanic white, hispanic of any race, and other, which includes native american, multi-racial, and asian-american and pacific islander.
+While people in neighborhoods of all racial compositions were affected by heat, people in majority black neighborhoods were much more affected than others  (See Fig \ref{fig:sub2}).
+Relative to an optimum temperature of 5 WBGT (12C/54F), as temperatures increase to 30 WBGT (42C/108F), the sentiment of tweets in majority black neighborhoods decreases four-times as much as people in other neighborhoods.
+Additionally, at levels of moderate heat from 10 WBGT (19C/66F) to 25 WBGT (36C/97F), people in majority hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
 
@@
   \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income.}
+  \caption{}
   \label{fig:sub1}
@@
   \includegraphics[width=\linewidth]{../res/wbgt-race_q.png}
-  \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood majority race.}
+  \caption{}
   \label{fig:sub2}
 \end{subfigure}
-\caption{How the relationship between WBGT and sentiment is moderated }
+\caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income (a) and race (b).}
 \label{fig:test}
@@
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.75\linewidth]{../res/map_wbgt.png}
-  \caption{Effects of rising temperatures on sentiment by Combined Statistical Areas (CSAs) with over 1 million people.
-The value shown is the predicted change in vader score for a 25-degree increase in temperatures.
-CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.}
-  \label{fig:map-wbgt}
-\end{figure}
+To examine whether local factors may determine the relationship between
 
@@
 \centering
-  \includegraphics[width=0.75\linewidth]{../res/map_wbgt_income.png}
-  \caption{Expected gap in sentiment scores between poor and rich neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
-The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 90th percentile for income.
-CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.}
-  \label{fig:map-wbgt-income}
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=\linewidth]{../res/map_wbgt.png}
+  \caption{}
+  \label{fig:map1}
+\end{subfigure}
+\begin{subfigure}{.6\textwidth}
+  \includegraphics[width=\linewidth]{../res/map_wbgt_income.png}
+  \caption{}
+  \label{fig:map2}
+\end{subfigure}
+\caption{Effects of rising temperatures on sentiment by Combined Statistical Areas (CSAs) with over 1 million people.
+The value shown is the predicted change in the size of the gap in vader score across between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b) as temperatures increase from 5 WBGT to 30 WBGT.
+CSAs with a statistically significant effect ($p < 0.05$) are outlined in black.}
+\label{fig:test}
 \end{figure}
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.75\linewidth]{../res/map_wbgt_race.png}
-  \caption{Expected gap in sentiment scores between largely minority and largely white neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
+Expected gap in sentiment scores between poor and rich neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
+The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 90th percentile for income.
+CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.
+E
+xpected gap in sentiment scores between largely minority and largely white neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
 The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 95th percentile for whiteness.
-CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.}
-  \label{fig:map-wbgt-race}
-\end{figure}
+CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.
 
@@
 
+% This could honestly be its own appendix section
+  The change from optimum temperature of 5WBGT (12C/54F) to a heatwave of 25WBGT (36C/97F) is associated with an overall decrease in the vader sentiment score of 0.01.
+This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372).
+Additionally, over the study period from 2009-2019, suicides followed a very similar weekly pattern, peaking on Mondays and declining over the course of the week until Saturday.
+From Saturday to Monday, suicides increase 23\%.
+Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment at some time scales.
+Moreover, both suicides and sentiment are affected by higher temperatures.
+Thus, while twitter sentiments are the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood and hourly effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations.
+
 \section{Data}
@@
 Appendix table. Sentiment scores across all types we considered.
+
+Weird effect of southwest - WBGT means this is more likely to be cultural than due to different heat effects.
 

### 2021-04-30 crime/model_day.R
play with models, settle on hurricanes

@@ -10,8 +10,9 @@ library(broom)
 dat <- fread('~/tweets/all.csv')
 
 dat$date <- paste0(dat$year, '-', dat$doy)
+dat$state <- substr(dat$fips + 100000, 2, 3)
 
-dat_day <- dat[ , .(afinn=mean(afinn), hedono=mean(hedono), vader=mean(vader), n=length(vader)), .(fips, date)]
+dat_day <- dat[ , .(afinn=mean(afinn), hedono=mean(hedono), vader=mean(vader), n=length(vader)), .(state, date)]
 
 ###############################
 # read crime data
@@ -19,6 +20,33 @@ dat_day <- dat[ , .(afinn=mean(afinn), hedono=mean(hedono), vader=mean(vader), n
 
 crim <- fread('~/tweets/crime/crime_cod_nibrs.csv')
 crim$date <- as.character(crim$date)
+crim$dow <- lubridate::wday(ymd(crim$date), label=TRUE)
+crim$state <- substr(crim$fips + 100000, 2, 3)
+
+crim$homicides <- rowSums(crim[ , c('homicides_nibrs', 'homicides_cod')], na.rm=T)
+crim$violent <- rowSums(crim[ , c('violent_nibrs', 'violent_cod')], na.rm=T)
+crim$assault <- rowSums(crim[ , c('assault_nibrs', 'assault_cod')], na.rm=T)
+
+crim2 <- crim[ , .(homicides_nibrs=sum(homicides_nibrs, na.rm=T),
+                   violent_nibrs=sum(violent_nibrs, na.rm=T),
+                   assault_nibrs=sum(assault_nibrs, na.rm=T),
+                   homicides_cod=sum(homicides_cod, na.rm=T),
+                   violent_cod=sum(violent_cod, na.rm=T),
+                   assault_cod=sum(assault_cod, na.rm=T)),
+             .(dow)]
+
+rescale <- function(x){
+  x <- x - min(x)
+  x <- x/max(x)
+  x
+}
+
+crim3 <- crim2 %>%
+  mutate_if(is.numeric, rescale) %>%
+  gather(key, value, -dow)
+
+ggplot(crim3) + 
+  geom_line(aes(x=dow, y=value, color=key, group=key))
 
 ######################################
 # Combine
@@ -26,12 +54,12 @@ crim$date <- as.character(crim$date)
 
 comb <- merge(crim, dat_day, all.x=T)
 comb$year <- as.numeric(substr(comb$date, 1, 4))
[…]

### 2021-04-30 csa/model_space.R

@@ -121,32 +121,6 @@ ggplot() +
 ggsave('~/temp-sentiment/res/map_wbgt_income.png', width=12, height=8)
 
 #############################
-# Nonwhite results
-###########################
-m <- merge(csa_sf, allres %>% filter(term == 'wbgt:race_white'))
-m$estimate <- 25*(m$estimate/(qs_white[2] - qs_white[20]))
-ggplot() + 
-  geom_sf(data = states) + 
-  geom_sf(data = m, aes(fill=est_change)) + 
-  geom_sf(data = states, fill=NA) + 
-  geom_sf(data = m, fill=NA, aes(color=p.value < 0.05, size=p.value < 0.05)) + 
-  scale_fill_gradient2(low='#d01c8b', mid='#f7f7f7', high='#4dac26', midpoint=0) + 
-  scale_color_manual(values=c('TRUE'='#000000', 'FALSE'=NA)) +
-  scale_size_manual(values=c(1, 1.1)) +
-  guides(color=FALSE, size=FALSE,
-         fill = guide_colourbar(title.position="top", title.hjust = 0.5)) + 
-  theme_void() + 
-  theme(legend.position=c(0.2, 0.1),
-        legend.direction='horizontal',
-        legend.key.width=unit(2, 'cm'),
-        legend.key.height=unit(1, 'cm'),
-        legend.text=element_text(size=rel(2)),
-        legend.title=element_text(size=rel(2))) + 
-  coord_sf(crs='epsg:2163') + 
-  labs(fill='Gap in Sentiment')
-ggsave('~/temp-sentiment/res/map_wbgt_race.png', width=12, height=8)
-
-#############################
 # Black results
 ###########################
 m <- merge(csa_sf, allres %>% filter(term == 'wbgt:race_black'))

### 2021-04-30 viz/temporal_dynamics.R

@@ -1,5 +1,7 @@
 library(data.table)
 library(tidyverse)
+library(lubridate)
+library(grid)
 
 setwd('~/tweets/')
 
@@ -28,14 +30,18 @@ rescale <- function(x){
 }
 
 # Day of Week
-dow_df <- data[ , .(vader = mean(vader)), .(dow)]
+dow_df <- data[ , .(vader = mean(vader), afinn=mean(afinn), hedono=mean(hedono)), .(dow)]
 
 m <- merge(dow_df, sui %>%
                      group_by(dow) %>%
                      summarize(suicides=sum(suicides))) %>%
   mutate(suicides = rescale(suicides),
          vader = rescale(vader),
-         vader = 1 - vader) %>%
+         afinn = rescale(afinn),
+         hedono = rescale(hedono),
+         vader = 1 - vader,
+         afinn = 1 - afinn,
+         hedono = 1 - hedono) %>%
   gather(key, val, -dow)
 m$dow <- factor(m$dow, levels=c('Sunday', 'Monday', 'Tuesday', 'Wednesday',
                                 'Thursday', 'Friday', 'Saturday'))
@@ -63,16 +69,132 @@ ggplot(m) +
 # State
 sui <- read.delim('suicide/Suicides by State.txt', nrows=49) %>%
   mutate(suicides = as.numeric(Crude.Rate),
-         state = as.numeric(State.Code)) %>%
-  select(state, suicides)
+         state = as.numeric(State.Code),
+         statename = State) %>%
+  select(state, suicides, statename)
 
 state_df <- data[ , .(vader = mean(vader)), .(state)]
 state_df$state <- as.numeric(state_df$state)
 
 m <- merge(state_df, sui)
+
+ggplot(m) + 
+  geom_text(aes(x=vader, y=suicides, label=statename))
+
 plot(m$vader, m$suicides)
 summary(lm(suicides ~ vader, data=m))
 
+
+
+#####################
+# Look at Hurricanes
+#####################
+
+data$date <- mdy(paste0(data$doy, '-', data$year))
+
+harvey_counties <- c(48007, 48025, 48039, 48057, 48071, 48157, 48167,
+  48175, 48201, 48239, 48245, 48255, 48273, 48291,
+  48321, 48391, 48409, 48469, 48481)
+
+harvey_date <- ymd('2017/08/26')
+
[…]

### 2021-04-30 viz/comparison_plot.R
make comparison plot

@@ -2,6 +2,7 @@ library(data.table)
 library(tidyverse)
 library(lubridate)
 library(grid)
+library(png)
 
 setwd('~/tweets/')
 
@@ -164,9 +165,9 @@ black_change <- 0.019
 
 
 plt <- data.frame(lab=c(# 'Celebrity\nSuicide', 
-                        'Houston\nDuring\nHurricane\nHarvey',
-                        'New York City\nDuring\nHurricane\nSandy',
-                    'Weekly Change\nFrom Saturday \nto Monday', 'Rich\n(95th %ile)',
+                        'Hurricane\nHarvey',
+                        'Hurricane\nSandy',
+                    'Saturday\nto Monday', 'Rich\n(95th %ile)',
                     'Poor\n(5th %ile)', 'Majority\nWhite',
                     'Majority\nBlack'),
                   val=-c(#robin_change,
@@ -174,29 +175,43 @@ plt <- data.frame(lab=c(# 'Celebrity\nSuicide',
                         rich_change, poor_change, white_change, black_change))
 
 write.csv(plt, 'compare_plot_dat.csv', row.names=F)
+plt <- read.csv('~/tweets/compare_plot_dat.csv')
 
-plt$lab <- factor(plt$lab, levels=c('Houston\nDuring\nHurricane\nHarvey', 'New York City\nDuring\nHurricane\nSandy', 'Weekly Change\nFrom Saturday \nto Monday', 'Majority\nWhite', 'Majority\nBlack', 'Rich\n(95th %ile)', 'Poor\n(5th %ile)'))
+plt$lab <- factor(plt$lab, levels=c('Hurricane\nHarvey', 'Hurricane\nSandy', 'Saturday\nto Monday', 'Majority\nWhite', 'Majority\nBlack', 'Rich\n(95th %ile)', 'Poor\n(5th %ile)'))
 
-annot <- data.frame(x='Majority White', y=0.1, label='}')
+img <- readPNG('~/temp-sentiment/viz/braceleft.png')
+g <- rasterGrob(img, height=0.1, width=1)
 
+annot <- data.frame(color=c(rep('Majority\nBlack', 2), 
+                        rep('Hurricane\nHarvey', 1)),
+                    x=c(5.5, 5.5, 2),
+                    y=c(0.006, 0.0025, 0.0025) + 0.00875,
+                    label=c('Heat Wave By', 'Neighborhood Characteristics',
+                            'Comparison Events'))
 ggplot() + 
[…]

### 2021-04-30 doc/temp-sentiment.tex

@@
 
+We ran models using multiple sentiment indicators, and found similar results no matter which sentiment metric was used (See Appendix).
+
 \subsection{Effects by Combined Statistical Area}
 
-To examine whether local factors may determine the relationship between
+To further explore our findings, we examined inequalities in the impacts of higher temperatures on sentiment by Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) with over 1 million people (hereafter: cities).
+We examined the change in size of the gap in sentiment scores between both poor and rich neighborhoods, as well as black and non-black neighborhoods, as temperatures change from optimum temperatures of 5 WBGT (12C/54F) to 30 WBGT (42C/108F).
+
+We found increasing inequalities in sentiment across income groups for 68.8\% of cities, with 35.6\% of these having a statistically significant effect (15 cities).
+Cities with a significantly unequal effect were found in the south, midwest, northeast, northwest, and southwest, although they were most common in the mid-atlantic region.
+Additionally, many cities in the southwest had an inverted effect, where higher temperatures actually narrowed the gap in sentiment between rich and poor neighborhoods, including Denver and Oklahoma City, which had a statistically significant inverted effect.
+
+Because we found a much stronger effect for the effects of temperature on sentiment in majority black neighborhoods, we also examined increases in sentiment gaps between black and non-black neighborhoods in cities with a black population greater than 5\% of the total population.
+We found increasing inequalities in sentiment for more black neighrborhoods in 64.7\% of cities, with a significant effect in 21.2\% of those (7 cities).
+Additionally, three sunbelt cities had a significant inverted effect.
+The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: cities with large and significant inequalities were found the mid-atlantic and midwest, as well as central florida.
 
@@
 \centering
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=\linewidth]{../res/map_wbgt.png}
+\begin{subfigure}{0.75\textwidth}
+  \includegraphics[width=\linewidth]{../res/map_wbgt_income.png}
   \caption{}
@@
 \end{subfigure}
-\begin{subfigure}{.6\textwidth}
-  \includegraphics[width=\linewidth]{../res/map_wbgt_income.png}
+\begin{subfigure}{0.75\textwidth}
+  \includegraphics[width=\linewidth]{../res/map_wbgt_black.png}
   \caption{}
@@
 \end{subfigure}
-\caption{Effects of rising temperatures on sentiment by Combined Statistical Areas (CSAs) with over 1 million people.
-The value shown is the predicted change in the size of the gap in vader score across between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b) as temperatures increase from 5 WBGT to 30 WBGT.
-CSAs with a statistically significant effect ($p < 0.05$) are outlined in black.}
+\caption{Inequalities in impacts of temperature on sentiment by CSAs and MSAs with over 1 million people.
+The value shown is the predicted change in the size of the gap in vader score as temperatures increase from 5 WBGT to 30 WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
+CSAs with a statistically significant effect ($p < 0.05$) are outlined in black.
+For racial inequalities (b), only MSAs and CSAs where black people make up >5\% of the population are included.}
 \label{fig:test}
 \end{figure}
-
-Expected gap in sentiment scores between poor and rich neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
-The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 90th percentile for income.
-CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.
-E
-xpected gap in sentiment scores between largely minority and largely white neighborhoods, by Combined Statistical Areas (CSAs) with over 1 million people.
-The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 95th percentile for whiteness.
-CSAs with a statistically significant effect ($p < 0.1$) are outlined in black.
 
@@
 \end{figure}
+
+\subsection{Relationship between fluctuations in sentiment and other indicators of heat stress and mental health}
+
+% Monday - Saturday
+% Depressed States - Happy States
+% Natural Disasters
+% Mass shootings
+% Celebrity suicides
+
+% four years after floods by 13.8 percent, from 12.1 to 13.8 per 100,000 (P<0.001);  
+% two years after hurricanes by 31.0 percent, from 12.0 to 15.7 per 100,000 (P<0.001)
+% first year after earthquakes by 62.9 percent, from 19.2 to 31.3 per 100,000 (P<0.001).
 

### 2021-04-30 doc/temp-sentiment.tex
Updates from Overleaf

@@
 \usepackage{lineno}
+\usepackage{xcolor}
 \usepackage[
@@
 Sentiment also declines with colder temperatures, but only slightly.
+We ran models using multiple sentiment indicators, and found similar results no matter which sentiment metric was used (See Appendix).
 
@@
 
-We ran models using multiple sentiment indicators, and found similar results no matter which sentiment metric was used (See Appendix).
+\subsection{Comparison With Other Events}
+
+Expressed sentiment is a widely used metric to assess mental health and well-being on social media.
+While, a variety of related metrics can be used to quantify sentiment, we chose to use the VADER metric that was specifically designed for microblogs like twitter \cite{hutto2014vader}.
+To give context to this metric, we compare the impacts of heat waves, defined as a change from 5 WBGT (12C/54F) to 25 WBGT (36C/97F), to the change in average sentiment from Saturday to Monday as well as the impacts on of the two most expensive hurricanes in the last decade in the United States (See Fig.
+\ref{fig:compare}.
+Saturday and Monday are the highest and lowest days for both twitter sentiment and suicides, and over the course of our study period, Mondays had 23\% more suicides than Saturdays.
+For the hurricanes, we compared the mean sentiment of counties affected by Hurricanes Harvey and Sandy on the day of landfall to the mean sentiment a week before.
+These hurricanes are associated with mental health effects including Post-Traumatic Stress Disorder (PTSD) \cite{Schwartz2017Aug, Schwartz2018May}, and similar disasters have been associated with subsequent increases in suicides \cite{Krug1998Feb}.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{../res/comparison_plot.png}
+  \caption{Decline in sentiment during a heat wave across several different neighborhoods, compared with other impacts on sentiment, such as major hurricanes, as well as the weekly variation in sentiment from the peak on Saturday to the low on Monday.}
+  \label{fig:compare}
+\end{figure}
+
+We found that the impacts of heatwaves on sentiment in rich and majority white neighborhoods were less than the average weekly change in sentiment from Saturday to Monday, but the effects of heat waves on majority black or poor neighborhoods were much greater than the average weekly change in sentiment.
+Additionally, the impacts of heat waves more vulnerable neighborhoods were near to the impacts of Hurricane Sandy.
+
+{\color{red} What other disasters/comparison events should we look at?}
 
@@
 
+In addition to providing data with high spatial resolution, twitter data also comes with very high temporal resolution as each tweet is time-stamped.
+Thus, we were able to examine the impact of heat on sentiment for each hour of the day (See Fig.
+\ref{fig:ts-wbgt}).
+We found that heat decreased expressed sentiment at nearly all hours of the day, and that heat had the largest effect on sentiment early in the morning, between 4am and 6am.
+Heat had the weakest effects on sentiment at midday, and at 3am.
+
 \begin{figure}[H]
@@
 The value shown is the predicted change in vader score for a 25-degree increase in temperatures.}
-  \label{fig:map-wbgt}
+  \label{fig:ts-wbgt}
 \end{figure}
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.75\linewidth]{../res/ts_income.png}
-  \caption{Expected gap in sentiment scores between largely minority and largely white neighborhoods, by hour of the day, with a 95\% confidence interval.
-The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 95th percentile for income.}
-  \label{fig:map-wbgt-income}
-\end{figure}
+\section{Discussion}
+We found that twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution.
+We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between heat and worsened sentiment.
+Moreover, we were able to show clear heterogeneities in vulnerability by neighborhood characteristics, overturning previous findings that found no heterogeneities at the county level \cite{Burke2018Aug, Mullins2019Dec}.
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.75\linewidth]{../res/ts_race.png}
-  \caption{Expected gap in sentiment scores between largely minority and largely white neighborhoods, by hour of the day, with a 95\% confidence interval.
-The value shown is the predicted gap in vader scores for a 25-degree increase in temperatures between neighborhoods at the 5th percentile and the 95th percentile for whiteness.}
-  \label{fig:map-wbgt-race}
-\end{figure}
-
-\subsection{Relationship between fluctuations in sentiment and other indicators of heat stress and mental health}
-
-% Monday - Saturday
-% Depressed States - Happy States
-% Natural Disasters
-% Mass shootings
-% Celebrity suicides
-
-% four years after floods by 13.8 percent, from 12.1 to 13.8 per 100,000 (P<0.001);  
-% two years after hurricanes by 31.0 percent, from 12.0 to 15.7 per 100,000 (P<0.001)
-% first year after earthquakes by 62.9 percent, from 19.2 to 31.3 per 100,000 (P<0.001).
-
-Figure of heatmap of tweet density across USA (pick a year? entire dataset of georeferenced tweets?)
-
-(sub-section) TEMPERATURE and HUMIDITY
-Figure of distribution of max and min temperatures
-Comment from Patrick Baylis on our "preliminary results". 2nd comment from Patrick
-
-(sub-section) INCOME
-
-(sub-section?) SENTIMENT
-Figure of distribution of observed sentiment scores (by Hedonometer, Vader, etc.)
-
-Comment from Patrick Baylis on our "preliminary results" doc
-
-\section{Discussion}
+However, one limitation of data from twitter is that a tweet is not a precise measurement of a mental health outcome, but rather only a rough indicator a user's mental state based on the vocabulary in the tweet.
+Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, accounting for negations (``\textit{not} good"), punctuation (``Good!!!"), emoticons and emojis (":-) or d grin
 
@@
 
+Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat, 
 \section{Data}

### 2021-05-01 viz/comparison_plot.R
clean up comparison plots, remove weather terms

@@ -7,85 +7,10 @@ library(png)
 setwd('~/tweets/')
 
 data <- fread('all.csv')
-data$month <- substr(data$statemonth, 4, 5)
-data$state <- substr(100000 + data$fips, 2, 3)
-
-# Export from https://wonder.cdc.gov/ucd-icd10.html
-# 1. Group Results by Month and Weekday
-# 2. Lower 48 states at all categories of urbanization
-# 3. All ages, genders, origins, races
-# 4. Years 2009-2019
-# 5. All weekdays, all values, all places
-# 6. Injry Intent and Mechanism: Injury Intent: Suicide, Mechanism: All Causes of Death
-sui <- read.delim('suicide/Suicides by Day of Week-Month.txt', nrows=1056) %>%
-  filter(Weekday != 'Unknown') %>%
-  mutate(month = substr(Month, 1, 3),
-         year = as.numeric(substr(Month.Code, 1, 4)),
-         suicides = as.numeric(Deaths)) %>%
-  select(month, year, dow=Weekday, suicides)
-
-rescale <- function(x){
-  x <- x - min(x)
-  x <- x/max(x)
-  x
-}
+data <- data[weather_term == 0, ]
 
 # Day of Week
 dow_df <- data[ , .(vader = mean(vader), afinn=mean(afinn), hedono=mean(hedono)), .(dow)]
-
-m <- merge(dow_df, sui %>%
-                     group_by(dow) %>%
-                     summarize(suicides=sum(suicides))) %>%
-  mutate(suicides = rescale(suicides),
-         vader = rescale(vader),
-         afinn = rescale(afinn),
-         hedono = rescale(hedono),
-         vader = 1 - vader,
-         afinn = 1 - afinn,
-         hedono = 1 - hedono) %>%
-  gather(key, val, -dow)
-m$dow <- factor(m$dow, levels=c('Sunday', 'Monday', 'Tuesday', 'Wednesday',
-                                'Thursday', 'Friday', 'Saturday'))
-
-ggplot(m) + 
-  geom_line(aes(x=dow, y=val, color=key, group=key))
-
-# Month of Year
-moy_df <- data[ , .(vader = mean(vader)), .(month)]
-moy_df$month <- month.abb[as.numeric(moy_df$month)]
-
-m <- merge(moy_df, sui %>%
-                     group_by(month) %>%
[…]

### 2021-05-01 viz/comparison_plot.R
local comparison plot

@@ -35,7 +35,9 @@ dow_df <- data[ , .(vader = mean(vader), afinn=mean(afinn), hedono=mean(hedono))
 
 m <- merge(dow_df, sui %>%
                      group_by(dow) %>%
-                     summarize(suicides=sum(suicides))) %>%
+                     summarize(suicides=sum(suicides))ptop ➜  temp-sentiment                                                  INSERT
+
+                   ) %>%
   mutate(suicides = rescale(suicides),
          vader = rescale(vader),
          afinn = rescale(afinn),

### 2021-05-01 doc/temp-sentiment.tex

@@
 ]{biblatex}
+\usepackage{scalerel,xparse}
 
 \addbibresource{TextDataClimateShocks.bib}
+
+% To be able to use emojis
+\NewDocumentCommand\emojismile{}{
+    \scalerel*{
+        \includegraphics{smiling-face-with-smiling-eyes_1f60a.png}
+    }{X}
+}
 
@@
 \ref{fig:wbgt}).
-Sentiment is highest at a WBGT of 5 degrees Celsius (a dry bulb temperature typically around 12C/54F).
-The greatest declines in sentiment are observed between 20-25 degrees WBGT (dry bulb 29-36C/84-97F).
+Sentiment is highest at a WBGT of 5\textdegree Celsius (a dry bulb temperature typically around 12\textdegree C/54\textdegree F).
+The greatest declines in sentiment are observed between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
 Sentiment also declines with colder temperatures, but only slightly.
@@
 We found that neighborhood income strongly moderates the relationship between temperature and sentiment.
-As temperatures increase to moderate heat of 20 WBGT (29C/84F), sentiment decreases for most neighborhoods, but increases in neighborhoods in the 95th income percentile.
-These rich do not see decreases in sentiment until temperatures are above 20 WBGT (29C/84F), at which point sentiment decreases evenly across income ranges, still leaving a large gap in expressed sentiment from low to high-income neighborhoods.
+As temperatures increase to moderate heat of 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment decreases for most neighborhoods, but increases in neighborhoods in the 95th income percentile.
+These rich do not see decreases in sentiment until temperatures are above 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point sentiment decreases evenly across income ranges, still leaving a large gap in expressed sentiment from low to high-income neighborhoods.
 
@@
 While people in neighborhoods of all racial compositions were affected by heat, people in majority black neighborhoods were much more affected than others  (See Fig \ref{fig:sub2}).
-Relative to an optimum temperature of 5 WBGT (12C/54F), as temperatures increase to 30 WBGT (42C/108F), the sentiment of tweets in majority black neighborhoods decreases four-times as much as people in other neighborhoods.
-Additionally, at levels of moderate heat from 10 WBGT (19C/66F) to 25 WBGT (36C/97F), people in majority hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decreases four-times as much as people in other neighborhoods.
+Additionally, at levels of moderate heat from 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
 
@@
 While, a variety of related metrics can be used to quantify sentiment, we chose to use the VADER metric that was specifically designed for microblogs like twitter \cite{hutto2014vader}.
-To give context to this metric, we compare the impacts of heat waves, defined as a change from 5 WBGT (12C/54F) to 25 WBGT (36C/97F), to the change in average sentiment from Saturday to Monday as well as the impacts on of the two most expensive hurricanes in the last decade in the United States (See Fig.
+To give context to this metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to the change in average sentiment from Saturday to Monday as well as the impacts on of the two most expensive hurricanes in the last decade in the United States (See Fig.
 \ref{fig:compare}.
@@
 To further explore our findings, we examined inequalities in the impacts of higher temperatures on sentiment by Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) with over 1 million people (hereafter: cities).
-We examined the change in size of the gap in sentiment scores between both poor and rich neighborhoods, as well as black and non-black neighborhoods, as temperatures change from optimum temperatures of 5 WBGT (12C/54F) to 30 WBGT (42C/108F).
+We examined the change in size of the gap in sentiment scores between both poor and rich neighborhoods, as well as black and non-black neighborhoods, as temperatures change from optimum temperatures of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 30\textdegree C WBGT (42\textdegree C/108\textdegree F).
 
@@
 \caption{Inequalities in impacts of temperature on sentiment by CSAs and MSAs with over 1 million people.
-The value shown is the predicted change in the size of the gap in vader score as temperatures increase from 5 WBGT to 30 WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
+The value shown is the predicted change in the size of the gap in vader score as temperatures increase from 5WBGT to 30WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
 CSAs with a statistically significant effect ($p < 0.05$) are outlined in black.
@@
   \centering
-  \includegraphics[width=0.75\linewidth]{../res/ts.png}
+  \includegraphics[width=0.5\linewidth]{../res/ts.png}
   \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
@@
 
-However, one limitation of data from twitter is that a tweet is not a precise measurement of a mental health outcome, but rather only a rough indicator a user's mental state based on the vocabulary in the tweet.
-Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, accounting for negations (``\textit{not} good"), punctuation (``Good!!!"), emoticons and emojis (":-) or d grin
+However, one limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
+Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, accounting for negations (``\textit{not} good"), punctuation (``Good!!!"), emoticons and emojis (``:-)" or ``\emojismile"), and degree modifiers (``very excellent" or ``kind of crappy") \cite{hutto2014vader}, the observed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
+We overcome this limitation by using an extremely large dataset of nearly 250 million tweets, allowing us to measure population-level changes in sentiment.
 
-% This could honestly be its own appendix section
-  The change from optimum temperature of 5WBGT (12C/54F) to a heatwave of 25WBGT (36C/97F) is associated with an overall decrease in the vader sentiment score of 0.01.
-This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372).
-Additionally, over the study period from 2009-2019, suicides followed a very similar weekly pattern, peaking on Mondays and declining over the course of the week until Saturday.
-From Saturday to Monday, suicides increase 23\%.
-Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment at some time scales.
-Moreover, both suicides and sentiment are affected by higher temperatures.
-Thus, while twitter sentiments are the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood and hourly effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations.
+We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the vader sentiment score of 0.01.
+This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a change in sentiment associated with a suicides increase of 23\% over the course of the study period \cite{CDC2021}.
+Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.
+Moreover, both suicides and sentiment are affected by higher temperatures \cite{baylis_weather_2018, Burke2018Aug}.
+Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood and hourly effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.
 
-Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat, 
+We examined our results relative to an optimal baseline temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as this was the temperature at which the highest sentiment scores were observed in the aggregated analysis (See Fig.
+\ref{fig:wbgt}).
+However, after examining heterogeneities by neighborhood race and income, we find different optima for different racial groups and income levels.
+For example, the optimum temperature for poor neighborhoods is 0\textdegree C WBGT (6\textdegree C/43\textdegree F) while the optimum temperature for rich neighborhoods is 20\textdegree C WBGT (29\textdegree C/84\textdegree F).
+These different varying optima show that even mild heat can affect human well-being, and the only reason that sentiment peaks as high as 20\textdegree C WBGT (29\textdegree C/84\textdegree F) in rich neighborhoods is because the rich are more likely to have air conditioning in their homes, transportation, and work spaces, as well as much greater choice in when they are outside and in the activities they perform outside.
+
+We found that majority black neighborhoods are much more affected by heat than neighborhoods with a majority of any other race.
+This is surprising, given that Hispanics can be marginalized in both income and housing.
+This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans.
+Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine specifically with respect to heat and mental health.
+Finally, in modeling the combined effects of race and income in moderating the impacts of heat on mental health, we find separate effect for these two variables (See Appendix).
+Neither race or income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by than than neighborhoods that are just poor or just black.
+
+In our spatial analysis, we found that most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, with noticeable regional patterns in the strength of the effect.
+Generally, the mid-west and mid-atlantic had the strongest inequalities, and the south-west had the weakest inequalities and an occasional inverted effect.
+Because we used Wet Bulb Globe Temperature, our metric accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
+Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different environmental conditions.
+
+Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis was able to examine this hypothesis.
+While our estimates were more uncertain at night due to a lower volume of tweets, we found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
+This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
+
+While these findings are robust to different model specifications and sentiment metrics, there are some caveats.
+For one, we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
+Many low-income and black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health.
+Moreover, census block income levels can give inaccurate estimates of the well-being of the people who are typically found in census blocks.
+For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may inhabit those spaces throughout the day.
+Additionally, neighborhoods of young college students are also typically estimated to have low income levels, even through college students are wealthier than the average American.
+Again, these issues mean we are likely under-estimating the true effect of heat on mental health.
+
+While climate change will have widespread and severe impacts on human well-being, these impacts are will be highly unevenly distributed.
+People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters \cite{bullard2012wrong}.
+While the physical health impacts of these shocks are more visible and easy to measure, there is no reason to believe that the mental health impacts of climate change will not also be highly unevenly distributed.
+Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales has not found such an effect.
+By using fine-scale twitter data, we showed that there are indeed stark difference in the mental health impacts of heat among neighborhoods in America.
+
 \section{Data}
@@
 To run XYZ analysis, we used Microsoft's ZYX package.
-
-%%KEY POINTS TO DISCUSS
-
-% Link between twitter mood and other very bad outcomes (crime, suicide)
-
-% Baselines: this could also be interpreted as: poor are less affected by the cold\textbf{:}
-
-% Caveats:
-    % Uncertainties from census blocks (low income, young areas).  This is why we used a linaer effect instead of a binned effect.
-    % We cant speak to types of mental heath that are worsening - depression, anxiety, and post-traumatic stress, etc.  Just a general overview
-
-% More nuanced interpretation: optimal temperature for poor & minority neighborhoods are much lower.
-
 \section{Conclusion}

### 2021-05-01 doc/temp-sentiment.tex
Updates from Overleaf

@@
 
+% https://medinform.jmir.org/2020/1/e16023/?utm_source=TrendMD&utm_medium=cpc&utm_campaign=JMIR_TrendMD_0
 \section{Background}
@@
 Higher temperatures are associated with many indicators of worsened mental health.
-Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2006Aug, Likhvar2011Jan}.
+Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
 In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mullins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
@@
 
+Finally, we modeled the combined effects of race and income in moderating the impacts of heat on mental health, and we found separate effects for these two variables (See Appendix).
+In other words, neither race or income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by than than neighborhoods that are just poor or just black.
+
 \begin{figure}[H]
@@
 \section{Discussion}
+% Benefits of twitter data/overview of findings
 We found that twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution.
@@
 
+% Sentiment is a fuzzy indicator but we have BIG DATA
 However, one limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
-Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, accounting for negations (``\textit{not} good"), punctuation (``Good!!!"), emoticons and emojis (``:-)" or ``\emojismile"), and degree modifiers (``very excellent" or ``kind of crappy") \cite{hutto2014vader}, the observed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large dataset of nearly 250 million tweets, allowing us to measure population-level changes in sentiment.
+Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a users expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
+We overcome this limitation by using an extremely large dataset of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
 
+% Compare weekly changes in sentiment to weekly changes in suicides
 We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the vader sentiment score of 0.01.
@@
 
+% Comment on interpretation of results
 We examined our results relative to an optimal baseline temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as this was the temperature at which the highest sentiment scores were observed in the aggregated analysis (See Fig.
@@
 
+% Comment on Race, why no effect on hispanics?
 We found that majority black neighborhoods are much more affected by heat than neighborhoods with a majority of any other race.
@@
 This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans.
-Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine specifically with respect to heat and mental health.
-Finally, in modeling the combined effects of race and income in moderating the impacts of heat on mental health, we find separate effect for these two variables (See Appendix).
-Neither race or income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by than than neighborhoods that are just poor or just black.
+Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
 
+% Interpret Spatial Analysis
 In our spatial analysis, we found that most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, with noticeable regional patterns in the strength of the effect.
@@
 
+% Interpret Time Series Analysis
 Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis was able to examine this hypothesis.
@@
 
+% Caveats
 While these findings are robust to different model specifications and sentiment metrics, there are some caveats.
 For one, we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
-Many low-income and black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health.
-Moreover, census block income levels can give inaccurate estimates of the well-being of the people who are typically found in census blocks.
-For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may inhabit those spaces throughout the day.
+Many low-income and black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
+Moreover, the income level of a census block may be only weakly correlated with the wealth of the people who are typically found that census block.
+For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
 Additionally, neighborhoods of young college students are also typically estimated to have low income levels, even through college students are wealthier than the average American.
 Again, these issues mean we are likely under-estimating the true effect of heat on mental health.
+A final issue with using twitter data is that, while twitter is used by more than one in five Americans, twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
+Nevertheless, we found a large volume of tweets across all neighborhood types.
 
+% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
 While climate change will have widespread and severe impacts on human well-being, these impacts are will be highly unevenly distributed.
@@
 
-\section{Data}
-A proposed list of tables and figures:
-Figures for location of USA Tweets (to show aggregation and coverage).
-Figures of income blocks and temperature extremes.
-Table 1. Variable descriptions for what's included in our models.
-Table 2. Descriptive statistics for datasets/variables we include.
-Table 3. Model results with variables included in the models coefficients, SEs, p-values, sample size.
-Appendix table. Sentiment scores across all types we considered.
+\section{Methods}
 
-Weird effect of southwest - WBGT means this is more likely to be cultural than due to different heat effects.
+% A proposed list of tables and figures:
+% Figures of income blocks and temperature extremes.
+% Table 1. Variable descriptions for what's included in our models.
+% Table 2. Descriptive statistics for datasets/variables we include.
+% Table 3. Model results with variables included in the models coefficients, SEs, p-values, sample size.
 
-\section{Methods}
-\subsection{Tweets}
-We used data from 243 million geolocated tweets from the lower 48 states from the years 2009 to 2019.
+\subsection{Tweets \& Sentiment}
+We used data from 243 million geo-located English-language tweets from the lower 48 states from the years 2009 to 2019.
+The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
+We only considered users’ original content, thus did not include retweets in our analysis.
+Additionally, we excluded all tweets that contained weather-related terms, to ensure that the sentiment expressed in the tweets was reflective of a user's mental state, and not a commentary on the weather.
 
-Text data used for this project come from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
-The "decahose" provides public access to a random 10\% of all public messages on Twitter.
-The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
-The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
-We used all geolocated Tweets from the Decahose for the period 2009-2019.
-We only considered users’ original content, thus did not include retweets in our analysis.
+We assessed the sentiment expressed in the tweets using the VADER (Valence Aware Dictionary for sEntiment Reasoning) sentiment corpus \cite{gilbert_vader_2014}.
+VADER is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media because of several features, as:
 
-We excluded all tweets that contained weather-related terms.
+\begin{itemize}
+  \item Incorporating lexical features common to informal media, such as slang (``sux"), and acronyms (``lol").
+  \item Negations (``\textit{not} good", ``\textit{wasn't} bad")
+  \item Punctuation (``Good!!!")
+  \item Word shape, such as capitalization (``The move was AMAZING")
+  \item Emoticons and emojis (``:-)", ``\emojismile")
+  \item Degree modifiers (``very excellent" or ``kind of crappy")
+\end{itemize}
+
+The VADER method yields a value for the mood of a tweet, with a score of 0 for neurtral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
+In addition to the vader method used in the body of this paper, we also conduct our analysis using the Hedonometer and AFINN sentiment analysis methods, with similar results (See Appendix).
+
+Some papers we might cite in this section (only for multilingual work): Vilares et al. 2017 , Lo et al. 2017, Dashtipour et al. 2016.
+
+Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! \#fail]), etc.
+
+Issues with what people say on social media being attenuated or mismatched from actual emotional state
 
@@
 
-\subsection{Sentiment}
-
-WE EXCLUDED WEATHER TERMS
-
-We assessed the sentiment expressed in text using several of the most commonly used measures of sentiment.
-Each of the measures uses validated dictionaries of words that have been assigned positive and negative scores and compares these words with words in the text corpora of interest.
-For measures of sentiment, we used the Hedonometer \href{https://hedonometer.org/timeseries/en_all/}{(link)}, Afinn \href{http://corpustext.com/reference/sentiment_afinn.html}{(link)}, Textblob \href{https://textblob.readthedocs.io/en/dev/}{(link)}, Vader , SentiwordNet and LIWC tools.
-These tools assign positive and negative scores to words, then average the scores for each text string.
-Our text data was multilingual, which can present challenges with analyzing text (cite).
-Given our interest only in the collection of words used (in aggregate) and not the sequence or nuance of the phrases, we translated all languages into English using Google Translate API and the Microsoft Azure API cognitive services (translator text API) before conducting the sentiment analyses.
-
-Some papers we might cite in this section (only for multilingual work): Vilares et al. 2017 , Lo et al. 2017, Dashtipour et al. 2016.
-
-Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! \#fail]), etc.
-
-Issues with sample population (non-representative)
-
-Issues with what people say on social media being attenuated or mismatched from actual emotional state
-
-\subsubsection{Hedonometer}
-The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
-The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2)using a naive algorithm for scaling up from individual words to texts.
-For the English corpus, Dodds and others \cite{dodds_temporal_2011} collect 10,222 unique words and used crowd-sourcing platform Amazon Mechanical Turk to get human evaluation of happiness degree for each word in an integer scale from 1 to 9, representing a sad to happy spectrum.
-Score 5 represents neutral words.
-Each word will be calculated average score and then the word and happiness score are compiled into a dataset (labMT 1.0).
-Some illustrative example of words are:
-
-\[h_{avg} (\text{laughter}) = 8.50 \]
-\[h_{avg} (\text{the}) = 4.98\]
-\[h_{avg} (\text{hate}) = 2.34\]
-
-The sentiment score for single text will be the mean happiness score for all words in the text.
-
-\subsubsection{VADER}
-The VADER sentiment corpus was proposed by Gilbert and Hutto \cite{gilbert_vader_2014} and stands for 'Valence Aware Dictionary for sentiment Reasoning'.
-It is is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media by incorporating lexical features common to sentiment expression in microblogs
-
 \subsection{Modeling}
@@
 
-\subsection{Data Processing}
-We analyzed meteorological data using the R statistics open-source program version 3.6.2  (R Core Team 2019) using R Studio version 1.2.1335 (RStudio Team, 2019) and the raster \href{https://www.rdocumentation.org/packages/raster/versions/3.3-13}{(link)}, sp \href{https://cran.r-project.org/web/packages/sp/index.html}{(link)}, sf \href{https://cran.r-project.org/web/packages/sf/index.html}{(link)}, and stars \href{https://cran.r-project.org/web/packages/stars/index.html}{(link)} processing and geospatial packages installed.
-We analyized sentiment data using python version x.x.x with pandas \href{https://pandas.pydata.org/}{(link)} and numpy \href{https://numpy.org/}{(link)} packages installed.
-We extracted meteorological data using the geolocated coordinates from the Twitter posts (latitude and longitude).
-We calculated mean climatological scores (z-scores)using baseline meteorological data from 1990 to present, calculated at a quarter degree resolution, and based on the long-term averages and standard deviation.
-We calculated z-scores for each day and the overall annual average z-score.
-For text data cleaning we used the python package tweet-preprocessor \href{https://pypi.org/project/tweet-preprocessor/}{(link)} to remove all text and characters not part of the main text phrase (e.g., URLs, mentions, GIFs, etc.).
-We converted all dates and times to UTC format.
-
-[NEED SOME TEXT ABOUT GOOGLE EARTH ENGINE AND MICROSOFT AI FOR EARTH] We conducted all data analyses on Microsoft AI for Earth's online cloud-computing environment using a combination of existing data analysis packages and scripts we developed.
-We used GitHub as a repository for all scripts we used in this project (link to the Git repo).
-To run ABC analysis, we used Microsoft's CBA package.
-To run XYZ analysis, we used Microsoft's ZYX package.
-\section{Conclusion}
-
 \section{Acknowledgements}
@@
 
+This data came from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
+The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
 \printbibliography

### 2021-05-01 doc/temp-sentiment-supplement.tex
Commits from Overleaf
Updates from Overleaf

@@
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{../res/Boston_Map.png}
+  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
+  \label{fig:timeseries}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=\linewidth]{../res/map_wbgt.png}
+  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
+  \label{fig:timeseries}
+\end{figure}
+
+\subsubsection{Hedonometer}
+The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
+The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2)using a naive algorithm for scaling up from individual words to texts.
+For the English corpus, Dodds and others \cite{dodds_temporal_2011} collect 10,222 unique words and used crowd-sourcing platform Amazon Mechanical Turk to get human evaluation of happiness degree for each word in an integer scale from 1 to 9, representing a sad to happy spectrum.
+Score 5 represents neutral words.
+Each word will be calculated average score and then the word and happiness score are compiled into a dataset (labMT 1.0).
+Some illustrative example of words are:
+
+\[h_{avg} (\text{laughter}) = 8.50 \]
+\[h_{avg} (\text{the}) = 4.98\]
+\[h_{avg} (\text{hate}) = 2.34\]
+
+The sentiment score for single text will be the mean happiness score for all words in the text.
+
 \end{document}

### 2021-05-01 doc/temp-sentiment.tex
Finish first draft of results, discussion, and methods.

@@
 Climate change is accelerating the rate and variability around meteorological norms like the timing, magnitude, intensity and duration of precipitation events and air temperature minima and maxima \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
-Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meterological conditions.
+Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meteorological conditions.
 Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to temperature extremes.
@@
 Higher temperature associated with climate change could influence mental health directly by exposing people to trauma.
-It could also indirectly influence mental health by affecting (1) physical health and (2)community wellbeing\citep{RN1314}.
+It could also indirectly influence mental health by affecting (1) physical health and (2)community well-being \citep{RN1314}.
 For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions(Anderson and Anderson 1998,Cheatwood 1995; Cohn et al.
@@
 In this research, we explore two primary questions.
-First, to what extent are amibent environmental conditions correlated with changes in expressed sentiment?
+First, to what extent are ambient environmental conditions correlated with changes in expressed sentiment?
 Second, is this effect moderated by wealth?
-and local land cover conditions?
 
@@
 Sentiment also declines with colder temperatures, but only slightly.
-We ran models using multiple sentiment indicators, and found similar results no matter which sentiment metric was used (See Appendix).
+We ran models using multiple sentiment indicators, and found similar results no matter which sentiment metric was used (See Supplement).
 
@@
 
-We examined patterns of temperature and sentiment for neighborhoods that had a majority population of four broad racial categories - non-hispanic black, non-hispanic white, hispanic of any race, and other, which includes native american, multi-racial, and asian-american and pacific islander.
+We examined patterns of temperature and sentiment for neighborhoods that had a majority population of four broad racial categories - non-Hispanic black, non-Hispanic white, Hispanic of any race, and other, which includes Native American, multi-racial, and Asian-American and Pacific Islander.
 While people in neighborhoods of all racial compositions were affected by heat, people in majority black neighborhoods were much more affected than others  (See Fig \ref{fig:sub2}).
 Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decreases four-times as much as people in other neighborhoods.
-Additionally, at levels of moderate heat from 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
+Additionally, at levels of moderate heat from 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
 
-Finally, we modeled the combined effects of race and income in moderating the impacts of heat on mental health, and we found separate effects for these two variables (See Appendix).
+Finally, we modeled the combined effects of race and income in moderating the impacts of heat on mental health, and we found separate effects for these two variables (See Supplement).
 In other words, neither race or income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by than than neighborhoods that are just poor or just black.
@@
 Expressed sentiment is a widely used metric to assess mental health and well-being on social media.
-While, a variety of related metrics can be used to quantify sentiment, we chose to use the VADER metric that was specifically designed for microblogs like twitter \cite{hutto2014vader}.
-To give context to this metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to the change in average sentiment from Saturday to Monday as well as the impacts on of the two most expensive hurricanes in the last decade in the United States (See Fig.
-\ref{fig:compare}.
-Saturday and Monday are the highest and lowest days for both twitter sentiment and suicides, and over the course of our study period, Mondays had 23\% more suicides than Saturdays.
+While a variety of related metrics can be used to quantify sentiment, we chose to use the VADER metric that was specifically designed for microblogs like twitter \cite{hutto2014vader}.
+To give context to this novel metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to other events that have large impacts on sentiment (See Fig.
+\ref{fig:compare}).
+Firstly, we examine the change in average sentiment from Saturday to Monday, the highest and lowest days for twitter sentiment as well as other mental health metrics, such as suicides.
+We also examine the decrease in sentiment associated with the two most expensive hurricanes in the last decade in the United States.
 For the hurricanes, we compared the mean sentiment of counties affected by Hurricanes Harvey and Sandy on the day of landfall to the mean sentiment a week before.
@@
 We found that the impacts of heatwaves on sentiment in rich and majority white neighborhoods were less than the average weekly change in sentiment from Saturday to Monday, but the effects of heat waves on majority black or poor neighborhoods were much greater than the average weekly change in sentiment.
-Additionally, the impacts of heat waves more vulnerable neighborhoods were near to the impacts of Hurricane Sandy.
+Additionally, the impacts of heat waves more vulnerable neighborhoods were close in degree to the impacts of both hurricanes.
 
@@
 We found increasing inequalities in sentiment across income groups for 68.8\% of cities, with 35.6\% of these having a statistically significant effect (15 cities).
-Cities with a significantly unequal effect were found in the south, midwest, northeast, northwest, and southwest, although they were most common in the mid-atlantic region.
+Cities with a significantly unequal effect were found in the south, Midwest, northeast, northwest, and southwest, although they were most common in the mid-Atlantic region.
 Additionally, many cities in the southwest had an inverted effect, where higher temperatures actually narrowed the gap in sentiment between rich and poor neighborhoods, including Denver and Oklahoma City, which had a statistically significant inverted effect.
@@
 Because we found a much stronger effect for the effects of temperature on sentiment in majority black neighborhoods, we also examined increases in sentiment gaps between black and non-black neighborhoods in cities with a black population greater than 5\% of the total population.
-We found increasing inequalities in sentiment for more black neighrborhoods in 64.7\% of cities, with a significant effect in 21.2\% of those (7 cities).
+We found increasing inequalities in sentiment for more black neighborhoods in 64.7\% of cities, with a significant effect in 21.2\% of those (7 cities).
 Additionally, three sunbelt cities had a significant inverted effect.
-The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: cities with large and significant inequalities were found the mid-atlantic and midwest, as well as central florida.
+The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: cities with large and significant inequalities were found the mid-Atlantic and Midwest, as well as central Florida.
 
@@
 \caption{Inequalities in impacts of temperature on sentiment by CSAs and MSAs with over 1 million people.
-The value shown is the predicted change in the size of the gap in vader score as temperatures increase from 5WBGT to 30WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
+The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5WBGT to 30WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
 CSAs with a statistically significant effect ($p < 0.05$) are outlined in black.
-For racial inequalities (b), only MSAs and CSAs where black people make up >5\% of the population are included.}
+For racial inequalities (b), only MSAs and CSAs where black people make up $> 5\%$ of the population are included.}
 \label{fig:test}
@@
   \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
-The value shown is the predicted change in vader score for a 25-degree increase in temperatures.}
+The value shown is the predicted change in VADER score for a 25-degree increase in temperatures.}
   \label{fig:ts-wbgt}
@@
 Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a users expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large dataset of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
+We overcome this limitation by using an extremely large data set of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
 
 % Compare weekly changes in sentiment to weekly changes in suicides
-We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the vader sentiment score of 0.01.
+We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the VADER sentiment score of 0.01.
 This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a change in sentiment associated with a suicides increase of 23\% over the course of the study period \cite{CDC2021}.
@@
 % NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
-While climate change will have widespread and severe impacts on human well-being, these impacts are will be highly unevenly distributed.
-People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters \cite{bullard2012wrong}.
-While the physical health impacts of these shocks are more visible and easy to measure, there is no reason to believe that the mental health impacts of climate change will not also be highly unevenly distributed.
+While climate change will have widespread and severe impacts on human well-being, it cannot be overemphasized just how highly unevenly these impacts will be distributed.
+People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \cite{bullard2012wrong}.
+While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they are not also highly unevenly distributed.
 Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales has not found such an effect.
-By using fine-scale twitter data, we showed that there are indeed stark difference in the mental health impacts of heat among neighborhoods in America.
+By using fine-scale twitter data, we showed that there are indeed stark difference in the mental health impacts of heat among neighborhoods in the United States.
+These findings have significant implications for urban planning, climate and environmental justice, and mental health.
 
 \section{Methods}
-
-% A proposed list of tables and figures:
-% Figures of income blocks and temperature extremes.
-% Table 1. Variable descriptions for what's included in our models.
-% Table 2. Descriptive statistics for datasets/variables we include.
-% Table 3. Model results with variables included in the models coefficients, SEs, p-values, sample size.
-
 \subsection{Tweets \& Sentiment}
@@
 The VADER method yields a value for the mood of a tweet, with a score of 0 for neurtral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
-In addition to the vader method used in the body of this paper, we also conduct our analysis using the Hedonometer and AFINN sentiment analysis methods, with similar results (See Appendix).
-
-Some papers we might cite in this section (only for multilingual work): Vilares et al. 2017 , Lo et al. 2017, Dashtipour et al. 2016.
-
-Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! \#fail]), etc.
-
-Issues with what people say on social media being attenuated or mismatched from actual emotional state
+In addition to the VADER method used in the body of this paper, we also conduct our analysis using the Hedonometer and AFINN sentiment analysis methods, with similar results (See Supplement).
 
@@
 \subsection{Socio-Econoimc Data}
-We used data from the American Communit   Survey administered by the US Census to income levels and the racial composition of neighborhoods where tweets were located.
+We used data from the American Community Survey (ACS) administered by the US Census to estimate income levels and the racial composition of neighborhoods where tweets were located.
 Data was at the level of the census block group, the smallest unit for which the census releases public data.
@@
 For example, tweets from 2014 were matched to data from the 2012-2016 ACS.
-Because the most recent available data was from 2014-2018, all tweets from year years greater than 2016 were matched to this dataset.
+Because the most recent available ACS was from 2014-2018, all tweets from year years greater than 2016 were matched to this dataset.
 Data was downloaded from the IPUMS NHGIS service provided by the University of Minnesota.
@@
 \subsection{Modeling}
-We assessed how expressed sentiment is affect by three weather variables: wet bulb globe temperature, precipitation, and sunshine.
-Moreover, we modeled how neighborhood income affects the relationship between these weather variables and sentiment.
+We assessed how expressed sentiment is affected by wet bulb globe temperature using segmented regressions, controlling for precipitation, and shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: the day of the week, the time of day, the day of the year, the year, the month, and the county.
 
-Our model takes the following form:
+Our initial model (for Fig. \ref{fig:wbgt} takes the following form:
+
 \begin{equation}
-    y = \beta_0 + f_t(t) + f_{mt}(m t) + \beta_p p + \beta_{mp} m s + \beta_s s + \beta_{ms} m s + \Phi + \epsilon
+    y = \beta_0 + f_t(t) + \beta_p p + \beta_s s + \Phi + \epsilon
 \end{equation}
 
-Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, and $m$ is the log-transformed average income in the census block where the tweet originated.
-We estimate the
+Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represented the segmented regression for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
 
-STANDARD ERRORS
+To examine how income and racial groups moderate the effect of heat on sentiment, we extend our model to the following form:
+
+\begin{equation}
+    y = \beta_0 + f_t(t) + f_{mt}(m t) + \beta_p p + \beta_{mp} m p + \beta_s s + \beta_{ms} m s + \Phi + \epsilon
+\end{equation}
+
+Where $m$ is either the log-transformed average income in the census block where the tweet originated, or a dummy variable for the four racial categories.
+We specify alternative models where income is a dummy variable in three bins, and were race is a continuous variable for the percent of a census block population that is what.
+These alternative specifications yield similar results (See Supplement).
+We also present the effects of rainfall and sunshine on sentiment by income and race in the Supplement.
+
+We estimate the 95\% confidence interval of our models using 80-fold bootstrapping.
+For our analyses by CSA and by time of day, we subset the data in the model and run the model with the same fixed effects, except we exclude time of day as a fixed effect in the temporal analysis.
 
 \section{Acknowledgements}
+%Official SESYNC wording
 This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
@@
 
+\section{Author Contributions}
+Author contributions: M.W.C.
+and P.B.
+designed research and modeling strategy; A.S.
+provided data; M.W.C and Z.L.
+prepared data; M.W.C.
+and J.O.G analyzed data; and M.W.C., J.O.G., J.L., and P.A.W.
+wrote the paper.
+
+This project was co-lead by M.W.C. and J.O.G.
+
 \end{document}

### 2021-05-12 doc/main.tex
add scientific reports files

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\title{Scientific Reports Title to see here}

\author[1,*]{Alice Author}
\author[2]{Bob Author}
\author[1,2,+]{Christine Author}
\author[2,+]{Derek Author}
\affil[1]{Affiliation, department, city, postcode, country}
\affil[2]{Affiliation, department, city, postcode, country}

\affil[*]{corresponding.author@email.example}

\affil[+]{these authors contributed equally to this work}

\keywords{Keyword1, Keyword2, Keyword3}

\begin{abstract}
Example Abstract. Abstract must not include subheadings or citations. Example Abstract. Abstract must not include subheadings or citations. Example Abstract. Abstract must not include subheadings or citations. Example Abstract. Abstract must not include subheadings or citations. Example Abstract. Abstract must not include subheadings or citations. Example Abstract. Abstract must not include subheadings or citations. Example Abstract. Abstract must not include subheadings or citations. Example Abstract. Abstract must not include subheadings or citations.
\end{abstract}
\begin{document}

\flushbottom
\maketitle
% * <john.hammersley@gmail.com> 2015-02-09T12:07:31.197Z:
%
%  Click the title above to edit the author information and abstract
%
\thispagestyle{empty}

\noindent Please note: Abbreviations should be introduced at the first mention in the main text – no abbreviations lists. Suggested structure of main text (not enforced) is provided below.

\section*{Introduction}

The Introduction section, of referenced text\cite{Figueredo:2009dg} expands on the background of the work (some overlap with the Abstract is acceptable). The introduction should not include subheadings.

\section*{Results}

Up to three levels of \textbf{subheading} are permitted. Subheadings should not be numbered.

\subsection*{Subsection}

Example text under a subsection. Bulleted lists may be used where appropriate, e.g.

\begin{itemize}
\item First item
\item Second item
\end{itemize}

\subsubsection*{Third-level section}

Topical subheadings are allowed.

\section*{Discussion}

The Discussion should be succinct and must not contain subheadings.

\section*{Methods}

Topical subheadings are allowed. Authors must ensure that their Methods section includes adequate experimental and characterization data necessary for others in the field to reproduce their work.

\bibliography{sample}

\noindent LaTeX formats citations and references automatically using the bibliography records in your .bib file, which you can edit via the project menu. Use the cite command for an inline citation, e.g.  \cite{Hao:gidmaps:2014}.

For data citations of datasets uploaded to e.g. \emph{figshare}, please use the \verb|howpublished| option in the bib entry to specify the platform and the link, as in the \verb|Hao:gidmaps:2014| example in the sample bibliography file.

\section*{Acknowledgements (not compulsory)}

Acknowledgements should be brief, and should not include thanks to anonymous referees and editors, or effusive comments. Grant or contribution numbers may be acknowledged.

\section*{Author contributions statement}

Must include all authors, identified by initials, for example:
A.A. conceived the experiment(s),  A.A. and B.A. conducted the experiment(s), C.A. and D.A. analysed the results.  All authors reviewed the manuscript.

\section*{Additional information}

To include, in this order: \textbf{Accession codes} (where applicable); \textbf{Competing interests} (mandatory statement).

The corresponding author is responsible for submitting a \href{http://www.nature.com/srep/policies/index.html#competing}{competing interests statement} on behalf of all authors of the paper. This statement must be included in the submitted article file.

\begin{figure}[ht]
\centering
\includegraphics[width=\linewidth]{stream}
\caption{Legend (350 words max). Example legend text.}
\label{fig:stream}
\end{figure}

\begin{table}[ht]
\centering
\begin{tabular}{|l|l|l|}
\hline
Condition & n & p \\
\hline
A & 5 & 0.1 \\
\hline
B & 10 & 0.01 \\
\hline
\end{tabular}
\caption{\label{tab:example}Legend (350 words max). Example legend text.}
\end{table}

Figures and tables can be referenced in LaTeX using the ref command, e.g. Figure \ref{fig:stream} and Table \ref{tab:example}.

\end{document}

### 2021-05-12 doc/old_intro.tex
Updates from Overleaf

--- 2021-05-01 doc/temp-sentiment.tex
@@
-\documentclass{article}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage[a4paper, margin=1.25in]{geometry}
-\usepackage{breakcites}
-\usepackage{subcaption}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
-\usepackage{textgreek}
-\usepackage{authblk}
-\usepackage{rotating}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{pdflscape}
-\usepackage{lineno}
-\usepackage{xcolor}
-\usepackage[
-  style=numeric,
-  citestyle=numeric-comp,
-  backend=biber,
-  doi=true,
-  natbib=true,
-  sorting=none
-]{biblatex}
-\usepackage{scalerel,xparse}
+% This is the intro section before applying major changes and cuts on 2021-05-04
 
-\addbibresource{TextDataClimateShocks.bib}
-
-% To be able to use emojis
-\NewDocumentCommand\emojismile{}{
-    \scalerel*{
-        \includegraphics{smiling-face-with-smiling-eyes_1f60a.png}
-    }{X}
-}
-
-\begin{document}
-%https://www.overleaf.com/project/5f918edd1afc380001feca0d
-\title{Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures}
-
-\author[1, *]{Matthew Cooper}
-\author[2]{Jeremiah Osborne-Gowey}
-\author[3]{Zheng Liu}
-\author[4]{Jie Liu}
-\author[5]{Portia Adade Williams}
-\author[6]{Aaron Schwartz}
-\author[7]{Patrick Baylis}
-
-\affil[1]{T.H. Chan School of Public Health, Harvard University}
-\affil[2]{University of Colorado Boulder}
-\affil[3]{Department of Geographical Sciences, University of Maryland College Park}
-\affil[4]{School of Business, East China University of Science and Technology}
-\affil[5]{University of Cape Town}
-\affil[6]{University of Colorado Boulder}
-\affil[7]{University of British Columbia}
-\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
-
-\maketitle
-
-\begin{abstract}
-%Limit 150 words: https://www.nature.com/nclimate/about/content
-%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
-Higher temperatures associated with climate change are expected to have major impacts on human mental health.
-Indeed, traditional analyses of heat and mental health outcomes using data collected by public agencies have found associations between elevated temperatures and increases in outcomes such as suicides and mental-health related hospitalizations.
-However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income.
-We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
-We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.
-
-\end{abstract}
-
-Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation, Suicide, Mortality, Morbidity, Self-harm,
+% Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation, Suicide, Mortality, Morbidity, Self-harm,
 
@@
 Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
-In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mullins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
+In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mullins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}, as well as reports of general mental health difficulties \cite{Obradovich2018Oct}.
 
@@
 Higher temperature associated with climate change could influence mental health directly by exposing people to trauma.
-It could also indirectly influence mental health by affecting (1) physical health and (2)community wellbeing\citep{RN1314}.
-For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions(Anderson and Anderson 1998,Cheatwood 1995; Cohn et al.
-2004,Bouchama et al.2007.).
-Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration((Bridger 2003; Kerslake 1972; Kjellstrom et al.2009d,Leithead and Lind 1964,Schrier et al.
-1970).
-The loss of work capacity would in turn result in loss of income, which could also cause mental health problems(Kjellstrom 2009b).%% more physiological terms
+It could also indirectly influence mental health by affecting (1) physical health and (2)community well-being \citep{RN1314}.
+For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions\citep{RN1314,RN1316,RN1317,RN1318}.
+Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration\citep{RN1319,RN1320,RN1321,RN1322}.
+The loss of work capacity would in turn result in loss of income, while the low income could also cause mental health problems\citep{Katz1997}.%% more physiological terms
 
-As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems(Miller et al.
-2009; Prince et al.
-2007), heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health((Berry 2007,World Health Organization 2008b).
+As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems\citep{RN1323,RN1324} heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health \citep{RN1325,RN1326}.
 Second，disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities.
-For example, extreme heat reduces the work capacity of laborers in farm fields(Kjellstrom et al.
-2009d), which further destroys agricultural-supported industries in local area(Berry et al.
-2008).
+For example, extreme heat reduces the work capacity of laborers in farm fields \citep{RN1320}, which further destroys agricultural-supported industries in local area\citep{RN1327}.
 The following economic pressures would undermine social capital and then lead to mental health problems.
@@
 [H2].
-
-\section{Results}
-
-\subsection{Overall Effect}
-We examined the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an indicator of heat stress that accounts for temperature, humidity, wind speed, and solar radiation.
-Controlling for a variety of fixed effects, we found that, across all tweets, higher temperatures are associated with lower sentiment (See Fig.
-\ref{fig:wbgt}).
-Sentiment is highest at a WBGT of 5\textdegree Celsius (a dry bulb temperature typically around 12\textdegree C/54\textdegree F).
-The greatest declines in sentiment are observed between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
-Sentiment also declines with colder temperatures, but only slightly.
-We ran models using multiple sentiment indicators, and found similar results no matter which sentiment metric was used (See Appendix).
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
-  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment in tweets.  As temperatures increase, sentiment declines.}
-  \label{fig:wbgt}
-\end{figure}
-
-We then explored how the racial and income characteristics of neighborhoods moderate the relationship between WBGT and sentiment (See Fig \ref{fig:sub1}).
-We found that neighborhood income strongly moderates the relationship between temperature and sentiment.
-As temperatures increase to moderate heat of 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment decreases for most neighborhoods, but increases in neighborhoods in the 95th income percentile.
-These rich do not see decreases in sentiment until temperatures are above 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point sentiment decreases evenly across income ranges, still leaving a large gap in expressed sentiment from low to high-income neighborhoods.
-
-We examined patterns of temperature and sentiment for neighborhoods that had a majority population of four broad racial categories - non-hispanic black, non-hispanic white, hispanic of any race, and other, which includes native american, multi-racial, and asian-american and pacific islander.
-While people in neighborhoods of all racial compositions were affected by heat, people in majority black neighborhoods were much more affected than others  (See Fig \ref{fig:sub2}).
-Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decreases four-times as much as people in other neighborhoods.
-Additionally, at levels of moderate heat from 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
-
-Finally, we modeled the combined effects of race and income in moderating the impacts of heat on mental health, and we found separate effects for these two variables (See Appendix).
-In other words, neither race or income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by than than neighborhoods that are just poor or just black.
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  \caption{}
-  \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-race_q.png}
-  \caption{}
-  \label{fig:sub2}
-\end{subfigure}
-\caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income (a) and race (b).}
-\label{fig:test}
-\end{figure}
-
-\subsection{Comparison With Other Events}
-
-Expressed sentiment is a widely used metric to assess mental health and well-being on social media.
-While, a variety of related metrics can be used to quantify sentiment, we chose to use the VADER metric that was specifically designed for microblogs like twitter \cite{hutto2014vader}.
-To give context to this metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to the change in average sentiment from Saturday to Monday as well as the impacts on of the two most expensive hurricanes in the last decade in the United States (See Fig.
-\ref{fig:compare}.
-Saturday and Monday are the highest and lowest days for both twitter sentiment and suicides, and over the course of our study period, Mondays had 23\% more suicides than Saturdays.
-For the hurricanes, we compared the mean sentiment of counties affected by Hurricanes Harvey and Sandy on the day of landfall to the mean sentiment a week before.
-These hurricanes are associated with mental health effects including Post-Traumatic Stress Disorder (PTSD) \cite{Schwartz2017Aug, Schwartz2018May}, and similar disasters have been associated with subsequent increases in suicides \cite{Krug1998Feb}.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/comparison_plot.png}
-  \caption{Decline in sentiment during a heat wave across several different neighborhoods, compared with other impacts on sentiment, such as major hurricanes, as well as the weekly variation in sentiment from the peak on Saturday to the low on Monday.}
-  \label{fig:compare}
-\end{figure}
-
-We found that the impacts of heatwaves on sentiment in rich and majority white neighborhoods were less than the average weekly change in sentiment from Saturday to Monday, but the effects of heat waves on majority black or poor neighborhoods were much greater than the average weekly change in sentiment.
-Additionally, the impacts of heat waves more vulnerable neighborhoods were near to the impacts of Hurricane Sandy.
-
-{\color{red} What other disasters/comparison events should we look at?}
-
-\subsection{Effects by Combined Statistical Area}
-
-To further explore our findings, we examined inequalities in the impacts of higher temperatures on sentiment by Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) with over 1 million people (hereafter: cities).
-We examined the change in size of the gap in sentiment scores between both poor and rich neighborhoods, as well as black and non-black neighborhoods, as temperatures change from optimum temperatures of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 30\textdegree C WBGT (42\textdegree C/108\textdegree F).
-
-We found increasing inequalities in sentiment across income groups for 68.8\% of cities, with 35.6\% of these having a statistically significant effect (15 cities).
-Cities with a significantly unequal effect were found in the south, midwest, northeast, northwest, and southwest, although they were most common in the mid-atlantic region.
-Additionally, many cities in the southwest had an inverted effect, where higher temperatures actually narrowed the gap in sentiment between rich and poor neighborhoods, including Denver and Oklahoma City, which had a statistically significant inverted effect.
-
-Because we found a much stronger effect for the effects of temperature on sentiment in majority black neighborhoods, we also examined increases in sentiment gaps between black and non-black neighborhoods in cities with a black population greater than 5\% of the total population.
-We found increasing inequalities in sentiment for more black neighrborhoods in 64.7\% of cities, with a significant effect in 21.2\% of those (7 cities).
-Additionally, three sunbelt cities had a significant inverted effect.
-The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: cities with large and significant inequalities were found the mid-atlantic and midwest, as well as central florida.
-
-\begin{figure}[H]
-\centering
-\begin{subfigure}{0.75\textwidth}
-  \includegraphics[width=\linewidth]{../res/map_wbgt_income.png}
-  \caption{}
-  \label{fig:map1}
-\end{subfigure}
-\begin{subfigure}{0.75\textwidth}
-  \includegraphics[width=\linewidth]{../res/map_wbgt_black.png}
-  \caption{}
-  \label{fig:map2}
-\end{subfigure}
-\caption{Inequalities in impacts of temperature on sentiment by CSAs and MSAs with over 1 million people.
-The value shown is the predicted change in the size of the gap in vader score as temperatures increase from 5WBGT to 30WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
-CSAs with a statistically significant effect ($p < 0.05$) are outlined in black.
-For racial inequalities (b), only MSAs and CSAs where black people make up >5\% of the population are included.}
-\label{fig:test}
-\end{figure}
-
-\subsection{Effects by Time of Day}
-
-In addition to providing data with high spatial resolution, twitter data also comes with very high temporal resolution as each tweet is time-stamped.
-Thus, we were able to examine the impact of heat on sentiment for each hour of the day (See Fig.
-\ref{fig:ts-wbgt}).
-We found that heat decreased expressed sentiment at nearly all hours of the day, and that heat had the largest effect on sentiment early in the morning, between 4am and 6am.
-Heat had the weakest effects on sentiment at midday, and at 3am.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.5\linewidth]{../res/ts.png}
-  \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
-The value shown is the predicted change in vader score for a 25-degree increase in temperatures.}
-  \label{fig:ts-wbgt}
-\end{figure}
-
-\section{Discussion}
-% Benefits of twitter data/overview of findings
-We found that twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution.
-We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between heat and worsened sentiment.
-Moreover, we were able to show clear heterogeneities in vulnerability by neighborhood characteristics, overturning previous findings that found no heterogeneities at the county level \cite{Burke2018Aug, Mullins2019Dec}.
-
-% Sentiment is a fuzzy indicator but we have BIG DATA
-However, one limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
-Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a users expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large dataset of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
-
-% Compare weekly changes in sentiment to weekly changes in suicides
-We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the vader sentiment score of 0.01.
-This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a change in sentiment associated with a suicides increase of 23\% over the course of the study period \cite{CDC2021}.
-Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.
-Moreover, both suicides and sentiment are affected by higher temperatures \cite{baylis_weather_2018, Burke2018Aug}.
-Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood and hourly effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.
-
-% Comment on interpretation of results
-We examined our results relative to an optimal baseline temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as this was the temperature at which the highest sentiment scores were observed in the aggregated analysis (See Fig.
-\ref{fig:wbgt}).
-However, after examining heterogeneities by neighborhood race and income, we find different optima for different racial groups and income levels.
-For example, the optimum temperature for poor neighborhoods is 0\textdegree C WBGT (6\textdegree C/43\textdegree F) while the optimum temperature for rich neighborhoods is 20\textdegree C WBGT (29\textdegree C/84\textdegree F).
-These different varying optima show that even mild heat can affect human well-being, and the only reason that sentiment peaks as high as 20\textdegree C WBGT (29\textdegree C/84\textdegree F) in rich neighborhoods is because the rich are more likely to have air conditioning in their homes, transportation, and work spaces, as well as much greater choice in when they are outside and in the activities they perform outside.
-
-% Comment on Race, why no effect on hispanics?
-We found that majority black neighborhoods are much more affected by heat than neighborhoods with a majority of any other race.
-This is surprising, given that Hispanics can be marginalized in both income and housing.
-This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans.
-Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
-
-% Interpret Spatial Analysis
-In our spatial analysis, we found that most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, with noticeable regional patterns in the strength of the effect.
-Generally, the mid-west and mid-atlantic had the strongest inequalities, and the south-west had the weakest inequalities and an occasional inverted effect.
-Because we used Wet Bulb Globe Temperature, our metric accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
-Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different environmental conditions.
-
-% Interpret Time Series Analysis
-Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis was able to examine this hypothesis.
-While our estimates were more uncertain at night due to a lower volume of tweets, we found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
-This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
-
-% Caveats
-While these findings are robust to different model specifications and sentiment metrics, there are some caveats.
-For one, we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
-Many low-income and black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
-Moreover, the income level of a census block may be only weakly correlated with the wealth of the people who are typically found that census block.
-For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
-Additionally, neighborhoods of young college students are also typically estimated to have low income levels, even through college students are wealthier than the average American.
-Again, these issues mean we are likely under-estimating the true effect of heat on mental health.
-A final issue with using twitter data is that, while twitter is used by more than one in five Americans, twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
-Nevertheless, we found a large volume of tweets across all neighborhood types.
-
-% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
-While climate change will have widespread and severe impacts on human well-being, these impacts are will be highly unevenly distributed.
-People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters \cite{bullard2012wrong}.
-While the physical health impacts of these shocks are more visible and easy to measure, there is no reason to believe that the mental health impacts of climate change will not also be highly unevenly distributed.
-Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales has not found such an effect.
-By using fine-scale twitter data, we showed that there are indeed stark difference in the mental health impacts of heat among neighborhoods in America.
-
-\section{Methods}
-
-% A proposed list of tables and figures:
-% Figures of income blocks and temperature extremes.
-% Table 1. Variable descriptions for what's included in our models.
-% Table 2. Descriptive statistics for datasets/variables we include.
-% Table 3. Model results with variables included in the models coefficients, SEs, p-values, sample size.
-
-\subsection{Tweets \& Sentiment}
-We used data from 243 million geo-located English-language tweets from the lower 48 states from the years 2009 to 2019.
-The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
-We only considered users’ original content, thus did not include retweets in our analysis.
-Additionally, we excluded all tweets that contained weather-related terms, to ensure that the sentiment expressed in the tweets was reflective of a user's mental state, and not a commentary on the weather.
-
-We assessed the sentiment expressed in the tweets using the VADER (Valence Aware Dictionary for sEntiment Reasoning) sentiment corpus \cite{gilbert_vader_2014}.
-VADER is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media because of several features, as:
-
-\begin{itemize}
-  \item Incorporating lexical features common to informal media, such as slang (``sux"), and acronyms (``lol").
-  \item Negations (``\textit{not} good", ``\textit{wasn't} bad")
-  \item Punctuation (``Good!!!")
-  \item Word shape, such as capitalization (``The move was AMAZING")
-  \item Emoticons and emojis (``:-)", ``\emojismile")
-  \item Degree modifiers (``very excellent" or ``kind of crappy")
-\end{itemize}
-
-The VADER method yields a value for the mood of a tweet, with a score of 0 for neurtral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
-In addition to the vader method used in the body of this paper, we also conduct our analysis using the Hedonometer and AFINN sentiment analysis methods, with similar results (See Appendix).
-
-Some papers we might cite in this section (only for multilingual work): Vilares et al. 2017 , Lo et al. 2017, Dashtipour et al. 2016.
-
-Discuss issues with text data (gtp) - that it doesn’t capture emoticons, code switching (e.g., [good job! \#fail]), etc.
-
-Issues with what people say on social media being attenuated or mismatched from actual emotional state
-
-\subsection{Weather}
-We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington.
-This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution, and integrates a large quantity of observation-based and modeled data  \cite{xia_continental-scale_2012}.
-For the exact hour and location of each tweet, we extracted temperature, specific humidity, air pressure, total precipitation, shortwave radiation, and wind speed.
-
-Because metrics of apparent temperature that take into account humidity and other factors can better account for the impacts of heat stress on human health and wellbeing, we calculate the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet.
-WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature would show due to evaporative cooling, and can be estimated given normal temperature, relative humidity, solar radiation, and wind speed.
-Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing.
-Metrics like WBGT that account for the effects of humidity and other factors on heat stress have been associated with diminished economic output \cite{rao2020projections}, increased crime \cite{hu2017impact}, increased mortality \cite{chien2016spatiotemporal, armstrong2019role}, and worsened mental health outcomes \cite{vida2012relationship, ding2016importance}.
-
-Using temperature, specific humidity, and pressure, we derived relative humidity using methods described by  \cite{bolton_computation_1980}.
-We then calculated the WBGT using the formula described by \cite{heo2019comparison}.
-
-\subsection{Socio-Econoimc Data}
-We used data from the American Communit   Survey administered by the US Census to income levels and the racial composition of neighborhoods where tweets were located.
-Data was at the level of the census block group, the smallest unit for which the census releases public data.
-
-ACS data is released to cover five-year periods.
-We therefore matched each tweet with census block group data from the year at the middle year of each survey's five-year range.
-For example, tweets from 2014 were matched to data from the 2012-2016 ACS.
-Because the most recent available data was from 2014-2018, all tweets from year years greater than 2016 were matched to this dataset.
-Data was downloaded from the IPUMS NHGIS service provided by the University of Minnesota.
-
-Mean annual income per capita is provided by the ACS, and we standardized this variable so that the values for each year were in 2019 dollars.
-For racial categories, we combined the various categories provided by the ACS into four racial groups: non-Hispanic white, non-Hispanic black, Hispanic of any race, and an "other" category for non-Hispanic people who were neither black nor white, such as Asian, Native American, or mixed-race people.
-
-\subsection{Modeling}
-We assessed how expressed sentiment is affect by three weather variables: wet bulb globe temperature, precipitation, and sunshine.
-Moreover, we modeled how neighborhood income affects the relationship between these weather variables and sentiment.
-
-Our model takes the following form:
-\begin{equation}
-    y = \beta_0 + f_t(t) + f_{mt}(m t) + \beta_p p + \beta_{mp} m s + \beta_s s + \beta_{ms} m s + \Phi + \epsilon
-\end{equation}
-
-Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, and $m$ is the log-transformed average income in the census block where the tweet originated.
-We estimate the
-
-STANDARD ERRORS
-
-\section{Acknowledgements}
-This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
-Additionally, cloud resources for this project were provided for the Microsoft Azure cloud by Microsoft's AI for Earth program, grant number .
-
-This data came from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
-The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
-\printbibliography
-
-\end{document}

### 2021-05-12 doc/temp-sentiment-supplement.tex
Finish first draft of results, discussion, and methods.
Updates from Overleaf

@@
 \usepackage{lineno}
+\usepackage{fancyhdr}
 \usepackage[
@@
 
-\printbibliography
-
-\end{document}
-
 \section*{Supplemental Info}
@@
 
+% Show results with hedonometer and AFINN
+
+% Show results with income binned and race continuous
+
+% Entire analysis by income-race together
+
+\subsection{Effects by Time of Day}
+
+In addition to providing data with high spatial resolution, twitter data also comes with very high temporal resolution as each tweet is time-stamped.
+Thus, we were able to examine the impact of heat on sentiment for each hour of the day (See Fig.
+\ref{fig:ts-wbgt}).
+We found that heat decreased expressed sentiment at nearly all hours of the day, and that heat had the largest effect on sentiment early in the morning, between 4am and 6am.
+Heat had the weakest effects on sentiment at midday, and at 3am.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\linewidth]{../res/ts.png}
+  \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted change in VADER score for a 25-degree increase in temperatures.}
+  \label{fig:ts-wbgt}
+\end{figure}
+
+% Interpret Time Series Analysis
+Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis was able to examine this hypothesis.
+While our estimates were more uncertain at night due to a lower volume of tweets, we found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
+This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
+
 \subsubsection{Hedonometer}
@@
 
+\printbibliography
+
 \end{document}

### 2021-05-12 doc/temp-sentiment.tex
Updates from Overleaf

@@
 Higher temperatures associated with climate change are expected to have major impacts on human mental health.
-Indeed, traditional analyses of heat and mental health outcomes using data collected by public agencies have found associations between elevated temperatures and increases in outcomes such as suicides and mental-health related hospitalizations.
-However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income.
+Indeed, traditional analyses of heat and mental health outcomes using data collected by public agencies have found strong associations between elevated temperatures and outcomes such as suicides and mental-health related hospitalizations.
+However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income or race.
 We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
@@
 
-Keywords: Text Data, Environmental Shocks, Social Media Mining, Sentiment Analysis, Topic Modeling, Social Media, Public Opinion, Perception, Sentiment, Media, News, Twitter, Climate Change, Science Communication, Policy, Planning, Methods, Text Mining, Natural Language Processing, NLP, Latent Dirichlet Allocation, LDA, Political Ecology, Politics, Resilience, Adaptation, Suicide, Mortality, Morbidity, Self-harm,
+\section{Main}
+%Overview
+Climate change is impacting many aspects of human well-being with these impacts projected to worsen in the coming years \cite{pachauri2014climate}.
+A large body of research has assessed the ongoing impacts on climate change on outcomes like agriculture, economic growth, and physical health (NEED CITATIONS).
+Academic and medical researchers have highlighted the relative paucity of empirical studies into climate impacts on mental health at multiple scales and recently called for more \cite{Berry2018Apr, hayes_climate_2018}.
+Researchers have begun to answer these calls, finding associations between deteriorating environmental conditions and a sense of "ecological grief" from their experienced loss \cite{Cunsolo2018Apr}, disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, as well as a variety of mental health impacts associated with warmer temperatures \cite{baylis_weather_2018, Mullins2019Dec} (PALINKAS and WONG 2020).
+MAYBE A SENTENCE HERE HIGHLIGHTING THE GAP IN KNOWLEDGE ABOUT DISPARITIES IN MENTAL HEALTH IMPACTS, VULNERABLE POPULATIONS, issues of scale for signal to noise, etc.
 
-%Cite this new paper: https://doi.org/10.1016/S2542-5196(20)30251-5
-% And: https://eos.org/articles/long-term-drought-harms-mental-health-in-rural-communities and this is the conf. paper citation https://agu.confex.com/agu/fm20/meetingapp.cgi/Paper/771750
-\section{Introduction}
-
-% Organize as follows:
-
-% Heatwaves are known to worsen mental health
-% AND mental health is strongly associated with income
-% AND wealthier better able to cope with heatwaves 
-% BUT we know little about the interaction between heatwaves, mental health and income
-% THEREFORE: we did this really novel study.
-
-% Jeremiah Osborne-Gowey: Randy Olsen (of "Don't be such a scientist" fame) and his ABT - And But Therefor framework for communicating - https://www.huffpost.com/entry/and-but-therefore-randy-o_b_8813330
-% Jeremiah Osborne-Gowey: https://blogs.scientificamerican.com/observations/how-the-word-but-could-save-the-world/
-
-This work builds on existing work by \citep{baylis_weather_2018}.
-
-%Intro Paragraph
-% I think this should be more about climate change and mental health broadly, as well as a bit about how income has been theorized to affect vulnerability, but there is no evidence yet.
-This can be shorter as well, as we have a lot of detail in the Background section
-% Cite this high-level study: https://www.nature.com/articles/s41558-018-0102-4/
-% Also this review article saying more research is needed https://www.sciencedirect.com/science/article/pii/S0033350618302130
-% -Matt
-
-Research has shown strong linkages between ambient environmental conditions and mental health, with extreme temperatures  frequently associated with poorer overall mental health.
-Individual's mental health is strongly associated with income with people in lower income levels frequently experiencing higher stress levels and less able to find help for coping with these stresses.
-Other research on how people cope with environmental temperature extremes indicate that wealthier are better able to cope with heatwaves, whether from traveling to cooler climes, accessing air conditioning, or other temperature ameliorating strategies that may require access to financial capital.
-Yet we know relatively little about the interaction between temperature, mental health and income.
-Furthermore, previous studies that did attempt to examine relationships between environmental extremes and health sometimes aggregated data at scales too coarse to examine whether income may be an ameliorating factor.
-This study examines the interactions between these three using relatively fine-scale, publicly available data.
-Our research expands on existing studies of temperature and sentiment - one measure of mental health - by examining finer-scale data and incorporating income data to examine correlations between temperature extremes and sentiment.
-
-Human moods and mental states are important aspects of overall well-being and can be influenced by ambient, persistent and fluctuating environmental conditions.
-Climate change is accelerating the rate and variability around meteorological norms like the timing, magnitude, intensity and duration of precipitation events and air temperature minima and maxima \href{https://www.ipcc.ch/site/assets/uploads/2018/03/SREX-Chap3_FINAL-1.pdf}{(link)}.
-Previous work \citep{baylis_weather_2018} indicate important linkages between human mood and meteorological conditions.
-Previous analyses of the effects of weather on mood (as expressed in sentiment of text-based message), however, were constructed at the aggregated city/day level and did not account for differences in socio-economic status which may affect access to resources (e.g., air conditioning) that can offset the effects of exposure to temperature extremes.
-Here, we build on these previous studies by examining the hourly effects of weather on human sentiment, an expression of mood, exploring geographic and economic heterogeneity at a finer-scale resolution than previous studies.
-
-% https://medinform.jmir.org/2020/1/e16023/?utm_source=TrendMD&utm_medium=cpc&utm_campaign=JMIR_TrendMD_0
-\section{Background}
-\subsection{Temperature and Mental Health}
-%Jeremiah Osborne-Gowey: We report on physiological and mental health in text here. Maybe just a heading of "health"?
-%I think physical health is beyond the scope of this paper -Matt
-
+%Heat and mental health
 %https://www.pnas.org/content/115/43/10953
 Higher temperatures are associated with many indicators of worsened mental health.
-Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and many others have demonstrated the same relationship around the world \citep{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
-In addition to suicide, other research has found effects temperature on hospitalization events related to mental health disorders such as bipolar disorder and schizophrenia \citep{Mullins2019Dec, Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
+Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May} with others demonstrating similar relationships in other locations around the world \citep{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
+Higher temperatures have also been correlated with increases in hospitalization events related to both general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and specific conditions like bipolar disorder and schizophrenia \citep{Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
 
@@
 Work on biological mechanisms emphasize that there may be mental health effects from maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.
-Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}
+
+Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}.
 
@@
 It could also indirectly influence mental health by affecting (1) physical health and (2)community well-being \citep{RN1314}.
-For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions(Anderson and Anderson 1998,Cheatwood 1995; Cohn et al.
-2004,Bouchama et al.2007.).
-Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration((Bridger 2003; Kerslake 1972; Kjellstrom et al.2009d,Leithead and Lind 1964,Schrier et al.
-1970).
-The loss of work capacity would in turn result in loss of income, which could also cause mental health problems(Kjellstrom 2009b).%% more physiological terms
-
-As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems(Miller et al.
-2009; Prince et al.
-2007), heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health((Berry 2007,World Health Organization 2008b).
-Second，disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities.
-For example, extreme heat reduces the work capacity of laborers in farm fields(Kjellstrom et al.
-2009d), which further destroys agricultural-supported industries in local area(Berry et al.
-2008).
+For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions \citep{RN1314,RN1316,RN1317,RN1318}.
+Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration \citep{RN1319,RN1320,RN1321,RN1322}.
+The loss of work capacity would in turn result in loss of income, while the low income could also cause mental health problems \citep{Katz1997}.
+As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems \citep{RN1323,RN1324} heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health \citep{RN1325,RN1326}.
+Second, disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities.
+For example, extreme heat reduces the work capacity of laborers in farm fields \citep{RN1320}, which further destroys agricultural-supported industries in local area\citep{RN1327}.
 The following economic pressures would undermine social capital and then lead to mental health problems.
 
-% Heatwaves are known to worsen mental health
+Recent research conducted on the effects of higher temperatures on mental health conducted at large, multi-city scales has found surprisingly uniform vulnerability.
+For example, focusing on suicides, Burke et al.
+found ``no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
+Similarly, Mullins et al.
+find no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status \cite{Mullins2019Dec}.
+One major challenge for these studies is that they rely on data from public health agencies that is aggregated to the municipality or city level.
+Thus, they must rely on between-county metrics of vulnerability and cannot take neighborhood effects into account.
 
-https://doi.org/10.1016/j.puhe.2018.06.008 (review study - will have more to cite)
-https://www.nature.com/articles/s41558-018-0222-x
-https://ehp.niehs.nih.gov/doi/full/10.1289/ehp.11339
-https://ijmhs.biomedcentral.com/articles/10.1186/s13033-018-0210-6
-https://www.nature.com/articles/s41558-018-0102-4/
-https://doi.org/10.1016/j.jhealeco.2019.102240
+%We would expect the poor to be more vulnerable, but this is not what we find
+The poor and disadvantaged are much more exposed to undesirable temperatures.
+They are more likely to work outside rather than indoors; they are more likely to rely on public transportation, bicycle, or walk instead of commute in their own air-conditioned car; and they are more likely to live in housing that is less insulated with poorer quality air-conditioning.
+Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, and these heterogeneities in vulnerability might be most pronounced at the neighborhood scale.
+Indeed this is what we find for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar}.
 
-[ADD SUICIDE DATA TO HEALTH SECTION?]
-https://www.psychologytoday.com/us/blog/greening-the-media/201809/global-warming-and-suicide
-
-\subsection{Mental Health and Vulnerability}
-% This is good, but probably should be shortened.  Maybe one para showing that income matters, one para discussing pathways, also maybe focus/change phrasing to be more on "vulnerability" -Matt
-
-% AND mental health is strongly associated with income
-% 1）Childhood 2）adulthood 3）gender
-% Perceived income level
-
-% Talk more about race
-% This is interesting: https://www.pnas.org/content/118/17/e2019624118.abstract?etoc
-
-The relationship between income, socioeconomic status (SES), well-being and mental health is one of the strongest established patterns in psycho-social literature \citep{Easterlin1974Jan, holzer1986increased, Perry1996Sep} (I cant find Ng et al.
-2014).
-Much of the research into these patterns focuses on the inequalities and distribution of the effects on individual well-being.
-Subsequent work on mental health outcomes, one measure of well-being, have established strong associations between SES, income and mental health with the most and least privileged most commonly associated with the best and worse mental health experiences and outcomes, respectively \citep{Sevenson2008Aug}(cites: Ng et al.
-2014).
-
-For example, results from a cross-sectional comparative study of socioeconomic factors and use of mental health services by people living in Ontario, Canada and the United States, indicate clear disparities in use of mental health services between those with the highest income levels and lowest mental health morbidity and those with the lowest income levels and highest mental health morbidity - those with higher incomes tended to have lower mental health morbidity issues while utilizing mental health services more than those with lower incomes and higher mental health issues \citep{Katz1997}.
-
-Results from a systematic review of the literature on associations between social inequalities and mental health disorders indicate a clear and prevailing link between  one or more indicators of less social privilege and higher prevalence of mental health disorders \citep{Fryers2003}, with low income one of the most consistent markers of and associations with increases in common mental health disorders.
-Collectively, results from the income and mental health literature indicate that socially disadvantaged populations experience significantly more frequent common mental health disorders.
-Scholars use mental health disparities to indicate the disproportionate amount of mental health disorders among persons of low SES \citep{RN1292}
-
-The established association between poverty and mental health revealed that that income distribution may have a significant influence upon mental health over and above the effect of poverty \citep{HANANDITA201459}.Poverty has been identified as one of the major risk factors in mental health.
-For instance, some systematic reviews have shown that socioeconomically disadvantaged children and adolescents were 2-3 times more likely to have mental health problems than others\citep{REISS201324}, while the household increases in financial resources are generally associated with an overall reduction of mental health problems\citep{2015Does}.
-Others have argued that increases in SES can serve as a buffer against the negative impacts of difficult life experiences on mental health, particularly for individuals already under substantive mental stress \citep{Kawachi2001Sep}.
-Thus, SES and income are associated with mental health improvements in SES and income can serve to decrease mental health problems.
-
-Several reasons could explain why poverty may effect mental health.
-1) the “social causation hypothesis” which suggests that stress or deprivation or decreasing the likelihood of people getting treatment may lead to poor mental health\citep{mills2015}, while social capital could reduce the likelihood of living in poverty and reduces the risk of mental disorders such as depressive symptoms and suicidal tendencies\citep{RN1291}.
-2) negative life events in a person’s life such as job loss are also associated with the risks of poverty\citep{RN1293} and mental health\citep{TAMPUBOLON201420}, and the causal inferences on the effect of poverty and mental health had also been made in several studies.
-For example, Tampubolon and Hanandita\citep{TAMPUBOLON201420} find that individual social capital is positively associated with mental health while adverse events were negatively associated; Chang.et.al \citep{RN1291} find that subjective and objective poverty is significantly associated with a higher risk of adverse life events, less social support and mental distress.
-Negative life events and social support in serial mediate the relationship between subjective poverty and mental health.
-
-% jieliu.cnah: Could delete
-For children, poverty and poor maternal mental health often co-exist and are two of the risk factors for child development\citep{LUND20111502}, and maternal mental health is significantly associated with child general psychopathology\citep{ RN1289}.
-Researchers also find that transition into poverty increased children’s socioemotional behavior problems and maternal psychological distress \citep{WICKHAM2017e141}.
-
-In general, the relationship between mental health and income level is well established.
-In the study, we explore whether income level would influence the link between weather and the sentiment in Twitter-posed text.
-
-https://jamanetwork.com/journals/jamapsychiatry/fullarticle/211213
-https://doi.org/10.1002/9781118410868.wbehibs570
-https://www.sciencedirect.com/science/article/abs/pii/S027795362030527X
-https://www.sciencedirect.com/science/article/pii/S2468266717300117
-https://doi.org/10.1016/S2468-2667(17)30011-7
-https://link.springer.com/article/10.1007/s00127-017-1370-4
- % Jeremiah Osborne-Gowey: Not as relevant.
-Saved the citation but didn't find enough relevant text to include reference to here in our manuscript.
-Maybe this: https://journals.plos.org/plosone/article?id=10.1371/journal.pone.0116820
-https://www.sciencedirect.com/science/article/abs/pii/S0143622814002537
-also this: https://www.sciencedirect.com/science/article/abs/pii/S0140673614614604
-
-\subsection{Income and Temperature}
-% AND wealthier better able to cope with heatwaves 
-https://www.nature.com/articles/nclimate3253
-https://www.sciencedirect.com/science/article/abs/pii/S0378778818321327
-https://www.mdpi.com/2225-1154/8/1/12/htm
-
-NOT CLEAR: https://doi.org/10.1016/j.jhealeco.2019.102240 and https://www.nature.com/articles/s41558-018-0222-x find little evidence of adaptation
-
-% BUT we know little about the interaction between heatwaves, mental health and income
-
-% THEREFORE: we did this really novel study.
-https://www.nature.com/articles/s41598-017-12961-9
-
-%OTHER LEFTOVER TEXT TO MAYBE BRING IN LATER
-
-Meteorological conditions can impact human physical (cite) and emotional states (cite).
-Emotional states and well-being are associated with physiological functioning and mental acuity which can affect social relationships, workplace productivity (cite) and health risks (cite).
-People that are more reliant on 1) livelihoods which require them to be outdoors (e.g., farming, agriculture, logging, etc.), 2) living and working in places where they are exposed to environmental minima and maxima, or 3) with prolonged exposure to ambient meteorological extremes are particularly vulnerable to changes in environmental exposure \citep{frimpong_heat_2017} and associated health risks (cite).
-Exposure to extreme environmental conditions can also have implications for workplace productivity and livelihoods \cite{kjellstrom_impact_2016}.
-For example, Nigerian maize farmers experience significant declines in productivity (2-8percent per degree above 17degC) along with substantive increases in health impacts including dehydration, muscle cramps, headaches and dizziness, heat exhaustion, sun stroke, and even death \cite{sadiq_impact_2019}.
-Other research indicates farm health and labor productivity were compromised under extreme heat and cold events in the Nepali Food Bowl region \cite{budhathoki_socio-economic_2019}.
-Similarly, Oregon farmers reported health impacts of working in both hot and cold conditions but the negative impacts were attenuated relative to the high heat situations \cite{bethel_heat-related_2014}.
-Other research found occupational injuries in Thailand increased with increasing heat exposure and occupational heat stress \cite{tawatsupa_association_2013}.
-Heat exposure risks are also a key factor in urban areas with particular attention needed for how the impacts of changing climate play out in health inequality \cite{friel_urban_2011}.
-
-Current estimates place annual workplace productivity losses due to heat exposure at 15-20percent with that rate potentially doubling by 2050 \cite{kjellstrom_heat_2016}.
-Projected increases in future ambient temperature suggest serious short- and long-term potential health consequences from exposure to heat stress (cite) with some estimates putting potential productivity losses at several percentage points by 2030 \cite{kjellstrom_heat_2016}, with middle- and low-income countries particularly vulnerable as they are often more reliant on physical work for their livelihoods.
-
-The level of changes in weather influences people’s mental health and patterns of emotions expressed.
-A study by Sun et al.
-2018 \href{file:///Users/portia/Downloads/ijerph-16-00086-v2 20(1).pdf.}{(Portia has this PDF?)} demonstrates varied relationships between haze and negative emotions of the public under different seasons of the year.
-Being sad or happy influences one's style and value of reasoning.
-Time of day, season, location, and climate allow aggregate prediction of sentiments \cite{hannak_tweetin_2012} \href{https://www.ccs.neu.edu/~amislove/publications/Weather-ICWSM.pdf}{(link to PDF)} while positive affect from an evaluative statement enhances potential responses \cite{clore_how_2007} \href{https://www.ncbi.nlm.nih.gov/pmc/articles/PMC2483304/pdf/nihms40349.pdf}{(link to PDF)}.
-For instance, changes in temperature affects people’s response about weather.
-Depending on a reference temperature for an  area and time of year, people are more likely to comment on unusual weather for a particular place and time than on the same weather considered typical in another place \cite{moore_rapidly_2019}.
-
-% Jeremiah Osborne-Gowey: Remove parts of this paragraph that relate to call for international research on these linkages as we now are only focusing on USA for Tweets.
-Although climate change is an international phenomenon experienced in every country, its discussions vary from country to country.
-According to Vu and others \cite{vu_nationalizing_2019}, a country’s climate severity, economic status and governance determine variations in its media discussions.
-% Jeremiah Osborne-Gowey: I think we can remove this entire paragraph now as it's outside the focus of this paper. Might be relevant for our 2nd paper!
-When removing the paragraph, cut from here and copy to the end of the doc where we have extra text saved. [DO NOT DELETE, just move from here]
-
-% Jeremiah Osborne-Gowey: This is perhaps the only piece of this paragraph still relevant for this paper...
-For example, Park and others \cite{park_mood_2013} \href{https://pdfs.semanticscholar.org/b282/feb759e57530b115dcc4bb080f96598a5246.pdf}{(link)} demonstrated that many people living in a state with higher mean temperature express more positive emotions on Twitter than those living in colder states in the US.
-Schmidt and others \cite{schmidt_media_2013} posit that climate severity and environmental factors related to carbon dependency influences the amount of coverage climate change receives in the media in countries like USA, Australia and Germany.
-Studies comparing how different countries portray climate change have been widely conducted in USA, UK, France and the Netherlands \cite{vu_nationalizing_2019}.
-Such comparisons are mainly between developed countries with developing countries outside the focus.
-Schafer and O’Neil \cite{schafer_what_2013} advance the need for academic scholars to investigate transnational contexts within climate communication.
-Wealthier countries with comparatively higher GDP are likely to frame climate change as a political issue as financial resources exist for exploring research on climate change.
-Conversely, coverage of climate change news from poorer developing countries focus mainly on international relations.
-Thus, the effects of national macro economic variables that affects a country’s sociopolitical and economic development such as GDP reflects a country’s governance system in media coverage \cite{vu_what_2018}.
-Cultural theorists assert that, understanding such differential variations depicts how different entities interpret danger and respond to risk \cite{tansey_cultural_1999}.
-Inclusion of developing countries in scientific research provides an avenue for international support to such countries in responding to the effects of climate change.Yet studies attempting to explain cross-national variation in climate change public opinion is limited \cite{knight_public_2016}.
-
-Public debates on the current consequences of climate change are discussed in both scientific and non-scientific mediums. 
-% Jeremiah Osborne-Gowey: Need a better transition here between meteorological conditions and effect on mood (expressed as sentiment).
-Feels like we're missing a discussion (new paragraph?) just above here that discusses this.
-This might be where we plug in the text PAW is working on?
-People can express their emotional states and feelings - sentiment - through physical, vocal and written expressions on mediums including newspapers, scientific articles, blogs and other online social media platforms.
-Twitter is one of the more popular social media platforms where people express their sentiments about any number of topics.
-Public posts on Twitter - called tweets - are widely available for public consumption and research, and offer a glimpse into collective social state.
-For example, text posted to these platforms can be used to gauge the public mood, assess opinions, measure brand affinity and for emergency planning and disaster response (cites).
-% Jeremiah Osborne-Gowey: Good to cite the top-cited public opinion papers, the Twitter flood study from CU, others from data science field including another Baylis paper
-Sentiment is frequently employed as a correlate of emotional state or mood (cites).
-Weather events focus public attention in different ways of expression (cites).
-These data offer a unique opportunity to analyze how people express and respond to events in people's daily lives, including their exposure to ambient environmental conditions.
-
-Here, we report on associations between meteorological conditions and expressed sentiment of public posts on Twitter from the United States of America (USA) between 2009 and 2019.
-This work builds on research from Baylis and others \cite{baylis_weather_2018} on weather impacts on expressed sentiment.
-In particular, we extend this work by including the temperature effects on sentiment in multiple regions across the USA while drawing in various climatic variables and socio-demographic data.
-In this research, we explore two primary questions.
-First, to what extent are ambient environmental conditions correlated with changes in expressed sentiment?
-Second, is this effect moderated by wealth?
-
-Building on previous findings from Baylis and others, we hypothesize that local ambient environmental conditions have a strong connection how people are feeling as expressed in the sentiment of text messages posted on Twitter [H1].
-Second, we hypothesize that the weather-sentiment link will be stronger in low-income areas.
-That is to say, people in lower income areas will be have lower sentiment scores than those in wealthier areas.
-[H2].
+While public health data on mental health outcomes is not available at the neighborhood scale, data from twitter can provide an important indicator of mental health at extremely precise spatial and temporal scales.
+Many studies have used the mood expressed in tweets as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
+Such studies have used twitter data to identify the onset of Post Traumatic Stress Disorder in individuals, even before their formal diagnosis \cite{Reece2017Oct}, while similar methods with facebook data have been used to predict the onset of depression \cite{Eichstaedt2018Oct}.
+Additionally, in London, day-to-day changes in mental health indicators derived from tweets have been associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}, while across the US more positive tweets are associated with a variety of metrics of human well being \cite{Mitchell2013May}.
+Finally, the mood expressed in tweets has been strongly associated with local weather \cite{baylis_weather_2018}, and has been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+Thus, we used a common sentiment metric to determine if neighborhood characteristics affect the vulnerability of mental health to higher temperatures.
 
@@
 
-\subsection{Effects by Time of Day}
-
-In addition to providing data with high spatial resolution, twitter data also comes with very high temporal resolution as each tweet is time-stamped.
-Thus, we were able to examine the impact of heat on sentiment for each hour of the day (See Fig.
-\ref{fig:ts-wbgt}).
-We found that heat decreased expressed sentiment at nearly all hours of the day, and that heat had the largest effect on sentiment early in the morning, between 4am and 6am.
-Heat had the weakest effects on sentiment at midday, and at 3am.
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=0.5\linewidth]{../res/ts.png}
-  \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
-The value shown is the predicted change in VADER score for a 25-degree increase in temperatures.}
-  \label{fig:ts-wbgt}
-\end{figure}
-
 \section{Discussion}
@@
 However, one limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
-Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a users expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
+Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
 We overcome this limitation by using an extremely large data set of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
@@
 We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the VADER sentiment score of 0.01.
-This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a change in sentiment associated with a suicides increase of 23\% over the course of the study period \cite{CDC2021}.
+This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with a increase in suicides of 23\% over the course of the study period \cite{CDC2021}.
 Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.
 Moreover, both suicides and sentiment are affected by higher temperatures \cite{baylis_weather_2018, Burke2018Aug}.
-Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood and hourly effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.
+Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.
+
+% Discuss the NOVELTY: quote Burke and Mullins, refute their interpretations, people have said XX, but we have shown ~XX.
+% Variations in vulnerability also indicates ADAPTATION
+Previous work conducted with county-scale data found no heterogeneities in mental health vulnerability by income.
+Such analyses have led to conclusions that (1) the mental health effects of climate change will be uniform, least in developed countries, and (2) that adaptation with technologies like air-conditioning will not be enough.
+For example, Mullins et al.
+conclude that ``high-temperature events will harm mental health everywhere" and that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}.
+Contrary to these analyses, we found large heterogeneities in vulnerability by neighborhood race and income.
+This suggests that the mental health effects of climate change will not be uniform and, like other impact, will fall disproportionately on the poor and vulnerable.
+More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar those of high-income Americans, the effects of high temperatures on mental health can be largely mitigated.
 
@@
 This is surprising, given that Hispanics can be marginalized in both income and housing.
-This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans.
+This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans, or may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in Spanish, which we excluded from our analysis.
 Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
@@
 % Interpret Spatial Analysis
-In our spatial analysis, we found that most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, with noticeable regional patterns in the strength of the effect.
-Generally, the mid-west and mid-atlantic had the strongest inequalities, and the south-west had the weakest inequalities and an occasional inverted effect.
-Because we used Wet Bulb Globe Temperature, our metric accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
-Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different environmental conditions.
+In our spatial analysis, we found that heat affects sentiment in cities across the US, suggesting that higher baselines temperatures do not reduce vulnerability.
+Additionally, most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, although the south-west had the weakest inequalities and an occasional inverted effect.
+While this could suggest that heat inequality is related humidity, this is unlikely because we used Wet Bulb Globe Temperature, which accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
+Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different baseline environmental conditions.
 
@@
 Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represented the segmented regression for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
+We estimate the 95\% confidence interval of our models using 80-fold bootstrapping.
 
@@
 We specify alternative models where income is a dummy variable in three bins, and were race is a continuous variable for the percent of a census block population that is what.
-These alternative specifications yield similar results (See Supplement).
-We also present the effects of rainfall and sunshine on sentiment by income and race in the Supplement.
+These alternative specifications yield similar results.
+We also examine the effects of rainfall and sunshine on sentiment by income and race.
+These alternative specifications and analysis of rainfall and sunshine effects are available in the Supplement.
 
-We estimate the 95\% confidence interval of our models using 80-fold bootstrapping.
-For our analyses by CSA and by time of day, we subset the data in the model and run the model with the same fixed effects, except we exclude time of day as a fixed effect in the temporal analysis.
+For our analyses by CSA, we subset the data in the model and run the model with the same fixed effects, although we fit a linear effect for $f_t()$.
 

### 2021-05-12 csa/model_space.R
update map results

@@ -8,16 +8,16 @@ library(USAboundaries)
 data <- fread('~/tweets/all.csv')
 csadat <- fread('~/tweets/csa/fips-csa.csv')
 
-qs_inc <- quantile(data$income_percap, seq(0, 1, by=0.05))
-qs_white <- quantile(data$race_white, seq(0, 1, by=0.05))
-qs_black <- quantile(data$race_black, seq(0, 1, by=0.05))
-
 data <- merge(data, csadat, on='fips', all=F)
 
 data$month <- substr(data$statemonth, 4, 5)
 data$income_percap <- log(data$income_percap)
 data <- data[weather_term == 0, ]
 data$raining <- data$prcp > 0
+
+qs_inc <- quantile(data$income_percap, seq(0, 1, by=0.05))
+qs_white <- quantile(data$race_white, seq(0, 1, by=0.05))
+qs_black <- quantile(data$race_black, seq(0, 1, by=0.05))
 
 csas <- tab(data$id)
 csas <- as.numeric(names(csas))[order(csas, decreasing=T)]
@@ -56,6 +56,7 @@ for (csasel in csas){
 }
 
 write.csv(allres, '~/tweets/csa/model_res_geo.csv', row.names=F)
+allres <- read.csv('~/tweets/csa/model_res_geo.csv')
 
 #########################
 # Visualize
@@ -101,7 +102,7 @@ m <- merge(csa_sf, allres %>% filter(term == 'wbgt:income_percap'))
 m$estimate <- 25*(m$estimate/(qs_inc[2] - qs_inc[20]))
 ggplot() + 
   geom_sf(data = states) + 
-  geom_sf(data = m, aes(fill=est_change)) + 
+  geom_sf(data = m, aes(fill=estimate)) + 
   geom_sf(data = states, fill=NA) + 
   geom_sf(data = m, fill=NA, aes(color=p.value < 0.05, size=p.value < 0.05)) + 
   scale_fill_gradient2(low='#e66101', mid='#f7f7f7', high='#5e3c99', midpoint=0) + 
@@ -124,7 +125,7 @@ ggsave('~/temp-sentiment/res/map_wbgt_income.png', width=12, height=8)
 # Black results
 ###########################
 m <- merge(csa_sf, allres %>% filter(term == 'wbgt:race_black'))
-m$estimate <- 25*(m$estimate/(qs_white[20] - qs_white[2]))
[…]

### 2021-05-12 csa/combine_maps.sh
make combination plots

@@ -1,0 +1,25 @@
+cd ~/temp-sentiment/res
+
+mkdir conv
+cd conv
+
+#Crop out all whitespace
+convert ../map_wbgt.png -trim +repage map_wbgt.png
+convert ../map_wbgt_income.png -trim +repage map_wbgt_income.png
+convert ../map_wbgt_black.png -trim +repage map_wbgt_black.png
+
+#Add some margins back (but narrower)
+convert map_wbgt.png -gravity center -background white -extent $(identify -format '%[fx:W+50]x%[fx:H+50]' map_wbgt.png) map_wbgt.png
+convert map_wbgt_income.png -gravity center -background white -extent $(identify -format '%[fx:W+50]x%[fx:H+50]' map_wbgt_income.png) map_wbgt_income.png
+convert map_wbgt_black.png -gravity center -background white -extent $(identify -format '%[fx:W+50]x%[fx:H+50]' map_wbgt_black.png) map_wbgt_black.png
+
+#Add labels
+convert map_wbgt.png -gravity NorthWest -pointsize 200 -font "Arial-Bold" -annotate +50+50 'A' map_wbgt.png
+convert map_wbgt_income.png -gravity NorthWest -pointsize 200 -font "Arial-Bold" -annotate +50+50 'B' map_wbgt_income.png
+convert map_wbgt_black.png -gravity NorthWest -pointsize 200 -font "Arial-Bold" -annotate +50+50 'C' map_wbgt_black.png
+
+#Stack
+convert -append map_wbgt.png map_wbgt_income.png map_wbgt_black.png ../map_combined.png
+
+cd ..
+rm -rf conv

### 2021-05-12 doc/temp-sentiment.tex

@@
-\documentclass{article}
+\documentclass[fleqn,10pt]{wlscirep}
+\usepackage[utf8]{inputenc}
+\usepackage[T1]{fontenc}
 \usepackage{graphicx}
 \usepackage{hyperref}
-\usepackage[a4paper, margin=1.25in]{geometry}
 \usepackage{breakcites}
@@
 \usepackage{xcolor}
-\usepackage[
-  style=numeric,
-  citestyle=numeric-comp,
-  backend=biber,
-  doi=true,
-  natbib=true,
-  sorting=none
-]{biblatex}
+% \usepackage[
+%   style=numeric,
+%   citestyle=numeric-comp,
+%   backend=biber,
+%   doi=true,
+%   natbib=true,
+%   sorting=none
+% ]{biblatex}
 \usepackage{scalerel,xparse}
-
-\addbibresource{TextDataClimateShocks.bib}
 
@@
 
-\begin{document}
 %https://www.overleaf.com/project/5f918edd1afc380001feca0d
@@
 
-\maketitle
-
 \begin{abstract}
@@
 We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.
+\end{abstract}
+\begin{document}
 
-\end{abstract}
+\flushbottom
+\maketitle
+\thispagestyle{empty}
 
-\section{Main}
+\section*{Main}
 %Overview
-Climate change is impacting many aspects of human well-being with these impacts projected to worsen in the coming years \cite{pachauri2014climate}.
+Climate change is impacting many aspects of human well-being with the impacts projected to worsen in the coming years \cite{pachauri2014climate}.
 A large body of research has assessed the ongoing impacts on climate change on outcomes like agriculture, economic growth, and physical health (NEED CITATIONS).
@@
 Higher temperatures are associated with many indicators of worsened mental health.
-Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May} with others demonstrating similar relationships in other locations around the world \citep{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
-Higher temperatures have also been correlated with increases in hospitalization events related to both general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and specific conditions like bipolar disorder and schizophrenia \citep{Lee2007Jan, Shapira2004Feb, Sung2013Feb, Gupta1992Jun, Hansen2008Oct}.
+Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other work demonstrating similar relationships in other locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
+Higher temperatures have also been correlated with increases in hospitalization events related to both general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb, Gupta1992Jun} as well as increased mortality among people with mental health disorders \cite{Hansen2008Oct}.
 
@@
 
-Additionally, recent studies have found that nighttime temperatures affects sleep quality \citep{Obradovich2017May, Mullins2019Dec}.
+Additionally, recent studies have found that nighttime temperatures affects sleep quality \cite{Obradovich2017May, Mullins2019Dec}.
 
 % The pathways
-Higher temperature associated with climate change could influence mental health directly by exposing people to trauma.
-It could also indirectly influence mental health by affecting (1) physical health and (2)community well-being \citep{RN1314}.
-For the direct influence pathway, extreme heat events or increasing temperatures have been associated with the increase in aggression,higher suicide rates and other hospital admissions \citep{RN1314,RN1316,RN1317,RN1318}.
-Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration \citep{RN1319,RN1320,RN1321,RN1322}.
-The loss of work capacity would in turn result in loss of income, while the low income could also cause mental health problems \citep{Katz1997}.
-As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems \citep{RN1323,RN1324} heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health \citep{RN1325,RN1326}.
+Higher temperature associated with climate change could influence mental health both directly and indirectly by exposing people to trauma (direct) (CITATIONS) or by affecting (1) physical health and (2)community well-being (indirect) \cite{BerryETAL2010}.
+Direct pathways can include extreme heat events or increasing temperatures associated with loss of workplace productivity from decreases in occupational health or increases in workplace injury \cite{kjellstrom_heat_2016, KjellstromETAL2009, kjellstrom_impact_2016} which can result in loss of income, all of which can exacerbate deteriorating mental health and lead to increased aggression, hospital admissions and suicide rates \cite{BerryETAL2010,CheatwoodDerral1995,CohnETAL2004,BouchamaETAL2007}.
+Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration \cite{KjellstromETAL2009}.
+The loss of work capacity would in turn result in loss of income, while the low income could also cause mental health problems \cite{Katz1997}.
+As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems \cite{MillerETAL2009,PrinceETAL2007} heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health \cite{Berry2007,WHO2007}.
 Second, disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities.
-For example, extreme heat reduces the work capacity of laborers in farm fields \citep{RN1320}, which further destroys agricultural-supported industries in local area\citep{RN1327}.
+For example, extreme heat reduces the work capacity of laborers in farm fields \cite{KjellstromETAL2009}, which further destroys agricultural-supported industries in local area\cite{BerryETAL2008}.
 The following economic pressures would undermine social capital and then lead to mental health problems.
@@
 
-\section{Results}
+\section*{Results}
 
-\subsection{Overall Effect}
+\subsection*{Overall Effect}
 We examined the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an indicator of heat stress that accounts for temperature, humidity, wind speed, and solar radiation.
@@
 
-\subsection{Comparison With Other Events}
+\subsection*{Comparison With Other Events}
 
@@
   \centering
-  \includegraphics[width=\linewidth]{../res/comparison_plot.png}
+  \includegraphics[width=0.66\linewidth]{../res/comparison_plot.png}
   \caption{Decline in sentiment during a heat wave across several different neighborhoods, compared with other impacts on sentiment, such as major hurricanes, as well as the weekly variation in sentiment from the peak on Saturday to the low on Monday.}
@@
 
-\subsection{Effects by Combined Statistical Area}
+\subsection*{Effects by Combined Statistical Area}
 
@@
 
-\section{Discussion}
+\section*{Discussion}
 % Benefits of twitter data/overview of findings
@@
 
-% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
+% NCC doesnt have a "conclusion" section*, but this is intended as an overall summary paragraph
 While climate change will have widespread and severe impacts on human well-being, it cannot be overemphasized just how highly unevenly these impacts will be distributed.
@@
 
-\section{Methods}
-\subsection{Tweets \& Sentiment}
+\section*{Methods}
+\subsection*{Tweets \& Sentiment}
 We used data from 243 million geo-located English-language tweets from the lower 48 states from the years 2009 to 2019.
@@
 
-We assessed the sentiment expressed in the tweets using the VADER (Valence Aware Dictionary for sEntiment Reasoning) sentiment corpus \cite{gilbert_vader_2014}.
+We assessed the sentiment expressed in the tweets using the VADER (Valence Aware Dictionary for sentiment Reasoning) sentiment corpus \cite{gilbert_vader_2014}.
 VADER is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media because of several features, as:
@@
 
-\subsection{Weather}
+\subsection*{Weather}
 We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington.
@@
 
-\subsection{Socio-Econoimc Data}
+\subsection*{Socio-Econoimc Data}
 We used data from the American Community Survey (ACS) administered by the US Census to estimate income levels and the racial composition of neighborhoods where tweets were located.
@@
 
-\subsection{Modeling}
+\subsection*{Modeling}
 We assessed how expressed sentiment is affected by wet bulb globe temperature using segmented regressions, controlling for precipitation, and shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: the day of the week, the time of day, the day of the year, the year, the month, and the county.
@@
 
-\section{Acknowledgements}
+\section*{Acknowledgements}
 %Official SESYNC wording
@@
 The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
-\printbibliography
 
-\section{Author Contributions}
-Author contributions: M.W.C.
+%\printbibliography
+\bibliography{TextDataClimateShocks}
+
+\section*{Author Contributions}
+Author contributions: M.W.C., J.O.G.
 and P.B.
-designed research and modeling strategy; A.S.
+designed the research and modeling strategy; A.S.
 provided data; M.W.C and Z.L.
@@
 
-This project was co-lead by M.W.C. and J.O.G.
+\noindent This project was co-lead by M.W.C. and J.O.G.
 

### 2021-05-18 viz/afinn/wbgt_income_bootstrap.R
add afinn results; resize graphs

@@ -1,30 +1,30 @@
-library(data.table)
-library(fixest)
-library(survey)
 library(tidyverse)
 library(cowplot)
+library(survey)
+library(data.table)
+
+MOD_RUN <- 'afinn_income_percap'
 
 options(scipen=100)
-
-MOD_RUN <- 'hedono_race_income'
 
 setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
 
 data <- fread('~/tweets/all.csv')
 data$income_percap <- log(data$income_percap)
 
+qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
+
 data <- data[weather_term == 0, ]
 
-inc_qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
-race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
+data$raining <- data$prcp > 0
 
 #Tempknots
 knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
                           max(data$wbgt)))
 
-data$raining <- data$prcp > 0
-
-#Define Segmenting Function
+######################################
+# Define Functions
+#####################################
 piece.formula <- function(var.name, knots, interact_var='') {
   knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
@@ -46,6 +46,7 @@ vcov.bootmod <- function(object){
 coef.bootmod <- function(object){
   return(object$coef)
 }
+
 
 fill_knots <- function(x, filln=10){
   vals <- NULL
@@ -76,27 +77,26 @@ for (t in seq(-20, 30, by=10)){
 }
 ref_temps <- round(ref_temps)
 
+
 ########################################
 # WBGT
 #########################################
 preddf <- data.frame(wbgt=seq(-22, 33))
-preddf <- make_groups(preddf, 'income_percap', inc_qs[c(2, 11, 20)])
-preddf <- make_groups(preddf, 'race_white', race_qs[c(2, 11, 20)])
+preddf <- make_groups(preddf, 'income_percap', qs[c(2, 11, 20)])
 
-preddf$hedono <- 1
+preddf$afinn <- 1
 
-mm <- model.matrix(as.formula(paste0("hedono ~ ", 
-                     piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
[…]

### 2021-05-18 viz/afinn/wbgt_nointeract_bootstrap.R

@@ -1,22 +1,17 @@
-library(data.table)
-library(fixest)
-library(survey)
 library(tidyverse)
 library(cowplot)
+library(survey)
+library(data.table)
 
 options(scipen=100)
 
-MOD_RUN <- 'hedono_race_income'
+MOD_RUN <- 'afinn_wbgt_nointeract'
 
-setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
+setwd(paste0('~/tweets/bootstrap/', MOD_RUN))
 
 data <- fread('~/tweets/all.csv')
-data$income_percap <- log(data$income_percap)
 
 data <- data[weather_term == 0, ]
-
-inc_qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
-race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
 
 #Tempknots
 knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
@@ -24,7 +19,13 @@ knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25,
 
 data$raining <- data$prcp > 0
 
-#Define Segmenting Function
+#Save memory by deleting useless columns!
+# data <- data[ , c('wbgt', 'raining', 'srad', 'afinn', 'fips', 'year', 
+#                   'statemonth', 'tod', 'doy', 'dow', 'temp')]
+
+######################################
+# Define Functions
+#####################################
 piece.formula <- function(var.name, knots, interact_var='') {
   knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
@@ -46,6 +47,7 @@ vcov.bootmod <- function(object){
 coef.bootmod <- function(object){
   return(object$coef)
 }
+
 
 fill_knots <- function(x, filln=10){
   vals <- NULL
@@ -76,25 +78,19 @@ for (t in seq(-20, 30, by=10)){
 }
 ref_temps <- round(ref_temps)
 
+
 ########################################
 # WBGT
 #########################################
 preddf <- data.frame(wbgt=seq(-22, 33))
-preddf <- make_groups(preddf, 'income_percap', inc_qs[c(2, 11, 20)])
-preddf <- make_groups(preddf, 'race_white', race_qs[c(2, 11, 20)])
 
-preddf$hedono <- 1
+preddf$afinn <- 1
 
[…]

### 2021-05-18 viz/afinn/wbgt_race_q_bootstrap.R

@@ -1,43 +1,48 @@
-library(data.table)
-library(fixest)
-library(survey)
 library(tidyverse)
 library(cowplot)
+library(survey)
+library(data.table)
+
+MOD_RUN <- 'afinn_race_q'
 
 options(scipen=100)
-
-MOD_RUN <- 'hedono_race_income'
 
 setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
 
 data <- fread('~/tweets/all.csv')
-data$income_percap <- log(data$income_percap)
 
 data <- data[weather_term == 0, ]
 
-inc_qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
-race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
+data$raining <- data$prcp > 0
 
 #Tempknots
 knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
                           max(data$wbgt)))
 
-data$raining <- data$prcp > 0
-
-#Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='') {
+######################################
+# Define Functions
+#####################################
+# DIFFERENT FOR GROUPING VARS
+piece.formula <- function(var.name, knots, interact_var='', interact_type=c('factor', 'continuous', 'none')) {
   knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
   formula.sign[knots < 0] <- " + "
-  if (interact_var == ''){
-    paste(var.name, "+",
+  if (interact_type == 'none'){
+    formula <- paste(var.name, "+",
         paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
               collapse = " + ", sep=""))
-  } else{
-    paste(interact_var, '*', var.name, "+",
-        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
   }
+  if (interact_type == 'continuous'){
+    formula <- paste(interact_var, '*', var.name, "+",
+          paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
+                collapse = " + ", sep=""))
+  } 
+  if (interact_type == 'factor'){
+    formula <- paste(interact_var, '*', var.name, "+",
[…]

### 2021-05-18 viz/hedono/wbgt_income_bootstrap.R

@@ -172,4 +172,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/hedono-wbgt-income.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/hedono-wbgt-income.png', width=4.5, height=4)

### 2021-05-18 viz/hedono/wbgt_income_q_bootstrap.R

@@ -173,4 +173,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/hedono-wbgt-income_q.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/hedono-wbgt-income_q.png', width=4.5, height=4)

### 2021-05-18 viz/hedono/wbgt_nointeract_bootstrap.R

@@ -152,5 +152,5 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/hedono-wbgt.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/hedono-wbgt.png', width=4.5, height=4)
 

### 2021-05-18 viz/hedono/wbgt_race_bootstrap.R

@@ -171,5 +171,5 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/hedono-wbgt-race.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/hedono-wbgt-race.png', width=4.5, height=4)
 

### 2021-05-18 viz/hedono/wbgt_race_q_bootstrap.R

@@ -168,4 +168,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/hedono-wbgt-race_q.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/hedono-wbgt-race_q.png', width=4.5, height=4)

### 2021-05-18 viz/vader/wbgt_income_bootstrap.R

@@ -132,6 +132,7 @@ preddf <- preddf %>%
   summarize(ymin = quantile(value, probs=0.025),
             ymax = quantile(value, probs=0.975),
             pred = quantile(value, probs=0.5))
+write.csv(preddf, '~/tweets/viz_cache/vader/wbgt-income.csv', row.names=F)
 
 curve <- ggplot(preddf) + 
   geom_line(aes(x=wbgt, y=pred, color=income_percap)) + 
@@ -173,7 +174,7 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt-income.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/wbgt-income.png', width=4.5, height=4)
 
 ########################################
 # Precip
@@ -226,6 +227,7 @@ preddf <- preddf %>%
   summarize(ymin = quantile(value, probs=0.025),
             ymax = quantile(value, probs=0.975),
             pred = quantile(value, probs=0.5))
+write.csv(preddf, '~/tweets/viz_cache/vader/precip-income.csv', row.names=F)
 
 ggplot(preddf %>% filter(raining)) + 
   geom_bar(aes(x=income_percap, y=pred, fill=income_percap), stat='identity') + 
@@ -240,7 +242,7 @@ ggplot(preddf %>% filter(raining)) +
   theme(axis.ticks.x = element_blank(),
         panel.grid.major = element_line(color = "lightgrey", size=0.5),
         panel.grid.minor = element_line(color = "lightgrey", size=0.25))
-ggsave('~/temp-sentiment/res/raining-income.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/raining-income.png', width=4.5, height=4)
 
 ########################################
 # srad
@@ -292,6 +294,7 @@ preddf <- preddf %>%
   summarize(ymin = quantile(value, probs=0.025),
             ymax = quantile(value, probs=0.975),
             pred = quantile(value, probs=0.5))
+write.csv(preddf, '~/tweets/viz_cache/vader/srad-income.csv', row.names=F)
 
 curve <- ggplot(preddf) + 
[…]

### 2021-05-18 viz/vader/wbgt_income_q_bootstrap.R

@@ -173,4 +173,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt-income_q.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/wbgt-income_q.png', width=4.5, height=4)

### 2021-05-18 viz/vader/wbgt_income_race_bootstrap.R

@@ -159,7 +159,7 @@ preddf <- preddf %>%
        color = 'Census Block\nIncome Per-Capita\n(Percentile)',
        linetype = 'Census Block\nPercent Minority') +
   theme_classic() + 
-  theme(legend.position=c(0.3, 0.4),
+  theme(#legend.position=c(0.3, 0.4),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -189,4 +189,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt-income-race.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/wbgt-income-race.png', width=5.5, height=4)

### 2021-05-18 viz/vader/wbgt_nointeract_bootstrap.R

@@ -116,6 +116,7 @@ preddf <- preddf %>%
   summarize(ymin = quantile(value, probs=0.025),
             ymax = quantile(value, probs=0.975),
             pred = quantile(value, probs=0.5))
+write.csv(preddf, '~/tweets/viz_cache/vader/wbgt.csv', row.names=F)
 
 curve <- ggplot(preddf) + 
   geom_line(aes(x=wbgt, y=pred)) + 
@@ -152,5 +153,5 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/wbgt.png', width=4.5, height=4)
 

### 2021-05-18 viz/vader/wbgt_race_bootstrap.R

@@ -171,5 +171,5 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt-race.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/wbgt-race.png', width=4.5, height=4)
 

### 2021-05-18 viz/vader/wbgt_race_q_bootstrap.R

@@ -128,6 +128,7 @@ preddf <- preddf %>%
   summarize(ymin = quantile(value, probs=0.025),
             ymax = quantile(value, probs=0.975),
             pred = quantile(value, probs=0.5))
+write.csv(preddf, '~/tweets/viz_cache/vader/wbgt-race_q.csv', row.names=F)
 
 curve <- ggplot(preddf) + 
   geom_line(aes(x=wbgt, y=pred, color=race_majority)) + 
@@ -169,7 +170,7 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt-race_q.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/wbgt-race_q.png', width=4.5, height=4)
 
 ########################################
 # Precip
@@ -226,7 +227,7 @@ ggplot(preddf %>% filter(raining)) +
   theme(axis.ticks.x = element_blank(),
         panel.grid.major = element_line(color = "lightgrey", size=0.5),
         panel.grid.minor = element_line(color = "lightgrey", size=0.25))
-ggsave('~/temp-sentiment/res/raining-race_q.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/raining-race_q.png', width=4.5, height=4)
 
 ########################################
 # srad
@@ -302,4 +303,4 @@ hist <- ggplot(data[sample(1:nrow(data), nrow(data)*0.01), 'srad']) +
 
 plot_grid(curve, hist, align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2))
-ggsave('~/temp-sentiment/res/srad-race_q.png', width=6, height=5)
+ggsave('~/temp-sentiment/res/srad-race_q.png', width=4.5, height=4)

### 2021-05-18 doc/temp-sentiment-supplement.tex
Updates from Overleaf

--- 2021-05-01 doc/temp-sentiment-supplement.tex
@@
 \usepackage{hyperref}
-\usepackage[a4paper, margin=1.25in]{geometry}
+\usepackage[a4paper, margin=1in]{geometry}
 \usepackage{breakcites}
@@
 
-%Supplement will have:
-
-%Equations used for RH and WBGT
-
-\section{Alternative Parameterizations}
-
-\section{Effects of Precipitation and Sunshine}
-
-\section{Alternative Sentiment Algorithm}
-
 \setcounter{table}{0}
@@
 
-\printbibliography
+\section{Alternative Sentiment Algorithm}
 
-\end{document}
+In addition to the Valence Aware Dictionary for sentiment Reasoning (VADER) sentiment analysis method \cite{gilbert_vader_2014}, we also used the hedonometer method \cite{Dodds2011Dec}.
+This sentiment metric yielded very similar findings to those presented in the main paper with the VADER method.
 
-\section*{Supplemental Info}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thesection}{S\arabic{section}}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/Boston_Map.png}
-  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
-  \label{fig:timeseries}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/map_wbgt.png}
-  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
-  \label{fig:timeseries}
-\end{figure}
-
-\subsubsection{Hedonometer}
 The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
@@
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt.png}
+  \caption{Effects of temperature on sentiment, measured with the hedonometer method.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt-income.png}
+  \caption{Effects of temperature on sentiment and moderated by income, measured with the hedonometer method.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt-race_q.png}
+  \caption{Effects of temperature on sentiment and moderated by race, measured with the hedonometer method.}
+\end{figure}
+
+\section{Alternative Interaction Variables}
+For our main analysis, we treated income as a continuous (log-transformed) variable and race as a categorical variable.
+We ran similar analyses with income as a categorical variable and race as a continuous variable (the percent of a census block's population that was non-white/minority).
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/wbgt-income_q.png}
+  \caption{Effect of temperature on sentiment and moderated by income, with income as a categorical variable.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/wbgt-race.png}
+  \caption{Effect of temperature on sentiment and moderated by race, with race as a continuous variable.}
+\end{figure}
+
+\section{Effects of Precipitation and Sunshine}
+While our main variable of interest was temperature as measured by Wet Bulb Globe Temperature, we also interacted our variables of interest with a binary variable for whether it was raining or not, as well as a continous variable for the amount of shortwave radiation in ($W/m^2$), or sunshine.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/raining-income.png}
+  \caption{Effect of rain on sentiment and moderated by income.}
+  \label{fig:timeseries}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/raining-race_q.png}
+  \caption{Effect of rain on sentiment and moderated by race.}
+  \label{fig:timeseries}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/srad-income.png}
+  \caption{Effect of sunshine on sentiment and moderated by income.}
+  \label{fig:timeseries}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/srad-race_q.png}
+  \caption{Effect of sunshine on sentiment and moderated by race.}
+  \label{fig:timeseries}
+\end{figure}
+
+\section{Combined Effects of Race and Income}
+
+In addition to our models exploring how income and race moderate the effect of heat on sentiment, we fit an additional model with an interaction terms for both race and income.
+Due to the complexity of a model with spline terms interacting with two different variables, using a categorical variable for race would have yielded a model matrix too large for the largest cloud virtual machine availale to us.
+Thus, we used a continuous metric for both income as well as race (percent non-white/minority).
+
+To examine how income and racial groups together moderate the effect of heat on sentiment, we extend our main model to the following form:
+
+\begin{equation}
+    y = \begin{tabular}{l}
+    \beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) + 
+    \beta_p p + \beta_{m_{r} p} m_{r} p + 
+    \beta_{m_{i} p} m_{i} p + \beta_{m_{i} m_{r} p} m_{i} m_{r} p + 
+    \\
+    \beta_s s + \beta_{m_{i} s} m_{i} s + 
+    \beta_{m_{r} s} m_{r} s + \beta_{m_{i} m_{r} s} m_{i} m_{r} s + 
+    \Phi + \epsilon
+    \end{tabular}
+    \label{mod:2}
+\end{equation}
+
+Identically to the model in the main body of the paper, $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represented the segmented regression for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
+We estimate the 95\% confidence interval of our models using 80-fold bootstrapping.
+
+In this new model, $m_{i}$ is the log-transformed average income in the census block where the tweet originated, while $m_{r}$ is the racial composition of the census block, in terms of the percent of the population that was a minority.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/wbgt-income-race.png}
+  \caption{Effect of temperature on sentiment and moderated by both race and income.}
+  \label{fig:timeseries}
+\end{figure}
+
+\section{Weather Terms}
+We excluded tweets that contained the following weather-related terms: aerovane, air, airstream, altocumulus, altostratus, anemometer, anemometers, anticyclone, anticyclones, arctic, arid, aridity, atmosphere, atmospheric, autumn, autumnal, balmy, baroclinic, barometer, barometers, barometric, blizzard, blizzards, blustering, blustery, blustery, breeze, breezes, breezy, brisk, calm, celsius, chill, chilled, chillier, chilliest, chilly, chinook, cirrocumulus, cirrostratus, cirrus, climate, climates, cloud, cloudburst, cloudbursts, cloudier, cloudiest, clouds, cloudy, cold, colder, coldest, condensation, contrail, contrails, cool, cooled, cooling, cools, cumulonimbus, cumulus, cyclone, cyclones, damp, damp, damper, damper, dampest, dampest, degree, degrees, deluge, dew, dews, dewy, doppler, downburst, downbursts, downdraft, downdrafts, downpour, downpours, dried, drier, dries, driest, drizzle, drizzled, drizzles, drizzly, drought, droughts, dry, dryline, fall, farenheit, flood, flooded, flooding, floods, flurries, flurry, fog, fogbow, fogbows, fogged, fogging, foggy, fogs, forecast, forecasted, forecasting, forecasts, freeze, freezes, freezing, frigid, frost, frostier, frostiest, frosts, frosty, froze, frozen, gale, gales, galoshes, gust, gusting, gusts, gusty, haboob, haboobs, hail, hailed, hailing, hails, haze, hazes, hazy, heat, heated, heating, heats, hoarfrost, hot, hotter, hottest, humid, humidity, hurricane, hurricanes, ice, iced, ices, icing, icy, inclement, landspout, landspouts, lightning, lightnings, macroburst, macrobursts, maelstrom, mercury, meteorologic, meteorologist, meteorologists, meteorology, microburst, microbursts, microclimate, microclimates, millibar, millibars, mist, misted, mists, misty, moist, moisture, monsoon, monsoons, mugginess, muggy, nexrad, nippy, NOAA, nor,’easter, nor,’easters, noreaster, noreasters, overcast, ozone, parched, parching, pollen, precipitate, precipitated, precipitates, precipitating, precipitation, psychrometer, radar, rain, rainboots, rainbow, rainbows, raincoat, raincoats, rained, rainfall, rainier, rainiest, raining, rains, rainy, sandstorm, sandstorms, scorcher, scorching, searing, shower, showering, showers, skiff, sleet, slicker, slickers, slush, slushy, smog, smoggier, smoggiest, smoggy, snow, snowed, snowier, snowiest, snowing, snowmageddon, snowpocalypse, snows, snowy, spring, sprinkle, sprinkles, sprinkling, squall, squalls, squally, storm, stormed, stormier, stormiest, storming, storms, stormy, stratocumulus, stratus, subtropical, summer, summery, sun, sunnier, sunniest, sunny, temperate, temperature, tempest, thaw, thawed, thawing, thaws, thermometer, thunder, thundered, thundering, thunders, thunderstorm, thunderstorms, tornadic, tornado, tornadoes, tropical, troposphere, tsunami, turbulent, twister, twisters, typhoon, typhoons, umbrella, umbrellas, vane, warm, warmed, warming, warms, warmth, waterspout, waterspouts, weather, wet, wetter, wettest, wind, windchill, windchills, windier, windiest, windspeed, windy, winter, wintery, and wintry.
+
+\printbibliography
+
 \end{document}

### 2021-05-18 doc/temp-sentiment.tex
Updates from Overleaf

@@
 \usepackage{breakcites}
-\usepackage{subcaption}
 \usepackage{float}
@@
 
-\author[1, *]{Matthew Cooper}
+\author[1,*]{Matthew Cooper}
 \author[2]{Jeremiah Osborne-Gowey}
@@
 \affil[1]{T.H. Chan School of Public Health, Harvard University}
-\affil[2]{University of Colorado Boulder}
+\affil[2]{Environmental Studies Program, University of Colorado Boulder}
 \affil[3]{Department of Geographical Sciences, University of Maryland College Park}
@@
 However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income or race.
-We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in 250 million geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
+We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in a quarter of a billion geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
 We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.
@@
 
-\flushbottom
+\raggedbottom
 \maketitle
@@
 %Overview
-Climate change is impacting many aspects of human well-being with the impacts projected to worsen in the coming years \cite{pachauri2014climate}.
-A large body of research has assessed the ongoing impacts on climate change on outcomes like agriculture, economic growth, and physical health (NEED CITATIONS).
-Academic and medical researchers have highlighted the relative paucity of empirical studies into climate impacts on mental health at multiple scales and recently called for more \cite{Berry2018Apr, hayes_climate_2018}.
-Researchers have begun to answer these calls, finding associations between deteriorating environmental conditions and a sense of "ecological grief" from their experienced loss \cite{Cunsolo2018Apr}, disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, as well as a variety of mental health impacts associated with warmer temperatures \cite{baylis_weather_2018, Mullins2019Dec} (PALINKAS and WONG 2020).
-MAYBE A SENTENCE HERE HIGHLIGHTING THE GAP IN KNOWLEDGE ABOUT DISPARITIES IN MENTAL HEALTH IMPACTS, VULNERABLE POPULATIONS, issues of scale for signal to noise, etc.
+Climate change is impacting many aspects of human well-being and a large body of research has assessed the ongoing impacts on climate change on outcomes like agriculture, economic growth, and physical health \cite{pachauri2014climate}.
+However, academic and medical researchers have highlighted the relative paucity of empirical studies on the potential impacts of climate change on mental health \cite{Berry2018Apr, hayes_climate_2018}.
+While mental health is under-studied, it makes up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}.
+Researchers have begun to answer the call more more research, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, as well as higher temperatures and a variety of mental health impacts \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}.
+Nevertheless, as global warming may reach 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of heat on mental health, especially in identifying the most vulnerable communities.
 
-%Heat and mental health
-%https://www.pnas.org/content/115/43/10953
-Higher temperatures are associated with many indicators of worsened mental health.
-Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other work demonstrating similar relationships in other locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
-Higher temperatures have also been correlated with increases in hospitalization events related to both general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb, Gupta1992Jun} as well as increased mortality among people with mental health disorders \cite{Hansen2008Oct}.
+%Theoritical: frameworks and pathways
+A number of frameworks have been proposed highlighting the pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr, BerryETAL2010}.
+Work on biological mechanisms emphasizes that there are mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul, sadiq_impact_2019}.
+Additionally, recent studies have found that nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}.
+It has also been noted that injuries and physical health impacts from heat can have consequences for mental health \cite{Berry2007, WHO2007}.
+Finally, higher temperatures affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
 
-Temperature has been hypothesized to impact mental health through a number of pathways.
-Work on biological mechanisms emphasize that there may be mental health effects from maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.
+%Empirical: Heat and mental health
+Building on these frameworks, an emerging literature has shown that higher temperatures are associated with many indicators of worsened mental health.
+Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other work demonstrating similar relationships in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
+Higher temperatures have also been correlated with increases in hospitalization events related to both general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb} as well as increased mortality among people with mental health disorders \cite{Hansen2008Oct}.
 
-Additionally, recent studies have found that nighttime temperatures affects sleep quality \cite{Obradovich2017May, Mullins2019Dec}.
-
-% The pathways
-Higher temperature associated with climate change could influence mental health both directly and indirectly by exposing people to trauma (direct) (CITATIONS) or by affecting (1) physical health and (2)community well-being (indirect) \cite{BerryETAL2010}.
-Direct pathways can include extreme heat events or increasing temperatures associated with loss of workplace productivity from decreases in occupational health or increases in workplace injury \cite{kjellstrom_heat_2016, KjellstromETAL2009, kjellstrom_impact_2016} which can result in loss of income, all of which can exacerbate deteriorating mental health and lead to increased aggression, hospital admissions and suicide rates \cite{BerryETAL2010,CheatwoodDerral1995,CohnETAL2004,BouchamaETAL2007}.
-Heat exposures in working environments could also reduce people's capacity to deal with physical and mental task, and increase the risk of accidents, because of the excessive core body temperature and dehydration \cite{KjellstromETAL2009}.
-The loss of work capacity would in turn result in loss of income, while the low income could also cause mental health problems \cite{Katz1997}.
-As for the indirect pathway, first, because of the reciprocal causation relationship between physical health problems and mental health problems \cite{MillerETAL2009,PrinceETAL2007} heat exposures associated with climate change along with other climate events and indirect health risks threatening physical health will directly influence mental health \cite{Berry2007,WHO2007}.
-Second, disordered temperature associated with climate change may destroy the economic in agricultural-production-dependent communities.
-For example, extreme heat reduces the work capacity of laborers in farm fields \cite{KjellstromETAL2009}, which further destroys agricultural-supported industries in local area\cite{BerryETAL2008}.
-The following economic pressures would undermine social capital and then lead to mental health problems.
-
-Recent research conducted on the effects of higher temperatures on mental health conducted at large, multi-city scales has found surprisingly uniform vulnerability.
+While the effects of heat on mental health are now well established, research conducted at national scales has found surprisingly little heterogeneity in impacts, with consequences for our understanding of vulnerability and adaptation.
 For example, focusing on suicides, Burke et al.
@@
 find no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status \cite{Mullins2019Dec}.
-One major challenge for these studies is that they rely on data from public health agencies that is aggregated to the municipality or city level.
-Thus, they must rely on between-county metrics of vulnerability and cannot take neighborhood effects into account.
+However, one major challenge for these studies is that they rely on data from public health agencies that is aggregated to the municipality or county level.
+As a consequence, they are restricted to between-county metrics of vulnerability and cannot take neighborhood effects into account.
 
-%We would expect the poor to be more vulnerable, but this is not what we find
-The poor and disadvantaged are much more exposed to undesirable temperatures.
-They are more likely to work outside rather than indoors; they are more likely to rely on public transportation, bicycle, or walk instead of commute in their own air-conditioned car; and they are more likely to live in housing that is less insulated with poorer quality air-conditioning.
-Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, and these heterogeneities in vulnerability might be most pronounced at the neighborhood scale.
-Indeed this is what we find for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar}.
+These findings of uniform vulnerability to heat from from national studies are surprising, because the poor and disadvantaged are much more exposed to undesirable temperatures.
+They are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycle, or walk instead of commute in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}.
+Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, and these heterogeneities in vulnerability might be most pronounced within cities rather than between cities.
+Indeed, neighborhoods have a large effect on vulnerability for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar}, as well as for the mental health impacts of other types of climate shocks, such as hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
 
-While public health data on mental health outcomes is not available at the neighborhood scale, data from twitter can provide an important indicator of mental health at extremely precise spatial and temporal scales.
+While public health data on mental health outcomes is not available at the neighborhood scale, data from the social media platform Twitter can provide an important indicator of mental health at extremely precise spatial and temporal scales.
 Many studies have used the mood expressed in tweets as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
-Such studies have used twitter data to identify the onset of Post Traumatic Stress Disorder in individuals, even before their formal diagnosis \cite{Reece2017Oct}, while similar methods with facebook data have been used to predict the onset of depression \cite{Eichstaedt2018Oct}.
-Additionally, in London, day-to-day changes in mental health indicators derived from tweets have been associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}, while across the US more positive tweets are associated with a variety of metrics of human well being \cite{Mitchell2013May}.
-Finally, the mood expressed in tweets has been strongly associated with local weather \cite{baylis_weather_2018}, and has been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
-Thus, we used a common sentiment metric to determine if neighborhood characteristics affect the vulnerability of mental health to higher temperatures.
+Such studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder in individuals, even before their formal diagnosis \cite{Reece2017Oct}, while similar methods with Facebook data have been used to predict the onset of depression \cite{Eichstaedt2018Oct}.
+Additionally, in London, day-to-day changes in mental health indicators derived from tweets have been associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}, while across the US, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
+Finally, the mood expressed in tweets has been strongly associated with local weather \cite{baylis_weather_2018, hannak_tweetin_2012}, and has been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+Thus, Twitter provides an indicator of mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
+We therefore draw on a data set of 243 million geolocated tweets from across the continental United States and covering the period 2009-2019, each paired local weather and neighborhood conditions, to explore how neighborhood conditions affect the relationship between heat and mental heath.
 
@@
 \subsection*{Overall Effect}
-We examined the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an indicator of heat stress that accounts for temperature, humidity, wind speed, and solar radiation.
-Controlling for a variety of fixed effects, we found that, across all tweets, higher temperatures are associated with lower sentiment (See Fig.
+We examine the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
+Controlling for a variety of fixed effects, we find that, across all tweets, higher temperatures are associated with lower sentiment (See Fig.
 \ref{fig:wbgt}).
-Sentiment is highest at a WBGT of 5\textdegree Celsius (a dry bulb temperature typically around 12\textdegree C/54\textdegree F).
-The greatest declines in sentiment are observed between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
+Sentiment is highest at a WBGT of 5\textdegree C (a dry bulb temperature typically around 12\textdegree C/54\textdegree F), and the greatest declines in sentiment are between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
 Sentiment also declines with colder temperatures, but only slightly.
-We ran models using multiple sentiment indicators, and found similar results no matter which sentiment metric was used (See Supplement).
+We run models using multiple sentiment indicators, and find similar results no matter which sentiment metric was used (See Supplement).
 
@@
 
-We then explored how the racial and income characteristics of neighborhoods moderate the relationship between WBGT and sentiment (See Fig \ref{fig:sub1}).
-We found that neighborhood income strongly moderates the relationship between temperature and sentiment.
-As temperatures increase to moderate heat of 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment decreases for most neighborhoods, but increases in neighborhoods in the 95th income percentile.
-These rich do not see decreases in sentiment until temperatures are above 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point sentiment decreases evenly across income ranges, still leaving a large gap in expressed sentiment from low to high-income neighborhoods.
+We then explore how the racial and income characteristics of neighborhoods moderate the relationship between WBGT and sentiment (See Fig \ref{fig:hetero}).
+We find that neighborhood income strongly moderates the relationship between temperature and sentiment.
+As temperatures increase to a moderate heat of 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment decreases for most neighborhoods, but increases in neighborhoods in the 95th income percentile.
+These rich areas do not see decreases in sentiment until temperatures are above 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point sentiment decreases evenly across income ranges, still leaving a large gap in expressed sentiment from low to high-income neighborhoods.
 
-We examined patterns of temperature and sentiment for neighborhoods that had a majority population of four broad racial categories - non-Hispanic black, non-Hispanic white, Hispanic of any race, and other, which includes Native American, multi-racial, and Asian-American and Pacific Islander.
-While people in neighborhoods of all racial compositions were affected by heat, people in majority black neighborhoods were much more affected than others  (See Fig \ref{fig:sub2}).
-Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decreases four-times as much as people in other neighborhoods.
+We also examine patterns of temperature and sentiment for neighborhoods that had a majority population of four broad racial categories - non-Hispanic black, non-Hispanic white, Hispanic of any race, and other, which includes Native American, multi-racial, and Asian-American and Pacific Islander.
+While people in neighborhoods of all racial compositions are affected by heat, people in majority black neighborhoods are much more affected than others  (See Fig \ref{fig:hetero}).
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decrease four times as much as people in other neighborhoods.
 Additionally, at levels of moderate heat from 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
 
-Finally, we modeled the combined effects of race and income in moderating the impacts of heat on mental health, and we found separate effects for these two variables (See Supplement).
-In other words, neither race or income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by than than neighborhoods that are just poor or just black.
+Finally, we model the combined effects of race and income in moderating the impacts of heat on mental health, and we find separate effects for these two variables (See Supplement).
+In other words, neither race nor income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by heat than than neighborhoods that are just poor or just black.
 
@@
 \centering
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-income.png}
-  \caption{}
-  \label{fig:sub1}
-\end{subfigure}%
-\begin{subfigure}{.5\textwidth}
-  \centering
-  \includegraphics[width=\linewidth]{../res/wbgt-race_q.png}
-  \caption{}
-  \label{fig:sub2}
-\end{subfigure}
-\caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income (a) and race (b).}
-\label{fig:test}
+  \includegraphics[width=\linewidth]{../res/wbgt_combined.png}
+    \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income (A) and race (B).}
+\label{fig:hetero}
 \end{figure}
@@
 Expressed sentiment is a widely used metric to assess mental health and well-being on social media.
-While a variety of related metrics can be used to quantify sentiment, we chose to use the VADER metric that was specifically designed for microblogs like twitter \cite{hutto2014vader}.
+While a variety of related metrics can be used to quantify sentiment, we use the VADER metric that was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
 To give context to this novel metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to other events that have large impacts on sentiment (See Fig.
 \ref{fig:compare}).
-Firstly, we examine the change in average sentiment from Saturday to Monday, the highest and lowest days for twitter sentiment as well as other mental health metrics, such as suicides.
+First, we examine the change in average sentiment from Saturday to Monday, the highest and lowest days for Twitter sentiment as well as other mental health metrics, such as suicide rates \cite{CDC2021}.
 We also examine the decrease in sentiment associated with the two most expensive hurricanes in the last decade in the United States.
-For the hurricanes, we compared the mean sentiment of counties affected by Hurricanes Harvey and Sandy on the day of landfall to the mean sentiment a week before.
+For the hurricanes, we compare the mean sentiment of counties affected by Hurricanes Harvey and Sandy on the day of landfall to the mean sentiment exactly one week before.
 These hurricanes are associated with mental health effects including Post-Traumatic Stress Disorder (PTSD) \cite{Schwartz2017Aug, Schwartz2018May}, and similar disasters have been associated with subsequent increases in suicides \cite{Krug1998Feb}.
@@
 
-We found that the impacts of heatwaves on sentiment in rich and majority white neighborhoods were less than the average weekly change in sentiment from Saturday to Monday, but the effects of heat waves on majority black or poor neighborhoods were much greater than the average weekly change in sentiment.
-Additionally, the impacts of heat waves more vulnerable neighborhoods were close in degree to the impacts of both hurricanes.
-
-{\color{red} What other disasters/comparison events should we look at?}
+We find that the impacts of heatwaves on sentiment in rich and majority white neighborhoods are less than the average weekly change in sentiment from Saturday to Monday, but the effects of heat waves on majority black or poor neighborhoods are much greater than the average weekly change in sentiment.
+Additionally, the impacts of heat waves on more vulnerable neighborhoods are close in severity to the impact of a major hurricane.
 
@@
 
-To further explore our findings, we examined inequalities in the impacts of higher temperatures on sentiment by Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) with over 1 million people (hereafter: cities).
-We examined the change in size of the gap in sentiment scores between both poor and rich neighborhoods, as well as black and non-black neighborhoods, as temperatures change from optimum temperatures of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 30\textdegree C WBGT (42\textdegree C/108\textdegree F).
+To further explore our findings, we examine the impact of heat on sentiment and sentiment inequalities by Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) with over 1 million people (hereafter: cities).
+We examine the impacts of heat alone, as well as the change in size of the gap in sentiment scores between both poor and rich neighborhoods and black and non-black neighborhoods.
+For our metric of inequality, we show the size of the gap in sentiment as temperatures increase from optimum temperatures of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 30\textdegree C WBGT (42\textdegree C/108\textdegree F).
 
-We found increasing inequalities in sentiment across income groups for 68.8\% of cities, with 35.6\% of these having a statistically significant effect (15 cities).
+Examining the effects of heat on sentiment by city, we find that heat is associated with worsened sentiment in 83.6\% of cities, and the effect is significant in 35.3\% of those.
+In no cities does heat significantly increase sentiment, and heat affects sentiment in cities across the US, from Florida to New England to California.
+
+We find that heat increases inequalities in sentiment across income groups for 68.9\% of cities, with 35.7\% of these having a statistically significant effect.
 Cities with a significantly unequal effect were found in the south, Midwest, northeast, northwest, and southwest, although they were most common in the mid-Atlantic region.
@@
 
-Because we found a much stronger effect for the effects of temperature on sentiment in majority black neighborhoods, we also examined increases in sentiment gaps between black and non-black neighborhoods in cities with a black population greater than 5\% of the total population.
+Because we found a much stronger effect of temperature on sentiment in majority black neighborhoods compared to other ethnic groups, we also examined increases in sentiment gaps between neighborhoods in the 95th percentile for black population (77.2\% black) and neighborhoods in the 5th percentile for black population (0\% black).
+We excluded cities with a black population less than 5\% of the total population.
 We found increasing inequalities in sentiment for more black neighborhoods in 64.7\% of cities, with a significant effect in 21.2\% of those (7 cities).
 Additionally, three sunbelt cities had a significant inverted effect.
-The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: cities with large and significant inequalities were found the mid-Atlantic and Midwest, as well as central Florida.
+The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: cities with large and significant inequalities were found the mid-Atlantic and Midwest.
 
@@
 \centering
-\begin{subfigure}{0.75\textwidth}
-  \includegraphics[width=\linewidth]{../res/map_wbgt_income.png}
+  \includegraphics[width=0.5\linewidth]{../res/map_combined.png}
   \caption{}
-  \label{fig:map1}
-\end{subfigure}
-\begin{subfigure}{0.75\textwidth}
-  \includegraphics[width=\linewidth]{../res/map_wbgt_black.png}
-  \caption{}
-  \label{fig:map2}
-\end{subfigure}
-\caption{Inequalities in impacts of temperature on sentiment by CSAs and MSAs with over 1 million people.
-The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5WBGT to 30WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
-CSAs with a statistically significant effect ($p < 0.05$) are outlined in black.
-For racial inequalities (b), only MSAs and CSAs where black people make up $> 5\%$ of the population are included.}
-\label{fig:test}
+  \label{fig:map}
+\caption{Figure A shows the change in sentiment for a 25-degree increase in WBGT for cities with a population of over 1 million.
+Figures B and C show inequalities in impacts of temperature on sentiment for cities with over 1 million people.
+The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5\textdegree C WBGT to 30\textdegree C WBGT between the 5th and 95 percentile for income (A) and the 95th and 5th percentile for percent black (B).
+Cities with a statistically significant effect ($p < 0.05$) are outlined in black.
+For racial inequalities (B), only cites where black people make up $> 5\%$ of the population are included.}
 \end{figure}
@@
 We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between heat and worsened sentiment.
-Moreover, we were able to show clear heterogeneities in vulnerability by neighborhood characteristics, overturning previous findings that found no heterogeneities at the county level \cite{Burke2018Aug, Mullins2019Dec}.
+Moreover, we were able to show heterogeneities in vulnerability by neighborhood characteristics, overturning previous findings that found no heterogeneities at the county level \cite{Burke2018Aug, Mullins2019Dec}.
 
 % Sentiment is a fuzzy indicator but we have BIG DATA
-However, one limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
+One limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
 Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large data set of nearly 250 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
+We overcome this limitation by using an extremely large data set of 243 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
 
@@
 Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.
-Moreover, both suicides and sentiment are affected by higher temperatures \cite{baylis_weather_2018, Burke2018Aug}.
+Moreover, like sentiment, suicides are affected by higher temperatures.
 Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.
@@
 Previous work conducted with county-scale data found no heterogeneities in mental health vulnerability by income.
-Such analyses have led to conclusions that (1) the mental health effects of climate change will be uniform, least in developed countries, and (2) that adaptation with technologies like air-conditioning will not be enough.
+Such analyses have led to conclusions that (1) the mental health effects of climate change will be uniform, least in developed countries, and (2) that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
 For example, Mullins et al.
@@
 Contrary to these analyses, we found large heterogeneities in vulnerability by neighborhood race and income.
-This suggests that the mental health effects of climate change will not be uniform and, like other impact, will fall disproportionately on the poor and vulnerable.
+This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the poor and vulnerable.
 More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar those of high-income Americans, the effects of high temperatures on mental health can be largely mitigated.
-
-% Comment on interpretation of results
-We examined our results relative to an optimal baseline temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as this was the temperature at which the highest sentiment scores were observed in the aggregated analysis (See Fig.
-\ref{fig:wbgt}).
-However, after examining heterogeneities by neighborhood race and income, we find different optima for different racial groups and income levels.
-For example, the optimum temperature for poor neighborhoods is 0\textdegree C WBGT (6\textdegree C/43\textdegree F) while the optimum temperature for rich neighborhoods is 20\textdegree C WBGT (29\textdegree C/84\textdegree F).
-These different varying optima show that even mild heat can affect human well-being, and the only reason that sentiment peaks as high as 20\textdegree C WBGT (29\textdegree C/84\textdegree F) in rich neighborhoods is because the rich are more likely to have air conditioning in their homes, transportation, and work spaces, as well as much greater choice in when they are outside and in the activities they perform outside.
 
@@
 This is surprising, given that Hispanics can be marginalized in both income and housing.
-This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans, or may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in Spanish, which we excluded from our analysis.
+This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans, or may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
 Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
+Finally, while race is correlated with income in the US, we found that race contributed to vulnerability independently of income.
+This suggests that racism and patterns of housing discrimination such as a legacy of redlining contribute to making black Americans more vulnerable to heat, event at similar levels of income.
 
 % Interpret Spatial Analysis
-In our spatial analysis, we found that heat affects sentiment in cities across the US, suggesting that higher baselines temperatures do not reduce vulnerability.
-Additionally, most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, although the south-west had the weakest inequalities and an occasional inverted effect.
-While this could suggest that heat inequality is related humidity, this is unlikely because we used Wet Bulb Globe Temperature, which accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
+In our spatial analysis, we found that heat affects sentiment in cities across the US, suggesting that higher baseline temperatures do not reduce vulnerability.
+Additionally, most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, although the southwest had the weakest inequalities and an occasional inverted effect.
+While unusual patterns of heat impacts in the southwest could suggest that heat inequality is related to environmental factors such as humidity, this is unlikely because we used Wet Bulb Globe Temperature, which accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
 Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different baseline environmental conditions.
 
-% Interpret Time Series Analysis
-Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis was able to examine this hypothesis.
-While our estimates were more uncertain at night due to a lower volume of tweets, we found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
-This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
+% Connect to bigger literature on vulnerability
+This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts, as well as the importance of measuring vulnerability at fine spatial scales.
+Within in relatively wealthy nation of the United States, we find large heterogeneities in vulnerability.
+This adds to the growing literature showing that the impacts of climate change will be highly unequal and disproportionately borne by the poor and marginalized \cite{Thomas2019Mar}.
+Moreover, given that more impoverished neighborhoods in the United States see greater mental health impacts during heatwaves, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much greater \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
 
@@
 Many low-income and black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
-Moreover, the income level of a census block may be only weakly correlated with the wealth of the people who are typically found that census block.
+Moreover, in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are typically found that census block.
 For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
 Additionally, neighborhoods of young college students are also typically estimated to have low income levels, even through college students are wealthier than the average American.
-Again, these issues mean we are likely under-estimating the true effect of heat on mental health.
+Again, these issues mean we are likely under-estimating the true effect of heat on mental health for poor and black people.
 A final issue with using twitter data is that, while twitter is used by more than one in five Americans, twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
@@
 
-% NCC doesnt have a "conclusion" section*, but this is intended as an overall summary paragraph
+% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
 While climate change will have widespread and severe impacts on human well-being, it cannot be overemphasized just how highly unevenly these impacts will be distributed.
 People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \cite{bullard2012wrong}.
-While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they are not also highly unevenly distributed.
-Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales has not found such an effect.
-By using fine-scale twitter data, we showed that there are indeed stark difference in the mental health impacts of heat among neighborhoods in the United States.
+While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they will not also be highly unevenly distributed.
+Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect.
+By using fine-scale twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States.
 These findings have significant implications for urban planning, climate and environmental justice, and mental health.
@@
 \subsection*{Tweets \& Sentiment}
-We used data from 243 million geo-located English-language tweets from the lower 48 states from the years 2009 to 2019.
+We used data from 243 million geo-located English-language tweets from the continental United States from the years 2009 to 2019.
 The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
@@
   \item Punctuation (``Good!!!")
-  \item Word shape, such as capitalization (``The move was AMAZING")
+  \item Word shape, such as capitalization (``The movie was AMAZING")
   \item Emoticons and emojis (``:-)", ``\emojismile")
@@
 
-The VADER method yields a value for the mood of a tweet, with a score of 0 for neurtral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
-In addition to the VADER method used in the body of this paper, we also conduct our analysis using the Hedonometer and AFINN sentiment analysis methods, with similar results (See Supplement).
+The VADER method yields a value for the mood of a tweet, with a score of 0 for neutral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
+In addition to the VADER method used in the body of this paper, we also conduct our analysis using the Hedonometer and XXXXX sentiment analysis methods, with similar results (See Supplement).
 
@@
 
-Because metrics of apparent temperature that take into account humidity and other factors can better account for the impacts of heat stress on human health and wellbeing, we calculate the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet.
+Because metrics of apparent temperature that take into account humidity and other factors can better account for the impacts of heat stress on human health and well-being, we calculate the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet.
 WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature would show due to evaporative cooling, and can be estimated given normal temperature, relative humidity, solar radiation, and wind speed.
-Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing.
+Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing \cite{budd2008wet}.
 Metrics like WBGT that account for the effects of humidity and other factors on heat stress have been associated with diminished economic output \cite{rao2020projections}, increased crime \cite{hu2017impact}, increased mortality \cite{chien2016spatiotemporal, armstrong2019role}, and worsened mental health outcomes \cite{vida2012relationship, ding2016importance}.
@@
 
+Throughout the paper, we give the dry bulb temp for reference as the mean value across our data set for a given wet bulb temp.
+
 \subsection*{Socio-Econoimc Data}
 We used data from the American Community Survey (ACS) administered by the US Census to estimate income levels and the racial composition of neighborhoods where tweets were located.
-Data was at the level of the census block group, the smallest unit for which the census releases public data.
+Data was at the level of the census block group, the smallest unit for which the US Census releases public data.
 
@@
 
+In this paper, we give income and racial percentiles based on the observed percentiles across all tweets in our data set.
+
 \subsection*{Modeling}
-We assessed how expressed sentiment is affected by wet bulb globe temperature using segmented regressions, controlling for precipitation, and shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: the day of the week, the time of day, the day of the year, the year, the month, and the county.
+We assessed how expressed sentiment is affected by Wet Bulb Globe Temperature using segmented regressions, controlling for precipitation, and shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: the day of the week, the time of day, the day of the year, the year, the month, and the county.
 
-Our initial model (for Fig. \ref{fig:wbgt} takes the following form:
+Our initial model (for Fig. \ref{fig:wbgt}) takes the following form:
 
@@
     y = \beta_0 + f_t(t) + \beta_p p + \beta_s s + \Phi + \epsilon
+    \label{mod:1}
 \end{equation}
@@
     y = \beta_0 + f_t(t) + f_{mt}(m t) + \beta_p p + \beta_{mp} m p + \beta_s s + \beta_{ms} m s + \Phi + \epsilon
+    \label{mod:2}
 \end{equation}
@@
 Where $m$ is either the log-transformed average income in the census block where the tweet originated, or a dummy variable for the four racial categories.
-We specify alternative models where income is a dummy variable in three bins, and were race is a continuous variable for the percent of a census block population that is what.
-These alternative specifications yield similar results.
+We specify alternative models where income is a dummy variable in three bins, and were race is a continuous variable for the percent of a census block population that is white These alternative specifications yield similar results.
 We also examine the effects of rainfall and sunshine on sentiment by income and race.
@@
 
-For our analyses by CSA, we subset the data in the model and run the model with the same fixed effects, although we fit a linear effect for $f_t()$.
+For our analyses by city, we run models with a similar form to Model \ref{mod:2}, including the same fixed effects, although we fit a linear effect for $f_t()$.
+Additionally, for exploring the vulnerability of black americans, we use a continuous variable for $m$ indicating the percentage of a neighborhoods population that is black.
+Finally, because we are not accounting for non-linearities in $f_t()$, we exclude tweets with a WBGT of less than 5\textdegree C, as our initial analysis showed that this is the temperature at which sentiment peaks.
 
@@
 This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
-Additionally, cloud resources for this project were provided for the Microsoft Azure cloud by Microsoft's AI for Earth program, grant number .
+Additionally, computational resources for this project were provided for the Microsoft Azure cloud by Microsoft's AI for Earth program.
 
@@
 \section*{Author Contributions}
-Author contributions: M.W.C., J.O.G.
+Author contributions: M.W.C., J.O.-G., A.S.
 and P.B.
@@
 prepared data; M.W.C.
-and J.O.G analyzed data; and M.W.C., J.O.G., J.L., and P.A.W.
+and J.O.-G.
+analyzed data; and M.W.C., J.O.-G., J.L., and P.A.W.
 wrote the paper.
 
-\noindent This project was co-lead by M.W.C. and J.O.G.
+\noindent This project was co-lead by M.W.C.
+and J.O.-G.
+and both contributed equally to this project.
+Authorship order for these two is listed alphabetically.
+Authorship order for others listed by proportion of contribution.
 

### 2021-05-18 viz/comparison_plot.R
improve comparison plot

@@ -1,151 +1,120 @@
 library(data.table)
 library(tidyverse)
 library(lubridate)
-library(grid)
-library(png)
 
 setwd('~/tweets/')
 
 data <- fread('all.csv')
 data <- data[weather_term == 0, ]
-
-# Day of Week
-dow_df <- data[ , .(vader = mean(vader), afinn=mean(afinn), hedono=mean(hedono)), .(dow)]
-
-doy_df$doy[doy_df$vader == max(doy_df$vader)]
-doy_df$doy[doy_df$vader == min(doy_df$vader[doy_df$doy != '02-29'])]
-max(doy_df$vader) - min(doy_df$vader[doy_df$doy != '02-29'])
-
 data$date <- mdy(paste0(data$doy, '-', data$year))
 
-data$month <- month(data$date)
+boostrap_diff <- function(x, y, n=100, q=0.95, gc=FALSE){
+  diffs <- c()
+  for (i in 1:n){
+    print(i/n)
+    diffs <- c(mean(sample(x, length(x), replace=T)) - mean(sample(y, length(y), replace=T)), 
+               diffs)
+    if(gc){
+      gc()
+    }
+  }
 
-mon_df <- data[ , .(vader = mean(vader), afinn=mean(afinn), hedono=mean(hedono)), .(month)]
+  min = quantile(diffs, probs=(1 - q)/2)
+  max = quantile(diffs, probs=(1 - (1 - q)/2))
+  med = quantile(diffs, probs=0.5)
 
-make_dif <- function(counties, event_date){
-  sel <- data %>%
-    filter(fips %in% counties, date == ymd(event_date) | date == ymd(event_date) - days(7)) %>%
-    mutate(bef_aft=ifelse(date >= ymd(event_date), 'after', 'before')) %>%
-    group_by(bef_aft) %>%
-    summarize(vader=mean(vader))
-  
-  change <- sel$vader[sel$bef_aft == 'before'] - sel$vader[sel$bef_aft == 'after']
-  
-  change
+  return(list(min=min, max=max, med=med))
 }
 
+######################################
+# Day of Week difference
+######################################
+mon <- data$vader[data$dow == 'Monday']
+sat <- data$vader[data$dow == 'Saturday']
 
-#####################
-# 
-#####################
+dow_diff <- boostrap_diff(mon, sat)
 
-#Harvey
-harvey_counties <- c(48007, 48025, 48039, 48057, 48071, 48157, 48167,
[…]

### 2021-05-18 doc/old_discussion.tex

% Random discussion paragraphs that were cut

% Comment on interpretation of results
We examined our results relative to an optimal baseline temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as this was the temperature at which the highest sentiment scores were observed in the aggregated analysis (See Fig. \ref{fig:wbgt}).  However, after examining heterogeneities by neighborhood race and income, we find different optima for different racial groups and income levels.  For example, the optimum temperature for poor neighborhoods is 0\textdegree C WBGT (6\textdegree C/43\textdegree F) while the optimum temperature for rich neighborhoods is 20\textdegree C WBGT (29\textdegree C/84\textdegree F).  These different varying optima show that even mild heat can affect human well-being, and the only reason that sentiment peaks as high as 20\textdegree C WBGT (29\textdegree C/84\textdegree F) in rich neighborhoods is because the rich are more likely to have air conditioning in their homes, transportation, and work spaces, as well as much greater choice in when they are outside and in the activities they perform outside.

### 2021-05-18 viz/comparison_plot.R
comparison events with errorbars

@@ -107,8 +107,7 @@ annot <- data.frame(color=c(rep('Majority\nBlack', 2),
                         rep('Hurricane\nSandy', 2)),
                     x=c(4.5, 4.5, 
                         1.5, 1.5),
-                    y=c(0.0003, 0.000015, 
-                        0.0003, 0.000015),
+                    y=c(0.0031, 0.00015, 0.0031, 0.00015) + 0.008,
                     label=c('Heat Wave Impact By', 'Neighborhood Characteristics',
                             'Impact of', 'Comparison Events'))
 
@@ -124,20 +123,19 @@ ggplot() +
                               'Hurricane\nSandy'="#636363", 
                               'Saturday\nto Monday'="#636363")) + 
   scale_color_manual(values=c('Majority\nBlack'='black',
-                              'Hurricane\nHarvey'='grey30')) + 
+                              'Hurricane\nSandy'='grey30')) + 
   scale_y_continuous(expand=expand_scale(mult=c(0.05, 0))) + 
   scale_x_discrete(position='top') + 
   labs(x='', y='Decrease in Sentiment') + 
   guides(fill=FALSE, color=FALSE) + 
   theme_classic() + 
   coord_cartesian(ylim=c(min(plt$min)*1.05, 0), clip='off') + 
-  annotate("segment", x = 2.75, xend = 6.25, y = 0.0006, yend = 0.0006,
+  annotate("segment", x = 2.75, xend = 6.25, y = 0.006, yend = 0.006,
     colour = "black", size=0.25) +
-  annotate("segment", x = 0.75, xend = 2.25, y = 0.0006, yend = 0.0006,
+  annotate("segment", x = 0.75, xend = 2.25, y = 0.006, yend = 0.006,
     colour = "grey30", size=0.25) +
   theme(axis.text.x = element_text(colour = c(rep('grey30',3), rep('black', 4))),
         #top, right, bottom, left
-        plot.margin = unit(c(5, 0.025, 0.25, 0.025), "cm"))
-
+        plot.margin = unit(c(0.65, 0.025, 0.25, 0.025), "cm"))
 ggsave('~/temp-sentiment/res/comparison_plot.png', width=5, height=2.5)
 

### 2021-05-18 viz/comparison_plot.R
fix off color

@@ -134,7 +134,7 @@ ggplot() +
     colour = "black", size=0.25) +
   annotate("segment", x = 0.75, xend = 2.25, y = 0.006, yend = 0.006,
     colour = "grey30", size=0.25) +
-  theme(axis.text.x = element_text(colour = c(rep('grey30',3), rep('black', 4))),
+  theme(axis.text.x = element_text(colour = c(rep('grey30', 2), rep('black', 4))),
         #top, right, bottom, left
         plot.margin = unit(c(0.65, 0.025, 0.25, 0.025), "cm"))
 ggsave('~/temp-sentiment/res/comparison_plot.png', width=5, height=2.5)

### 2021-05-18 doc/temp-sentiment.tex

@@
   \includegraphics[width=\linewidth]{../res/wbgt_combined.png}
-    \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income (A) and race (B).}
+    \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income (a) and race (b).}
 \label{fig:hetero}
@@
   \includegraphics[width=0.5\linewidth]{../res/map_combined.png}
-  \caption{}
   \label{fig:map}
-\caption{Figure A shows the change in sentiment for a 25-degree increase in WBGT for cities with a population of over 1 million.
-Figures B and C show inequalities in impacts of temperature on sentiment for cities with over 1 million people.
-The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5\textdegree C WBGT to 30\textdegree C WBGT between the 5th and 95 percentile for income (A) and the 95th and 5th percentile for percent black (B).
+    \caption{Figure (a) shows the change in sentiment for a 25-degree increase in WBGT for cities with a population of over 1 million.
+Figures (b) and (c) show inequalities in impacts of temperature on sentiment for cities with over 1 million people.
+The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5\textdegree C WBGT to 30\textdegree C WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
 Cities with a statistically significant effect ($p < 0.05$) are outlined in black.
-For racial inequalities (B), only cites where black people make up $> 5\%$ of the population are included.}
+For racial inequalities (b), only cites where black people make up $> 5\%$ of the population are included.}
 \end{figure}

### 2021-05-18 viz/vader/wbgt_income_bootstrap.R
fix second x axis

@@ -12,9 +12,9 @@ setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
 data <- fread('~/tweets/all.csv')
 data$income_percap <- log(data$income_percap)
 
-qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
-
 data <- data[weather_term == 0, ]
+
+qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
 
 data$raining <- data$prcp > 0
 
@@ -173,7 +173,7 @@ x <- get_x_axis(x2)
 xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
-          rel_heights=c(0.8, 0.2, 0.04, 0.04))
+          rel_heights=c(0.8, 0.2, 0.04, 0.08))
 ggsave('~/temp-sentiment/res/wbgt-income.png', width=4.5, height=4)
 
 ########################################

### 2021-05-18 viz/vader/wbgt_nointeract_bootstrap.R

@@ -152,6 +152,6 @@ x <- get_x_axis(x2)
 xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
-          rel_heights=c(0.8, 0.2, 0.04, 0.04))
+          rel_heights=c(0.8, 0.2, 0.04, 0.08))
 ggsave('~/temp-sentiment/res/wbgt.png', width=4.5, height=4)
 

### 2021-05-18 viz/vader/wbgt_race_q_bootstrap.R

@@ -169,7 +169,7 @@ x <- get_x_axis(x2)
 xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
-          rel_heights=c(0.8, 0.2, 0.04, 0.04))
+          rel_heights=c(0.8, 0.2, 0.04, 0.08))
 ggsave('~/temp-sentiment/res/wbgt-race_q.png', width=4.5, height=4)
 
 ########################################

### 2021-05-18 csa/model_space_bootstrap.R
add map bootstrap script; run later

@@ -1,0 +1,187 @@
+library(data.table)
+library(tidyverse)
+library(fixest)
+library(broom)
+library(sf)
+library(USAboundaries)
+
+data <- fread('~/tweets/all.csv')
+csadat <- fread('~/tweets/csa/fips-csa.csv')
+
+data <- merge(data, csadat, on='fips', all=F)
+
+data$month <- substr(data$statemonth, 4, 5)
+data$income_percap <- log(data$income_percap)
+data <- data[weather_term == 0, ]
+data$raining <- data$prcp > 0
+
+qs_inc <- quantile(data$income_percap, seq(0, 1, by=0.05))
+qs_white <- quantile(data$race_white, seq(0, 1, by=0.05))
+qs_black <- quantile(data$race_black, seq(0, 1, by=0.05))
+
+csas <- tab(data$id)
+csas <- as.numeric(names(csas))[order(csas, decreasing=T)]
+
+#############################################
+# Now effects of temp*income and temp*race
+#############################################
+
+allres <- data.frame()
+for (csasel in csas){
+  print(which(csas == csasel)/length(csas))
+
+  sel <- data[data$id == csasel & data$wbgt > 5, ]
+
+  ####################
+  #### Baseline
+  ########################
+  all_base <- c()
+  for (i in 1:80){
+    print(i)
+    mod <- feols(vader ~ wbgt + raining + srad | dow + doy + tod + fips + year + month, 
+                 data=sel[sample(1:nrow(sel), nrow(sel), replace=T), ])
+    base_res <- tidy(mod) %>%
+      filter(term =='wbgt')
+    all_base <- c(base_res$estimate, all_base)
+  }
+  base_res <- data.frame(min = quantile(all_base, probs=0.025),
+                         max = quantile(all_base, probs=0.975),
+                         med = quantile(all_base, probs=0.5)) %>%
+    mutate(term = 'wbgt',
+           id = csasel)
+
+  ####################
+  ### Income
+  ####################
+  all_inc <- c()
+  for (i in 1:80){
+    print(i)
+    mod <- feols(vader ~ wbgt*income_percap + raining*income_percap + srad*income_percap | dow + doy + tod + fips + year + month, 
+                 data=sel[sample(1:nrow(sel), nrow(sel), replace=T), ])
+    inc_res <- tidy(mod) %>%
[…]

### 2021-05-19 doc/temp-sentiment.tex
Updates from Overleaf

@@
 We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in a quarter of a billion geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
-We find that increased temperatures worsen expressed sentiment in all areas, but that this effects is much stronger in poor and black neighborhoods.
+We find that increased temperatures worsen expressed sentiment in all areas, but that this effect is much stronger in poor and black neighborhoods.
 \end{abstract}
@@
 %Overview
-Climate change is impacting many aspects of human well-being and a large body of research has assessed the ongoing impacts on climate change on outcomes like agriculture, economic growth, and physical health \cite{pachauri2014climate}.
-However, academic and medical researchers have highlighted the relative paucity of empirical studies on the potential impacts of climate change on mental health \cite{Berry2018Apr, hayes_climate_2018}.
-While mental health is under-studied, it makes up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}.
-Researchers have begun to answer the call more more research, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, as well as higher temperatures and a variety of mental health impacts \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}.
-Nevertheless, as global warming may reach 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of heat on mental health, especially in identifying the most vulnerable communities.
+Climate change is impacting many aspects of human well-being.
+A large body of research has explored the ongoing impacts of climate change on agricultural outcomes, economic growth, and physical health \cite{pachauri2014climate}.
+However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr, hayes_climate_2018}.
+While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \cite{Berry2018Apr, Collins2011Jul}.
+Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, and linkages between higher temperatures and a variety of mental health impacts \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}.
+Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of heat on mental health, especially on identifying the most vulnerable communities.
 
-%Theoritical: frameworks and pathways
-A number of frameworks have been proposed highlighting the pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr, BerryETAL2010}.
-Work on biological mechanisms emphasizes that there are mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul, sadiq_impact_2019}.
-Additionally, recent studies have found that nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}.
-It has also been noted that injuries and physical health impacts from heat can have consequences for mental health \cite{Berry2007, WHO2007}.
-Finally, higher temperatures affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
+%Theoretical: frameworks and pathways
+Researchers have proposed a number of frameworks that focus on pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr, BerryETAL2010}.
+Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul, sadiq_impact_2019}.
+Additionally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}.
+Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can have compounding consequences for exacerbate mental health issues \cite{Berry2007, WHO2007}.
+Finally, exposure to higher temperatures can affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health impacts and outcomes \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
 
 %Empirical: Heat and mental health
-Building on these frameworks, an emerging literature has shown that higher temperatures are associated with many indicators of worsened mental health.
-Multiple studies have found strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other work demonstrating similar relationships in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
-Higher temperatures have also been correlated with increases in hospitalization events related to both general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb} as well as increased mortality among people with mental health disorders \cite{Hansen2008Oct}.
+Results from an emerging literature indicate higher temperatures are also associated with indicators of deteriorating mental health.
+Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May}.
+Other studies find similar relationships in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
+Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb}.
+Higher temperatures are also linked to increased mortality among people with mental health disorders \cite{Hansen2008Oct}.
 
-While the effects of heat on mental health are now well established, research conducted at national scales has found surprisingly little heterogeneity in impacts, with consequences for our understanding of vulnerability and adaptation.
-For example, focusing on suicides, Burke et al.
-found ``no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
-Similarly, Mullins et al.
-find no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status \cite{Mullins2019Dec}.
-However, one major challenge for these studies is that they rely on data from public health agencies that is aggregated to the municipality or county level.
-As a consequence, they are restricted to between-county metrics of vulnerability and cannot take neighborhood effects into account.
+The effects of heat exposure on mental health are relatively well established, yet results from research conducted at national scales indicate surprisingly little heterogeneity in impacts, with consequences for our understanding of who may be most or least vulnerable to heat stress and ways we plan for and adapt to changing environmental conditions.
+For example, a large study of temperature and suicide in the USA and Mexico found ``no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
+Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \cite{Mullins2019Dec}.
+A major challenge for these studies, however, is their reliance on data from public health agencies, data which is typically aggregated to the municipality or county level.
+As a consequence, these studies are often restricted to between-county metrics of vulnerability and cannot take into account neighborhood effects.
 
-These findings of uniform vulnerability to heat from from national studies are surprising, because the poor and disadvantaged are much more exposed to undesirable temperatures.
-They are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycle, or walk instead of commute in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}.
-Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, and these heterogeneities in vulnerability might be most pronounced within cities rather than between cities.
-Indeed, neighborhoods have a large effect on vulnerability for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar}, as well as for the mental health impacts of other types of climate shocks, such as hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+Findings from large-scale studies of relatively uniform vulnerability to heat are surprising given minority groups and poorer people are more likely to be exposed to undesirable temperatures and otherwise less able to mitigate the effects of higher heat.
+For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality or no air-conditioning \cite{Samuelson2020Jun}.
+Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, with these heterogeneities in vulnerability being more pronounced within cities rather than between cities.
+Indeed, other studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar} and mental health for other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
 
-While public health data on mental health outcomes is not available at the neighborhood scale, data from the social media platform Twitter can provide an important indicator of mental health at extremely precise spatial and temporal scales.
-Many studies have used the mood expressed in tweets as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
-Such studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder in individuals, even before their formal diagnosis \cite{Reece2017Oct}, while similar methods with Facebook data have been used to predict the onset of depression \cite{Eichstaedt2018Oct}.
-Additionally, in London, day-to-day changes in mental health indicators derived from tweets have been associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}, while across the US, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
-Finally, the mood expressed in tweets has been strongly associated with local weather \cite{baylis_weather_2018, hannak_tweetin_2012}, and has been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
-Thus, Twitter provides an indicator of mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
+While public health data on mental health outcomes is not available at the neighborhood scale, spontaneous, in situ data from the social media platform Twitter can provide an important, inexpensive and timely indicator of mental health at extremely precise spatial and temporal scales.
+Many studies use the mood expressed in social media as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
+For example, studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder (PTSD) in individuals even before their formal diagnosis \cite{Reece2017Oct}.
+Similar methods applied to posts on Facebook can predict the onset of depression \cite{Eichstaedt2018Oct}.
+In London, day-to-day changes in mental health indicators derived from the text of Tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
+Across the USA, more positive Tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
+Finally, the mood expressed in Tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+Thus, posts on Twitter provide an indicator of mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
 We therefore draw on a data set of 243 million geolocated tweets from across the continental United States and covering the period 2009-2019, each paired local weather and neighborhood conditions, to explore how neighborhood conditions affect the relationship between heat and mental heath.
@@
 \subsection*{Overall Effect}
-We examine the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), an temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
-Controlling for a variety of fixed effects, we find that, across all tweets, higher temperatures are associated with lower sentiment (See Fig.
+We examine the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
+Controlling for a variety of fixed effects across all Tweet, we find higher temperatures are associated with lower sentiment (i.e., worsening mood; See Fig.
 \ref{fig:wbgt}).
-Sentiment is highest at a WBGT of 5\textdegree C (a dry bulb temperature typically around 12\textdegree C/54\textdegree F), and the greatest declines in sentiment are between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
-Sentiment also declines with colder temperatures, but only slightly.
-We run models using multiple sentiment indicators, and find similar results no matter which sentiment metric was used (See Supplement).
+Sentiment is highest at 5\textdegree C WBGT (a dry bulb temperature typically around 12\textdegree C/54\textdegree F) with the largest declines in sentiment between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
+Sentiment also declines with colder temperatures (below 5\textdegree C WBGT), although only slightly.
+We ran models using multiple measures of sentiment and find similar results and patterns across all sentiment measures (see Supplement).
 
@@
   \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
-  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment in tweets.  As temperatures increase, sentiment declines.}
+  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment from text in Tweets.  As temperatures increase above 5\textdegree C WBGT, sentiment rapidly declines.}
   \label{fig:wbgt}
@@
 
-We then explore how the racial and income characteristics of neighborhoods moderate the relationship between WBGT and sentiment (See Fig \ref{fig:hetero}).
-We find that neighborhood income strongly moderates the relationship between temperature and sentiment.
-As temperatures increase to a moderate heat of 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment decreases for most neighborhoods, but increases in neighborhoods in the 95th income percentile.
-These rich areas do not see decreases in sentiment until temperatures are above 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point sentiment decreases evenly across income ranges, still leaving a large gap in expressed sentiment from low to high-income neighborhoods.
+We find median neighborhood income strongly moderates the relationship between temperature and sentiment (see Fig \ref{fig:hetero}, panel A), with large differences in sentiment between the poorest and wealthiest neighborhoods.
+As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment increases in the wealthiest neighborhoods (95th income percentile) but decreases in the median and lowest income neighborhoods.
+The wealthiest neighborhoods do not see decreases in sentiment until temperatures exceed 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point sentiment decreases relatively evenly regardless of neighborhood income percentile.
 
-We also examine patterns of temperature and sentiment for neighborhoods that had a majority population of four broad racial categories - non-Hispanic black, non-Hispanic white, Hispanic of any race, and other, which includes Native American, multi-racial, and Asian-American and Pacific Islander.
-While people in neighborhoods of all racial compositions are affected by heat, people in majority black neighborhoods are much more affected than others  (See Fig \ref{fig:hetero}).
-Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decrease four times as much as people in other neighborhoods.
-Additionally, at levels of moderate heat from 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap closes with high temperatures.
+We explored the relationship between neighborhood racial group characteristics and temperature and sentiment and find the effects of heat are felt disproportionately in neighborhoods that are majority black (see Fig \ref{fig:hetero}).
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decrease four times as much as the sentiment of people in other neighborhoods.
+Additionally, at modest to moderate temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
 
-Finally, we model the combined effects of race and income in moderating the impacts of heat on mental health, and we find separate effects for these two variables (See Supplement).
-In other words, neither race nor income alone account for all the heterogeneity in vulnerability, and neighborhoods that are poor and black are more affected by heat than than neighborhoods that are just poor or just black.
+Exploring the combined effects of neighborhood percent minority race and income in moderating the impacts of heat on mental health, we find separate effects for these two variables (see Supplement).
+In other words, neither neither percent minority population nor income alone account for the heterogeneity in vulnerability as neighborhoods that are both poorer and higher percent black are more affected by heat than than neighborhoods that are just poorer or higher proportion black.
 
@@
 
-\subsection*{Comparison With Other Events}
-
+\subsection*{Sentiment Comparison With Other Events}
 Expressed sentiment is a widely used metric to assess mental health and well-being on social media.
 While a variety of related metrics can be used to quantify sentiment, we use the VADER metric that was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
-To give context to this novel metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to other events that have large impacts on sentiment (See Fig.
+To give context to this novel metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to two other changes that have large impacts on sentiment (See Fig.
 \ref{fig:compare}).
 First, we examine the change in average sentiment from Saturday to Monday, the highest and lowest days for Twitter sentiment as well as other mental health metrics, such as suicide rates \cite{CDC2021}.
-We also examine the decrease in sentiment associated with the two most expensive hurricanes in the last decade in the United States.
-For the hurricanes, we compare the mean sentiment of counties affected by Hurricanes Harvey and Sandy on the day of landfall to the mean sentiment exactly one week before.
-These hurricanes are associated with mental health effects including Post-Traumatic Stress Disorder (PTSD) \cite{Schwartz2017Aug, Schwartz2018May}, and similar disasters have been associated with subsequent increases in suicides \cite{Krug1998Feb}.
+We also examine the decrease in sentiment associated with one of the most expensive hurricanes in the last decade in the United States: Hurricane Sandy.
+Specifically, we compare the mean sentiment of counties affected by Sandy on the week after landfall to the mean sentiment for the week previous.
+Hurricanes like Sandy are associated with mental health effects including Post-Traumatic Stress Disorder (PTSD) \cite{Schwartz2017Aug, Schwartz2018May}, and similar disasters have been associated with subsequent increases in suicides \cite{Krug1998Feb}.
 
@@
 Data was at the level of the census block group, the smallest unit for which the US Census releases public data.
+Following Census categories, we report racial characteristics of neighborhoods as majority population of four broad racial categories - non-Hispanic black, non-Hispanic white, Hispanic of any race, and other, which includes Native American, multi-racial, Asian-American and Pacific Islander.
 
@@
 prepared data; M.W.C.
-and J.O.-G.
 analyzed data; and M.W.C., J.O.-G., J.L., and P.A.W.
@@
 
-\noindent This project was co-lead by M.W.C.
-and J.O.-G.
-and both contributed equally to this project.
-Authorship order for these two is listed alphabetically.
-Authorship order for others listed by proportion of contribution.
+\noindent This paper was the result of a SESYNC graduate pursuits project co-lead by M.W.C. and J.O.-G.
 

### 2021-06-08 model/vader/temporal.R
re-run income-race analysis

@@ -1,31 +1,21 @@
-#Use E48
-
 library(data.table)
 library(fixest)
-library(survey)
 library(tidyverse)
-library(cowplot)
 
-options(scipen=100)
-
-MOD_RUN <- 'race_income2'
+RUN <- 'temporal'
 
 setwd('~/tweets/')
 
 data <- fread('all.csv')
-data$income_percap <- log(data$income_percap)
 
 data <- data[weather_term == 0, ]
 
-#Tempknots
-knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
-                          max(data$wbgt)))
-
 data$raining <- data$prcp > 0
+
+data$tod_cont <- as.numeric(substr(data$tod, 1, 2)) + as.numeric(substr(data$tod, 4, 4))/6
 
 #Define Segmenting Function
 piece.formula <- function(var.name, knots, interact_var='') {
-  knots <- knots[c(-1, -length(knots))]
   formula.sign <- rep(" - ", length(knots))
   formula.sign[knots < 0] <- " + "
   if (interact_var == ''){
@@ -40,12 +30,20 @@ piece.formula <- function(var.name, knots, interact_var='') {
 }
 
 formula <- paste0("vader ~ ", 
-                     piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], "race_black"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     'race_black*income_percap*raining + race_black*srad*income_percap',
+                     piece.formula("tod_cont", seq(0, 24, 2), "wbgt"), ' + ',
+                     'raining + srad',
                      " | dow + doy + tod + fips + year + statemonth")
+
+mod <- feols(as.formula(formula), data)
+
+cf <- coef(mod)
+vc <- vcov(mod)
+myobj <- list(vcov=vc, coef=cf)
+class(myobj) <- 'bootmod'
+saveRDS(myobj, file=paste0('bootstrap/', RUN, '/', Sys.time()))
+system('telegram "Did a business "')
+
+
 
 
 for (i in 1:80){
@@ -54,16 +52,13 @@ for (i in 1:80){
[…]

### 2021-06-08 viz/vader/wbgt_income_race_bootstrap.R

@@ -6,7 +6,7 @@ library(cowplot)
 
 options(scipen=100)
 
-MOD_RUN <- 'race_income'
+MOD_RUN <- 'race_income2'
 
 setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
 
@@ -16,7 +16,7 @@ data$income_percap <- log(data$income_percap)
 data <- data[weather_term == 0, ]
 
 inc_qs <- quantile(data$income_percap, seq(0, 1, by=0.05))
-race_qs <- quantile(data$race_white, seq(0, 1, by=0.05))
+race_qs <- quantile(data$race_black, seq(0, 1, by=0.05))
 
 #Tempknots
 knots = list("wbgt"= c(min(data$wbgt), -10, 0, 5, 10, 15, 20, 25, 
@@ -81,25 +81,25 @@ ref_temps <- round(ref_temps)
 #########################################
 preddf <- data.frame(wbgt=seq(-22, 33))
 preddf <- make_groups(preddf, 'income_percap', inc_qs[c(2, 11, 20)])
-preddf <- make_groups(preddf, 'race_white', race_qs[c(2, 11, 20)])
+preddf <- make_groups(preddf, 'race_black', race_qs[c(2, 11, 20)])
 
 preddf$vader <- 1
 
 mm <- model.matrix(as.formula(paste0("vader ~ ", 
                      piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
                      piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], "race_white"), ' + ',
+                     piece.formula("wbgt", knots[['wbgt']], "race_black"), ' + ',
                      piece.formula("wbgt", knots[['wbgt']], ""))),
                    data=preddf)
 
 mm <- mm[ , colnames(mm) != '(Intercept)']
 mm <- mm[ , colnames(mm) != 'income_percap']
-mm <- mm[ , colnames(mm) != 'race_white']
+mm <- mm[ , colnames(mm) != 'race_black']
 
 mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
 
 #Iterate over models
-mods <- list.files()
+mods <- list.files(pattern='06-04')
 for (modf in mods){
 
   mod <- readRDS(modf)
@@ -111,7 +111,7 @@ for (modf in mods){
 }
 
[…]

### 2021-06-08 doc/temp-sentiment.tex

@@
 %https://www.overleaf.com/project/5f918edd1afc380001feca0d
-\title{Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures}
+\title{Fine-scale Twitter Data Reveal Neighborhood Inequalities in Mental Health Effects of High Temperatures}
 
@@
 We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in a quarter of a billion geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
-We find that increased temperatures worsen expressed sentiment in all areas, but that this effect is much stronger in poor and black neighborhoods.
+We find that increased temperatures worsen expressed sentiment in all areas, but that this effect is much stronger in poor and Black neighborhoods.
 \end{abstract}
@@
 For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality or no air-conditioning \cite{Samuelson2020Jun}.
-Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, with these heterogeneities in vulnerability being more pronounced within cities rather than between cities.
+Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, with these heterogeneities in vulnerability being more pronounced within city regions rather than between city regions.
 Indeed, other studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar} and mental health for other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
@@
 Similar methods applied to posts on Facebook can predict the onset of depression \cite{Eichstaedt2018Oct}.
-In London, day-to-day changes in mental health indicators derived from the text of Tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
-Across the USA, more positive Tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
-Finally, the mood expressed in Tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
+Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
+Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
 Thus, posts on Twitter provide an indicator of mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
@@
 We examine the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
-Controlling for a variety of fixed effects across all Tweet, we find higher temperatures are associated with lower sentiment (i.e., worsening mood; See Fig.
+Controlling for a variety of fixed effects across all tweets, we find higher temperatures are associated with lower sentiment (i.e., worsening mood; See Fig.
 \ref{fig:wbgt}).
@@
   \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
-  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment from text in Tweets.  As temperatures increase above 5\textdegree C WBGT, sentiment rapidly declines.}
+  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment from text in tweets.  As temperatures increase above 5\textdegree C WBGT, sentiment rapidly declines.}
   \label{fig:wbgt}
@@
 
-We explored the relationship between neighborhood racial group characteristics and temperature and sentiment and find the effects of heat are felt disproportionately in neighborhoods that are majority black (see Fig \ref{fig:hetero}).
-Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority black neighborhoods decrease four times as much as the sentiment of people in other neighborhoods.
+We explored the relationship between neighborhood racial group characteristics and temperature and sentiment and find the effects of heat are felt disproportionately in neighborhoods that are majority Black (see Fig \ref{fig:hetero}).
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority Black neighborhoods decrease four times as much as the sentiment of people in other neighborhoods.
 Additionally, at modest to moderate temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
@@
 Exploring the combined effects of neighborhood percent minority race and income in moderating the impacts of heat on mental health, we find separate effects for these two variables (see Supplement).
-In other words, neither neither percent minority population nor income alone account for the heterogeneity in vulnerability as neighborhoods that are both poorer and higher percent black are more affected by heat than than neighborhoods that are just poorer or higher proportion black.
+In other words, neither neither percent minority population nor income alone account for the heterogeneity in vulnerability as neighborhoods that are both poorer and higher percent Black are more affected by heat than than neighborhoods that are just poorer or higher proportion Black.
 
@@
 \subsection*{Sentiment Comparison With Other Events}
-Expressed sentiment is a widely used metric to assess mental health and well-being on social media.
-While a variety of related metrics can be used to quantify sentiment, we use the VADER metric that was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
-To give context to this novel metric, we compare the impacts of heat waves, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), to two other changes that have large impacts on sentiment (See Fig.
+We compared the expressed sentiment from exposure to temperatures to other events that have large impacts on sentiment (see Fig.
 \ref{fig:compare}).
-First, we examine the change in average sentiment from Saturday to Monday, the highest and lowest days for Twitter sentiment as well as other mental health metrics, such as suicide rates \cite{CDC2021}.
-We also examine the decrease in sentiment associated with one of the most expensive hurricanes in the last decade in the United States: Hurricane Sandy.
-Specifically, we compare the mean sentiment of counties affected by Sandy on the week after landfall to the mean sentiment for the week previous.
-Hurricanes like Sandy are associated with mental health effects including Post-Traumatic Stress Disorder (PTSD) \cite{Schwartz2017Aug, Schwartz2018May}, and similar disasters have been associated with subsequent increases in suicides \cite{Krug1998Feb}.
+In particular, we examine change in sentiment from lower to higher temperatures, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F) and compare to changes in sentiment from Saturday to Monday and Hurricane Sandy the week before and after landfall.
+Previous work indicates sentiment expressed on Twitter are generally highest on Saturdays and lowest on Mondays, while conversely suicide rates and other metrics of mental health are lowest and highest, respectively \cite{CDC2021}.
+Hurricane Sandy was one of the most damaging and expensive hurricanes to make landfall in the USA in the last decade and hurricanes are associated with mental health effects including PTSD \cite{Schwartz2017Aug, Schwartz2018May}.
+Similar disasters have also been associated with subsequent increases in suicides \cite{Krug1998Feb}.
 
@@
   \includegraphics[width=0.66\linewidth]{../res/comparison_plot.png}
-  \caption{Decline in sentiment during a heat wave across several different neighborhoods, compared with other impacts on sentiment, such as major hurricanes, as well as the weekly variation in sentiment from the peak on Saturday to the low on Monday.}
+  \caption{Decline in sentiment during exposure to change in temperature from lower to higher across several different neighborhoods, compared with other impacts on sentiment, such as major hurricanes, as well as the weekly variation in sentiment from the peak on Saturday to the low on Monday.}
   \label{fig:compare}
@@
 
-We find that the impacts of heatwaves on sentiment in rich and majority white neighborhoods are less than the average weekly change in sentiment from Saturday to Monday, but the effects of heat waves on majority black or poor neighborhoods are much greater than the average weekly change in sentiment.
-Additionally, the impacts of heat waves on more vulnerable neighborhoods are close in severity to the impact of a major hurricane.
+We find that the decline in sentiment in wealthier and majority white neighborhoods from exposure to increasing temperatures is less than the average weekly change in sentiment from Saturday to Monday (see Fig.
+\ref{fig:compare}).
+Conversely, the effects of exposure to increasing temperatures in poorer and majority Black neighborhoods are much larger than the average weekly change in sentiment.
+Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on sentiment from experiencing a major hurricane.
 
-\subsection*{Effects by Combined Statistical Area}
+\subsection*{Temperature Effects by Combined Statistical Area}
+We examine the impact of temperature on sentiment across Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) containing over 1 million people (hereafter: city regions).
+We find that higher temperatures are associated with declining sentiment in 83.6\% of city regions (see Fig.
+\ref{fig:map}) and the effect is significant in 35.3\% of these.
+Furthermore, we find no correlation between higher temperatures and increased sentiment in any of the city regions.
 
-To further explore our findings, we examine the impact of heat on sentiment and sentiment inequalities by Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) with over 1 million people (hereafter: cities).
-We examine the impacts of heat alone, as well as the change in size of the gap in sentiment scores between both poor and rich neighborhoods and black and non-black neighborhoods.
-For our metric of inequality, we show the size of the gap in sentiment as temperatures increase from optimum temperatures of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 30\textdegree C WBGT (42\textdegree C/108\textdegree F).
+Increasing exposure from lower to higher temperatures - 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 30\textdegree C WBGT (42\textdegree C/108\textdegree F) - increases disparities in sentiment across income groups for 68.9\% of city regions, with 35.7\% of these having a statistically significant effect.
+City regions with a significantly unequal effect were located across the USA although they were most common in the mid-Atlantic region.
+Additionally, many city regions in the southwest had an inverted effect, where higher temperatures actually narrowed the gap in sentiment between wealthier and poorer neighborhoods, including Denver and Oklahoma City areas, which had a statistically significant inverted effect.
 
-Examining the effects of heat on sentiment by city, we find that heat is associated with worsened sentiment in 83.6\% of cities, and the effect is significant in 35.3\% of those.
-In no cities does heat significantly increase sentiment, and heat affects sentiment in cities across the US, from Florida to New England to California.
-
-We find that heat increases inequalities in sentiment across income groups for 68.9\% of cities, with 35.7\% of these having a statistically significant effect.
-Cities with a significantly unequal effect were found in the south, Midwest, northeast, northwest, and southwest, although they were most common in the mid-Atlantic region.
-Additionally, many cities in the southwest had an inverted effect, where higher temperatures actually narrowed the gap in sentiment between rich and poor neighborhoods, including Denver and Oklahoma City, which had a statistically significant inverted effect.
-
-Because we found a much stronger effect of temperature on sentiment in majority black neighborhoods compared to other ethnic groups, we also examined increases in sentiment gaps between neighborhoods in the 95th percentile for black population (77.2\% black) and neighborhoods in the 5th percentile for black population (0\% black).
-We excluded cities with a black population less than 5\% of the total population.
-We found increasing inequalities in sentiment for more black neighborhoods in 64.7\% of cities, with a significant effect in 21.2\% of those (7 cities).
-Additionally, three sunbelt cities had a significant inverted effect.
-The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: cities with large and significant inequalities were found the mid-Atlantic and Midwest.
+Given the strong effect of temperature on sentiment in majority Black neighborhoods compared to other non-majority Black neighborhoods, we also examined increases in sentiment gaps between neighborhoods in the 95th percentile for Black population (77.2\% Black) and neighborhoods in the 5th percentile for Black population (0\% Black).
+We excluded city regions with a Black population less than 5\% of the total population.
+We found increasing inequalities in sentiment for more Black neighborhoods in 64.7\% of city regions, with a significant effect in 21.2\% of those (7 city regions).
+Additionally, three sunbelt city regions had a significant inverted effect.
+The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: city regions with large and significant inequalities were located in the mid-Atlantic and Midwest.
 
@@
   \label{fig:map}
-    \caption{Figure (a) shows the change in sentiment for a 25-degree increase in WBGT for cities with a population of over 1 million.
-Figures (b) and (c) show inequalities in impacts of temperature on sentiment for cities with over 1 million people.
-The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5\textdegree C WBGT to 30\textdegree C WBGT between the 5th and 95 percentile for income (a) and the 95th and 5th percentile for percent black (b).
-Cities with a statistically significant effect ($p < 0.05$) are outlined in black.
-For racial inequalities (b), only cites where black people make up $> 5\%$ of the population are included.}
+    \caption{Figure (a) shows the change in sentiment for a 25-degree increase in WBGT for city regions with a population of over 1 million.
+Figures (b) and (c) show inequalities in impacts of temperature on sentiment for city regions with over 1 million people.
+The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5\textdegree C WBGT to 30\textdegree C WBGT between the 5th and 95 percentile for income (b) and the 95th and 5th percentile for percent Black (c).
+City regions with a statistically significant effect ($p < 0.05$) are outlined in Black.
+For racial inequalities (c), only cities where the proportion of Black people is $> 5\%$ of the population are included.}
 \end{figure}
@@
 \section*{Discussion}
-% Benefits of twitter data/overview of findings
-We found that twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution.
-We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between heat and worsened sentiment.
+% Benefits of Twitter data/overview of findings
+We find that Twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution.
+We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between increasing temperatures and declining sentiment.
 Moreover, we were able to show heterogeneities in vulnerability by neighborhood characteristics, overturning previous findings that found no heterogeneities at the county level \cite{Burke2018Aug, Mullins2019Dec}.
 
-% Sentiment is a fuzzy indicator but we have BIG DATA
-One limitation of data from twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
-Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large data set of 243 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
-
 % Compare weekly changes in sentiment to weekly changes in suicides
-We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a heatwave of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in the VADER sentiment score of 0.01.
-This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with a increase in suicides of 23\% over the course of the study period \cite{CDC2021}.
+We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a higher temperature of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in sentiment of 0.01.
+This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}.
 Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.
 Moreover, like sentiment, suicides are affected by higher temperatures.
-Thus, while twitter sentiment is the only mental health indicator available at fine enough spatial and temporal scales to examine neighborhood effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although this analysis cannot speak to those outcomes specifically.
+Thus, while Twitter sentiment is a mental health indicator at fine enough spatial and temporal scales to examine neighborhood effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although our analyses here do not incorporate these outcomes.
 
@@
 This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the poor and vulnerable.
-More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar those of high-income Americans, the effects of high temperatures on mental health can be largely mitigated.
+More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health can be addressed and mitigated.
 
 % Comment on Race, why no effect on hispanics?
-We found that majority black neighborhoods are much more affected by heat than neighborhoods with a majority of any other race.
+We found that majority Black neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
 This is surprising, given that Hispanics can be marginalized in both income and housing.
-This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than black Americans, or may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
+This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than Black Americans, or may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
 Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
-Finally, while race is correlated with income in the US, we found that race contributed to vulnerability independently of income.
-This suggests that racism and patterns of housing discrimination such as a legacy of redlining contribute to making black Americans more vulnerable to heat, event at similar levels of income.
+Finally, while race is correlated with income in the US, we found that race contributed to vulnerability to heat, independently of income.
+This suggests that racism and patterns of housing discrimination like the legacy of redlining contribute to making Black Americans more vulnerable to heat, event at similar levels of income.
 
 % Interpret Spatial Analysis
-In our spatial analysis, we found that heat affects sentiment in cities across the US, suggesting that higher baseline temperatures do not reduce vulnerability.
-Additionally, most cities exhibited both racial and income inequalities in the vulnerability of mental health to heat, although the southwest had the weakest inequalities and an occasional inverted effect.
-While unusual patterns of heat impacts in the southwest could suggest that heat inequality is related to environmental factors such as humidity, this is unlikely because we used Wet Bulb Globe Temperature, which accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
+In our spatial analysis, we found that higher temperatures affect sentiment in city regions across the USA, suggesting that higher baseline temperatures in warmer regions do not necessarily reduce vulnerability.
+Additionally, most city regions exhibited both racial and income inequalities in the vulnerability of mental health to high temperatures, although the Southwest had the weakest inequalities and an occasional inverted effect.
+While unusual patterns of temperature impacts in the Southwest could suggest that temperature inequality is related to environmental factors such as humidity, this is unlikely because we used Wet Bulb Globe Temperature (WBGT) which accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
 Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different baseline environmental conditions.
@@
 % Caveats
-While these findings are robust to different model specifications and sentiment metrics, there are some caveats.
-For one, we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
-Many low-income and black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
+% Sentiment is a fuzzy indicator but we have BIG DATA
+While these findings are robust to different model specifications and sentiment metrics, there are some caveats and limitations.
+One limitation of data from Twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
+Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
+We overcome this limitation by using an extremely large data set of 243 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
+
+One caveat associated with this data is that we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
+Many low-income and Black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
 Moreover, in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are typically found that census block.
@@
 Additionally, neighborhoods of young college students are also typically estimated to have low income levels, even through college students are wealthier than the average American.
-Again, these issues mean we are likely under-estimating the true effect of heat on mental health for poor and black people.
-A final issue with using twitter data is that, while twitter is used by more than one in five Americans, twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
+Again, these issues mean we are likely under-estimating the true effect of heat on mental health for poor and Black people.
+A final issue with using Twitter data is that, while Twitter is used by more than one in five Americans, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
 Nevertheless, we found a large volume of tweets across all neighborhood types.
@@
 % NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
-While climate change will have widespread and severe impacts on human well-being, it cannot be overemphasized just how highly unevenly these impacts will be distributed.
+While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how highly unevenly these impacts will be distributed.
 People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \cite{bullard2012wrong}.
@@
 Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect.
-By using fine-scale twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States.
+By using fine-scale Twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States.
 These findings have significant implications for urban planning, climate and environmental justice, and mental health.
@@
 We used data from 243 million geo-located English-language tweets from the continental United States from the years 2009 to 2019.
-The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
+The Twitter data consists of publicly posted messages, or tweets, that are short status messages users post to the platform.
 We only considered users’ original content, thus did not include retweets in our analysis.
@@
 
-We assessed the sentiment expressed in the tweets using the VADER (Valence Aware Dictionary for sentiment Reasoning) sentiment corpus \cite{gilbert_vader_2014}.
-VADER is a lexicon and rule-based sentiment analysis tool that is specifically attuned to sentiments expressed in social media because of several features, as:
+Expressed sentiment is a widely used metric to assess mental health and well-being from posts on social media platforms.
+We use the VADER metric to assess the sentiment of Twitter posts as it was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
+The VADER (Valence Aware Dictionary for sentiment Reasoning) sentiment corpus \cite{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to sentiments expressed on social media and includes several features, as:
 
@@
 Data was at the level of the census block group, the smallest unit for which the US Census releases public data.
-Following Census categories, we report racial characteristics of neighborhoods as majority population of four broad racial categories - non-Hispanic black, non-Hispanic white, Hispanic of any race, and other, which includes Native American, multi-racial, Asian-American and Pacific Islander.
+Following Census categories, we report racial characteristics of neighborhoods as majority population of four broad racial categories - non-Hispanic Black, non-Hispanic white, Hispanic of any race, and other, which includes Native American, multi-racial, Asian-American and Pacific Islander.
 
@@
 Mean annual income per capita is provided by the ACS, and we standardized this variable so that the values for each year were in 2019 dollars.
-For racial categories, we combined the various categories provided by the ACS into four racial groups: non-Hispanic white, non-Hispanic black, Hispanic of any race, and an "other" category for non-Hispanic people who were neither black nor white, such as Asian, Native American, or mixed-race people.
+For racial categories, we combined the various categories provided by the ACS into four racial groups: non-Hispanic white, non-Hispanic Black, Hispanic of any race, and an "other" category for non-Hispanic people who were neither Black nor white, such as Asian, Native American, or mixed-race people.
 
@@
 For our analyses by city, we run models with a similar form to Model \ref{mod:2}, including the same fixed effects, although we fit a linear effect for $f_t()$.
-Additionally, for exploring the vulnerability of black americans, we use a continuous variable for $m$ indicating the percentage of a neighborhoods population that is black.
+Additionally, for exploring the vulnerability of Black americans, we use a continuous variable for $m$ indicating the percentage of a neighborhoods population that is Black.
 Finally, because we are not accounting for non-linearities in $f_t()$, we exclude tweets with a WBGT of less than 5\textdegree C, as our initial analysis showed that this is the temperature at which sentiment peaks.
@@
 prepared data; M.W.C.
+and J.O.-G.
 analyzed data; and M.W.C., J.O.-G., J.L., and P.A.W.

### 2021-06-14 viz/vader/temporal.R
viz temporal analysis

@@ -1,0 +1,117 @@
+library(data.table)
+library(fixest)
+library(survey)
+library(tidyverse)
+library(cowplot)
+
+options(scipen=100)
+
+MOD_RUN <- 'temporal'
+
+setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
+
+data <- fread('~/tweets/all.csv')
+data <- data[weather_term == 0, ]
+data$raining <- data$prcp > 0
+
+data$tod_cont <- as.numeric(substr(data$tod, 1, 2)) + as.numeric(substr(data$tod, 4, 4))/6
+
+####################################
+#Functions
+####################################
+
+vcov.bootmod <- function(object){
+  return(object$vcov)
+}
+coef.bootmod <- function(object){
+  return(object$coef)
+}
+
+fill_knots <- function(x, filln=10){
+  vals <- NULL
+  for (i in 2:length(x)){
+    vals <- c(vals, seq(x[i-1], x[i], length.out=filln))
+  }
+  vals <- unique(vals)
+  vals <- vals[order(vals)]
+  vals
+}
+
+make_groups <- function(df, label, values){
+  final <- data.frame()
+  for (v in values){
+    df[ , label] <- v
+    final <- bind_rows(df, final)
+  }
+  final
+}
+
+
+#Define Segmenting Function
+piece.formula <- function(var.name, knots, interact_var='') {
+  formula.sign <- rep(" - ", length(knots))
+  formula.sign[knots < 0] <- " + "
+  if (interact_var == ''){
+    paste(var.name, "+",
+        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
+  } else{
+    paste(interact_var, '*', var.name, "+",
+        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
+              collapse = " + ", sep=""))
+  }
+}
+
+########################################
+# WBGT
+#########################################
+preddf <- data.frame(wbgt=c(0, 1))
+preddf <- make_groups(preddf, 'tod_cont', seq(0, 24, 0.1))
+preddf$vader <- 1
+
+mm <- model.matrix(as.formula(paste0("vader ~ ", 
[…]

### 2021-06-15 model/vader/temporal.R
better cyclic spline temporal models

@@ -1,50 +1,38 @@
 library(data.table)
-library(fixest)
+library(survey)
 library(tidyverse)
+library(fixest)
+library(splines)
 
-RUN <- 'temporal'
+RUN <- 'temporal2'
 
 setwd('~/tweets/')
 
 data <- fread('all.csv')
 
 data <- data[weather_term == 0, ]
+data <- data[data$wbgt > 5, ]
 
 data$raining <- data$prcp > 0
 
 data$tod_cont <- as.numeric(substr(data$tod, 1, 2)) + as.numeric(substr(data$tod, 4, 4))/6
 
-#Define Segmenting Function
-piece.formula <- function(var.name, knots, interact_var='') {
-  formula.sign <- rep(" - ", length(knots))
-  formula.sign[knots < 0] <- " + "
-  if (interact_var == ''){
-    paste(var.name, "+",
-        paste("I(pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
-  } else{
-    paste(interact_var, '*', var.name, "+",
-        paste("I(", interact_var, "*pmax(", var.name, formula.sign, abs(knots), ", 0))",
-              collapse = " + ", sep=""))
-  }
+knots <- seq(0, 24, by=2)
+
+make_cyclic_splines <- function(knots, x){
+  df <- data.frame(splineDesign(knots, x, ord=2, outer.ok=TRUE))
+  names(df) <- paste0('s', 1:ncol(df))
+  df$s0 <- ifelse(x < knots[2], 1 - df$s1,
+                  ifelse(x > knots[length(knots) - 1], 1 - df[ , ncol(df)],
+                         0))
+  return(df)
 }
 
-formula <- paste0("vader ~ ", 
-                     piece.formula("tod_cont", seq(0, 24, 2), "wbgt"), ' + ',
-                     'raining + srad',
-                     " | dow + doy + tod + fips + year + statemonth")
+data <- cbind(data, make_cyclic_splines(knots, data$tod_cont))
 
-mod <- feols(as.formula(formula), data)
-
-cf <- coef(mod)
-vc <- vcov(mod)
-myobj <- list(vcov=vc, coef=cf)
-class(myobj) <- 'bootmod'
-saveRDS(myobj, file=paste0('bootstrap/', RUN, '/', Sys.time()))
-system('telegram "Did a business "')
-
-
-
+formula <- paste0('vader ~ ', 
[…]

### 2021-06-15 viz/vader/temporal.R

@@ -3,22 +3,39 @@ library(fixest)
 library(survey)
 library(tidyverse)
 library(cowplot)
+library(splines)
+library(lubridate)
 
 options(scipen=100)
 
-MOD_RUN <- 'temporal'
+MOD_RUN <- 'temporal2'
 
 setwd(paste0('~/tweets/bootstrap/', MOD_RUN, '/'))
 
 data <- fread('~/tweets/all.csv')
+
 data <- data[weather_term == 0, ]
+data <- data[data$wbgt > 5, ]
+
 data$raining <- data$prcp > 0
 
 data$tod_cont <- as.numeric(substr(data$tod, 1, 2)) + as.numeric(substr(data$tod, 4, 4))/6
 
+knots <- seq(0, 24, by=2)
+
 ####################################
 #Functions
 ####################################
+
+make_cyclic_splines <- function(knots, x){
+  df <- data.frame(splineDesign(knots, x, ord=2, outer.ok=TRUE))
+  names(df) <- paste0('s', 1:ncol(df))
+  df$s0 <- ifelse(x < knots[2], 1 - df$s1,
+                  ifelse(x > knots[length(knots) - 1], 1 - df[ , ncol(df)],
+                         0))
+  return(df)
+}
+
 
 vcov.bootmod <- function(object){
   return(object$vcov)
@@ -65,53 +82,87 @@ piece.formula <- function(var.name, knots, interact_var='') {
 ########################################
 # WBGT
 #########################################
-preddf <- data.frame(wbgt=c(0, 1))
+preddf <- data.frame(wbgt=1)
 preddf <- make_groups(preddf, 'tod_cont', seq(0, 24, 0.1))
 preddf$vader <- 1
 
-mm <- model.matrix(as.formula(paste0("vader ~ ", 
-                     piece.formula("tod_cont", knots=seq(0, 24, 2), "wbgt"))),
+make_cyclic_splines <- function(knots, x){
+  df <- data.frame(splineDesign(knots, x, ord=2, outer.ok=TRUE))
+  names(df) <- paste0('s', 1:ncol(df))
+  df$s0 <- ifelse(x < knots[2], 1 - df$s1,
+                  ifelse(x > knots[length(knots) - 1], 1 - df[ , ncol(df)],
+                         0))
+  return(df)
+}
+
[…]

### 2021-06-15 viz/vader/temporal.R
Final graphs for temporal and new race/income analysis

@@ -86,15 +86,6 @@ preddf <- data.frame(wbgt=1)
 preddf <- make_groups(preddf, 'tod_cont', seq(0, 24, 0.1))
 preddf$vader <- 1
 
-make_cyclic_splines <- function(knots, x){
-  df <- data.frame(splineDesign(knots, x, ord=2, outer.ok=TRUE))
-  names(df) <- paste0('s', 1:ncol(df))
-  df$s0 <- ifelse(x < knots[2], 1 - df$s1,
-                  ifelse(x > knots[length(knots) - 1], 1 - df[ , ncol(df)],
-                         0))
-  return(df)
-}
-
 preddf <- cbind(preddf, make_cyclic_splines(knots, preddf$tod_cont))
 
 formula <- paste0('vader ~ ', 
@@ -108,7 +99,7 @@ mm <- mm[ , colnames(mm) != '(Intercept)']
 mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
 
 #Iterate over models
-mods <- list.files()[81:160]
+mods <- list.files()
 for (modf in mods){
 
   mod <- readRDS(modf)
@@ -134,13 +125,18 @@ preddf2$tod <- ymd('2021-01-01') +
 data$tod2 <- ymd_hms(paste0('2021-01-01 ', data$tod, '0:00'))
 
 curve <- ggplot(preddf2) + 
-  geom_ribbon(aes(x=tod, ymin=ymin, ymax=ymax), fill='#FF0000', alpha=0.5) + 
-  geom_line(aes(x=tod, y=pred)) + 
+  geom_ribbon(aes(x=tod, ymin=ymin*10, ymax=ymax*10), fill='#FF0000', alpha=0.5) + 
+  geom_line(aes(x=tod, y=pred*10)) + 
   geom_hline(aes(yintercept=0), linetype=2) + 
-  scale_x_datetime(expand=c(0, 0), limits=c(ymd_hms('2021-01-01 00:00:00'),
-                                            ymd_hms('2021-01-02 00:00:00'))) + 
+  scale_x_datetime(expand=c(0, 0), 
+                   limits=c(ymd_hms('2021-01-01 00:00:00'),
+                            ymd_hms('2021-01-02 00:00:00')),
+                   label=function(x){paste0(hour(x), ':00')},
+                   breaks=seq(ymd_hms('2021-01-01 00:00:00'),
+                              ymd_hms('2021-01-01 21:00:00'),
+                              by='3 hours')) + 
   labs(x='Time of Day', 
[…]

### 2021-06-15 viz/vader/wbgt_income_race_bootstrap.R

@@ -81,7 +81,7 @@ ref_temps <- round(ref_temps)
 #########################################
 preddf <- data.frame(wbgt=seq(-22, 33))
 preddf <- make_groups(preddf, 'income_percap', inc_qs[c(2, 11, 20)])
-preddf <- make_groups(preddf, 'race_black', race_qs[c(2, 11, 20)])
+preddf <- make_groups(preddf, 'race_black', c(0, .25, 1))
 
 preddf$vader <- 1
 
@@ -99,7 +99,7 @@ mm <- mm[ , colnames(mm) != 'race_black']
 mmp <- lapply(seq_len(nrow(mm)), function(i) mm[i,])
 
 #Iterate over models
-mods <- list.files(pattern='06-04')
+mods <- list.files()[81:160]
 for (modf in mods){
 
   mod <- readRDS(modf)
@@ -124,16 +124,11 @@ levels(preddf$income_percap) <- preddf$income_percap %>%
 levels(preddf$income_percap) <- paste0(levels(preddf$income_percap), ' (', 
                                        names(inc_qs[c(2, 11, 20)]), ')')
 
-
-conv <- function(x){
-  #Convert % black to % minority
-  100*(1 - x)
-}
-
+m100 <- function(x){x*100}
 levels(preddf$race_black) <- preddf$race_black %>%
                                    levels %>%
                                    as.numeric %>%
-                                   conv %>%
+                                   m100 %>%
                                    round(1) %>%
                                    paste0(., '%')
 
@@ -151,13 +146,15 @@ preddf <- preddf %>%
 
 (curve <- ggplot(preddf) + 
   geom_line(aes(x=wbgt, y=pred, color=income_percap, linetype=race_black)) + 
+  scale_linetype_manual(values=c('0%'=3, '25%'=2, '100%'=1)) + 
   geom_ribbon(aes(x=wbgt, ymin=ymin, ymax=ymax, fill=income_percap, grou=race_black), alpha=0.2) + 
   scale_x_continuous(expand=c(0, 0), limits=c(-22, 33)) + 
-  #scale_y_continuous(breaks=seq(0, -0.02, -0.005)) + 
+  scale_fill_manual(values=c("#e31a1c", "#ff7f00", "#33a02c")) + 
+  scale_color_manual(values=c("#e31a1c", "#ff7f00", "#33a02c")) + 
   labs(x='', y='Change in Mood of Tweet',
[…]

### 2021-06-23 viz/comparison_plot.R
change "Heat Wave" to "Heat Impact"

@@ -108,7 +108,7 @@ annot <- data.frame(color=c(rep('Majority\nBlack', 2),
                     x=c(4.5, 4.5, 
                         1.5, 1.5),
                     y=c(0.0031, 0.00015, 0.0031, 0.00015) + 0.008,
-                    label=c('Heat Wave Impact By', 'Neighborhood Characteristics',
+                    label=c('Impact of Heat By', 'Neighborhood Characteristics',
                             'Impact of', 'Comparison Events'))
 
 ggplot() + 

### 2021-06-23 doc/temp-sentiment.tex

@@
 Findings from large-scale studies of relatively uniform vulnerability to heat are surprising given minority groups and poorer people are more likely to be exposed to undesirable temperatures and otherwise less able to mitigate the effects of higher heat.
-For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality or no air-conditioning \cite{Samuelson2020Jun}.
-Thus, there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes, with these heterogeneities in vulnerability being more pronounced within city regions rather than between city regions.
-Indeed, other studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar} and mental health for other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec} Also, while minority groups and poorer people could be expanding their livelihood activities to increase their income levels by engaging in a range of activities, exclusions based on factors like race and gender limit such opportunities (Hallegatte et al.
+2015).
+Diverse dimensions of the factors of inequalities have implications on coping abilities and further aggravate vulnerability.
+Thus, particular groups and conditions, for example race/ethnicity and density (urbanization) have been identified as having differential exposure or vulnerability to extreme events (Cardona et al 2012).
+For instance, there is evidence for increased vulnerability to extreme events among migrant groups because of language barriers hence inability to understand extreme event-related information and also their prioritization of finding employment and housing (Thomas et al 2019).
+Heterogeneities in vulnerability are also more pronounced within city regions rather than between cities.
+In the United States, considerable regional variability is evident (Cardona et al 2012).
+Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar} and mental health for other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+So there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes.
+with these .
+Importantly, different degrees of inequality combine and interact for systems that relate directly to differential levels of social vulnerability, both in normal times and in the context of a disaster (Kuran et al 2020).
+This refers to the ‘concept of intersectionality’, where according to Kuran et al 2020, a factor may be a significant variable but only when allied with other variables.
+It primarily concerns how intersecting factors be it socioeconomic factors affect individuals who face multiple social inequities, with consequent multiple marginalisation.
 
@@
 \subsection*{Sentiment Comparison With Other Events}
-We compared the expressed sentiment from exposure to temperatures to other events that have large impacts on sentiment (see Fig.
+We compared the expressed sentiment from exposure to temperatures to two other events associated with impacts on sentiment and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact off a major natural disaster (see Fig.
 \ref{fig:compare}).
-In particular, we examine change in sentiment from lower to higher temperatures, defined as a change from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F) and compare to changes in sentiment from Saturday to Monday and Hurricane Sandy the week before and after landfall.
-Previous work indicates sentiment expressed on Twitter are generally highest on Saturdays and lowest on Mondays, while conversely suicide rates and other metrics of mental health are lowest and highest, respectively \cite{CDC2021}.
-Hurricane Sandy was one of the most damaging and expensive hurricanes to make landfall in the USA in the last decade and hurricanes are associated with mental health effects including PTSD \cite{Schwartz2017Aug, Schwartz2018May}.
-Similar disasters have also been associated with subsequent increases in suicides \cite{Krug1998Feb}.
+For this comparison, we define heat impact as the changes in sentiment associated with temperatures increasing from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F); we define weekly changes as the difference between the mean Saturday sentiment and the mean Monday sentiment, calculated across the entire dataset; and we define the impact of Hurricane Sandy as the change in sentiment from the week before the hurricane made landfall to the week after, calculated for only impacted counties.
+Both of these events are associated with mental health impacts: expressed sentiment on Twitter is highest on Saturdays and lowest on Mondays, a weekly pattern that mirrors the dynamics of suicide rates \cite{CDC2021}, while Hurricane Sandy was associated with mental health effects on victims including increased anxiety, PTSD, and depression \cite{Schwartz2017Aug, Lieberman-Cribbin2017}.
 
@@
 Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on sentiment from experiencing a major hurricane.
+
+\subsection{Temperature Effects by Time of Day}
+In addition to providing data with high spatial resolution, twitter data also comes with very high temporal resolution as each tweet is time-stamped.
+Thus, we were able to examine the impact of heat on sentiment over the course of the day (See Fig.
+\ref{fig:ts-wbgt}).
+We found that heat is associated with improved sentiment for a brief period in the late morning, and then has a consistently negative effect on sentiment during the latter half of the day, from noon until 9pm.
+This effect weakens in the early evening through midnight, and then increases greatly throughout the night.
+Heat has the greatest effect on expressed sentiment at 6am.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.5\linewidth]{../res/ts_heat.png}
+  \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted change in VADER score for a 1\textdegree C WBGT increase in temperatures.}
+  \label{fig:ts-wbgt}
+\end{figure}
 
@@
 
-% Comment on Race, why no effect on hispanics?
-We found that majority Black neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
-This is surprising, given that Hispanics can be marginalized in both income and housing.
-This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than Black Americans, or may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
+Our study found disproportionate effects of climate change on identifiable and marginalized populations.
+We however found that majority Black neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
+This is surprising, given that Hispanics can be marginalized in income, language and housing.
+This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than Black Americans.
+It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
 Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
@@
 This suggests that racism and patterns of housing discrimination like the legacy of redlining contribute to making Black Americans more vulnerable to heat, event at similar levels of income.
+Thus, our study shows that, the degree of vulnerability could be related to a category of different dimensions (like combinations of race and housing) and not to a single dimension (race or income).
+This explains how complex vulnerability can be assessed.
+It further concurs with the ‘concept of intersectionality’ where the degree of vulnerability does not depend on a single dimensional attribution but results from a complex relationship between different factors such as a combination of race, housing and income (Kuran et al 2020).
 

### 2021-06-23 doc/temp-sentiment-supplement.tex
Updates from Overleaf
Edits from Overleaf

--- 2021-05-01 doc/temp-sentiment-supplement.tex
@@
 \fancyhf{}
-\lfoot{Supplemental Materials for \textit{Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures}}
+\lfoot{Supplemental Materials for \textit{Fine-scale Data Reveals Neighborhood Inequalities Effects of Heat on Expressed Mood}}
 \rfoot{\thepage}
@@
 
-\printbibliography
+\section{Alternative Sentiment Algorithms}
 
-\end{document}
+\subsection{Hedonometer}
 
-\section*{Supplemental Info}
-\setcounter{table}{0}
-\setcounter{figure}{0}
-\setcounter{section}{0}
-\renewcommand{\thetable}{S\arabic{table}}
-\renewcommand{\thefigure}{S\arabic{figure}}
-\renewcommand{\thesection}{S\arabic{section}}
+In addition to the Valence Aware Dictionary for sentiment Reasoning (VADER) sentiment analysis method \cite{gilbert_vader_2014}, we also used the hedonometer method \cite{Dodds2011Dec}.
+This sentiment metric yielded very similar findings to those presented in the main paper with the VADER method.
 
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/Boston_Map.png}
-  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
-  \label{fig:timeseries}
-\end{figure}
-
-\begin{figure}[H]
-  \centering
-  \includegraphics[width=\linewidth]{../res/map_wbgt.png}
-  \caption{Here is a map of all tweets in boston, overlaid with income by census block}
-  \label{fig:timeseries}
-\end{figure}
-
-\subsubsection{Hedonometer}
 The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
@@
 
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt.png}
+  \caption{Effects of temperature on sentiment, measured with the hedonometer method.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt-income.png}
+  \caption{Effects of temperature on sentiment and moderated by income, measured with the hedonometer method.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt-race_q.png}
+  \caption{Effects of temperature on sentiment and moderated by race, measured with the hedonometer method.}
+\end{figure}
+
+\subsection{AFINN}
+
+AFINN is another wordlist used in sentiment analysis and designed for microblogs like twitter \cite{nielsen2011new}.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/afinn-wbgt.png}
+  \caption{Effects of temperature on sentiment, measured with the AFINN method.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/afinn-wbgt-income.png}
+  \caption{Effects of temperature on sentiment and moderated by income, measured with the AFINN method.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/afinn-wbgt-race_q.png}
+  \caption{Effects of temperature on sentiment and moderated by race, measured with the AFINN method.}
+\end{figure}
+
+\section{Alternative Interaction Variables}
+For our main analysis, we treated income as a continuous (log-transformed) variable and race as a categorical variable.
+We ran similar analyses with income as a categorical variable and race as a continuous variable (the percent of a census block's population that was non-white/minority).
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/wbgt-income_q.png}
+  \caption{Effect of temperature on sentiment and moderated by income, with income as a categorical variable.}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/wbgt-race.png}
+  \caption{Effect of temperature on sentiment and moderated by race, with race as a continuous variable.}
+\end{figure}
+
+\section{Effects of Precipitation and Sunshine}
+While our main variable of interest was temperature as measured by Wet Bulb Globe Temperature, we also interacted our variables of interest with a binary variable for whether it was raining or not, as well as a continous variable for the amount of shortwave radiation in ($W/m^2$), or sunshine.
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/raining-income.png}
+  \caption{Effect of rain on sentiment and moderated by income.}
+  \label{fig:timeseries}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/raining-race_q.png}
+  \caption{Effect of rain on sentiment and moderated by race.}
+  \label{fig:timeseries}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/srad-income.png}
+  \caption{Effect of sunshine on sentiment and moderated by income.}
+  \label{fig:timeseries}
+\end{figure}
+
+\begin{figure}[H]
+  \centering
+  \includegraphics[width=0.6\linewidth]{../res/srad-race_q.png}
+  \caption{Effect of sunshine on sentiment and moderated by race.}
+  \label{fig:timeseries}
+\end{figure}
+
+\section*{Temperature Effects by Combined Statistical Area}
+To explore whether cultural or environmental baselines may affect the relatinoship between heat and sentiment, we examine the impact of temperature on sentiment across Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) containing over 1 million people (hereafter: city regions).
+We find that higher temperatures are associated with declining sentiment in 83.6\% of city regions (see Fig.
+\ref{fig:map}) and the effect is significant in 35.3\% of these.
+Furthermore, we find no correlation between higher temperatures and increased sentiment in any of the city regions.
+
+For this analysis by city, we run models with a similar form to Model \ref{mod:2}, including the same fixed effects, although we fit a linear effect for $f_t()$.
+Because we are not accounting for non-linearities in $f_t()$, we exclude tweets with a WBGT of less than 5\textdegree C, as our initial analysis showed that this is the temperature at which sentiment peaks.
+
+\begin{figure}[H]
+\centering
+  \includegraphics[width=\linewidth]{../res/map_wbgt.png}
+  \label{fig:map}
+    \caption{The change in sentiment for a 25-degree increase in WBGT for city regions with a population of over 1 million.
+For cities shown with a dark outline, the effect is significant at a 95\% confidence interval.}
+\end{figure}
+
+\section{Weather Terms}
+For all of our analysis, we excluded tweets that contained the following weather-related terms: aerovane, air, airstream, altocumulus, altostratus, anemometer, anemometers, anticyclone, anticyclones, arctic, arid, aridity, atmosphere, atmospheric, autumn, autumnal, balmy, baroclinic, barometer, barometers, barometric, blizzard, blizzards, blustering, blustery, blustery, breeze, breezes, breezy, brisk, calm, celsius, chill, chilled, chillier, chilliest, chilly, chinook, cirrocumulus, cirrostratus, cirrus, climate, climates, cloud, cloudburst, cloudbursts, cloudier, cloudiest, clouds, cloudy, cold, colder, coldest, condensation, contrail, contrails, cool, cooled, cooling, cools, cumulonimbus, cumulus, cyclone, cyclones, damp, damp, damper, damper, dampest, dampest, degree, degrees, deluge, dew, dews, dewy, doppler, downburst, downbursts, downdraft, downdrafts, downpour, downpours, dried, drier, dries, driest, drizzle, drizzled, drizzles, drizzly, drought, droughts, dry, dryline, fall, farenheit, flood, flooded, flooding, floods, flurries, flurry, fog, fogbow, fogbows, fogged, fogging, foggy, fogs, forecast, forecasted, forecasting, forecasts, freeze, freezes, freezing, frigid, frost, frostier, frostiest, frosts, frosty, froze, frozen, gale, gales, galoshes, gust, gusting, gusts, gusty, haboob, haboobs, hail, hailed, hailing, hails, haze, hazes, hazy, heat, heated, heating, heats, hoarfrost, hot, hotter, hottest, humid, humidity, hurricane, hurricanes, ice, iced, ices, icing, icy, inclement, landspout, landspouts, lightning, lightnings, macroburst, macrobursts, maelstrom, mercury, meteorologic, meteorologist, meteorologists, meteorology, microburst, microbursts, microclimate, microclimates, millibar, millibars, mist, misted, mists, misty, moist, moisture, monsoon, monsoons, mugginess, muggy, nexrad, nippy, NOAA, nor,’easter, nor,’easters, noreaster, noreasters, overcast, ozone, parched, parching, pollen, precipitate, precipitated, precipitates, precipitating, precipitation, psychrometer, radar, rain, rainboots, rainbow, rainbows, raincoat, raincoats, rained, rainfall, rainier, rainiest, raining, rains, rainy, sandstorm, sandstorms, scorcher, scorching, searing, shower, showering, showers, skiff, sleet, slicker, slickers, slush, slushy, smog, smoggier, smoggiest, smoggy, snow, snowed, snowier, snowiest, snowing, snowmageddon, snowpocalypse, snows, snowy, spring, sprinkle, sprinkles, sprinkling, squall, squalls, squally, storm, stormed, stormier, stormiest, storming, storms, stormy, stratocumulus, stratus, subtropical, summer, summery, sun, sunnier, sunniest, sunny, temperate, temperature, tempest, thaw, thawed, thawing, thaws, thermometer, thunder, thundered, thundering, thunders, thunderstorm, thunderstorms, tornadic, tornado, tornadoes, tropical, troposphere, tsunami, turbulent, twister, twisters, typhoon, typhoons, umbrella, umbrellas, vane, warm, warmed, warming, warms, warmth, waterspout, waterspouts, weather, wet, wetter, wettest, wind, windchill, windchills, windier, windiest, windspeed, windy, winter, wintery, and wintry.
+
+\printbibliography
+
 \end{document}

### 2021-06-23 doc/temp-sentiment.tex
Edits from Overleaf

@@
 %https://www.overleaf.com/project/5f918edd1afc380001feca0d
-\title{Fine-scale Twitter Data Reveal Neighborhood Inequalities in Mental Health Effects of High Temperatures}
+\title{Fine-scale Data Reveals Neighborhood Inequalities Effects of Heat on Expressed Mood}
 
@@
 However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income or race.
-We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in a quarter of a billion geolocated tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
+We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in a quarter of a billion geolocated Tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
 We find that increased temperatures worsen expressed sentiment in all areas, but that this effect is much stronger in poor and Black neighborhoods.
@@
 Findings from large-scale studies of relatively uniform vulnerability to heat are surprising given minority groups and poorer people are more likely to be exposed to undesirable temperatures and otherwise less able to mitigate the effects of higher heat.
-For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec} Also, while minority groups and poorer people could be expanding their livelihood activities to increase their income levels by engaging in a range of activities, exclusions based on factors like race and gender limit such opportunities (Hallegatte et al.
-2015).
-Diverse dimensions of the factors of inequalities have implications on coping abilities and further aggravate vulnerability.
-Thus, particular groups and conditions, for example race/ethnicity and density (urbanization) have been identified as having differential exposure or vulnerability to extreme events (Cardona et al 2012).
-For instance, there is evidence for increased vulnerability to extreme events among migrant groups because of language barriers hence inability to understand extreme event-related information and also their prioritization of finding employment and housing (Thomas et al 2019).
-Heterogeneities in vulnerability are also more pronounced within city regions rather than between cities.
-In the United States, considerable regional variability is evident (Cardona et al 2012).
-Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical health  \cite{Belanger2015Mar, Uejio2011Mar} and mental health for other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}.
+Minority groups and poorer peoples frequently face other systemic and societal exclusions that can limit opportunities for advancement which can limit coping abilities and further aggravate and compound vulnerabilities \cite{hallegatte2016shock}, seomtimes referred to as 'intersectionality' \cite{Kuran2020Nov}.
+Thus, particular groups and conditions like race/ethnicity and density (urbanization) can lead to compounding and differential exposures or vulnerabilities of these groups to extreme events \cite{Cardona2012Jan}.
+For instance, migrant groups are sometimes especially vulnerable to heat waves because language barriers make it difficult or impractical to interpret public safety information, compounded during heat-related events if these groups are prioritize finding employment and housing \cite{Thomas2019Mar}.
+Heterogeneities in vulnerability are also more pronounced within city regions rather than between cities, particularly in the United States where there may be considerable regional variability \cite{Cardona2012Jan}.
+Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical \cite{Belanger2015Mar, Uejio2011Mar} and mental health for other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+Importantly, different degrees of inequality combine and interact for systems that relate directly to differential levels of social vulnerability, both in normal times and in the context of a disaster \cite{Kuran2020Nov}.
 So there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes.
-with these .
-Importantly, different degrees of inequality combine and interact for systems that relate directly to differential levels of social vulnerability, both in normal times and in the context of a disaster (Kuran et al 2020).
-This refers to the ‘concept of intersectionality’, where according to Kuran et al 2020, a factor may be a significant variable but only when allied with other variables.
 It primarily concerns how intersecting factors be it socioeconomic factors affect individuals who face multiple social inequities, with consequent multiple marginalisation.
@@
 Similar methods applied to posts on Facebook can predict the onset of depression \cite{Eichstaedt2018Oct}.
-In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
-Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
-Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+In London, day-to-day changes in mental health indicators derived from the text of Tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
+Across the USA, more positive Tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
+Finally, the mood expressed in Tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
 Thus, posts on Twitter provide an indicator of mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
-We therefore draw on a data set of 243 million geolocated tweets from across the continental United States and covering the period 2009-2019, each paired local weather and neighborhood conditions, to explore how neighborhood conditions affect the relationship between heat and mental heath.
+We therefore draw on a data set of 243 million geolocated Tweets from across the continental United States and covering the period 2009-2019, each paired local weather and neighborhood conditions, to explore how neighborhood conditions affect the relationship between heat and mental heath.
 
@@
 \subsection*{Overall Effect}
-We examine the relationship between the sentiment expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
-Controlling for a variety of fixed effects across all tweets, we find higher temperatures are associated with lower sentiment (i.e., worsening mood; See Fig.
+We examine the relationship between the sentiment expressed in Tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
+Controlling for a variety of fixed effects across all Tweets, we find higher temperatures are associated with lower sentiment (i.e., worsening mood; See Fig.
 \ref{fig:wbgt}).
@@
   \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
-  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment from text in tweets.  As temperatures increase above 5\textdegree C WBGT, sentiment rapidly declines.}
+  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment from text in Tweets.  As temperatures increase above 5\textdegree C WBGT, sentiment rapidly declines.}
   \label{fig:wbgt}
@@
 We explored the relationship between neighborhood racial group characteristics and temperature and sentiment and find the effects of heat are felt disproportionately in neighborhoods that are majority Black (see Fig \ref{fig:hetero}).
-Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of tweets in majority Black neighborhoods decrease four times as much as the sentiment of people in other neighborhoods.
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of Tweets in majority Black neighborhoods decrease four times as much as the sentiment of people in other neighborhoods.
 Additionally, at modest to moderate temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
-
-Exploring the combined effects of neighborhood percent minority race and income in moderating the impacts of heat on mental health, we find separate effects for these two variables (see Supplement).
-In other words, neither neither percent minority population nor income alone account for the heterogeneity in vulnerability as neighborhoods that are both poorer and higher percent Black are more affected by heat than than neighborhoods that are just poorer or higher proportion Black.
 
@@
   \includegraphics[width=\linewidth]{../res/wbgt_combined.png}
-    \caption{Effect of changes in wet-bulb globe temperature on expressed sentiment relationship as moderated by neighborhood income (a) and race (b).}
+    \caption{Effect of changes in WBGT on expressed sentiment as moderated by neighborhood income (a) and race (b).}
 \label{fig:hetero}
 \end{figure}
+
+Finally, given that both low-income and majority black neighborhoods are more vulnerable to heat waves, we model the combined effects of race and sentiment together to explore the intersection of race and income.
+We model how the relationship between temperature and sentiment is moderated by the per-capita income of a neighborhood as well as the percentage of a neighborhood's inhabitants that are black.
+Additionally, we compare AICs for models with no term for vulnerability, a race term for vulnerability, an income term for vulnerability, and both race and income terms for vulnerability.
+
+\begin{minipage}{.65\textwidth}
+\centering
+\includegraphics[width=\textwidth]{../res/wbgt-income-race.png}
+\captionof{figure}{Relationship between WBGT and sentiment from text in Tweets, moderated by the interaction of race and income.}
+\label{fig:race-income}
+\end{minipage}\hfill
+\begin{minipage}{.35\textwidth}
+\centering
+\begin{tabular}{| l | l |}
+\hline
+Terms & AIC \\
+\hline
+Neither & 277270561 \\
+Income  & 277182239 \\
+Race    & 277143579 \\
+Both    & 277108676 \\
+\hline
+\end{tabular}
+\label{tab:aics}
+\captionof{table}{Akaike information criterion (AIC) of models with interaction terms for race, income, both race and income, as well as neither race nor income.
+The model with the best fit by AIC includes both race and income interaction terms.}
+\end{minipage}
+
+Graphing the combined effects of race and income, we find that race has a larger impact on vulnerability than income (see Fig.
+\ref{fig:race-income}).
+A fully black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants.
+Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and sentiment.
+Using AIC as an indicator of how different terms affect model quality (see Tab.
+\ref{tab:aics}), we find that a model including both terms performs the best, indicating that both race and income matter for the impact of heat on sentiment.
+Moreover, including only a race term improves the model more than including only an income term, further supporting the conclusion that race is more important than income for explaining the vulnerability of sentiment to heat.
 
@@
 \subsection{Temperature Effects by Time of Day}
-In addition to providing data with high spatial resolution, twitter data also comes with very high temporal resolution as each tweet is time-stamped.
+In addition to providing data with high spatial resolution, Twitter data also comes with very high temporal resolution as each tweet is time-stamped.
 Thus, we were able to examine the impact of heat on sentiment over the course of the day (See Fig.
@@
   \label{fig:ts-wbgt}
-\end{figure}
-
-\subsection*{Temperature Effects by Combined Statistical Area}
-We examine the impact of temperature on sentiment across Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) containing over 1 million people (hereafter: city regions).
-We find that higher temperatures are associated with declining sentiment in 83.6\% of city regions (see Fig.
-\ref{fig:map}) and the effect is significant in 35.3\% of these.
-Furthermore, we find no correlation between higher temperatures and increased sentiment in any of the city regions.
-
-Increasing exposure from lower to higher temperatures - 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 30\textdegree C WBGT (42\textdegree C/108\textdegree F) - increases disparities in sentiment across income groups for 68.9\% of city regions, with 35.7\% of these having a statistically significant effect.
-City regions with a significantly unequal effect were located across the USA although they were most common in the mid-Atlantic region.
-Additionally, many city regions in the southwest had an inverted effect, where higher temperatures actually narrowed the gap in sentiment between wealthier and poorer neighborhoods, including Denver and Oklahoma City areas, which had a statistically significant inverted effect.
-
-Given the strong effect of temperature on sentiment in majority Black neighborhoods compared to other non-majority Black neighborhoods, we also examined increases in sentiment gaps between neighborhoods in the 95th percentile for Black population (77.2\% Black) and neighborhoods in the 5th percentile for Black population (0\% Black).
-We excluded city regions with a Black population less than 5\% of the total population.
-We found increasing inequalities in sentiment for more Black neighborhoods in 64.7\% of city regions, with a significant effect in 21.2\% of those (7 city regions).
-Additionally, three sunbelt city regions had a significant inverted effect.
-The patterns of inequalities in the impacts of heat on sentiment by race were similar to inequalities by income: city regions with large and significant inequalities were located in the mid-Atlantic and Midwest.
-
-\begin{figure}[H]
-\centering
-  \includegraphics[width=0.5\linewidth]{../res/map_combined.png}
-  \label{fig:map}
-    \caption{Figure (a) shows the change in sentiment for a 25-degree increase in WBGT for city regions with a population of over 1 million.
-Figures (b) and (c) show inequalities in impacts of temperature on sentiment for city regions with over 1 million people.
-The value shown is the predicted change in the size of the gap in VADER score as temperatures increase from 5\textdegree C WBGT to 30\textdegree C WBGT between the 5th and 95 percentile for income (b) and the 95th and 5th percentile for percent Black (c).
-City regions with a statistically significant effect ($p < 0.05$) are outlined in Black.
-For racial inequalities (c), only cities where the proportion of Black people is $> 5\%$ of the population are included.}
 \end{figure}
@@
 This explains how complex vulnerability can be assessed.
-It further concurs with the ‘concept of intersectionality’ where the degree of vulnerability does not depend on a single dimensional attribution but results from a complex relationship between different factors such as a combination of race, housing and income (Kuran et al 2020).
+It further concurs with the ‘concept of intersectionality’ where the degree of vulnerability does not depend on a single dimensional attribution but results from a complex relationship between different factors such as a combination of race, housing and income \cite{Kuran2020Nov}.
 
@@
 Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large data set of 243 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
+We overcome this limitation by using an extremely large data set of 243 million Tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
 
-One caveat associated with this data is that we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
+One caveat associated with this data is that we were only able to locate the Tweets within the census block where the tweet was sent - we did not to infer where the person sending the Tweets typically lived or where they had been.
 Many low-income and Black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
@@
 A final issue with using Twitter data is that, while Twitter is used by more than one in five Americans, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
-Nevertheless, we found a large volume of tweets across all neighborhood types.
+Nevertheless, we found a large volume of Tweets across all neighborhood types.
 
@@
 \subsection*{Tweets \& Sentiment}
-We used data from 243 million geo-located English-language tweets from the continental United States from the years 2009 to 2019.
-The Twitter data consists of publicly posted messages, or tweets, that are short status messages users post to the platform.
-We only considered users’ original content, thus did not include retweets in our analysis.
-Additionally, we excluded all tweets that contained weather-related terms, to ensure that the sentiment expressed in the tweets was reflective of a user's mental state, and not a commentary on the weather.
+We used data from 243 million geo-located English-language Tweets from the continental United States from the years 2009 to 2019.
+The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
+We only considered users’ original content, thus did not include reTweets in our analysis.
+Additionally, we excluded all Tweets that contained weather-related terms, to ensure that the sentiment expressed in the Tweets was reflective of a user's mental state, and not a commentary on the weather.
 
@@
 
-The VADER method yields a value for the mood of a tweet, with a score of 0 for neutral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
-In addition to the VADER method used in the body of this paper, we also conduct our analysis using the Hedonometer and XXXXX sentiment analysis methods, with similar results (See Supplement).
+The VADER method yields a value for the mood of a tweet, with a score of 0 for neutral Tweets, a score $> 0$ for positive-mood Tweets and a score $< 0$ for negative-mood Tweets.
+In addition to the VADER method used in the body of this paper, we also conduct our analysis using the Hedonometer and AFINN sentiment analysis methods, with similar results (See Supplement).
 
@@
 \subsection*{Socio-Econoimc Data}
-We used data from the American Community Survey (ACS) administered by the US Census to estimate income levels and the racial composition of neighborhoods where tweets were located.
+We used data from the American Community Survey (ACS) administered by the US Census to estimate income levels and the racial composition of neighborhoods where Tweets were located.
 Data was at the level of the census block group, the smallest unit for which the US Census releases public data.
@@
 We therefore matched each tweet with census block group data from the year at the middle year of each survey's five-year range.
-For example, tweets from 2014 were matched to data from the 2012-2016 ACS.
-Because the most recent available ACS was from 2014-2018, all tweets from year years greater than 2016 were matched to this dataset.
+For example, Tweets from 2014 were matched to data from the 2012-2016 ACS.
+Because the most recent available ACS was from 2014-2018, all Tweets from year years greater than 2016 were matched to this dataset.
 Data was downloaded from the IPUMS NHGIS service provided by the University of Minnesota.
@@
 
-In this paper, we give income and racial percentiles based on the observed percentiles across all tweets in our data set.
+In this paper, we give income and racial percentiles based on the observed percentiles across all Tweets in our data set.
 
 \subsection*{Modeling}
-We assessed how expressed sentiment is affected by Wet Bulb Globe Temperature using segmented regressions, controlling for precipitation, and shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: the day of the week, the time of day, the day of the year, the year, the month, and the county.
+We assessed how expressed sentiment is affected by Wet Bulb Globe Temperature using segmented regressions, controlling for precipitation, shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: the day of the week, the time of day, the day of the year, the year, the month, and the county.
 
@@
 
-Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represented the segmented regression for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
-We estimate the 95\% confidence interval of our models using 80-fold bootstrapping.
+Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represents the segmented effect for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
+We estimate the 95\% confidence interval of all our models using 80-fold bootstrapping.
 
@@
 
-For our analyses by city, we run models with a similar form to Model \ref{mod:2}, including the same fixed effects, although we fit a linear effect for $f_t()$.
-Additionally, for exploring the vulnerability of Black americans, we use a continuous variable for $m$ indicating the percentage of a neighborhoods population that is Black.
-Finally, because we are not accounting for non-linearities in $f_t()$, we exclude tweets with a WBGT of less than 5\textdegree C, as our initial analysis showed that this is the temperature at which sentiment peaks.
+In addition to our models exploring how income and race moderate the effect of heat on sentiment individually, we fit an additional model with an interaction terms for both race and income.
+Due to the complexity of a model with segmented terms interacting with two different variables, using a categorical variable for race would have yielded a model matrix too large for the largest cloud virtual machine available to us.
+Thus, we used a continuous metric for both income as well as race (percent black).
+This model extends our main model to the following form:
+
+\begin{equation}
+    y = \begin{tabular}{l}
+    $\beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) + 
+    \beta_p p + \beta_{m_{r} p} m_{r} p + 
+    \beta_{m_{i} p} m_{i} p + \beta_{m_{i} m_{r} p} m_{i} m_{r} p + $
+    \\
+    $\beta_s s + \beta_{m_{i} s} m_{i} s + 
+    \beta_{m_{r} s} m_{r} s + \beta_{m_{i} m_{r} s} m_{i} m_{r} s + 
+    \Phi + \epsilon$
+    \end{tabular}
+    \label{mod:race-inc}
+\end{equation}
+
+In this new model, $m_{i}$ is average income in the census block where the tweet originated, while $m_{r}$ is the racial composition of the census block.
+
+Finally, we fit a model for how heat affects sentiment by time of day.
+For this analysis, we used a varying-coefficient model, where the effect of heat is linear, but the coefficient for the effect varies non-linearly as a function of the time of day.
+Because we modeled the effect of heat as linear, we excluded all tweets with an observed temperatures less than 5\textdegree C WBGT so that we were only including values at which our previous analyses suggested the heat-sentiment relationship is monotonic.
+This final model took the following form:
+
+\begin{equation}
+    y = \beta_0 + f_h(h)t + \beta_p p + \beta_s s + \Phi + \epsilon
+    \label{mod:tod}
+\end{equation}
+
+In this final model, $f_{h}$ is the coefficient for temperature, $t$, and varies non-linearly by time of day, $h$.
+To ensure that the effect $f_{h}$ is cyclical, we estimated $f_{h}$ using 2nd-order B-splines, also known as a tent function basis \cite[Chapter~4.2]{wood2017generalized}, with knots every two hours.
 
@@
 \section*{Author Contributions}
-Author contributions: M.W.C., J.O.-G., A.S.
+Author contributions: M.C., J.O.-G., A.S.
 and P.B.
 designed the research and modeling strategy; A.S.
-provided data; M.W.C and Z.L.
-prepared data; M.W.C.
+provided data; M.C and Z.L.
+prepared data; M.C.
 and J.O.-G.
-analyzed data; and M.W.C., J.O.-G., J.L., and P.A.W.
+analyzed data; and M.C., J.O.-G., J.L., and P.A.W.
 wrote the paper.
 
-\noindent This paper was the result of a SESYNC graduate pursuits project co-lead by M.W.C. and J.O.-G.
+\noindent This paper was the result of a SESYNC graduate pursuits project co-lead by M.C. and J.O.-G.
 

### 2021-06-23 viz/Black_vs_Income.R
change sentiment -> mood in figures

@@ -1,0 +1,46 @@
+library(sf)
+library(tidyverse)
+
+setwd('~/tweets/')
+
+#Because it is 5-year average, use the middle year of the 5.  So 2011 is 2007-2011, so most accurate to 2009
+
+#Convert all to 2019 dollars
+#https://data.bls.gov/cgi-bin/cpicalc.pl?cost1=100&year1=201106&year2=201906
+convert <- data.frame(year=2011:2018,
+                      conv=c(1.134772, 1.1161985, 1.0969534, 1.0746823, 
+                             1.0733538, 1.0627546, 1.0456737, 1.0164848))
+
+dat <- read.csv('wealth/nhgis0003_csv/nhgis0003_ds206_20145_2014_blck_grp_E.csv')
+
+names(dat)[38:94] <- c("TOTAL_POPULATION", "RACE_TOT", "RACE_NHISP_TOTXX", "RACE_WHITE", 
+                       "RACE_BLACK", "RACE_INDIAN", "RACE_ASIAN", "RACE_OTHER1", "RACE_OTHER2",
+                       "RACE_OTHER3", "RACE_OTHER4", "RACE_OTHER5", "RACE_HISP", 
+                       "RACE_HISPXX1", "RACE_HISPXX2", "RACE_HISPXX3", "RACE_HISPXX4", 
+                       "RACE_HISPXX5", "RACE_HISPXX6", "RACE_HISPXX7", "RACE_HISPXX8", 
+                       "RACE_HISPXX9", "TRAN_TOT", "TRAN_CAR1", "TRAN_CAR2", "TRAN_CAR3", 
+                       "TRAN_CAR4", "TRAN_CAR5", "TRAN_CAR6", "TRAN_CAR7", "TRAN_CAR8", 
+                       "TRAN_PUB1", "TRAN_PUB2", "TRAN_PUB3", "TRAN_PUB4", "TRAN_PUB5", 
+                       "TRAN_PUB6", "TRAN_CAR9", "TRAN_EXP1", "TRAN_EXP2", "TRAN_EXP3", 
+                       "TRAN_OTHER", "TRAN_WFH", "TIME_TOT", "TIME_LT5", "TIME_5_9",
+                       "TIME_10_14", "TIME_15_19", "TIME_20_24", "TIME_25_29", "TIME_30_34", 
[…]

### 2021-06-23 viz/comparison_plot.R

@@ -126,7 +126,7 @@ ggplot() +
                               'Hurricane\nSandy'='grey30')) + 
   scale_y_continuous(expand=expand_scale(mult=c(0.05, 0))) + 
   scale_x_discrete(position='top') + 
-  labs(x='', y='Decrease in Sentiment') + 
+  labs(x='', y='Decrease in Mood') + 
   guides(fill=FALSE, color=FALSE) + 
   theme_classic() + 
   coord_cartesian(ylim=c(min(plt$min)*1.05, 0), clip='off') + 

### 2021-06-23 doc/temp-sentiment.tex
Updates from Overleaf

@@
 Higher temperatures associated with climate change are expected to have major impacts on human mental health.
-Indeed, traditional analyses of heat and mental health outcomes using data collected by public agencies have found strong associations between elevated temperatures and outcomes such as suicides and mental-health related hospitalizations.
-However, these studies, which typically use data reported at the city or county level, have found no difference in vulnerability based on income or race.
-We overturn this finding and show that there are stark differences in vulnerability using a novel data source: expressed sentiment in a quarter of a billion geolocated Tweets, matched with prevailing weather as well as neighborhood economic and demographic conditions.
-We find that increased temperatures worsen expressed sentiment in all areas, but that this effect is much stronger in poor and Black neighborhoods.
+Indeed, previous work has found strong associations between heat and outcomes such as suicides and mental health related hospitalizations.
+These studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race at coarse spatial scales.
+However, using expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status, we show that there are stark differences in the effects of heat depending on neighborhood characteristics.
+We find that increased temperatures worsen expressed mood in all areas, but that this effect is much stronger in poor and Black neighborhoods.
+We also find that the effect of heat on expressed mood is greatest in the early morning, supporting the hypothesis that heat affects mental health through a sleep quality pathway.
+This paper presents the first evidence of heterogeneities in the vulnerability of mental health to heat.
 \end{abstract}
@@
 While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \cite{Berry2018Apr, Collins2011Jul}.
-Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, and linkages between higher temperatures and a variety of mental health impacts \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}.
+Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, and linkages between higher temperatures and a variety of mental health outcomes \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}.
 Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of heat on mental health, especially on identifying the most vulnerable communities.
@@
 Additionally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}.
-Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can have compounding consequences for exacerbate mental health issues \cite{Berry2007, WHO2007}.
-Finally, exposure to higher temperatures can affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health impacts and outcomes \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
+Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007, WHO2007}.
+Finally, exposure to higher temperatures can affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health outcomes \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
 
 %Empirical: Heat and mental health
-Results from an emerging literature indicate higher temperatures are also associated with indicators of deteriorating mental health.
-Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May}.
-Other studies find similar relationships in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
+Results from an emerging literature are beginning to demonstrate that higher temperatures can lead to deteriorating mental health.
+Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and other studies find a similar relationship in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
 Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb}.
-Higher temperatures are also linked to increased mortality among people with mental health disorders \cite{Hansen2008Oct}.
+Higher temperatures are also linked to increased mortality among people with pre-existing mental health disorders \cite{Hansen2008Oct}.
 
-The effects of heat exposure on mental health are relatively well established, yet results from research conducted at national scales indicate surprisingly little heterogeneity in impacts, with consequences for our understanding of who may be most or least vulnerable to heat stress and ways we plan for and adapt to changing environmental conditions.
+The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicate surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as adaptation to future heat stress.
 For example, a large study of temperature and suicide in the USA and Mexico found ``no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
 Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \cite{Mullins2019Dec}.
-A major challenge for these studies, however, is their reliance on data from public health agencies, data which is typically aggregated to the municipality or county level.
+A major challenge for these studies, however, is their reliance on data that is typically aggregated to the municipality or county level.
 As a consequence, these studies are often restricted to between-county metrics of vulnerability and cannot take into account neighborhood effects.
@@
 
-While public health data on mental health outcomes is not available at the neighborhood scale, spontaneous, in situ data from the social media platform Twitter can provide an important, inexpensive and timely indicator of mental health at extremely precise spatial and temporal scales.
+While public health data on mental health outcomes is not available at the neighborhood scale, spontaneous, in situ data from the social media platform Twitter can provide an important, inexpensive and timely proxy for mental health at extremely precise spatial and temporal scales.
 Many studies use the mood expressed in social media as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.

### 2021-06-23 viz/combine_plots.sh
tweak figures

@@ -4,8 +4,8 @@ mkdir conv
 cd conv
 
 #Add labels
-convert ../wbgt-income.png -gravity NorthWest -pointsize 125 -font "Arial-Bold" -annotate +25+25 'a' wbgt-income.png
-convert ../wbgt-race_q.png -gravity NorthWest -pointsize 125 -font "Arial-Bold" -annotate +25+25 'b' wbgt-race_q.png
+convert ../wbgt-income.png -gravity NorthWest -pointsize 125 -font "Arial-Bold" -annotate +10+10 'a' wbgt-income.png
+convert ../wbgt-race_q.png -gravity NorthWest -pointsize 125 -font "Arial-Bold" -annotate +10+10 'b' wbgt-race_q.png
 
 #Stack
 convert +append wbgt-income.png wbgt-race_q.png ../wbgt_combined.png

### 2021-06-23 doc/temp-sentiment.tex
Updates from Overleaf

@@
 \subsection*{Overall Effect}
-We examine the relationship between the sentiment expressed in Tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
-Controlling for a variety of fixed effects across all Tweets, we find higher temperatures are associated with lower sentiment (i.e., worsening mood; See Fig.
+We examine the relationship between the mood expressed in Tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
+Controlling for a variety of fixed effects across all Tweets, we find higher temperatures are associated with worsened mood (Fig.
 \ref{fig:wbgt}).
-Sentiment is highest at 5\textdegree C WBGT (a dry bulb temperature typically around 12\textdegree C/54\textdegree F) with the largest declines in sentiment between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
-Sentiment also declines with colder temperatures (below 5\textdegree C WBGT), although only slightly.
-We ran models using multiple measures of sentiment and find similar results and patterns across all sentiment measures (see Supplement).
+Mood is highest at 5\textdegree C WBGT (a dry bulb temperature typically around 12\textdegree C/54\textdegree F) with the largest declines between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
+Mood also declines with colder temperatures (below 5\textdegree C WBGT), although only slightly.
+We estimated mood using the Valence Aware Dictionary for Sentiment Reasoning (VADER) method in our main analysis \cite{gilbert_vader_2014}, although using other methods yield similar results (see Supplement).
 
@@
   \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
-  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and sentiment from text in Tweets.  As temperatures increase above 5\textdegree C WBGT, sentiment rapidly declines.}
+  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and expressed mood in Tweets.  As temperatures increase above 5\textdegree C WBGT, mood rapidly declines.}
   \label{fig:wbgt}
@@
 
-We find median neighborhood income strongly moderates the relationship between temperature and sentiment (see Fig \ref{fig:hetero}, panel A), with large differences in sentiment between the poorest and wealthiest neighborhoods.
+We find that mean neighborhood income strongly moderates the relationship between temperature and sentiment (Fig.
+\ref{fig:hetero}, panel A), with large differences in sentiment between the poorest and wealthiest neighborhoods.
 As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment increases in the wealthiest neighborhoods (95th income percentile) but decreases in the median and lowest income neighborhoods.

### 2021-06-24 doc/temp-sentiment-supplement.tex
Final draft for coauthors approval

@@
 
-\section{Alternative Sentiment Algorithms}
+\section{Alternative Sentiment Algorithms to Estimate Mood}
 
@@
 
-In addition to the Valence Aware Dictionary for sentiment Reasoning (VADER) sentiment analysis method \cite{gilbert_vader_2014}, we also used the hedonometer method \cite{Dodds2011Dec}.
-This sentiment metric yielded very similar findings to those presented in the main paper with the VADER method.
+In addition to the Valence Aware Dictionary for Sentiment Reasoning (VADER) method \cite{gilbert_vader_2014}, we also used the hedonometer method \cite{Dodds2011Dec}.
+This metric yielded very similar findings to those presented in the main paper with the VADER method.
 
-The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get sentiment score from multi-lingual texts.
-The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2)using a naive algorithm for scaling up from individual words to texts.
+The Hedonometer \cite{dodds_temporal_2011} is a corpus-based technique to get mood scores from multi-lingual texts.
+The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2) using an algorithm for scaling up from individual words to texts.
 For the English corpus, Dodds and others \cite{dodds_temporal_2011} collect 10,222 unique words and used crowd-sourcing platform Amazon Mechanical Turk to get human evaluation of happiness degree for each word in an integer scale from 1 to 9, representing a sad to happy spectrum.
 Score 5 represents neutral words.
-Each word will be calculated average score and then the word and happiness score are compiled into a dataset (labMT 1.0).
 Some illustrative example of words are:
@@
 
-The sentiment score for single text will be the mean happiness score for all words in the text.
+The score for single text will be the mean happiness score for all words in the text.
 
@@
   \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt.png}
-  \caption{Effects of temperature on sentiment, measured with the hedonometer method.}
+  \caption{Effects of temperature on mood, measured with the hedonometer method.}
 \end{figure}
@@
   \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt-income.png}
-  \caption{Effects of temperature on sentiment and moderated by income, measured with the hedonometer method.}
+  \caption{Effects of temperature on mood and moderated by income, measured with the hedonometer method.}
 \end{figure}
@@
   \includegraphics[width=0.6\linewidth]{../res/hedono-wbgt-race_q.png}
-  \caption{Effects of temperature on sentiment and moderated by race, measured with the hedonometer method.}
+  \caption{Effects of temperature on mood and moderated by race, measured with the hedonometer method.}
 \end{figure}
@@
 
-AFINN is another wordlist used in sentiment analysis and designed for microblogs like twitter \cite{nielsen2011new}.
+AFINN is another wordlist used in mood analysis and designed for microblogs like twitter \cite{nielsen2011new}.
 
@@
   \includegraphics[width=0.6\linewidth]{../res/afinn-wbgt.png}
-  \caption{Effects of temperature on sentiment, measured with the AFINN method.}
+  \caption{Effects of temperature on mood, measured with the AFINN method.}
 \end{figure}
@@
   \includegraphics[width=0.6\linewidth]{../res/afinn-wbgt-income.png}
-  \caption{Effects of temperature on sentiment and moderated by income, measured with the AFINN method.}
+  \caption{Effects of temperature on mood and moderated by income, measured with the AFINN method.}
 \end{figure}
@@
   \includegraphics[width=0.6\linewidth]{../res/afinn-wbgt-race_q.png}
-  \caption{Effects of temperature on sentiment and moderated by race, measured with the AFINN method.}
+  \caption{Effects of temperature on mood and moderated by race, measured with the AFINN method.}
 \end{figure}
@@
   \includegraphics[width=0.6\linewidth]{../res/wbgt-income_q.png}
-  \caption{Effect of temperature on sentiment and moderated by income, with income as a categorical variable.}
+  \caption{Effect of temperature on mood and moderated by income, with income as a categorical variable.}
 \end{figure}
@@
   \includegraphics[width=0.6\linewidth]{../res/wbgt-race.png}
-  \caption{Effect of temperature on sentiment and moderated by race, with race as a continuous variable.}
+  \caption{Effect of temperature on mood and moderated by race, with race as a continuous variable.}
 \end{figure}
@@
   \includegraphics[width=0.6\linewidth]{../res/raining-income.png}
-  \caption{Effect of rain on sentiment and moderated by income.}
+  \caption{Effect of rain on mood and moderated by income.}
   \label{fig:timeseries}
@@
   \includegraphics[width=0.6\linewidth]{../res/raining-race_q.png}
-  \caption{Effect of rain on sentiment and moderated by race.}
+  \caption{Effect of rain on mood and moderated by race.}
   \label{fig:timeseries}
@@
   \includegraphics[width=0.6\linewidth]{../res/srad-income.png}
-  \caption{Effect of sunshine on sentiment and moderated by income.}
+  \caption{Effect of sunshine on mood and moderated by income.}
   \label{fig:timeseries}
@@
   \includegraphics[width=0.6\linewidth]{../res/srad-race_q.png}
-  \caption{Effect of sunshine on sentiment and moderated by race.}
+  \caption{Effect of sunshine on mood and moderated by race.}
   \label{fig:timeseries}
@@
 \section*{Temperature Effects by Combined Statistical Area}
-To explore whether cultural or environmental baselines may affect the relatinoship between heat and sentiment, we examine the impact of temperature on sentiment across Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) containing over 1 million people (hereafter: city regions).
-We find that higher temperatures are associated with declining sentiment in 83.6\% of city regions (see Fig.
+To explore whether cultural or environmental baselines may affect the relationship between heat and mood, we examine the impact of temperature on mood across Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) containing over 1 million people (hereafter: city regions).
+We find that higher temperatures are associated with declining mood in 83.6\% of city regions (see Fig.
 \ref{fig:map}) and the effect is significant in 35.3\% of these.
-Furthermore, we find no correlation between higher temperatures and increased sentiment in any of the city regions.
+Furthermore, we find no significant association between higher temperatures and increased mood in any of the cities.
 
-For this analysis by city, we run models with a similar form to Model \ref{mod:2}, including the same fixed effects, although we fit a linear effect for $f_t()$.
-Because we are not accounting for non-linearities in $f_t()$, we exclude tweets with a WBGT of less than 5\textdegree C, as our initial analysis showed that this is the temperature at which sentiment peaks.
+For this analysis by city, we run models with a similar form to Model 2, including the same fixed effects, although we fit a linear effect for $f_t()$.
+Because we are not accounting for non-linearities in $f_t()$, we exclude tweets with a WBGT of less than 5\textdegree C, as our initial analysis showed that this is the temperature at which mood peaks.
 
@@
   \label{fig:map}
-    \caption{The change in sentiment for a 25-degree increase in WBGT for city regions with a population of over 1 million.
+    \caption{The change in mood for a 25-degree increase in WBGT for city regions with a population of over 1 million.
 For cities shown with a dark outline, the effect is significant at a 95\% confidence interval.}

### 2021-06-30 viz/vader/wbgt_income_race_bootstrap.R
tweak race-income figure

@@ -81,13 +81,13 @@ ref_temps <- round(ref_temps)
 #########################################
 preddf <- data.frame(wbgt=seq(-22, 33))
 preddf <- make_groups(preddf, 'income_percap', inc_qs[c(2, 11, 20)])
-preddf <- make_groups(preddf, 'race_black', c(0, .25, 1))
+preddf <- make_groups(preddf, 'race_black', c(0.05, 0.5, 0.95))
 
 preddf$vader <- 1
 
 mm <- model.matrix(as.formula(paste0("vader ~ ", 
                      piece.formula("wbgt", knots[['wbgt']], "income_percap"), ' + ',
-                     piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
+                     #piece.formula("wbgt", knots[['wbgt']], ""), ' + ',
                      piece.formula("wbgt", knots[['wbgt']], "race_black"), ' + ',
                      piece.formula("wbgt", knots[['wbgt']], ""))),
                    data=preddf)
@@ -133,6 +133,10 @@ levels(preddf$race_black) <- preddf$race_black %>%
                                    paste0(., '%')
 
 
+#preddf <- preddf %>%
+#  filter(race_black == '100%')
+
+
 preddf <- preddf %>%
   gather(key, value, -wbgt, -income_percap, -race_black, -vader) %>%
   #Normalize so that graph shows difference from wbgt = X
@@ -146,7 +150,7 @@ preddf <- preddf %>%
 
 (curve <- ggplot(preddf) + 
   geom_line(aes(x=wbgt, y=pred, color=income_percap, linetype=race_black)) + 
-  scale_linetype_manual(values=c('0%'=3, '25%'=2, '100%'=1)) + 
+  scale_linetype_manual(values=c(3, 2, 1)) + 
   geom_ribbon(aes(x=wbgt, ymin=ymin, ymax=ymax, fill=income_percap, grou=race_black), alpha=0.2) + 
   scale_x_continuous(expand=c(0, 0), limits=c(-22, 33)) + 
   scale_fill_manual(values=c("#e31a1c", "#ff7f00", "#33a02c")) + 

### 2021-06-30 doc/temp-sentiment.tex

@@
 %https://www.overleaf.com/project/5f918edd1afc380001feca0d
-\title{Fine-scale Data Reveals Neighborhood Inequalities Effects of Heat on Expressed Mood}
+\title{Fine-scale Data Reveals Neighborhood Inequalities In Effects of Heat on Expressed Mood}
 
@@
 \author[3]{Zheng Liu}
-\author[4]{Jie Liu}
-\author[5]{Portia Adade Williams}
+\author[4]{Portia Adade Williams}
+\author[5]{Jie Liu}
 \author[6]{Aaron Schwartz}
@@
 \affil[3]{Department of Geographical Sciences, University of Maryland College Park}
-\affil[4]{School of Business, East China University of Science and Technology}
-\affil[5]{University of Cape Town}
+\affil[4]{University of Cape Town}
+\affil[5]{School of Business, East China University of Science and Technology}
 \affil[6]{University of Colorado Boulder}
@@
 Higher temperatures associated with climate change are expected to have major impacts on human mental health.
-Indeed, previous work has found strong associations between heat and outcomes such as suicides and mental health related hospitalizations.
+Indeed, previous work has found strong associations between heat and mental health outcomes.
 These studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race at coarse spatial scales.
@@
 We also find that the effect of heat on expressed mood is greatest in the early morning, supporting the hypothesis that heat affects mental health through a sleep quality pathway.
-This paper presents the first evidence of heterogeneities in the vulnerability of mental health to heat.
+This paper presents the first evidence of heterogeneity in the vulnerability of mental health to heat.
 \end{abstract}
@@
 Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, and linkages between higher temperatures and a variety of mental health outcomes \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}.
-Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of heat on mental health, especially on identifying the most vulnerable communities.
+Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of increased heat on mental health \cite{Berry2018Apr}, especially for identifying the most vulnerable communities.
 
@@
 Results from an emerging literature are beginning to demonstrate that higher temperatures can lead to deteriorating mental health.
-Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May}, and other studies find a similar relationship in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
+Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
 Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb}.
@@
 
-The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicate surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as adaptation to future heat stress.
-For example, a large study of temperature and suicide in the USA and Mexico found ``no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
+The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicate surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as potential adaptations to future heat stress.
+For example, a large study of temperature and suicide in the USA and Mexico found "no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
 Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \cite{Mullins2019Dec}.
@@
 
-Findings from large-scale studies of relatively uniform vulnerability to heat are surprising given minority groups and poorer people are more likely to be exposed to undesirable temperatures and otherwise less able to mitigate the effects of higher heat.
-For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}.
-Minority groups and poorer peoples frequently face other systemic and societal exclusions that can limit opportunities for advancement which can limit coping abilities and further aggravate and compound vulnerabilities \cite{hallegatte2016shock}, seomtimes referred to as 'intersectionality' \cite{Kuran2020Nov}.
-Thus, particular groups and conditions like race/ethnicity and density (urbanization) can lead to compounding and differential exposures or vulnerabilities of these groups to extreme events \cite{Cardona2012Jan}.
-For instance, migrant groups are sometimes especially vulnerable to heat waves because language barriers make it difficult or impractical to interpret public safety information, compounded during heat-related events if these groups are prioritize finding employment and housing \cite{Thomas2019Mar}.
-Heterogeneities in vulnerability are also more pronounced within city regions rather than between cities, particularly in the United States where there may be considerable regional variability \cite{Cardona2012Jan}.
-Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical \cite{Belanger2015Mar, Uejio2011Mar} and mental health for other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
-Importantly, different degrees of inequality combine and interact for systems that relate directly to differential levels of social vulnerability, both in normal times and in the context of a disaster \cite{Kuran2020Nov}.
-So there are strong reasons to expect heterogeneities in the impact of heat on mental health outcomes.
-It primarily concerns how intersecting factors be it socioeconomic factors affect individuals who face multiple social inequities, with consequent multiple marginalisation.
+Findings from large-scale studies of relatively uniform vulnerability to heat are surprising given that minority groups and low-income people are more likely to be exposed to undesirable temperatures and less able to mitigate the effects of higher heat.
+For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality or no air-conditioning \cite{Samuelson2020Jun}.
+Many of these factors that expose low-income people to heat may be exacerbated and compounded for racial minorities.
+For example, Black Americans face discrimination in employment \cite{Reid2003, Kang2016, Penner2008} and housing access \cite{Desmond2015, Akbar2019}, forcing them into jobs and homes that may expose them to more heat.
+Additionally, policymakers have historically been less likely to take measures to address environmental hazards in Black neighborhoods \cite{Banzhaf2019}.
+Notably, there is often an intersectionality between different forms of vulnerability and oppression, with income and race likely having a compounding effect on vulnerability to environmental risks \cite{Kuran2020Nov}.
+Heterogeneity in vulnerability is also more pronounced within city regions rather than between cities, particularly in the United States where there can be considerable regional variability \cite{Cardona2012Jan}.
+Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+Thus, there are strong reasons to expect heterogeneity in the impact of heat on mental health outcomes for different populations, particularly where socioeconomic, racial and other social inequities intersect, although there is little evidence of this to date.
 
 While public health data on mental health outcomes is not available at the neighborhood scale, spontaneous, in situ data from the social media platform Twitter can provide an important, inexpensive and timely proxy for mental health at extremely precise spatial and temporal scales.
-Many studies use the mood expressed in social media as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
+Many studies use the mood expressed in social media posts as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
 For example, studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder (PTSD) in individuals even before their formal diagnosis \cite{Reece2017Oct}.
 Similar methods applied to posts on Facebook can predict the onset of depression \cite{Eichstaedt2018Oct}.
-In London, day-to-day changes in mental health indicators derived from the text of Tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
-Across the USA, more positive Tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
-Finally, the mood expressed in Tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
-Thus, posts on Twitter provide an indicator of mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
-We therefore draw on a data set of 243 million geolocated Tweets from across the continental United States and covering the period 2009-2019, each paired local weather and neighborhood conditions, to explore how neighborhood conditions affect the relationship between heat and mental heath.
+In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
+Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
+Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+Thus, posts on Twitter provide an proxy for mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
+To explore how neighborhood conditions affect the relationship between heat and mental heath, we draw on a data set of 243 million geolocated tweets from across the continental United States covering the period 2009-2019, each paired with data on local weather and neighborhood conditions.
 
@@
 \subsection*{Overall Effect}
-We examine the relationship between the mood expressed in Tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
-Controlling for a variety of fixed effects across all Tweets, we find higher temperatures are associated with worsened mood (Fig.
-\ref{fig:wbgt}).
-Mood is highest at 5\textdegree C WBGT (a dry bulb temperature typically around 12\textdegree C/54\textdegree F) with the largest declines between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
+We examine the relationship between the mood expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
+We find higher temperatures are associated with worsened mood (Fig.
+\ref{fig:wbgt}) after controlling for a variety of fixed effects across all tweets.
+Mood is highest at 5\textdegree C WBGT (associated in our data set with a median dry bulb temperature on 12\textdegree C/54\textdegree F) with the largest declines between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
 Mood also declines with colder temperatures (below 5\textdegree C WBGT), although only slightly.
-We estimated mood using the Valence Aware Dictionary for Sentiment Reasoning (VADER) method in our main analysis \cite{gilbert_vader_2014}, although using other methods yield similar results (see Supplement).
+Dis-aggregating our analysis by metropolitan area and using alternative methods to estimate expressed mood all yield similar results (see Supplement).
 
@@
  \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
-  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and expressed mood in Tweets.  As temperatures increase above 5\textdegree C WBGT, mood rapidly declines.}
+ \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and expressed mood in tweets. As temperatures increase above 5\textdegree C WBGT, mood rapidly declines.}
  \label{fig:wbgt}
@@
 
-We find that mean neighborhood income strongly moderates the relationship between temperature and sentiment (Fig.
-\ref{fig:hetero}, panel A), with large differences in sentiment between the poorest and wealthiest neighborhoods.
-As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), sentiment increases in the wealthiest neighborhoods (95th income percentile) but decreases in the median and lowest income neighborhoods.
-The wealthiest neighborhoods do not see decreases in sentiment until temperatures exceed 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point sentiment decreases relatively evenly regardless of neighborhood income percentile.
+We find that mean neighborhood income strongly moderates the relationship between temperature and mood (Fig.
+\ref{fig:hetero}), with large differences in mood between the poorest (5th percentile) and wealthiest (95th percentile) neighborhoods.
+As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), mood increases in the wealthiest neighborhoods but decreases in the median and lowest income neighborhoods.
+The wealthiest neighborhoods do not see decreases in mood until temperatures exceed 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point mood decreases relatively evenly regardless of neighborhood income percentile.
 
-We explored the relationship between neighborhood racial group characteristics and temperature and sentiment and find the effects of heat are felt disproportionately in neighborhoods that are majority Black (see Fig \ref{fig:hetero}).
-Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the sentiment of Tweets in majority Black neighborhoods decrease four times as much as the sentiment of people in other neighborhoods.
-Additionally, at modest to moderate temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower sentiment than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
+We also explored how neighborhood racial characteristics affect the relationship between temperature and mood, and we find the effects of heat are felt disproportionately in neighborhoods that are majority Black (Fig.
+\ref{fig:hetero}).
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the mood of tweets in majority Black neighborhoods decrease four times as much as the mood of people in other neighborhoods.
+Additionally, at mild to warm temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower mood than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
 
@@
  \includegraphics[width=\linewidth]{../res/wbgt_combined.png}
-    \caption{Effect of changes in WBGT on expressed sentiment as moderated by neighborhood income (a) and race (b).}
+ \caption{Effect of changes in WBGT on expressed mood as moderated by neighborhood income (a) and race (b).}
 \label{fig:hetero}
@@
 
-Finally, given that both low-income and majority black neighborhoods are more vulnerable to heat waves, we model the combined effects of race and sentiment together to explore the intersection of race and income.
-We model how the relationship between temperature and sentiment is moderated by the per-capita income of a neighborhood as well as the percentage of a neighborhood's inhabitants that are black.
+Finally, given that both low-income and majority black neighborhoods are more vulnerable to heat waves, we model the combined effects of race and income together to explore the intersection of these two types of vulnerability.
+Specifically, we model how the relationship between temperature and mood is moderated by the interaction between a neighborhood's per-capita income and the percentage of a neighborhood's inhabitants that are black.
 Additionally, we compare AICs for models with no term for vulnerability, a race term for vulnerability, an income term for vulnerability, and both race and income terms for vulnerability.
@@
 \includegraphics[width=\textwidth]{../res/wbgt-income-race.png}
-\captionof{figure}{Relationship between WBGT and sentiment from text in Tweets, moderated by the interaction of race and income.}
+\captionof{figure}{Relationship between WBGT and mood from text in tweets, moderated by the interaction of race and income.}
 \label{fig:race-income}
@@
 \end{tabular}
-\label{tab:aics}
 \captionof{table}{Akaike information criterion (AIC) of models with interaction terms for race, income, both race and income, as well as neither race nor income.
 The model with the best fit by AIC includes both race and income interaction terms.}
+\label{tab:aics}
 \end{minipage}
@@
 A fully black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants.
-Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and sentiment.
+Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and mood.
 Using AIC as an indicator of how different terms affect model quality (see Tab.
-\ref{tab:aics}), we find that a model including both terms performs the best, indicating that both race and income matter for the impact of heat on sentiment.
-Moreover, including only a race term improves the model more than including only an income term, further supporting the conclusion that race is more important than income for explaining the vulnerability of sentiment to heat.
+\ref{tab:aics}), we find that a model including both terms performs the best, indicating that both race and income matter for the impact of heat on mood.
+However, including only a race term improves the model more than including only an income term, further supporting the conclusion that race is more important than income for explaining the vulnerability of mood to heat.
 
-\subsection*{Sentiment Comparison With Other Events}
-We compared the expressed sentiment from exposure to temperatures to two other events associated with impacts on sentiment and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact off a major natural disaster (see Fig.
+\subsection*{Comparison With Other Events}
+We compared the expressed mood from exposure to temperatures to two other events associated with impacts on mood and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact of a major natural disaster (see Fig.
 \ref{fig:compare}).
-For this comparison, we define heat impact as the changes in sentiment associated with temperatures increasing from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F); we define weekly changes as the difference between the mean Saturday sentiment and the mean Monday sentiment, calculated across the entire dataset; and we define the impact of Hurricane Sandy as the change in sentiment from the week before the hurricane made landfall to the week after, calculated for only impacted counties.
-Both of these events are associated with mental health impacts: expressed sentiment on Twitter is highest on Saturdays and lowest on Mondays, a weekly pattern that mirrors the dynamics of suicide rates \cite{CDC2021}, while Hurricane Sandy was associated with mental health effects on victims including increased anxiety, PTSD, and depression \cite{Schwartz2017Aug, Lieberman-Cribbin2017}.
+For this comparison, we define heat impact as the changes in mood associated with temperatures increasing from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F); we define weekly changes as the difference between the mean Saturday mood and the mean Monday mood, calculated across the entire data set; and we define the impact of Hurricane Sandy as the change in mood from the week before the hurricane made landfall to the week after, calculated for only impacted counties.
+Both of these events are associated with mental health impacts: the decrease in expressed mood on Twitter from Saturdays to Mondays mirrors mirrors the weekly dynamics of suicide rates \cite{CDC2021}, while Hurricane Sandy was associated with mental health effects on victims including increased anxiety, PTSD, and depression \cite{Schwartz2017Aug, Lieberman-Cribbin2017}.
 
@@
  \includegraphics[width=0.66\linewidth]{../res/comparison_plot.png}
-  \caption{Decline in sentiment during exposure to change in temperature from lower to higher across several different neighborhoods, compared with other impacts on sentiment, such as major hurricanes, as well as the weekly variation in sentiment from the peak on Saturday to the low on Monday.}
+ \caption{Decline in mood as temperature changes from mild to high across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.}
  \label{fig:compare}
@@
 
-We find that the decline in sentiment in wealthier and majority white neighborhoods from exposure to increasing temperatures is less than the average weekly change in sentiment from Saturday to Monday (see Fig.
+We find that the decline in mood in wealthier and majority white neighborhoods from exposure to increasing temperatures is less than the average weekly change in mood from Saturday to Monday (see Fig.
 \ref{fig:compare}).
-Conversely, the effects of exposure to increasing temperatures in poorer and majority Black neighborhoods are much larger than the average weekly change in sentiment.
-Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on sentiment from experiencing a major hurricane.
+Conversely, the effects of exposure to increasing temperatures in poorer and majority Black neighborhoods are much larger than the average weekly change in mood.
+Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on mood from experiencing a major hurricane.
 
-\subsection{Temperature Effects by Time of Day}
+\subsection*{Temperature Effects by Time of Day}
 In addition to providing data with high spatial resolution, Twitter data also comes with very high temporal resolution as each tweet is time-stamped.
-Thus, we were able to examine the impact of heat on sentiment over the course of the day (See Fig.
+Thus, we were able to examine the impact of heat on mood over the course of the day (Fig.
 \ref{fig:ts-wbgt}).
-We found that heat is associated with improved sentiment for a brief period in the late morning, and then has a consistently negative effect on sentiment during the latter half of the day, from noon until 9pm.
-This effect weakens in the early evening through midnight, and then increases greatly throughout the night.
-Heat has the greatest effect on expressed sentiment at 6am.
+We found that heat is associated with improved mood for a brief period in the late morning, and then has a negative effect on mood during the latter half of the day, with a consistent effect from noon until 9pm.
+This effect weakens in the early evening through midnight, and then increases substantially throughout the night.
+Heat has the greatest effect on expressed mood at 6am, an effect over twice as large as during the day.
 
@@
  \includegraphics[width=0.5\linewidth]{../res/ts_heat.png}
-  \caption{Effects of rising temperatures on sentiment by hour of the day, with a 95\% confidence interval.
-The value shown is the predicted change in VADER score for a 1\textdegree C WBGT increase in temperatures.}
+ \caption{Effects of rising temperatures on mood by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted change in VADER score for a 10\textdegree C WBGT increase in temperatures.}
  \label{fig:ts-wbgt}
@@
 We find that Twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution.
-We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between increasing temperatures and declining sentiment.
-Moreover, we were able to show heterogeneities in vulnerability by neighborhood characteristics, overturning previous findings that found no heterogeneities at the county level \cite{Burke2018Aug, Mullins2019Dec}.
+We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between increasing temperatures and declining mood.
+Moreover, we were able to show heterogeneity in vulnerability by neighborhood characteristics, improving on previous findings that found no heterogeneity at the county level \cite{Burke2018Aug, Mullins2019Dec}.
 
-% Compare weekly changes in sentiment to weekly changes in suicides
-We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a higher temperature of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in sentiment of 0.01.
-This is similar to the degree of change in sentiment over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}.
-Thus, while research on linkages between sentiment and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of sentiment and suicides at some time scales.
-Moreover, like sentiment, suicides are affected by higher temperatures.
-Thus, while Twitter sentiment is a mental health indicator at fine enough spatial and temporal scales to examine neighborhood effects, these changes in sentiment are representative of real human suffering, and are also very likely indicative of medical outcomes like suicides and hospitalizations, although our analyses here do not incorporate these outcomes.
+% Compare weekly changes in mood to weekly changes in suicides
+We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a higher temperature of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in mood of 0.01.
+This is similar to the degree of change in mood over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}.
+Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of mood and suicides at some time scales.
+Moreover, like mood, suicides are affected by higher temperatures.
+While the mood expressed on Twitter is only a proxy for mental health and not a direct measurement, the two phenomena have similar temporal patterns and respond similarly to heat.
+Thus, the changes in expressed mood we observe at neighborhood scales are very likely indicative of mental health outcomes such as suicides and hospitalizations.
 
@@
 % Variations in vulnerability also indicates ADAPTATION
-Previous work conducted with county-scale data found no heterogeneities in mental health vulnerability by income.
-Such analyses have led to conclusions that (1) the mental health effects of climate change will be uniform, least in developed countries, and (2) that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
+Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income.
+Such analyses have led to conclusions that the mental health effects of climate change will be uniform, least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
 For example, Mullins et al.
-conclude that ``high-temperature events will harm mental health everywhere" and that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}.
-Contrary to these analyses, we found large heterogeneities in vulnerability by neighborhood race and income.
+conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}.
+Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income.
 This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the poor and vulnerable.
-More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health can be addressed and mitigated.
+More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health may be addressable.
+Additionally, our results suggest that reducing existing discrimination in housing and employment would contribute narrowing racial gaps in vulnerability and be an important:w
+adaptation strategy.
 
-Our study found disproportionate effects of climate change on identifiable and marginalized populations.
-We however found that majority Black neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
-This is surprising, given that Hispanics can be marginalized in income, language and housing.
-This finding may be because the racial category of Hispanic is more broad and encompasses many more groups with more diverse histories and income levels than Black Americans.
+Counter to previous studies, we find that the highest temperatures disproportionately affect poor and marginalized populations.
+Moreover, we find that majority Black neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
+This is surprising, given that Hispanic peoples can be marginalized in income, language and housing.
+This may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans.
 It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
 Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
-Finally, while race is correlated with income in the US, we found that race contributed to vulnerability to heat, independently of income.
-This suggests that racism and patterns of housing discrimination like the legacy of redlining contribute to making Black Americans more vulnerable to heat, event at similar levels of income.
-Thus, our study shows that, the degree of vulnerability could be related to a category of different dimensions (like combinations of race and housing) and not to a single dimension (race or income).
-This explains how complex vulnerability can be assessed.
-It further concurs with the ‘concept of intersectionality’ where the degree of vulnerability does not depend on a single dimensional attribution but results from a complex relationship between different factors such as a combination of race, housing and income \cite{Kuran2020Nov}.
-
-% Interpret Spatial Analysis
-In our spatial analysis, we found that higher temperatures affect sentiment in city regions across the USA, suggesting that higher baseline temperatures in warmer regions do not necessarily reduce vulnerability.
-Additionally, most city regions exhibited both racial and income inequalities in the vulnerability of mental health to high temperatures, although the Southwest had the weakest inequalities and an occasional inverted effect.
-While unusual patterns of temperature impacts in the Southwest could suggest that temperature inequality is related to environmental factors such as humidity, this is unlikely because we used Wet Bulb Globe Temperature (WBGT) which accounts for how differences in humidity and wind speed affect the human body's ability to cool itself through perspiration.
-Thus, this regional differences in vulnerability are likely more due to cultural or infrastructural differences than to different baseline environmental conditions.
+Finally, while race is correlated with income in the US, we find that neighborhood racial characteristics more directly contributed to vulnerability to heat, independent of income.
+This suggests that racism and patterns of housing discrimination like the legacy of redlining contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups.
+Thus, our study shows that the degree of vulnerability to heat stress is less related to a single category or variable (e.g., race or income) but is more likely related to a variety of complex, intersecting dimensions like combinations of race and housing \cite{Kuran2020Nov}.
 
 % Connect to bigger literature on vulnerability
-This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts, as well as the importance of measuring vulnerability at fine spatial scales.
-Within in relatively wealthy nation of the United States, we find large heterogeneities in vulnerability.
+This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts, as well as the importance of measuring impacts at fine spatial scales.
+Within in relatively wealthy nation of the United States, we find large heterogeneity in vulnerability.
 This adds to the growing literature showing that the impacts of climate change will be highly unequal and disproportionately borne by the poor and marginalized \cite{Thomas2019Mar}.
-Moreover, given that more impoverished neighborhoods in the United States see greater mental health impacts during heatwaves, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much greater \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
+Moreover, given that more impoverished neighborhoods in the United States see greater mental health impacts during heatwaves, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
+
+In addition to exploring heterogeneities in vulnerability, using twitter data lets us explore how the effect of heat varies by time of date and infer possible important pathways in the heat-mental health relationship.
+Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis largely supports this hypothesis.
+We found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
+This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
 
 % Caveats
-% Sentiment is a fuzzy indicator but we have BIG DATA
-While these findings are robust to different model specifications and sentiment metrics, there are some caveats and limitations.
-One limitation of data from Twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
-Moreover, while sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed sentiment in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large data set of 243 million Tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in sentiment in relation to the weather.
+% Mood is a fuzzy indicator but we have BIG DATA
+While these findings are robust to different model specifications and metrics of mood, a limitation of data from Twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
+Moreover, while mood/sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed mood in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
+We overcome this limitation by using an extremely large data set of 243 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in mood in relation to the weather.
 
-One caveat associated with this data is that we were only able to locate the Tweets within the census block where the tweet was sent - we did not to infer where the person sending the Tweets typically lived or where they had been.
+One caveat associated with this data is that we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
 Many low-income and Black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
@@
 A final issue with using Twitter data is that, while Twitter is used by more than one in five Americans, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
-Nevertheless, we found a large volume of Tweets across all neighborhood types.
+Nevertheless, we found a large volume of tweets across all neighborhood types.
 
@@
 \section*{Methods}
-\subsection*{Tweets \& Sentiment}
-We used data from 243 million geo-located English-language Tweets from the continental United States from the years 2009 to 2019.
-The Twitter data consists of publicly posted messages, or Tweets, that are short status messages users post to the platform.
-We only considered users’ original content, thus did not include reTweets in our analysis.
-Additionally, we excluded all Tweets that contained weather-related terms, to ensure that the sentiment expressed in the Tweets was reflective of a user's mental state, and not a commentary on the weather.
+\subsection*{Tweets \& Mood}
+We used data from 243 million geo-located English-language tweets from the continental United States from the years 2009 to 2019.
+The Twitter data consists of publicly posted messages, or tweets, that are short status messages users post to the platform.
+We only considered users’ original content, thus did not include retweets in our analysis.
+Additionally, we excluded all tweets that contained weather-related terms, to ensure that the mood expressed in the tweets was reflective of a user's mental state, and not a commentary on the weather.
 
-Expressed sentiment is a widely used metric to assess mental health and well-being from posts on social media platforms.
-We use the VADER metric to assess the sentiment of Twitter posts as it was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
-The VADER (Valence Aware Dictionary for sentiment Reasoning) sentiment corpus \cite{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to sentiments expressed on social media and includes several features, as:
+Expressed mood is a widely used metric to assess mental health and well-being from posts on social media platforms.
+We use the VADER metric to assess the mood of Twitter posts as it was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
+The VADER (Valence Aware Dictionary for Sentiment Reasoning) corpus \cite{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to moods expressed on social media and includes several features, as:
 
@@
 
-The VADER method yields a value for the mood of a tweet, with a score of 0 for neutral Tweets, a score $> 0$ for positive-mood Tweets and a score $< 0$ for negative-mood Tweets.
-In addition to the VADER method used in the body of this paper, we also conduct our analysis using the Hedonometer and AFINN sentiment analysis methods, with similar results (See Supplement).
+The VADER method yields a value for the mood of a tweet, with a score of 0 for neutral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
+While we ultimately settled on using VADER for assigning sentiment scores, we also conducted analyses using the Hedonometer and AFINN mood analysis methods, with similar results (See Supplement).
 
@@
 This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution, and integrates a large quantity of observation-based and modeled data \cite{xia_continental-scale_2012}.
-For the exact hour and location of each tweet, we extracted temperature, specific humidity, air pressure, total precipitation, shortwave radiation, and wind speed.
+We extracted temperature, specific humidity, air pressure, total precipitation, shortwave radiation, and wind speed for the exact hour and location of each tweet.
 
-Because metrics of apparent temperature that take into account humidity and other factors can better account for the impacts of heat stress on human health and well-being, we calculate the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet.
-WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature would show due to evaporative cooling, and can be estimated given normal temperature, relative humidity, solar radiation, and wind speed.
+We calculated the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet as temperature metrics that account humidity and other factors can better account for the impacts of heat stress on human health and well-being.
+The WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature due to evaporative cooling.
+The WBGT can be estimated given normal temperature, relative humidity, solar radiation, and wind speed.
 Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing \cite{budd2008wet}.
@@
 
-Using temperature, specific humidity, and pressure, we derived relative humidity using methods described by  \cite{bolton_computation_1980}.
-We then calculated the WBGT using the formula described by \cite{heo2019comparison}.
+We derived relative humidity using methods described by Bolton et al.
+\cite{bolton_computation_1980} which use temperature, specific humidity, and pressure.
+We then calculated the WBGT using the formula described by Heo et al.
+\cite{heo2019comparison}.
 
-Throughout the paper, we give the dry bulb temp for reference as the mean value across our data set for a given wet bulb temp.
+Throughout the paper we give the median dry bulb temperature observed at each given wet bulb temp degree across our data set.
+It should be noted, however, that a wide range of dry bulb temperatures can be associated with a given wet bulb globe temperature, depending on humidity, wind speed, and sunlight.
 
-\subsection*{Socio-Econoimc Data}
-We used data from the American Community Survey (ACS) administered by the US Census to estimate income levels and the racial composition of neighborhoods where Tweets were located.
+\subsection*{Socio-Economic Data}
+We used data from the American Community Survey (ACS) administered by the US Census to estimate income levels and the racial composition of neighborhoods where tweets were located.
 Data was at the level of the census block group, the smallest unit for which the US Census releases public data.
@@
 
-ACS data is released to cover five-year periods.
+The ACS data covers five-year periods.
 We therefore matched each tweet with census block group data from the year at the middle year of each survey's five-year range.
-For example, Tweets from 2014 were matched to data from the 2012-2016 ACS.
-Because the most recent available ACS was from 2014-2018, all Tweets from year years greater than 2016 were matched to this dataset.
-Data was downloaded from the IPUMS NHGIS service provided by the University of Minnesota.
+For example, tweets from 2014 were matched to data from the 2012-2016 ACS.
+Because the most recent available ACS was from 2014-2018, all tweets from year years greater than 2016 were matched to this dataset.
+We used data from the Integrated Public Use Microdata Series (IPUMS USA) National Historical Geographic Information System (NHGIS) service provided by the University of Minnesota \cite{ruggles2018ipums}.
 
-Mean annual income per capita is provided by the ACS, and we standardized this variable so that the values for each year were in 2019 dollars.
+We use the mean annual income per capita from the ACS, standardized so that values for each year are in 2019 dollars.
 For racial categories, we combined the various categories provided by the ACS into four racial groups: non-Hispanic white, non-Hispanic Black, Hispanic of any race, and an "other" category for non-Hispanic people who were neither Black nor white, such as Asian, Native American, or mixed-race people.
 
-In this paper, we give income and racial percentiles based on the observed percentiles across all Tweets in our data set.
+We present income and racial percentiles based on the observed percentiles across all tweets in our data set.
 
 \subsection*{Modeling}
-We assessed how expressed sentiment is affected by Wet Bulb Globe Temperature using segmented regressions, controlling for precipitation, shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: the day of the week, the time of day, the day of the year, the year, the month, and the county.
+We assessed how expressed mood is affected by Wet Bulb Globe Temperature using segmented regressions, controlling for precipitation, shortwave solar radiation (sunshine), as well as the following spatio-temporal fixed effects: day of the week, time of day, day of the year, year, month, and county.
 
@@
 
-Where $y$ is the sentiment of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represents the segmented effect for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
+Where $y$ is the mood of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represents the segmented effect for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
 We estimate the 95\% confidence interval of all our models using 80-fold bootstrapping.
 
-To examine how income and racial groups moderate the effect of heat on sentiment, we extend our model to the following form:
+To examine how income and racial groups moderate the effect of heat on mood, we extend our model to the following form:
 
@@
 Where $m$ is either the log-transformed average income in the census block where the tweet originated, or a dummy variable for the four racial categories.
-We specify alternative models where income is a dummy variable in three bins, and were race is a continuous variable for the percent of a census block population that is white These alternative specifications yield similar results.
-We also examine the effects of rainfall and sunshine on sentiment by income and race.
+We specify alternative models where income is a dummy variable in three bins and were race is a continuous variable for the percent of a census block population that is white These alternative specifications yield similar results.
+We also examine the effects of rainfall and sunshine on mood by income and race.
 These alternative specifications and analysis of rainfall and sunshine effects are available in the Supplement.
 
-In addition to our models exploring how income and race moderate the effect of heat on sentiment individually, we fit an additional model with an interaction terms for both race and income.
+In addition to our models exploring how income and race moderate the effect of heat on mood individually, we fit an additional model with an interaction terms for both race and income.
 Due to the complexity of a model with segmented terms interacting with two different variables, using a categorical variable for race would have yielded a model matrix too large for the largest cloud virtual machine available to us.
@@
 
-Finally, we fit a model for how heat affects sentiment by time of day.
+Finally, we fit a model for how heat affects mood by time of day.
 For this analysis, we used a varying-coefficient model, where the effect of heat is linear, but the coefficient for the effect varies non-linearly as a function of the time of day.
-Because we modeled the effect of heat as linear, we excluded all tweets with an observed temperatures less than 5\textdegree C WBGT so that we were only including values at which our previous analyses suggested the heat-sentiment relationship is monotonic.
+Because we modeled the effect of heat as linear, we excluded all tweets with an observed temperatures less than 5\textdegree C WBGT so that we were only including values at which our previous analyses suggested the heat-mood relationship is monotonic.
 This final model took the following form:
@@
 This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
-Additionally, computational resources for this project were provided for the Microsoft Azure cloud by Microsoft's AI for Earth program.
+Additionally, computational resources for this project were provided for the Microsoft Azure cloud through Microsoft's AI for Earth program.
 
 This data came from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
-The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and we have complied with all the terms of service for Twitter and UVM.
+The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and this work complies with all the terms of service for Twitter and UVM.
 
@@
 and J.O.-G.
-analyzed data; and M.C., J.O.-G., J.L., and P.A.W.
+analyzed data; and M.C., J.O.-G., P.A.W., and J.L.
 wrote the paper.

### 2021-06-30 doc/temp-sentiment.tex
Updates from Overleaf

@@
 \begin{abstract}
-%Limit 150 words: https://www.nature.com/nclimate/about/content
+%Limit 150 words:
+
 %Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
@@
 However, using expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status, we show that there are stark differences in the effects of heat depending on neighborhood characteristics.
-We find that increased temperatures worsen expressed mood in all areas, but that this effect is much stronger in poor and Black neighborhoods.
+Here we show that increased temperatures worsen expressed mood in all areas, but that this effect is much stronger in poor and Black neighborhoods.
 We also find that the effect of heat on expressed mood is greatest in the early morning, supporting the hypothesis that heat affects mental health through a sleep quality pathway.
@@
 %Empirical: Heat and mental health
-Results from an emerging literature are beginning to demonstrate that higher temperatures can lead to deteriorating mental health.
+Drawing on these frameworks, results from an emerging literature are beginning to document linkages between higher temperatures worsened mental health outcomes.
 Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
@@
 
-Findings from large-scale studies of relatively uniform vulnerability to heat are surprising given that minority groups and low-income people are more likely to be exposed to undesirable temperatures and less able to mitigate the effects of higher heat.
-For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality or no air-conditioning \cite{Samuelson2020Jun}.
+These findings of uniform vulnerability to heat are surprising given that minority groups and low-income people are more likely to be exposed to undesirable temperatures and less able to mitigate the effects of higher heat.
+For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}.
 Many of these factors that expose low-income people to heat may be exacerbated and compounded for racial minorities.
@@
 Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
-Thus, there are strong reasons to expect heterogeneity in the impact of heat on mental health outcomes for different populations, particularly where socioeconomic, racial and other social inequities intersect, although there is little evidence of this to date.
+Thus, there are theoretical reasons to expect heterogeneity in the impact of heat on mental health outcomes for different populations, particularly where socioeconomic, racial and other social inequities intersect, although there is little evidence of this to date.
 
@@
 
-\section*{Results}
-
-\subsection*{Overall Effect}
+\section*{Effect of Heat on Sentiment}
 We examine the relationship between the mood expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
@@
 
-\subsection*{Comparison With Other Events}
+\section*{Comparison With Other Events}
 We compared the expressed mood from exposure to temperatures to two other events associated with impacts on mood and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact of a major natural disaster (see Fig.
@@
 
-\subsection*{Temperature Effects by Time of Day}
+\section*{Effect of Heat by Time of Day}
 In addition to providing data with high spatial resolution, Twitter data also comes with very high temporal resolution as each tweet is time-stamped.
@@
 \section*{Methods}
+% from NCC: methods section should be written as concisely as possible and "typically do not exceed 3,000 words"
 \subsection*{Tweets \& Mood}

### 2021-06-30 doc/temp-sentiment.tex
more tweaks from overleaf

@@
 \affil[3]{Department of Geographical Sciences, University of Maryland College Park}
-\affil[4]{University of Cape Town}
+\affil[4]{CSIR-Science and Technology Policy Research Institute, Ghana}
 \affil[5]{School of Business, East China University of Science and Technology}
@@
 %Theoretical: frameworks and pathways
-Researchers have proposed a number of frameworks that focus on pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr, BerryETAL2010}.
+Researchers have proposed a number pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr, BerryETAL2010}.
 Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul, sadiq_impact_2019}.
-Additionally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}.
 Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007, WHO2007}.
-Finally, exposure to higher temperatures can affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health outcomes \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
+Additionally, exposure to higher temperatures can affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health outcomes \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
 
@@
 Higher temperatures are also linked to increased mortality among people with pre-existing mental health disorders \cite{Hansen2008Oct}.
+Finally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}, although these studies have relied on retrospective survey data and have not used data from across the diurnal cycle.
 
@@
 In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
-Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}.
+Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}, and twitter has provided insight into population-level patterns of happiness over the course of the day \cite{Dodds2011, Dzogang2017}.
 Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
@@
 \hline
-Neither & 277270561 \\
-Income & 277182239 \\
-Race & 277143579 \\
-Both & 277108676 \\
+Neither & 277,270,561 \\
+Income & 277,182,239 \\
+Race & 277,143,579 \\
+Both & 277,108,676 \\
 \hline
@@
 \ref{fig:race-income}).
-A fully black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants.
+A Black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants.
 Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and mood.
@@
  \includegraphics[width=0.66\linewidth]{../res/comparison_plot.png}
- \caption{Decline in mood as temperature changes from mild to high across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.}
+ \caption{Decline in mood as temperature changes from mild (5\textdegree C WBGT) high (25\textdegree C WBGT) across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.}
  \label{fig:compare}
@@
 
-One caveat associated with this data is that we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been.
-Many low-income and Black people commute during the day to work in the service sector in higher-income areas, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
+One caveat associated with this data is that we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
 Moreover, in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are typically found that census block.
@@
 Again, these issues mean we are likely under-estimating the true effect of heat on mental health for poor and Black people.
-A final issue with using Twitter data is that, while Twitter is used by more than one in five Americans, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
+A final issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
 Nevertheless, we found a large volume of tweets across all neighborhood types.
@@
 
-Where $y$ is the mood of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the income shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represents the segmented effect for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
+Where $y$ is the mood of a tweet, $t$ is the wet bulb globe temperature at the hour of the tweet, $p$ is a binary variable indicating whether it rained at the hour of the tweet, $s$ is the incoming shortwave radiation, or sunshine, in $W/m^2$, at the hour of the tweet, $\Phi$ is the spatio-temporal fixed effects, $\epsilon$ is the normally-distributed errors, and $f_t$ represents the segmented effect for $t$, with knots at -10\textdegree, 0\textdegree, 5\textdegree, 10\textdegree, 15\textdegree, 20\textdegree, and 25\textdegree C WBGT.
 We estimate the 95\% confidence interval of all our models using 80-fold bootstrapping.

### 2021-06-30 viz/afinn/wbgt_income_bootstrap.R
move figure legends for supplement

@@ -142,7 +142,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block\nIncome Per-Capita\n(Percentile)',
        color = 'Census Block\nIncome Per-Capita\n(Percentile)') +
   theme_classic() + 
-  theme(legend.position=c(0.3, 0.4),
+  theme(#legend.position=c(0.3, 0.4),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -172,4 +172,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/afinn-wbgt-income.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/afinn-wbgt-income.png', width=5.5, height=4)

### 2021-06-30 viz/afinn/wbgt_race_q_bootstrap.R

@@ -138,7 +138,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block\nMajority Race',
        color = 'Census Block\nMajority Race') +
   theme_classic() + 
-  theme(legend.position=c(0.3, 0.4),
+  theme(#legend.position=c(0.3, 0.4),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -168,4 +168,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/afinn-wbgt-race_q.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/afinn-wbgt-race_q.png', width=5.5, height=4)

### 2021-06-30 viz/hedono/wbgt_income_bootstrap.R

@@ -142,7 +142,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block\nIncome Per-Capita\n(Percentile)',
        color = 'Census Block\nIncome Per-Capita\n(Percentile)') +
   theme_classic() + 
-  theme(legend.position=c(0.3, 0.4),
+  theme(#legend.position=c(0.3, 0.4),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -172,4 +172,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/hedono-wbgt-income.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/hedono-wbgt-income.png', width=5.5, height=4)

### 2021-06-30 viz/hedono/wbgt_race_q_bootstrap.R

@@ -138,7 +138,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block\nMajority Race',
        color = 'Census Block\nMajority Race') +
   theme_classic() + 
-  theme(legend.position=c(0.3, 0.4),
+  theme(#legend.position=c(0.3, 0.4),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -168,4 +168,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/hedono-wbgt-race_q.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/hedono-wbgt-race_q.png', width=5.5, height=4)

### 2021-06-30 viz/vader/wbgt_income_bootstrap.R

@@ -306,7 +306,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block\nIncome Per-Capita\n(Percentile)',
        color = 'Census Block\nIncome Per-Capita\n(Percentile)') +
   theme_classic() + 
-  theme(legend.position=c(0.8, 0.2),
+  theme(#legend.position=c(0.8, 0.2),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -327,4 +327,4 @@ hist <- ggplot(data[sample(1:nrow(data), nrow(data)*0.01), 'srad']) +
 
 plot_grid(curve, hist, align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2))
-ggsave('~/temp-sentiment/res/srad-income.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/srad-income.png', width=5.5, height=4)

### 2021-06-30 viz/vader/wbgt_income_q_bootstrap.R

@@ -143,7 +143,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block\nIncome Per-Capita\nTercile',
        color = 'Census Block\nIncome Per-Capita\nTercile') +
   theme_classic() + 
-  theme(legend.position=c(0.3, 0.4),
+  theme(#legend.position=c(0.3, 0.4),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -173,4 +173,4 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt-income_q.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/wbgt-income_q.png', width=5.5, height=4)

### 2021-06-30 viz/vader/wbgt_race_bootstrap.R

@@ -141,7 +141,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block\nPercent Minority',
        color = 'Census Block\nPercent Minority') +
   theme_classic() + 
-  theme(legend.position=c(0.3, 0.4),
+  theme(#legend.position=c(0.3, 0.4),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -171,5 +171,5 @@ xl <- get_plot_component(x2, "xlab-b")
 
 plot_grid(curve, hist, ggdraw(x), ggdraw(xl), align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2, 0.04, 0.04))
-ggsave('~/temp-sentiment/res/wbgt-race.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/wbgt-race.png', width=5.5, height=4)
 

### 2021-06-30 viz/vader/wbgt_race_q_bootstrap.R

@@ -282,7 +282,7 @@ curve <- ggplot(preddf) +
        fill = 'Census Block Majority Race',
        color = 'Census Block Majority Race') +
   theme_classic() + 
-  theme(legend.position=c(0.2, 0.8),
+  theme(#legend.position=c(0.2, 0.8),
         axis.text.x = element_blank(),
         axis.ticks.x = element_blank(),
         plot.margin = unit(c(0.025, 0.025, -1, 0.025), "cm"),
@@ -303,4 +303,4 @@ hist <- ggplot(data[sample(1:nrow(data), nrow(data)*0.01), 'srad']) +
 
 plot_grid(curve, hist, align='v', axis='rl', ncol=1, 
           rel_heights=c(0.8, 0.2))
-ggsave('~/temp-sentiment/res/srad-race_q.png', width=4.5, height=4)
+ggsave('~/temp-sentiment/res/srad-race_q.png', width=5.5, height=4)

### 2021-06-30 doc/temp-sentiment-summary.tex

\documentclass[fleqn,10pt]{wlscirep}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{graphicx}
\usepackage{hyperref}
\usepackage{breakcites}
\usepackage{float}
\usepackage{textcomp}
\usepackage{amsmath}
\usepackage{textgreek}
\usepackage{authblk}
\usepackage{rotating}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{pdflscape}
\usepackage{lineno}
\usepackage{xcolor}
% \usepackage[
% style=numeric,
% citestyle=numeric-comp,
% backend=biber,
% doi=true,
% natbib=true,
% sorting=none
% ]{biblatex}
\usepackage{scalerel,xparse}

\title{Fine-scale Data Reveals Neighborhood Inequalities In Effects of Heat on Expressed Mood}

\author[1,*]{Matthew Cooper}
\author[2]{Jeremiah Osborne-Gowey}
\author[3]{Zheng Liu}
\author[4]{Portia Adade Williams}
\author[5]{Jie Liu}
\author[6]{Aaron Schwartz}
\author[7]{Patrick Baylis}

\affil[1]{T.H. Chan School of Public Health, Harvard University}
\affil[2]{Environmental Studies Program, University of Colorado Boulder}
\affil[3]{Department of Geographical Sciences, University of Maryland College Park}
\affil[4]{University of Cape Town}
\affil[5]{School of Business, East China University of Science and Technology}
\affil[6]{University of Colorado Boulder}
\affil[7]{University of British Columbia}
\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}

\begin{abstract}
%Summary paragraph word limit "ideally ideally of no more than 200 words" unless the editor finds the summary is compellingly enhanced by the inclusion of the broader perspective concluding sentences: https://www.nature.com/nature/for-authors/formatting-guide

% This summary paragraph should be structured as follows: 2-3 sentences of basic-level introduction to the field; a brief account of the background and rationale of the work; a statement of the main conclusions (introduced by the phrase 'Here we show' or its equivalent); and finally, 2-3 sentences putting the main findings into general context so it is clear how the results described in the paper have moved the field forwards.

% See summary paragraph example from NCC: https://www.nature.com/documents/nature-summary-paragraph.pdf

Previous research indicate strong associations between heat and human mental health \cite{Berry2018Apr, hayes_climate_2018}. Mental health represents a major share of global human suffering, accounting for up to 13\% of the total global burden of disease \cite{Collins2011Jul}. Higher temperatures associated with climate change are expected to have major impacts on mental health with researchers and medical experts sounding a clarion call for more research into potential linkages between climate change and mental health outcomes \cite{Collins2011Jul, Cardona2012Jan, Berry2018Apr}. Researchers have started responding to these calls finding associations between various environmental conditions a variety of mental health outcomes \cite{Waite2017Dec, Raker2019Dec, Cunsolo2018Apr, baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}. Yet there remains a relative paucity of empirical studies in this space. What studies exist typically use relatively coarse-scale data (e.g., city or county level) and, to date, have found no differential vulnerability for changing levels of income or race. Here, and in stark contrast to previous studies, we use finer-scale census block and climate data linked with spontaneous, in-situ observations of mood as expressed in a quarter of a billion geo-located posts on Twitter from 2009-2018, and find strong heterogeneity between heat exposure and mood with the poorest and highest proportion Black neighborhoods disproportionately negatively affected by higher temperatures. These results demonstrate some of the first and strongest to date linkages between increasing temperatures and deteriorating mental health vulnerability, as expressed through spontaneous posts about mood, and the first study that demonstrates heterogeneous effects across various income and racial and ethnic categories, particularly at finer spatial scales than previous reported. We anticipate these findings will help address the needs and risks of the most vulnerable groups and align these with research policy agendas and mental health treatment programs.
\end{abstract}

\begin{document}

\end{document}

### 2021-06-30 doc/temp-sentiment.tex

@@
 %Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
-Higher temperatures associated with climate change are expected to have major impacts on human mental health.
-Indeed, previous work has found strong associations between heat and mental health outcomes.
-These studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race at coarse spatial scales.
-However, using expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status, we show that there are stark differences in the effects of heat depending on neighborhood characteristics.
+Higher temperatures associated with climate change are expected to have major impacts on human mental health, and previous work has found strong associations between heat and mental well-being.
+However, these studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race.
+We therefore use expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status available at fine spatial and temporal scales, and find stark differences in the effects of heat depending on neighborhood characteristics.
 Here we show that increased temperatures worsen expressed mood in all areas, but that this effect is much stronger in poor and Black neighborhoods.
 We also find that the effect of heat on expressed mood is greatest in the early morning, supporting the hypothesis that heat affects mental health through a sleep quality pathway.
-This paper presents the first evidence of heterogeneity in the vulnerability of mental health to heat.
+This paper presents the first evidence of neighborhood heterogeneity in the vulnerability of mental health to heat.
 \end{abstract}
@@
 A large body of research has explored the ongoing impacts of climate change on agricultural outcomes, economic growth, and physical health \cite{pachauri2014climate}.
-However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr, hayes_climate_2018}.
+However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr}.
 While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \cite{Berry2018Apr, Collins2011Jul}.
-Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Waite2017Dec, Raker2019Dec}, deteriorating environmental conditions and a sense of "ecological grief" \cite{Cunsolo2018Apr}, and linkages between higher temperatures and a variety of mental health outcomes \cite{baylis_weather_2018, Mullins2019Dec, Li2020Mar, Obradovich2018Oct}.
-Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of increased heat on mental health \cite{Berry2018Apr}, especially for identifying the most vulnerable communities.
+Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Schwartz2017Aug}, as well as linkages between higher temperatures and a variety of mental health outcomes \cite{baylis_weather_2018, Mullins2019Dec, Obradovich2018Oct}.
+Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of increased heat on mental health, especially for identifying the most vulnerable communities.
 
 %Theoretical: frameworks and pathways
-Researchers have proposed a number pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr, BerryETAL2010}.
-Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul, sadiq_impact_2019}.
-Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007, WHO2007}.
-Additionally, exposure to higher temperatures can affect productivity and income \cite{kjellstrom_impact_2016, Burke2015Nov}, leading to second-order impacts on mental health outcomes \cite{Katz1997, CohnETAL2004, BouchamaETAL2007}.
+Researchers have proposed a number pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr}.
+Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.
+Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007}.
+Additionally, exposure to higher temperatures can affect productivity and income \cite{Burke2015Nov}, leading to second-order impacts on mental health outcomes.
 
@@
 Drawing on these frameworks, results from an emerging literature are beginning to document linkages between higher temperatures worsened mental health outcomes.
-Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \cite{Qi2014Dec, Page2007Aug, Likhvar2011Jan}.
+Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \cite{Qi2014Dec, Likhvar2011Jan}.
 Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb}.
@@
 
-The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicate surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as potential adaptations to future heat stress.
+The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicates surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as potential adaptations to future heat stress.
 For example, a large study of temperature and suicide in the USA and Mexico found "no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
@@
 Many of these factors that expose low-income people to heat may be exacerbated and compounded for racial minorities.
-For example, Black Americans face discrimination in employment \cite{Reid2003, Kang2016, Penner2008} and housing access \cite{Desmond2015, Akbar2019}, forcing them into jobs and homes that may expose them to more heat.
+For example, Black Americans face discrimination in employment \cite{Kang2016, Penner2008} and housing access \cite{Desmond2015, Akbar2019}, forcing them into jobs and homes that may expose them to more heat.
 Additionally, policymakers have historically been less likely to take measures to address environmental hazards in Black neighborhoods \cite{Banzhaf2019}.
-Notably, there is often an intersectionality between different forms of vulnerability and oppression, with income and race likely having a compounding effect on vulnerability to environmental risks \cite{Kuran2020Nov}.
-Heterogeneity in vulnerability is also more pronounced within city regions rather than between cities, particularly in the United States where there can be considerable regional variability \cite{Cardona2012Jan}.
 Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
@@
 In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
-Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}, and twitter has provided insight into population-level patterns of happiness over the course of the day \cite{Dodds2011, Dzogang2017}.
-Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018, hannak_tweetin_2012} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}, and twitter has provided insight into population-level patterns of happiness over the course of the day \cite{Dodds2011}.
+Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
 Thus, posts on Twitter provide an proxy for mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
@@
 \section*{Discussion}
-% Benefits of Twitter data/overview of findings
-We find that Twitter data offers significant advantages for observing environmental effects on human well-being due to its very fine spatio-temporal resolution.
-We were able to pair each tweet with the local temperature at the time of the tweet, and found a clear association between increasing temperatures and declining mood.
-Moreover, we were able to show heterogeneity in vulnerability by neighborhood characteristics, improving on previous findings that found no heterogeneity at the county level \cite{Burke2018Aug, Mullins2019Dec}.
-
 % Compare weekly changes in mood to weekly changes in suicides
-We found that the change from optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to a higher temperature of 25\textdegree C WBGT (36\textdegree C/97\textdegree F) is associated with an overall decrease in mood of 0.01.
-This is similar to the degree of change in mood over the course of a week from a Monday nadir (0.1268) to a Saturday peak (0.1372), a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}.
-Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of mood and suicides at some time scales.
-Moreover, like mood, suicides are affected by higher temperatures.
-While the mood expressed on Twitter is only a proxy for mental health and not a direct measurement, the two phenomena have similar temporal patterns and respond similarly to heat.
-Thus, the changes in expressed mood we observe at neighborhood scales are very likely indicative of mental health outcomes such as suicides and hospitalizations.
+We found that the change from an optimum temperature to a higher temperature is associated with an overall decrease in mood similar in magnitude to the degree of change in mood from Monday to Saturday, a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}, and that the same level of heat in a poor or Black neighborhood is associated with a change in mood nearly matching the impact of a major Hurricane, an event associated with suicides and PTSD \cite{Schwartz2017Aug, Lieberman-Cribbin2017}.
+Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of expressed mood and mental health at specific time scales and following certain events.
+This suggests that the changes in expressed mood we observe at neighborhood scales is indicative of other mental health outcomes such as suicides and hospitalizations.
 
@@
 Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income.
-This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the poor and vulnerable.
+This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the vulnerable and marginalized.
 More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health may be addressable.
-Additionally, our results suggest that reducing existing discrimination in housing and employment would contribute narrowing racial gaps in vulnerability and be an important:w
-adaptation strategy.
 
-Counter to previous studies, we find that the highest temperatures disproportionately affect poor and marginalized populations.
-Moreover, we find that majority Black neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
-This is surprising, given that Hispanic peoples can be marginalized in income, language and housing.
-This may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans.
+Examining the role of race in vulnerability, we find that majority Black non-Hispanic neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
+Surprisingly, we did not find an effect for people with Hispanic ethnicity, given that Hispanic peoples can be marginalized in employment and housing.
+These findings may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans.
 It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
-Additionally, there are many other marginalized groups in America, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
-Finally, while race is correlated with income in the US, we find that neighborhood racial characteristics more directly contributed to vulnerability to heat, independent of income.
-This suggests that racism and patterns of housing discrimination like the legacy of redlining contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups.
-Thus, our study shows that the degree of vulnerability to heat stress is less related to a single category or variable (e.g., race or income) but is more likely related to a variety of complex, intersecting dimensions like combinations of race and housing \cite{Kuran2020Nov}.
+Additionally, there are many other marginalized groups in the United States, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
+Examining the effects of both race and income, we find that race is more important for vulnerability than income.
+This probably due in part to the fact that Black Americans have historically been forced through redlining into neighborhoods with fewer trees that are much more impacted by heat \cite{Hoffman2020, Locke2021}.
+This suggests that racism and patterns of housing discrimination contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups, and that reducing existing discrimination in housing and employment is an important adaptation strategy.
 
 % Connect to bigger literature on vulnerability
-This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts, as well as the importance of measuring impacts at fine spatial scales.
-Within in relatively wealthy nation of the United States, we find large heterogeneity in vulnerability.
-This adds to the growing literature showing that the impacts of climate change will be highly unequal and disproportionately borne by the poor and marginalized \cite{Thomas2019Mar}.
-Moreover, given that more impoverished neighborhoods in the United States see greater mental health impacts during heatwaves, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
+This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts.
+Additionally, it shows how measuring outcomes at fine spatial scales necessary for understanding how factors like race and income matter for vulnerability \cite{Tong2021}.
+Even within a relatively wealthy nation, we find that high temperatures affect certain populations to the same degree as a major hurricane.
+Thus, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
 
@@
 % Mood is a fuzzy indicator but we have BIG DATA
-While these findings are robust to different model specifications and metrics of mood, a limitation of data from Twitter is that a tweet is not a precise measurement of a discrete mental health outcome and only provides a rough indicator a user's mental state based on the vocabulary in the tweet.
-Moreover, while mood/sentiment analysis algorithms have gotten increasingly sophisticated at estimating the mood in a body of text, a user's expressed mood in a tweet is highly variable and is mostly affected by factors like current events and the user's personal life, with local weather conditions having only a small impact.
-We overcome this limitation by using an extremely large data set of 243 million tweets, allowing us to control for a wide variety of spatial and temporal effects to estimate population-level changes in mood in relation to the weather.
-
-One caveat associated with this data is that we were only able to locate the tweets within the census block where the tweet was sent - we did not to infer where the person sending the tweets typically lived or where they had been, so these results may in fact under-estimate the impacts of higher temperatures on mental health for poor and minority individuals.
-Moreover, in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are typically found that census block.
+While these findings are robust to different model specifications and metrics of mood, there are important considerations in this analysis.
+One caveat associated with this data is that we were only able to locate the tweets within the census block from which the tweet was sent, and in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are found that census block.
 For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
-Additionally, neighborhoods of young college students are also typically estimated to have low income levels, even through college students are wealthier than the average American.
-Again, these issues mean we are likely under-estimating the true effect of heat on mental health for poor and Black people.
-A final issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
+These issue may in fact lead us to under-estimate the true effect of heat on mental health for poor and Black people.
+Another issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
 Nevertheless, we found a large volume of tweets across all neighborhood types.
@@
 % NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
-While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how highly unevenly these impacts will be distributed.
+While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how unevenly these impacts will be distributed.
 People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \cite{bullard2012wrong}.
@@
 
-We calculated the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet as temperature metrics that account humidity and other factors can better account for the impacts of heat stress on human health and well-being.
+We calculated the Wet Bulb Globe Temperature (WBGT) at the time and location of each tweet.
 The WBGT is the temperature that a wet globe thermometer would read in direct sunlight, and gives a reading lower than a dry bulb temperature due to evaporative cooling.
@@
 \cite{heo2019comparison}.
-
 Throughout the paper we give the median dry bulb temperature observed at each given wet bulb temp degree across our data set.
@@
 
-We present income and racial percentiles based on the observed percentiles across all tweets in our data set.
+We present income and percentiles based on the observed percentiles across all tweets in our data set.
 
@@
 Thus, we used a continuous metric for both income as well as race (percent black).
-This model extends our main model to the following form:
+This model extends our main model to the following form, interacting income and race with a temperature segmented spline as well as precipitation in sunshine:
 
 \begin{equation}
- y = \begin{tabular}{l}
- $\beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) + 
- \beta_p p + \beta_{m_{r} p} m_{r} p + 
- \beta_{m_{i} p} m_{i} p + \beta_{m_{i} m_{r} p} m_{i} m_{r} p + $
- \\
- $\beta_s s + \beta_{m_{i} s} m_{i} s + 
- \beta_{m_{r} s} m_{r} s + \beta_{m_{i} m_{r} s} m_{i} m_{r} s + 
- \Phi + \epsilon$
- \end{tabular}
+ y = \beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) +  \beta_p p + \beta_{m_{r} p} m_{r} p + \beta_{m_{i} p} m_{i} p + \beta_s s + \beta_{m_{i} s} m_{i} s +  \beta_{m_{r} s} m_{r} s + 
+ \Phi + \epsilon
  \label{mod:race-inc}

### 2021-07-01 doc/temp-sentiment-supplement.tex
submitted versions of files

@@
   \includegraphics[width=\linewidth]{../res/map_wbgt.png}
-  \label{fig:map}
     \caption{The change in mood for a 25-degree increase in WBGT for city regions with a population of over 1 million.
 For cities shown with a dark outline, the effect is significant at a 95\% confidence interval.}
+  \label{fig:map}
 \end{figure}
@@
 \section{Weather Terms}
-For all of our analysis, we excluded tweets that contained the following weather-related terms: aerovane, air, airstream, altocumulus, altostratus, anemometer, anemometers, anticyclone, anticyclones, arctic, arid, aridity, atmosphere, atmospheric, autumn, autumnal, balmy, baroclinic, barometer, barometers, barometric, blizzard, blizzards, blustering, blustery, blustery, breeze, breezes, breezy, brisk, calm, celsius, chill, chilled, chillier, chilliest, chilly, chinook, cirrocumulus, cirrostratus, cirrus, climate, climates, cloud, cloudburst, cloudbursts, cloudier, cloudiest, clouds, cloudy, cold, colder, coldest, condensation, contrail, contrails, cool, cooled, cooling, cools, cumulonimbus, cumulus, cyclone, cyclones, damp, damp, damper, damper, dampest, dampest, degree, degrees, deluge, dew, dews, dewy, doppler, downburst, downbursts, downdraft, downdrafts, downpour, downpours, dried, drier, dries, driest, drizzle, drizzled, drizzles, drizzly, drought, droughts, dry, dryline, fall, farenheit, flood, flooded, flooding, floods, flurries, flurry, fog, fogbow, fogbows, fogged, fogging, foggy, fogs, forecast, forecasted, forecasting, forecasts, freeze, freezes, freezing, frigid, frost, frostier, frostiest, frosts, frosty, froze, frozen, gale, gales, galoshes, gust, gusting, gusts, gusty, haboob, haboobs, hail, hailed, hailing, hails, haze, hazes, hazy, heat, heated, heating, heats, hoarfrost, hot, hotter, hottest, humid, humidity, hurricane, hurricanes, ice, iced, ices, icing, icy, inclement, landspout, landspouts, lightning, lightnings, macroburst, macrobursts, maelstrom, mercury, meteorologic, meteorologist, meteorologists, meteorology, microburst, microbursts, microclimate, microclimates, millibar, millibars, mist, misted, mists, misty, moist, moisture, monsoon, monsoons, mugginess, muggy, nexrad, nippy, NOAA, nor,’easter, nor,’easters, noreaster, noreasters, overcast, ozone, parched, parching, pollen, precipitate, precipitated, precipitates, precipitating, precipitation, psychrometer, radar, rain, rainboots, rainbow, rainbows, raincoat, raincoats, rained, rainfall, rainier, rainiest, raining, rains, rainy, sandstorm, sandstorms, scorcher, scorching, searing, shower, showering, showers, skiff, sleet, slicker, slickers, slush, slushy, smog, smoggier, smoggiest, smoggy, snow, snowed, snowier, snowiest, snowing, snowmageddon, snowpocalypse, snows, snowy, spring, sprinkle, sprinkles, sprinkling, squall, squalls, squally, storm, stormed, stormier, stormiest, storming, storms, stormy, stratocumulus, stratus, subtropical, summer, summery, sun, sunnier, sunniest, sunny, temperate, temperature, tempest, thaw, thawed, thawing, thaws, thermometer, thunder, thundered, thundering, thunders, thunderstorm, thunderstorms, tornadic, tornado, tornadoes, tropical, troposphere, tsunami, turbulent, twister, twisters, typhoon, typhoons, umbrella, umbrellas, vane, warm, warmed, warming, warms, warmth, waterspout, waterspouts, weather, wet, wetter, wettest, wind, windchill, windchills, windier, windiest, windspeed, windy, winter, wintery, and wintry.
+For all of our analysis, we excluded tweets that contained the following weather-related terms: aerovane, air, airstream, altocumulus, altostratus, anemometer, anemometers, anticyclone, anticyclones, arctic, arid, aridity, atmosphere, atmospheric, autumn, autumnal, balmy, baroclinic, barometer, barometers, barometric, blizzard, blizzards, blustering, blustery, blustery, breeze, breezes, breezy, brisk, calm, celsius, chill, chilled, chillier, chilliest, chilly, chinook, cirrocumulus, cirrostratus, cirrus, climate, climates, cloud, cloudburst, cloudbursts, cloudier, cloudiest, clouds, cloudy, cold, colder, coldest, condensation, contrail, contrails, cool, cooled, cooling, cools, cumulonimbus, cumulus, cyclone, cyclones, damp, damp, damper, damper, dampest, dampest, degree, degrees, deluge, dew, dews, dewy, doppler, downburst, downbursts, downdraft, downdrafts, downpour, downpours, dried, drier, dries, driest, drizzle, drizzled, drizzles, drizzly, drought, droughts, dry, dryline, fall, farenheit, flood, flooded, flooding, floods, flurries, flurry, fog, fogbow, fogbows, fogged, fogging, foggy, fogs, forecast, forecasted, forecasting, forecasts, freeze, freezes, freezing, frigid, frost, frostier, frostiest, frosts, frosty, froze, frozen, gale, gales, galoshes, gust, gusting, gusts, gusty, haboob, haboobs, hail, hailed, hailing, hails, haze, hazes, hazy, heat, heated, heating, heats, hoarfrost, hot, hotter, hottest, humid, humidity, hurricane, hurricanes, ice, iced, ices, icing, icy, inclement, landspout, landspouts, lightning, lightnings, macroburst, macrobursts, maelstrom, mercury, meteorologic, meteorologist, meteorologists, meteorology, microburst, microbursts, microclimate, microclimates, millibar, millibars, mist, misted, mists, misty, moist, moisture, monsoon, monsoons, mugginess, muggy, nexrad, nippy, NOAA, nor'easter, nor'easters, noreaster, noreasters, overcast, ozone, parched, parching, pollen, precipitate, precipitated, precipitates, precipitating, precipitation, psychrometer, radar, rain, rainboots, rainbow, rainbows, raincoat, raincoats, rained, rainfall, rainier, rainiest, raining, rains, rainy, sandstorm, sandstorms, scorcher, scorching, searing, shower, showering, showers, skiff, sleet, slicker, slickers, slush, slushy, smog, smoggier, smoggiest, smoggy, snow, snowed, snowier, snowiest, snowing, snowmageddon, snowpocalypse, snows, snowy, spring, sprinkle, sprinkles, sprinkling, squall, squalls, squally, storm, stormed, stormier, stormiest, storming, storms, stormy, stratocumulus, stratus, subtropical, summer, summery, sun, sunnier, sunniest, sunny, temperate, temperature, tempest, thaw, thawed, thawing, thaws, thermometer, thunder, thundered, thundering, thunders, thunderstorm, thunderstorms, tornadic, tornado, tornadoes, tropical, troposphere, tsunami, turbulent, twister, twisters, typhoon, typhoons, umbrella, umbrellas, vane, warm, warmed, warming, warms, warmth, waterspout, waterspouts, weather, wet, wetter, wettest, wind, windchill, windchills, windier, windiest, windspeed, windy, winter, wintery, and wintry.
 

### 2021-07-01 doc/temp-sentiment.tex
submitted versions of files

@@
 
-\author[1,*]{Matthew Cooper}
-\author[2]{Jeremiah Osborne-Gowey}
-\author[3]{Zheng Liu}
-\author[4]{Portia Adade Williams}
-\author[5]{Jie Liu}
-\author[6]{Aaron Schwartz}
-\author[7]{Patrick Baylis}
-
-\affil[1]{T.H. Chan School of Public Health, Harvard University}
-\affil[2]{Environmental Studies Program, University of Colorado Boulder}
-\affil[3]{Department of Geographical Sciences, University of Maryland College Park}
-\affil[4]{CSIR-Science and Technology Policy Research Institute, Ghana}
-\affil[5]{School of Business, East China University of Science and Technology}
-\affil[6]{University of Colorado Boulder}
-\affil[7]{University of British Columbia}
-\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
+\author{}
+% \author[1,*]{Matthew Cooper}
+% \author[2]{Jeremiah Osborne-Gowey}
+% \author[3]{Zheng Liu}
+% \author[4]{Portia Adade Williams}
+% \author[5]{Jie Liu}
+% \author[6]{Aaron Schwartz}
+% \author[7]{Patrick Baylis} 
+% 
+% \affil[1]{T.H. Chan School of Public Health, Harvard University, U.S.A.}
+% \affil[2]{Environmental Studies Program, University of Colorado Boulder, U.S.A.}
+% \affil[3]{Department of Geographical Sciences, University of Maryland College Park, U.S.A.}
+% \affil[4]{CSIR-Science and Technology Policy Research Institute, Ghana}
+% \affil[5]{School of Business, East China University of Science and Technology, China} 
+% \affil[6]{Department of Ecology \& Evolutionary Biology, University of Colorado Boulder, U.S.A.}
+% \affil[7]{Vancouver School of Economics, University of British Columbia, Canada}
+% \affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
 
@@
 
-%\printbibliography
-\bibliography{TextDataClimateShocks}
+%\bibliography{TextDataClimateShocks}
 
-\section*{Author Contributions}
-Author contributions: M.C., J.O.-G., A.S.
+\begin{thebibliography}{10}
+\urlstyle{rm}
+\expandafter\ifx\csname url\endcsname\relax
+  \def\url#1{\texttt{#1}}\fi
+\expandafter\ifx\csname urlprefix\endcsname\relax\def\urlprefix{URL }\fi
+\expandafter\ifx\csname doiprefix\endcsname\relax\def\doiprefix{DOI: }\fi
+\providecommand{\bibinfo}[2]{#2}
+\providecommand{\eprint}[2][]{\url{#2}}
+
+\bibitem{pachauri2014climate}
+\bibinfo{author}{Pachauri, R.~K.} \emph{et~al.}
+\newblock \emph{\bibinfo{title}{Climate change 2014: synthesis report.
+  Contribution of Working Groups I, II and III to the fifth assessment report
+  of the Intergovernmental Panel on Climate Change}}
+  (\bibinfo{publisher}{Ipcc}, \bibinfo{year}{2014}).
+
+\bibitem{Berry2018Apr}
+\bibinfo{author}{Berry, H.~L.}, \bibinfo{author}{Waite, T.~D.},
+  \bibinfo{author}{Dear, K. B.~G.}, \bibinfo{author}{Capon, A.~G.} \&
+  \bibinfo{author}{Murray, V.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{The case for systems thinking
+  about climate change and mental health}}}.
+\newblock {\emph{\JournalTitle{Nat. Clim. Change}}}
+  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{282--290},
+  \doiprefix\url{10.1038/s41558-018-0102-4} (\bibinfo{year}{2018}).
+
+\bibitem{Collins2011Jul}
+\bibinfo{author}{Collins, P.~Y.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Grand challenges in global mental
+  health}}}.
+\newblock {\emph{\JournalTitle{Nature}}} \textbf{\bibinfo{volume}{475}},
+  \bibinfo{pages}{27--30}, \doiprefix\url{10.1038/475027a}
+  (\bibinfo{year}{2011}).
+
+\bibitem{Schwartz2017Aug}
+\bibinfo{author}{Schwartz, R.~M.}, \bibinfo{author}{Gillezeau, C.~N.},
+  \bibinfo{author}{Liu, B.}, \bibinfo{author}{Lieberman-Cribbin, W.} \&
+  \bibinfo{author}{Taioli, E.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Longitudinal Impact of Hurricane
+  Sandy Exposure on Mental Health Symptoms}}}.
+\newblock {\emph{\JournalTitle{Int. J. Environ. Res. Public Health}}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{957},
+  \doiprefix\url{10.3390/ijerph14090957} (\bibinfo{year}{2017}).
+
+\bibitem{baylis_weather_2018}
+\bibinfo{author}{Baylis, P.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Weather impacts expressed
+  sentiment}}.
+\newblock {\emph{\JournalTitle{PLOS One}}} \textbf{\bibinfo{volume}{13}},
+  \bibinfo{pages}{e0195750}, \doiprefix\url{10.1371/journal.pone.0195750}
+  (\bibinfo{year}{2018}).
+
+\bibitem{Mullins2019Dec}
+\bibinfo{author}{Mullins, J.~T.} \& \bibinfo{author}{White, C.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Temperature and mental health:
+  Evidence from the spectrum of mental health outcomes}}}.
+\newblock {\emph{\JournalTitle{J. Health Econ.}}}
+  \textbf{\bibinfo{volume}{68}}, \bibinfo{pages}{102240},
+  \doiprefix\url{10.1016/j.jhealeco.2019.102240} (\bibinfo{year}{2019}).
+
+\bibitem{Obradovich2018Oct}
+\bibinfo{author}{Obradovich, N.}, \bibinfo{author}{Migliorini, R.},
+  \bibinfo{author}{Paulus, M.~P.} \& \bibinfo{author}{Rahwan, I.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Empirical evidence of mental
+  health risks posed by climate change}}}.
+\newblock {\emph{\JournalTitle{Proc. Natl. Acad. Sci. U.S.A.}}}
+  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{10953--10958},
+  \doiprefix\url{10.1073/pnas.1801528115} (\bibinfo{year}{2018}).
+
+\bibitem{allen2019technical}
+\bibinfo{author}{Allen, M.} \emph{et~al.}
+\newblock \bibinfo{title}{Technical summary: Global warming of 1.5° c. an ipcc
+  special report on the impacts of global warming of 1.5° c above
+  pre-industrial levels and related global greenhouse gas emission pathways, in
+  the context of strengthening the global response to the threat of climate
+  change, sustainable development, and efforts to eradicate poverty}
+  (\bibinfo{year}{2019}).
+
+\bibitem{Palinkas2020Apr}
+\bibinfo{author}{Palinkas, L.~A.} \& \bibinfo{author}{Wong, M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Global climate change and mental
+  health}}}.
+\newblock {\emph{\JournalTitle{Current Opinion in Psychology}}}
+  \textbf{\bibinfo{volume}{32}}, \bibinfo{pages}{12--16},
+  \doiprefix\url{10.1016/j.copsyc.2019.06.023} (\bibinfo{year}{2020}).
+
+\bibitem{Lohmus2018Jul}
+\bibinfo{author}{L{\ifmmode\tilde{o}\else\~{o}\fi}hmus, M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Possible Biological Mechanisms
+  Linking Mental Health and Heat{\ifmmode---\else\textemdash\fi}A Contemplative
+  Review}}}.
+\newblock {\emph{\JournalTitle{Int. J. Environ. Res. Public Health}}}
+  \textbf{\bibinfo{volume}{15}}, \bibinfo{pages}{1515},
+  \doiprefix\url{10.3390/ijerph15071515} (\bibinfo{year}{2018}).
+
+\bibitem{Berry2007}
+\bibinfo{author}{Berry, H.~L.}
+\newblock \bibinfo{journal}{\bibinfo{title}{'crowded suburbs' and 'killer
+  cities': a brief review of the relationship between urban environments and
+  mental health}}.
+\newblock {\emph{\JournalTitle{N S W Public Health Bull}}}
+  \textbf{\bibinfo{volume}{18}}, \bibinfo{pages}{222--7},
+  \doiprefix\url{10.1071/nb07024} (\bibinfo{year}{2007}).
+
+\bibitem{Burke2015Nov}
+\bibinfo{author}{Burke, M.}, \bibinfo{author}{Hsiang, S.~M.} \&
+  \bibinfo{author}{Miguel, E.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Global non-linear effect of
+  temperature on economic production}}}.
+\newblock {\emph{\JournalTitle{Nature}}} \textbf{\bibinfo{volume}{527}},
+  \bibinfo{pages}{235--239}, \doiprefix\url{10.1038/nature15725}
+  (\bibinfo{year}{2015}).
+
+\bibitem{Burke2018Aug}
+\bibinfo{author}{Burke, M.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Higher temperatures increase
+  suicide rates in the United States and Mexico}}}.
+\newblock {\emph{\JournalTitle{Nat. Clim. Change}}}
+  \textbf{\bibinfo{volume}{8}}, \bibinfo{pages}{723--729},
+  \doiprefix\url{10.1038/s41558-018-0222-x} (\bibinfo{year}{2018}).
+
+\bibitem{Dixon2007May}
+\bibinfo{author}{Dixon, P.~G.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Effects of temperature variation
+  on suicide in five U.S. counties, 1991{\textendash}2001}}}.
+\newblock {\emph{\JournalTitle{Int. J. Biometeorol.}}}
+  \textbf{\bibinfo{volume}{51}}, \bibinfo{pages}{395--403},
+  \doiprefix\url{10.1007/s00484-006-0081-4} (\bibinfo{year}{2007}).
+
+\bibitem{Qi2014Dec}
+\bibinfo{author}{Qi, X.}, \bibinfo{author}{Hu, W.}, \bibinfo{author}{Mengersen,
+  K.} \& \bibinfo{author}{Tong, S.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Socio-environmental drivers and
+  suicide in Australia: Bayesian spatial analysis}}}.
+\newblock {\emph{\JournalTitle{BMC Public Health}}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{1--10},
+  \doiprefix\url{10.1186/1471-2458-14-681} (\bibinfo{year}{2014}).
+
+\bibitem{Likhvar2011Jan}
+\bibinfo{author}{Likhvar, V.}, \bibinfo{author}{Honda, Y.} \&
+  \bibinfo{author}{Ono, M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Relation between temperature and
+  suicide mortality in Japan in the presence of other confounding factors using
+  time-series analysis with a semiparametric approach}}}.
+\newblock {\emph{\JournalTitle{Environ. Health Preventive Med.}}}
+  \textbf{\bibinfo{volume}{16}}, \bibinfo{pages}{36--43},
+  \doiprefix\url{10.1007/s12199-010-0163-0} (\bibinfo{year}{2011}).
+
+\bibitem{Lee2007Jan}
+\bibinfo{author}{Lee, H.-C.}, \bibinfo{author}{Tsai, S.-Y.} \&
+  \bibinfo{author}{Lin, H.-C.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Seasonal variations in bipolar
+  disorder admissions and the association with climate: A population-based
+  study}}}.
+\newblock {\emph{\JournalTitle{J. Affect. Disord.}}}
+  \textbf{\bibinfo{volume}{97}}, \bibinfo{pages}{61--69},
+  \doiprefix\url{10.1016/j.jad.2006.06.026} (\bibinfo{year}{2007}).
+
+\bibitem{Sung2013Feb}
+\bibinfo{author}{Sung, T.-I.}, \bibinfo{author}{Chen, M.-J.} \&
+  \bibinfo{author}{Su, H.-J.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{A positive relationship between
+  ambient temperature and bipolar disorder identified using a national cohort
+  of psychiatric inpatients}}}.
+\newblock {\emph{\JournalTitle{Soc. Psychiatry Psychiatr. Epidemiol.}}}
+  \textbf{\bibinfo{volume}{48}}, \bibinfo{pages}{295--302},
+  \doiprefix\url{10.1007/s00127-012-0542-5} (\bibinfo{year}{2013}).
+
+\bibitem{Hansen2008Oct}
+\bibinfo{author}{Hansen, A.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{The Effect of Heat Waves on Mental
+  Health in a Temperate Australian City}}}.
+\newblock {\emph{\JournalTitle{Environ. Health Perspect.}}}
+  (\bibinfo{year}{2008}).
+
+\bibitem{Obradovich2017May}
+\bibinfo{author}{Obradovich, N.}, \bibinfo{author}{Migliorini, R.},
+  \bibinfo{author}{Mednick, S.~C.} \& \bibinfo{author}{Fowler, J.~H.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Nighttime temperature and human
+  sleep loss in a changing climate}}}.
+\newblock {\emph{\JournalTitle{Sci. Adv.}}} \textbf{\bibinfo{volume}{3}},
+  \bibinfo{pages}{e1601555}, \doiprefix\url{10.1126/sciadv.1601555}
+  (\bibinfo{year}{2017}).
+
+\bibitem{Gubernot2014Oct}
+\bibinfo{author}{Gubernot, D.~M.}, \bibinfo{author}{Anderson, G.~B.} \&
+  \bibinfo{author}{Hunting, K.~L.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{The epidemiology of occupational
+  heat exposure in the United States: a review of the literature and assessment
+  of research needs in a changing climate}}}.
+\newblock {\emph{\JournalTitle{Int. J. Biometeorol.}}}
+  \textbf{\bibinfo{volume}{58}}, \bibinfo{pages}{1779--1788},
+  \doiprefix\url{10.1007/s00484-013-0752-x} (\bibinfo{year}{2014}).
+
+\bibitem{Karner2015Dec}
+\bibinfo{author}{Karner, A.}, \bibinfo{author}{Hondula, D.~M.} \&
+  \bibinfo{author}{Vanos, J.~K.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Heat exposure during non-motorized
+  travel: Implications for transportation policy under climate change}}}.
+\newblock {\emph{\JournalTitle{Journal of Transport {\&} Health}}}
+  \textbf{\bibinfo{volume}{2}}, \bibinfo{pages}{451--459},
+  \doiprefix\url{10.1016/j.jth.2015.10.001} (\bibinfo{year}{2015}).
+
+\bibitem{Samuelson2020Jun}
+\bibinfo{author}{Samuelson, H.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Housing as a critical determinant
+  of heat vulnerability and health}}}.
+\newblock {\emph{\JournalTitle{Sci. Total Environ.}}}
+  \textbf{\bibinfo{volume}{720}}, \bibinfo{pages}{137296},
+  \doiprefix\url{10.1016/j.scitotenv.2020.137296} (\bibinfo{year}{2020}).
+
+\bibitem{Kang2016}
+\bibinfo{author}{Kang, S.~K.}, \bibinfo{author}{DeCelles, K.~A.},
+  \bibinfo{author}{Tilcsik, A.} \& \bibinfo{author}{Jun, S.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Whitened
+  R{\ifmmode\acute{e}\else\'{e}\fi}sum{\ifmmode\acute{e}\else\'{e}\fi}s: Race
+  and Self-Presentation in the Labor Market}}}.
+\newblock {\emph{\JournalTitle{Adm. Sci. Q.}}} \textbf{\bibinfo{volume}{61}},
+  \bibinfo{pages}{469--502}, \doiprefix\url{10.1177/0001839216639577}
+  (\bibinfo{year}{2016}).
+
+\bibitem{Penner2008}
+\bibinfo{author}{Penner, A.~M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Race and Gender Differences in
+  Wages: The Role of Occupational Sorting at the Point of Hire}}}.
+\newblock {\emph{\JournalTitle{Sociological Quarterly}}}
+  \textbf{\bibinfo{volume}{49}}, \bibinfo{pages}{597--614},
+  \doiprefix\url{10.1111/j.1533-8525.2008.00129.x} (\bibinfo{year}{2008}).
+
+\bibitem{Desmond2015}
+\bibinfo{author}{Desmond, M.} \& \bibinfo{author}{Shollenberger, T.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Forced Displacement From Rental
+  Housing: Prevalence and Neighborhood Consequences}}}.
+\newblock {\emph{\JournalTitle{Demography}}} \textbf{\bibinfo{volume}{52}},
+  \bibinfo{pages}{1751--1772}, \doiprefix\url{10.1007/s13524-015-0419-9}
+  (\bibinfo{year}{2015}).
+
+\bibitem{Akbar2019}
+\bibinfo{author}{Akbar, P.~A.}, \bibinfo{author}{Li, S.},
+  \bibinfo{author}{Shertzer, A.} \& \bibinfo{author}{Walsh, R.~P.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Racial Segregation in Housing
+  Markets and the Erosion of Black Wealth}}}.
+\newblock {\emph{\JournalTitle{NBER}}} \doiprefix\url{10.3386/w25805}
+  (\bibinfo{year}{2019}).
+
+\bibitem{Banzhaf2019}
+\bibinfo{author}{Banzhaf, S.}, \bibinfo{author}{Ma, L.} \&
+  \bibinfo{author}{Timmins, C.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Environmental Justice: The
+  Economics of Race, Place, and Pollution}}}.
+\newblock {\emph{\JournalTitle{J. Econ. Perspect.}}}
+  \textbf{\bibinfo{volume}{33}}, \bibinfo{pages}{185--208},
+  \doiprefix\url{10.1257/jep.33.1.185} (\bibinfo{year}{2019}).
+
+\bibitem{Belanger2015Mar}
+\bibinfo{author}{B{\ifmmode\acute{e}\else\'{e}\fi}langer, D.},
+  \bibinfo{author}{Gosselin, P.}, \bibinfo{author}{Valois, P.} \&
+  \bibinfo{author}{Abdous, B.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Neighbourhood and dwelling
+  characteristics associated with the self-reported adverse health effects of
+  heat in most deprived urban areas: A cross-sectional study in 9 cities}}}.
+\newblock {\emph{\JournalTitle{Health {\&} Place}}}
+  \textbf{\bibinfo{volume}{32}}, \bibinfo{pages}{8--18},
+  \doiprefix\url{10.1016/j.healthplace.2014.12.014} (\bibinfo{year}{2015}).
+
+\bibitem{Uejio2011Mar}
+\bibinfo{author}{Uejio, C.~K.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Intra-urban societal vulnerability
+  to extreme heat: The role of heat exposure and the built environment,
+  socioeconomics, and neighborhood stability}}}.
+\newblock {\emph{\JournalTitle{Health {\&} Place}}}
+  \textbf{\bibinfo{volume}{17}}, \bibinfo{pages}{498--507},
+  \doiprefix\url{10.1016/j.healthplace.2010.12.005} (\bibinfo{year}{2011}).
+
+\bibitem{ferre2019hurricane}
+\bibinfo{author}{Ferr{\'e}, I.~M.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Hurricane maria’s impact on punta
+  santiago, puerto rico: community needs and mental health assessment six
+  months postimpact}}.
+\newblock {\emph{\JournalTitle{Disaster Med Public Health Prep}}}
+  \textbf{\bibinfo{volume}{13}}, \bibinfo{pages}{18--23}
+  (\bibinfo{year}{2019}).
+
+\bibitem{Gruebner2015Jun}
+\bibinfo{author}{Gruebner, O.}, \bibinfo{author}{Lowe, S.~R.},
+  \bibinfo{author}{Sampson, L.} \& \bibinfo{author}{Galea, S.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{The geography of post-disaster
+  mental health: spatial patterning of psychological vulnerability and
+  resilience factors in New York City after Hurricane Sandy}}}.
+\newblock {\emph{\JournalTitle{Int. J. Health Geographics}}}
+  \textbf{\bibinfo{volume}{14}}, \bibinfo{pages}{16--13},
+  \doiprefix\url{10.1186/s12942-015-0008-6} (\bibinfo{year}{2015}).
+
+\bibitem{Edo-Osagie2020Jul}
+\bibinfo{author}{Edo-Osagie, O.}, \bibinfo{author}{De~La~Iglesia, B.},
+  \bibinfo{author}{Lake, I.} \& \bibinfo{author}{Edeghere, O.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{A scoping review of the use of
+  Twitter for public health research}}}.
+\newblock {\emph{\JournalTitle{Comput. Biol. Med.}}}
+  \textbf{\bibinfo{volume}{122}}, \bibinfo{pages}{103770},
+  \doiprefix\url{10.1016/j.compbiomed.2020.103770} (\bibinfo{year}{2020}).
+
+\bibitem{Sinnenberg2016Dec}
+\bibinfo{author}{Sinnenberg, L.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Twitter as a Tool for Health
+  Research: A Systematic Review}}}.
+\newblock {\emph{\JournalTitle{American Public Health Association (APHA)
+  publications}}}  (\bibinfo{year}{2016}).
+
+\bibitem{Reece2017Oct}
+\bibinfo{author}{Reece, A.~G.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Forecasting the onset and course
+  of mental illness with Twitter data}}}.
+\newblock {\emph{\JournalTitle{Sci. Rep.}}} \textbf{\bibinfo{volume}{7}},
+  \bibinfo{pages}{1--11}, \doiprefix\url{10.1038/s41598-017-12961-9}
+  (\bibinfo{year}{2017}).
+
+\bibitem{Eichstaedt2018Oct}
+\bibinfo{author}{Eichstaedt, J.~C.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Facebook language predicts
+  depression in medical records}}}.
+\newblock {\emph{\JournalTitle{Proc. Natl. Acad. Sci. U.S.A.}}}
+  \textbf{\bibinfo{volume}{115}}, \bibinfo{pages}{11203--11208},
+  \doiprefix\url{10.1073/pnas.1802331115} (\bibinfo{year}{2018}).
+
+\bibitem{Kolliakou2020Feb}
+\bibinfo{author}{Kolliakou, A.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Mental health-related
+  conversations on social media and crisis episodes: a time-series regression
+  analysis}}}.
+\newblock {\emph{\JournalTitle{Sci. Rep.}}} \textbf{\bibinfo{volume}{10}},
+  \bibinfo{pages}{1--7}, \doiprefix\url{10.1038/s41598-020-57835-9}
+  (\bibinfo{year}{2020}).
+
+\bibitem{Mitchell2013May}
+\bibinfo{author}{Mitchell, L.}, \bibinfo{author}{Frank, M.~R.},
+  \bibinfo{author}{Harris, K.~D.}, \bibinfo{author}{Dodds, P.~S.} \&
+  \bibinfo{author}{Danforth, C.~M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{The Geography of Happiness:
+  Connecting Twitter Sentiment and Expression, Demographics, and Objective
+  Characteristics of Place}}}.
+\newblock {\emph{\JournalTitle{PLOS One}}} \textbf{\bibinfo{volume}{8}},
+  \bibinfo{pages}{e64417}, \doiprefix\url{10.1371/journal.pone.0064417}
+  (\bibinfo{year}{2013}).
+
+\bibitem{Dodds2011}
+\bibinfo{author}{Dodds, P.~S.}, \bibinfo{author}{Harris, K.~D.},
+  \bibinfo{author}{Kloumann, I.~M.}, \bibinfo{author}{Bliss, C.~A.} \&
+  \bibinfo{author}{Danforth, C.~M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Temporal Patterns of Happiness and
+  Information in a Global Social Network: Hedonometrics and Twitter}}}.
+\newblock {\emph{\JournalTitle{PLoS One}}} \textbf{\bibinfo{volume}{6}},
+  \bibinfo{pages}{e26752}, \doiprefix\url{10.1371/journal.pone.0026752}
+  (\bibinfo{year}{2011}).
+
+\bibitem{budd2008wet}
+\bibinfo{author}{Budd, G.~M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Wet-bulb globe temperature
+  (wbgt)—its history and its limitations}}.
+\newblock {\emph{\JournalTitle{Journal of Science and Medicine in Sport}}}
+  \textbf{\bibinfo{volume}{11}}, \bibinfo{pages}{20--32}
+  (\bibinfo{year}{2008}).
+
+\bibitem{CDC2021}
+\bibinfo{author}{for Disease Control Wide-ranging ONline Data~for Epidemiologic
+  Research (CDC~WONDER), C.}
+\newblock \bibinfo{title}{Underlying cause of death} (\bibinfo{year}{2021}).
+
+\bibitem{Lieberman-Cribbin2017}
+\bibinfo{author}{Lieberman-Cribbin, W.}, \bibinfo{author}{Liu, B.},
+  \bibinfo{author}{Schneider, S.}, \bibinfo{author}{Schwartz, R.} \&
+  \bibinfo{author}{Taioli, E.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Self-Reported and FEMA Flood
+  Exposure Assessment after Hurricane Sandy: Association with Mental Health
+  Outcomes}}}.
+\newblock {\emph{\JournalTitle{PLOS One}}} \textbf{\bibinfo{volume}{12}},
+  \bibinfo{pages}{e0170965}, \doiprefix\url{10.1371/journal.pone.0170965}
+  (\bibinfo{year}{2017}).
+
+\bibitem{Hoffman2020}
+\bibinfo{author}{Hoffman, J.~S.}, \bibinfo{author}{Shandas, V.} \&
+  \bibinfo{author}{Pendleton, N.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{The Effects of Historical Housing
+  Policies on Resident Exposure to Intra-Urban Heat: A Study of 108 US Urban
+  Areas}}}.
+\newblock {\emph{\JournalTitle{Climate}}} \textbf{\bibinfo{volume}{8}},
+  \bibinfo{pages}{12}, \doiprefix\url{10.3390/cli8010012}
+  (\bibinfo{year}{2020}).
+
+\bibitem{Locke2021}
+\bibinfo{author}{Locke, D.~H.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Residential housing segregation
+  and urban tree canopy in 37 US Cities}}}.
+\newblock {\emph{\JournalTitle{npj Urban Sustain.}}}
+  \textbf{\bibinfo{volume}{1}}, \bibinfo{pages}{1--9},
+  \doiprefix\url{10.1038/s42949-021-00022-0} (\bibinfo{year}{2021}).
+
+\bibitem{Tong2021}
+\bibinfo{author}{Tong, K.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Measuring social equity in urban
+  energy use and interventions using fine-scale data}}}.
+\newblock {\emph{\JournalTitle{Proc. Natl. Acad. Sci. U.S.A.}}}
+  \textbf{\bibinfo{volume}{118}}, \bibinfo{pages}{e2023554118},
+  \doiprefix\url{10.1073/pnas.2023554118} (\bibinfo{year}{2021}).
+
+\bibitem{Raymond2020May}
+\bibinfo{author}{Raymond, C.}, \bibinfo{author}{Matthews, T.} \&
+  \bibinfo{author}{Horton, R.~M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{The emergence of heat and humidity
+  too severe for human tolerance}}}.
+\newblock {\emph{\JournalTitle{Sci. Adv.}}} \textbf{\bibinfo{volume}{6}},
+  \bibinfo{pages}{eaaw1838}, \doiprefix\url{10.1126/sciadv.aaw1838}
+  (\bibinfo{year}{2020}).
+
+\bibitem{Harrington2020Sep}
+\bibinfo{author}{Harrington, L.~J.} \& \bibinfo{author}{Otto, F. E.~L.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Reconciling theory with the
+  reality of African heatwaves}}}.
+\newblock {\emph{\JournalTitle{Nat. Clim. Change}}}
+  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{796--798},
+  \doiprefix\url{10.1038/s41558-020-0851-8} (\bibinfo{year}{2020}).
+
+\bibitem{Biardeau2020Jan}
+\bibinfo{author}{Biardeau, L.~T.}, \bibinfo{author}{Davis, L.~W.},
+  \bibinfo{author}{Gertler, P.} \& \bibinfo{author}{Wolfram, C.}
+\newblock \bibinfo{journal}{\bibinfo{title}{{Heat exposure and global air
+  conditioning}}}.
+\newblock {\emph{\JournalTitle{Nat. Sustainability}}}
+  \textbf{\bibinfo{volume}{3}}, \bibinfo{pages}{25--28},
+  \doiprefix\url{10.1038/s41893-019-0441-9} (\bibinfo{year}{2020}).
+
+\bibitem{Pew2020Sep}
+\bibinfo{title}{{10 facts about Americans and Twitter}} (\bibinfo{year}{2020}).
+\newblock \bibinfo{note}{[Online; accessed 1. May 2021]}.
+
+\bibitem{bullard2012wrong}
+\bibinfo{author}{Bullard, R.~D.} \& \bibinfo{author}{Wright, B.}
+\newblock \emph{\bibinfo{title}{The wrong complexion for protection: How the
+  government response to disaster endangers African American communities}}
+  (\bibinfo{publisher}{NYU Press}, \bibinfo{year}{2012}).
+
+\bibitem{hutto2014vader}
+\bibinfo{author}{Hutto, C.} \& \bibinfo{author}{Gilbert, E.}
+\newblock \bibinfo{title}{Vader: A parsimonious rule-based model for sentiment
+  analysis of social media text}.
+\newblock In \emph{\bibinfo{booktitle}{Proceedings of the International AAAI
+  Conference on Web and Social Media}}, vol.~\bibinfo{volume}{8}
+  (\bibinfo{year}{2014}).
+
+\bibitem{gilbert_vader_2014}
+\bibinfo{author}{Gilbert, C. H.~E.} \& \bibinfo{author}{Hutto, E.}
+\newblock \bibinfo{title}{Vader: {A} parsimonious rule-based model for
+  sentiment analysis of social media text}.
+\newblock In \emph{\bibinfo{booktitle}{Eighth {International} {Conference} on
+  {Weblogs} and {Social} {Media} ({ICWSM}-14). {Available} at (20/04/16)
+  http://comp. social. gatech. edu/papers/icwsm14. vader. hutto. pdf}},
+  vol.~\bibinfo{volume}{81}, \bibinfo{pages}{82} (\bibinfo{year}{2014}).
+
+\bibitem{xia_continental-scale_2012}
+\bibinfo{author}{Xia, Y.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Continental-scale water and energy
+  flux analysis and validation for the {North} {American} {Land} {Data}
+  {Assimilation} {System} project phase 2 ({NLDAS}-2): 1. {Intercomparison} and
+  application of model products}}.
+\newblock {\emph{\JournalTitle{Journal of Geophysical Research: Atmospheres}}}
+  \textbf{\bibinfo{volume}{117}},
+  \doiprefix\url{https://doi.org/10.1029/2011JD016048} (\bibinfo{year}{2012}).
+\newblock \bibinfo{note}{\_eprint:
+  https://agupubs.onlinelibrary.wiley.com/doi/pdf/10.1029/2011JD016048}.
+
+\bibitem{rao2020projections}
+\bibinfo{author}{Rao, K.~K.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{projections of heat stress and
+  associated work performance over india in response to global warming}}.
+\newblock {\emph{\JournalTitle{Scientific reports}}}
+  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{1--14} (\bibinfo{year}{2020}).
+
+\bibitem{hu2017impact}
+\bibinfo{author}{Hu, X.}, \bibinfo{author}{Wu, J.}, \bibinfo{author}{Chen, P.},
+  \bibinfo{author}{Sun, T.} \& \bibinfo{author}{Li, D.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Impact of climate variability and
+  change on crime rates in tangshan, china}}.
+\newblock {\emph{\JournalTitle{Science of the total environment}}}
+  \textbf{\bibinfo{volume}{609}}, \bibinfo{pages}{1041--1048}
+  (\bibinfo{year}{2017}).
+
+\bibitem{chien2016spatiotemporal}
+\bibinfo{author}{Chien, L.-C.}, \bibinfo{author}{Guo, Y.} \&
+  \bibinfo{author}{Zhang, K.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Spatiotemporal analysis of heat and
+  heat wave effects on elderly mortality in texas, 2006--2011}}.
+\newblock {\emph{\JournalTitle{Science of the Total Environment}}}
+  \textbf{\bibinfo{volume}{562}}, \bibinfo{pages}{845--851}
+  (\bibinfo{year}{2016}).
+
+\bibitem{armstrong2019role}
+\bibinfo{author}{Armstrong, B.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{The role of humidity in
+  associations of high temperature with mortality: a multicountry, multicity
+  study}}.
+\newblock {\emph{\JournalTitle{Environmental health perspectives}}}
+  \textbf{\bibinfo{volume}{127}}, \bibinfo{pages}{097007}
+  (\bibinfo{year}{2019}).
+
+\bibitem{vida2012relationship}
+\bibinfo{author}{Vida, S.}, \bibinfo{author}{Durocher, M.},
+  \bibinfo{author}{Ouarda, T.~B.} \& \bibinfo{author}{Gosselin, P.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Relationship between ambient
+  temperature and humidity and visits to mental health emergency departments in
+  qu{\'e}bec}}.
+\newblock {\emph{\JournalTitle{Psychiatric Services}}}
+  \textbf{\bibinfo{volume}{63}}, \bibinfo{pages}{1150--1153}
+  (\bibinfo{year}{2012}).
+
+\bibitem{ding2016importance}
+\bibinfo{author}{Ding, N.}, \bibinfo{author}{Berry, H.~L.} \&
+  \bibinfo{author}{Bennett, C.~M.}
+\newblock \bibinfo{journal}{\bibinfo{title}{The importance of humidity in the
+  relationship between heat and population mental health: evidence from
+  australia}}.
+\newblock {\emph{\JournalTitle{PLOS One}}} \textbf{\bibinfo{volume}{11}},
+  \bibinfo{pages}{e0164190} (\bibinfo{year}{2016}).
+
+\bibitem{bolton_computation_1980}
+\bibinfo{author}{Bolton, D.}
+\newblock \bibinfo{journal}{\bibinfo{title}{The {Computation} of {Equivalent}
+  {Potential} {Temperature}}}.
+\newblock {\emph{\JournalTitle{Monthly Weather Review}}}
+  \textbf{\bibinfo{volume}{108}}, \bibinfo{pages}{1046--1053},
+  \doiprefix\url{10.1175/1520-0493(1980)108<1046:TCOEPT>2.0.CO;2}
+  (\bibinfo{year}{1980}).
+\newblock \bibinfo{note}{Publisher: American Meteorological Society}.
+
+\bibitem{heo2019comparison}
+\bibinfo{author}{Heo, S.}, \bibinfo{author}{Bell, M.~L.} \&
+  \bibinfo{author}{Lee, J.-T.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Comparison of health risks by heat
+  wave definition: Applicability of wet-bulb globe temperature for heat wave
+  criteria}}.
+\newblock {\emph{\JournalTitle{Environmental research}}}
+  \textbf{\bibinfo{volume}{168}}, \bibinfo{pages}{158--170}
+  (\bibinfo{year}{2019}).
+
+\bibitem{ruggles2018ipums}
+\bibinfo{author}{Ruggles, S.} \emph{et~al.}
+\newblock \bibinfo{journal}{\bibinfo{title}{Ipums usa: Version 8.0 [dataset]}}.
+\newblock {\emph{\JournalTitle{Minneapolis, MN: IPUMS}}}
+  \textbf{\bibinfo{volume}{10}}, \bibinfo{pages}{D010} (\bibinfo{year}{2018}).
+
+\bibitem{wood2017generalized}
+\bibinfo{author}{Wood, S.~N.}
+\newblock \emph{\bibinfo{title}{Generalized additive models: an introduction
+  with R}} (\bibinfo{publisher}{CRC press}, \bibinfo{year}{2017}).
+
+\end{thebibliography}
+
+% \section*{Author Contributions}
+% Author contributions: M.C., J.O.-G., A.S.
 and P.B.

### 2021-08-19 doc/pnas/PNAS-template-main.tex
pnas submission

--- 2021-06-30 doc/temp-sentiment.tex
@@
-\documentclass[fleqn,10pt]{wlscirep}
-\usepackage[utf8]{inputenc}
-\usepackage[T1]{fontenc}
-\usepackage{graphicx}
-\usepackage{hyperref}
-\usepackage{breakcites}
-\usepackage{float}
-\usepackage{textcomp}
-\usepackage{amsmath}
-\usepackage{textgreek}
-\usepackage{authblk}
-\usepackage{rotating}
-\usepackage{booktabs}
-\usepackage{longtable}
-\usepackage{pdflscape}
-\usepackage{lineno}
-\usepackage{xcolor}
-% \usepackage[
-% style=numeric,
-% citestyle=numeric-comp,
-% backend=biber,
-% doi=true,
-% natbib=true,
-% sorting=none
-% ]{biblatex}
-\usepackage{scalerel,xparse}
+\documentclass[9pt,twocolumn,twoside,lineno]{pnas-new}
+% Use the lineno option to display guide line numbers if required.
 
-% To be able to use emojis
-\NewDocumentCommand\emojismile{}{
- \scalerel*{
- \includegraphics{smiling-face-with-smiling-eyes_1f60a.png}
- }{X}
-}
+\templatetype{pnasresearcharticle} % Choose template 
+% {pnasresearcharticle} = Template for a two-column research article
+% {pnasmathematics} %= Template for a one-column mathematics article
+% {pnasinvited} %= Template for a PNAS invited submission
 
-%https://www.overleaf.com/project/5f918edd1afc380001feca0d
 \title{Fine-scale Data Reveals Neighborhood Inequalities In Effects of Heat on Expressed Mood}
 
-\author[1,*]{Matthew Cooper}
-\author[2]{Jeremiah Osborne-Gowey}
-\author[3]{Zheng Liu}
-\author[4]{Portia Adade Williams}
-\author[5]{Jie Liu}
-\author[6]{Aaron Schwartz}
-\author[7]{Patrick Baylis}
+% Use letters for affiliations, numbers to show equal authorship (if applicable) and to indicate the corresponding author
+\author[a]{Matthew Cooper}
+\author[b]{Jeremiah Osborne-Gowey}
+\author[c]{Zheng Liu}
+\author[d]{Portia Adade Williams}
+\author[e]{Jie Liu}
+\author[f]{Aaron Schwartz}
+\author[g]{Patrick Baylis}
 
-\affil[1]{T.H. Chan School of Public Health, Harvard University}
-\affil[2]{Environmental Studies Program, University of Colorado Boulder}
-\affil[3]{Department of Geographical Sciences, University of Maryland College Park}
-\affil[4]{CSIR-Science and Technology Policy Research Institute, Ghana}
-\affil[5]{School of Business, East China University of Science and Technology}
-\affil[6]{University of Colorado Boulder}
-\affil[7]{University of British Columbia}
-\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
+\affil[a]{T.H. Chan School of Public Health, Harvard University, 677 Huntington Avenue, Boston, MA 02115, USA}
+\affil[b]{Environmental Studies Program, University of Colorado Boulder, 4001 Discoverty Drive, Boulder, CO 80303, USA}
+\affil[c]{Department of Geographical Sciences, University of Maryland College Park, 7251 Preinkert Drive, College Park, MD 20742, USA}
+\affil[d]{CSIR-Science and Technology Policy Research Institute, Agostinho Neto Road, Accra, Ghana}
+\affil[e]{School of Business, East China University of Science and Technology, Meilonglu Street 130, Shanghai 2000237, China} 
+\affil[f]{Department of Ecology \& Evolutionary Biology, University of Colorado Boulder, 1900 Pleasant Street, Boulder, CO 80309, USA}
+\affil[g]{Vancouver School of Economics, University of British Columbia, 6000 Iona Drive
+Vancouver, BC, V6T 1L4, Canada}
+
+% Please give the surname of the lead author for the running footer
+\leadauthor{Cooper}
+
+% Please add here a significance statement to explain the relevance of your work
+\significancestatement{Higher temperatures are known affect many mental health outcomes, including suicides, hospitalizations, and expressed mood in social media.
+However, work to date has found that heat affects mental health uniformly across all populations.
+This is surprising, given that people vary greatly in their exposure to heat based on their working, commuting, and housing conditions.
+We draw on fine-scale twitter data to show that temperature does not affect the mental health of all populations the same, and that people in poor and Black neighbors are much more affected by the heat.}
+
+% Please include corresponding author, author contribution and author declaration information
+\authorcontributions{M.C., J.O.-G., A.S.
+and P.B.
+designed the research and modeling strategy; A.S.
+provided data; M.C and Z.L.
+prepared data; M.C.
+and J.O.-G.
+analyzed data; and M.C., J.O.-G., P.A.W., and J.L.
+wrote the paper.}
+\authordeclaration{We have no competing interests to decalre.}
+\correspondingauthor{\textsuperscript{2}To whom correspondence should be addressed. E-mail: mcooper@hsph.harvard.edu}
+
+% Keywords are not mandatory, but authors are strongly encouraged to provide them. If provided, please include two to five keywords, separated by the pipe symbol, e.g:
+\keywords{Heat $|$ Mental Health $|$ Inequity $|$ Big Data}
 
 \begin{abstract}
-%Limit 150 words:
-
-%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
+  %Up to 250 words
 Higher temperatures associated with climate change are expected to have major impacts on human mental health, and previous work has found strong associations between heat and mental well-being.
-However, these studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race.
-We therefore use expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status available at fine spatial and temporal scales, and find stark differences in the effects of heat depending on neighborhood characteristics.
-Here we show that increased temperatures worsen expressed mood in all areas, but that this effect is much stronger in poor and Black neighborhoods.
+However, these previous studies typically used data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race.
+We therefore use expressed mood in a quarter of a billion geolocated tweets as a proxy for mental health status available at fine spatial and temporal scales, and pair each tweet with data on local temperature and neighborhood demographics.
+We show that heat affects expressed mood in all areas, but find that this effect is much stronger in poor and Black neighborhoods.
 We also find that the effect of heat on expressed mood is greatest in the early morning, supporting the hypothesis that heat affects mental health through a sleep quality pathway.
@@
 \end{abstract}
+
+\dates{This manuscript was compiled on \today}
+\doi{\url{www.pnas.org/cgi/doi/10.1073/pnas.XXXXXXXXXX}}
+
 \begin{document}
 
-\raggedbottom
 \maketitle
-\thispagestyle{empty}
+\thispagestyle{firststyle}
+\ifthenelse{\boolean{shortarticle}}{\ifthenelse{\boolean{singlecolumn}}{\abscontentformatted}{\abscontent}}{}
 
-\section*{Main}
+% If your first paragraph (i.e.
+with the \dropcap) contains a list environment (quote, quotation, theorem, definition, enumerate, itemize...), the line after the list may have some extra indentation.
+If this is the case, add \parshape=0 to the end of the list environment.
+
 %Overview
-Climate change is impacting many aspects of human well-being.
+\dropcap{C}limate change is impacting many aspects of human well-being.
 A large body of research has explored the ongoing impacts of climate change on agricultural outcomes, economic growth, and physical health \cite{pachauri2014climate}.
-However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr}.
+However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr, Berrang-Ford2021}.
 While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \cite{Berry2018Apr, Collins2011Jul}.
@@
 Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.
-Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007}.
-Additionally, exposure to higher temperatures can affect productivity and income \cite{Burke2015Nov}, leading to second-order impacts on mental health outcomes.
+Other studies point to worsened mental health as a second-order effect of the direct impacts of increased temperature on overall physical health and risk of injury \cite{Berry2007}, as well as productivity and income \cite{Burke2015Nov}.
 
@@
 The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicates surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as potential adaptations to future heat stress.
-For example, a large study of temperature and suicide in the USA and Mexico found "no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
+For example, a large study of temperature and suicide in the USA and Mexico found ``no significant difference in suicide response to temperature between rich and poor municipalities or counties'' \cite{Burke2018Aug}.
 Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \cite{Mullins2019Dec}.
 A major challenge for these studies, however, is their reliance on data that is typically aggregated to the municipality or county level.
-As a consequence, these studies are often restricted to between-county metrics of vulnerability and cannot take into account neighborhood effects.
+As a consequence, these studies are often restricted to between-county metrics of vulnerability and cannot take into account heterogeneity at more local scales.
 
 These findings of uniform vulnerability to heat are surprising given that minority groups and low-income people are more likely to be exposed to undesirable temperatures and less able to mitigate the effects of higher heat.
-For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}.
+For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}; more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}; and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}.
 Many of these factors that expose low-income people to heat may be exacerbated and compounded for racial minorities.
@@
 Additionally, policymakers have historically been less likely to take measures to address environmental hazards in Black neighborhoods \cite{Banzhaf2019}.
-Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
 Thus, there are theoretical reasons to expect heterogeneity in the impact of heat on mental health outcomes for different populations, particularly where socioeconomic, racial and other social inequities intersect, although there is little evidence of this to date.
@@
 
-\section*{Effect of Heat on Sentiment}
+\section*{Results}
+\subsection*{Effect of Heat on Sentiment}
 We examine the relationship between the mood expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
@@
  \centering
- \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
+ \includegraphics[width=\linewidth]{../../res/wbgt.png}
  \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and expressed mood in tweets. As temperatures increase above 5\textdegree C WBGT, mood rapidly declines.}
@@
 
-We find that mean neighborhood income strongly moderates the relationship between temperature and mood (Fig.
-\ref{fig:hetero}), with large differences in mood between the poorest (5th percentile) and wealthiest (95th percentile) neighborhoods.
+We find that mean neighborhood income strongly moderates the relationship between temperature and mood (Figs.
+\ref{fig:income}), with large differences in mood between the poorest (5th percentile) and wealthiest (95th percentile) neighborhoods.
 As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), mood increases in the wealthiest neighborhoods but decreases in the median and lowest income neighborhoods.
@@
 We also explored how neighborhood racial characteristics affect the relationship between temperature and mood, and we find the effects of heat are felt disproportionately in neighborhoods that are majority Black (Fig.
-\ref{fig:hetero}).
+\ref{fig:race}).
 Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the mood of tweets in majority Black neighborhoods decrease four times as much as the mood of people in other neighborhoods.
@@
 \centering
- \includegraphics[width=\linewidth]{../res/wbgt_combined.png}
- \caption{Effect of changes in WBGT on expressed mood as moderated by neighborhood income (a) and race (b).}
-\label{fig:hetero}
+ \includegraphics[width=0.8\linewidth]{../../res/wbgt-income.png}
+ \caption{Effect of changes in WBGT on expressed mood as moderated by neighborhood income per capita.}
+\label{fig:income}
 \end{figure}
 
-Finally, given that both low-income and majority black neighborhoods are more vulnerable to heat waves, we model the combined effects of race and income together to explore the intersection of these two types of vulnerability.
-Specifically, we model how the relationship between temperature and mood is moderated by the interaction between a neighborhood's per-capita income and the percentage of a neighborhood's inhabitants that are black.
-Additionally, we compare AICs for models with no term for vulnerability, a race term for vulnerability, an income term for vulnerability, and both race and income terms for vulnerability.
+\begin{figure}[H]
+\centering
+ \includegraphics[width=0.8\linewidth]{../../res/wbgt-race_q.png}
+ \caption{Effect of changes in WBGT on expressed mood as moderated by neighborhood majority race.}
+\label{fig:race}
+\end{figure}
 
-\begin{minipage}{.65\textwidth}
-\centering
-\includegraphics[width=\textwidth]{../res/wbgt-income-race.png}
-\captionof{figure}{Relationship between WBGT and mood from text in tweets, moderated by the interaction of race and income.}
-\label{fig:race-income}
-\end{minipage}\hfill
-\begin{minipage}{.35\textwidth}
-\centering
-\begin{tabular}{| l | l |}
-\hline
-Terms & AIC \\
-\hline
-Neither & 277,270,561 \\
-Income & 277,182,239 \\
-Race & 277,143,579 \\
-Both & 277,108,676 \\
-\hline
-\end{tabular}
-\captionof{table}{Akaike information criterion (AIC) of models with interaction terms for race, income, both race and income, as well as neither race nor income.
-The model with the best fit by AIC includes both race and income interaction terms.}
-\label{tab:aics}
-\end{minipage}
-
-Graphing the combined effects of race and income, we find that race has a larger impact on vulnerability than income (see Fig.
-\ref{fig:race-income}).
-A Black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants.
-Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and mood.
-Using AIC as an indicator of how different terms affect model quality (see Tab.
-\ref{tab:aics}), we find that a model including both terms performs the best, indicating that both race and income matter for the impact of heat on mood.
-However, including only a race term improves the model more than including only an income term, further supporting the conclusion that race is more important than income for explaining the vulnerability of mood to heat.
-
-\section*{Comparison With Other Events}
-We compared the expressed mood from exposure to temperatures to two other events associated with impacts on mood and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact of a major natural disaster (see Fig.
+\subsection*{Comparison With Other Events}
+We compared the change in expressed mood from heat exposure to two other events associated with impacts on mood and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact of a major natural disaster (see Fig.
 \ref{fig:compare}).
@@
  \centering
- \includegraphics[width=0.66\linewidth]{../res/comparison_plot.png}
- \caption{Decline in mood as temperature changes from mild (5\textdegree C WBGT) high (25\textdegree C WBGT) across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.}
+ \includegraphics[width=0.9\linewidth]{../../res/comparison_plot.png}
+ \caption{Decline in mood as temperature changes from mild (5\textdegree C WBGT) to high (25\textdegree C WBGT) across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.}
  \label{fig:compare}
@@
 Conversely, the effects of exposure to increasing temperatures in poorer and majority Black neighborhoods are much larger than the average weekly change in mood.
-Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on mood from experiencing a major hurricane.
+Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the average impact on mood of experiencing a major hurricane.
 
-\section*{Effect of Heat by Time of Day}
+\subsection*{Effect of Heat by Time of Day}
 In addition to providing data with high spatial resolution, Twitter data also comes with very high temporal resolution as each tweet is time-stamped.
@@
  \centering
- \includegraphics[width=0.5\linewidth]{../res/ts_heat.png}
+ \includegraphics[width=0.8\linewidth]{../../res/ts_heat.png}
  \caption{Effects of rising temperatures on mood by hour of the day, with a 95\% confidence interval.
@@
 Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of expressed mood and mental health at specific time scales and following certain events.
-This suggests that the changes in expressed mood we observe at neighborhood scales is indicative of other mental health outcomes such as suicides and hospitalizations.
+This suggests that the changes in expressed mood we observe at neighborhood scales is indicative of other mental health outcomes such as suicides and hospitalizations that cannot be measured with neighborhood-scale precision.
 
@@
 Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income.
-Such analyses have led to conclusions that the mental health effects of climate change will be uniform, least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
+Such analyses have led to conclusions that the mental health effects of climate change will be uniform, at least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
 For example, Mullins et al.
-conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}.
+conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health'' \cite{Mullins2019Dec}.
 Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income.
@@
 This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts.
-Additionally, it shows how measuring outcomes at fine spatial scales necessary for understanding how factors like race and income matter for vulnerability \cite{Tong2021}.
+Additionally, it shows how measuring outcomes at fine spatial scales is necessary for understanding how factors like race and income matter for vulnerability \cite{Tong2021}.
 Even within a relatively wealthy nation, we find that high temperatures affect certain populations to the same degree as a major hurricane.
-Thus, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
+Thus, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heat waves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
 
-In addition to exploring heterogeneities in vulnerability, using twitter data lets us explore how the effect of heat varies by time of date and infer possible important pathways in the heat-mental health relationship.
+In addition to exploring heterogeneities in vulnerability, using twitter data lets us explore how the effect of heat varies by time of day and infer possible important pathways in the heat-mental health relationship.
 Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis largely supports this hypothesis.
@@
 For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
-These issue may in fact lead us to under-estimate the true effect of heat on mental health for poor and Black people.
+These issues may in fact lead us to under-estimate the true effect of heat on mental health for poor and Black people.
 Another issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
@@
 While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they will not also be highly unevenly distributed.
-Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect.
+Thus, there is strong theoretical justification behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect.
 By using fine-scale Twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States.
@@
 
-\section*{Methods}
+\section*{Materials and methods}
 % from NCC: methods section should be written as concisely as possible and "typically do not exceed 3,000 words"
@@
 
-Expressed mood is a widely used metric to assess mental health and well-being from posts on social media platforms.
+Expressed mood (also called sentiment) is a widely used metric to assess mental health and well-being from posts on social media platforms.
 We use the VADER metric to assess the mood of Twitter posts as it was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
-The VADER (Valence Aware Dictionary for Sentiment Reasoning) corpus \cite{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to moods expressed on social media and includes several features, as:
-
-\begin{itemize}
- \item Incorporating lexical features common to informal media, such as slang (``sux"), and acronyms (``lol").
- \item Negations (``\textit{not} good", ``\textit{wasn't} bad")
- \item Punctuation (``Good!!!")
- \item Word shape, such as capitalization (``The movie was AMAZING")
- \item Emoticons and emojis (``:-)", ``\emojismile")
- \item Degree modifiers (``very excellent" or ``kind of crappy")
-\end{itemize}
+The VADER (Valence Aware Dictionary for Sentiment Reasoning) corpus \cite{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to moods expressed on social media and includes several features, as: incorporating lexical features common to informal media, such as slang (``sux''), and acronyms (``lol''); negations (``\textit{not} good'', ``\textit{wasn't} bad''); punctuation (``Good!!!''); word shape, such as capitalization (``The movie was AMAZING''); emoticons and emojis (``:-)''); and degree modifiers (``very excellent'' or ``kind of crappy'').
 
 The VADER method yields a value for the mood of a tweet, with a score of 0 for neutral tweets, a score $> 0$ for positive-mood tweets and a score $< 0$ for negative-mood tweets.
-While we ultimately settled on using VADER for assigning sentiment scores, we also conducted analyses using the Hedonometer and AFINN mood analysis methods, with similar results (See Supplement).
+We also conducted analyses using the Hedonometer and AFINN mood analysis methods, with similar results (See Supplement).
 
@@
 \cite{heo2019comparison}.
-Throughout the paper we give the median dry bulb temperature observed at each given wet bulb temp degree across our data set.
+Throughout the paper we give the median dry bulb temperature observed at each given wet bulb degree across our data set.
 It should be noted, however, that a wide range of dry bulb temperatures can be associated with a given wet bulb globe temperature, depending on humidity, wind speed, and sunlight.
@@
 We use the mean annual income per capita from the ACS, standardized so that values for each year are in 2019 dollars.
-For racial categories, we combined the various categories provided by the ACS into four racial groups: non-Hispanic white, non-Hispanic Black, Hispanic of any race, and an "other" category for non-Hispanic people who were neither Black nor white, such as Asian, Native American, or mixed-race people.
+For racial categories, we combined the various categories provided by the ACS into four racial groups: non-Hispanic white, non-Hispanic Black, Hispanic of any race, and an ``other'' category for non-Hispanic people who were neither Black nor white, such as Asian, Native American, or mixed-race people.
 
@@
 \begin{equation}
- y = \beta_0 + f_t(t) + f_{mt}(m t) + \beta_p p + \beta_{mp} m p + \beta_s s + \beta_{ms} m s + \Phi + \epsilon
+  \begin{align*}
+    y = & \beta_0 + f_t(t) + f_{mt}(m t) + \beta_p p + \beta_{mp} m p + \\
+       & \beta_s s + \beta_{ms} m s + \Phi + \epsilon
  \label{mod:2}
+  \end{align*}
 \end{equation}
+
+% \begin{align*}
+% (x+y)^3&=(x+y)(x+y)^2\\
+%        &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
+%        &=x^3+3x^2y+3xy^3+x^3. 
+% \end{align*}
 
@@
 These alternative specifications and analysis of rainfall and sunshine effects are available in the Supplement.
-
-In addition to our models exploring how income and race moderate the effect of heat on mood individually, we fit an additional model with an interaction terms for both race and income.
-Due to the complexity of a model with segmented terms interacting with two different variables, using a categorical variable for race would have yielded a model matrix too large for the largest cloud virtual machine available to us.
-Thus, we used a continuous metric for both income as well as race (percent black).
-This model extends our main model to the following form, interacting income and race with a temperature segmented spline as well as precipitation in sunshine:
-
-\begin{equation}
- y = \beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) +  \beta_p p + \beta_{m_{r} p} m_{r} p + \beta_{m_{i} p} m_{i} p + \beta_s s + \beta_{m_{i} s} m_{i} s +  \beta_{m_{r} s} m_{r} s + 
- \Phi + \epsilon
- \label{mod:race-inc}
-\end{equation}
-
-In this new model, $m_{i}$ is average income in the census block where the tweet originated, while $m_{r}$ is the racial composition of the census block.
 
@@
 
-\section*{Acknowledgements}
-%Official SESYNC wording
+\section*{Acknowledgments}
 This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
@@
 
-This data came from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
+This data came from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API.
 The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and this work complies with all the terms of service for Twitter and UVM.
 
-%\printbibliography
+\section*{References}
+
+% \section*{Guide to using this template on Overleaf}
+% 
+% Please note that whilst this template provides a preview of the typeset manuscript for submission, to help in this preparation, it will not necessarily be the final publication layout.
+For more detailed information please see the \href{http://www.pnas.org/site/authors/format.xhtml}{PNAS Information for Authors}.
+% 
+% If you have a question while using this template on Overleaf, please use the help menu (``?'') on the top bar to search for \href{https://www.overleaf.com/help}{help and tutorials}.
+You can also \href{https://www.overleaf.com/contact}{contact the Overleaf support team} at any time with specific questions about your manuscript or feedback on the template.
+% 
+% \subsection*{Submitting Manuscripts}
+% 
+% All authors must submit their articles at \href{http://www.pnascentral.org/cgi-bin/main.plex}{PNAScentral}.
+If you are using Overleaf to write your article, you can use the ``Submit to PNAS'' option in the top bar of the editor window.
+% 
+% \subsection*{Manuscript Length}
+% 
+% PNAS generally uses a two-column format averaging 67 characters, including spaces, per line.
+The maximum length of a Direct Submission research article is six pages and a Direct Submission Plus research article is ten pages including all text, spaces, and the number of characters displaced by figures, tables, and equations.
+When submitting tables, figures, and/or equations in addition to text, keep the text for your manuscript under 39,000 characters (including spaces) for Direct Submissions and 72,000 characters (including spaces) for Direct Submission Plus.
+% 
+% \subsection*{References}
+% 
+% References should be cited in numerical order as they appear in text; this will be done automatically via bibtex, e.g.
+\cite{belkin2002using} and \cite{berard1994embedding,coifman2005geometric}.
+All references should be included in the main manuscript file.
+% 
+% \begin{figure}%[tbhp]
+% \centering
+% \includegraphics[width=.8\linewidth]{frog}
+% \caption{Placeholder image of a frog with a long example caption to show justification setting.}
+% \label{fig:frog}
+% \end{figure}
+% 
+% 
+% \begin{SCfigure*}[\sidecaptionrelwidth][t]
+% \centering
+% \includegraphics[width=11.4cm,height=11.4cm]{frog}
+% \caption{This caption would be placed at the side of the figure, rather than below it.}\label{fig:side}
+% \end{SCfigure*}
+% 
+% \subsection*{Digital Figures}
+% 
+% Only TIFF, EPS, and high-resolution PDF for Mac or PC are allowed for figures that will appear in the main text, and images must be final size.
+Authors may submit U3D or PRC files for 3D images; these must be accompanied by 2D representations in TIFF, EPS, or high-resolution PDF format.
+Color images must be in RGB (red, green, blue) mode.
+Include the font files for any text.
+% 
+% Figures and Tables should be labelled and referenced in the standard way using the \verb|\label{}| and \verb|\ref{}| commands.
+% 
+% Figure \ref{fig:frog} shows an example of how to insert a column-wide figure.
+To insert a figure wider than one column, please use the \verb|\begin{figure*}...\end{figure*}| environment.
+Figures wider than one column should be sized to 11.4 cm or 17.8 cm wide.
+Use \verb|\begin{SCfigure*}...\end{SCfigure*}| for a wide figure with side captions.
+% 
+% \subsection*{Tables}
+% In addition to including your tables within this manuscript file, PNAS requires that each table be uploaded to the submission separately as a “Table” file.
+Please ensure that each table .tex file contains a preamble, the \verb|\begin{document}| command, and the \verb|\end{document}| command.
+This is necessary so that the submission system can convert each file to PDF.
+% 
+% \subsection*{Single column equations}
+% 
+% Authors may use 1- or 2-column equations in their article, according to their preference.
+% 
+% To allow an equation to span both columns, use the \verb|\begin{figure*}...\end{figure*}| environment mentioned above for figures.
+% 
+% Note that the use of the \verb|widetext| environment for equations is not recommended, and should not be used. 
+% 
+% \begin{figure*}[bt!]
+% \begin{align*}
+% (x+y)^3&=(x+y)(x+y)^2\\
+%        &=(x+y)(x^2+2xy+y^2) \numberthis \label{eqn:example} \\
+%        &=x^3+3x^2y+3xy^3+x^3. 
+% \end{align*}
+% \end{figure*}
+% 
+% 
+% \begin{table}%[tbhp]
+% \centering
+% \caption{Comparison of the fitted potential energy surfaces and ab initio benchmark electronic energy calculations}
+% \begin{tabular}{lrrr}
+% Species & CBS & CV & G3 \\
+% \midrule
+% 1. Acetaldehyde & 0.0 & 0.0 & 0.0 \\
+% 2. Vinyl alcohol & 9.1 & 9.6 & 13.5 \\
+% 3. Hydroxyethylidene & 50.8 & 51.2 & 54.0\\
+% \bottomrule
+% \end{tabular}
+% 
+% \addtabletext{nomenclature for the TSs refers to the numbered species in the table.}
+% \end{table}
+% 
+% \subsection*{Supporting Information (SI)}
+% 
+% Authors should submit SI as a single separate PDF file, combining all text, figures, tables, movie legends, and SI references.
+PNAS will publish SI uncomposed, as the authors have provided it.
+Additional details can be found here: \href{http://www.pnas.org/page/authors/journal-policies}{policy on SI}.
+For SI formatting instructions click \href{https://www.pnascentral.org/cgi-bin/main.plex?form_type=display_auth_si_instructions}{here}.
+The PNAS Overleaf SI template can be found \href{https://www.overleaf.com/latex/templates/pnas-template-for-supplementary-information/wqfsfqwyjtsd}{here}.
+Refer to the SI Appendix in the manuscript at an appropriate point in the text.
+Number supporting figures and tables starting with S1, S2, etc.
+% 
+% Authors who place detailed materials and methods in an SI Appendix must provide sufficient detail in the main text methods to enable a reader to follow the logic of the procedures and results and also must reference the SI methods.
+If a paper is fundamentally a study of a new method or technique, then the methods must be described completely in the main text.
+% 
+% \subsubsection*{SI Datasets} 
+% 
+% Supply Excel (.xls), RTF, or PDF files. This file type will be published in raw format and will not be edited or composed.
+% 
+% 
+% \subsubsection*{SI Movies}
+% 
+% Supply Audio Video Interleave (avi), Quicktime (mov), Windows Media (wmv), animated GIF (gif), or MPEG files and submit a brief legend for each movie in a Word or RTF file.
+All movies should be submitted at the desired reproduction size and length.
+Movies should be no more than 10 MB in size.
+% 
+% 
+% \subsubsection*{3D Figures}
+% 
+% Supply a composable U3D or PRC file so that it may be edited and composed. Authors may submit a PDF file but please note it will be published in raw format and will not be edited or composed.
+% 
+% 
+% \matmethods{Please describe your materials and methods here.
+This can be more than one paragraph, and may contain subsections and equations as required.
+Authors should include a statement in the methods section describing how readers will be able to access the data in the paper.
+% 
+% \subsection*{Subsection for Method}
+% Example text for subsection.
+% }
+% 
+% \showmatmethods{} % Display the Materials and Methods section
+% 
+% \acknow{Please include your acknowledgments here, set in a single paragraph. Please do not include any acknowledgments in the Supporting Information, or anywhere else in the manuscript.}
+% 
+% \showacknow{} % Display the acknowledgments section
+
+% Bibliography
 \bibliography{TextDataClimateShocks}
 
-\section*{Author Contributions}
-Author contributions: M.C., J.O.-G., A.S.
-and P.B.
-designed the research and modeling strategy; A.S.
-provided data; M.C and Z.L.
-prepared data; M.C.
-and J.O.-G.
-analyzed data; and M.C., J.O.-G., P.A.W., and J.L.
-wrote the paper.
-
-\noindent This paper was the result of a SESYNC graduate pursuits project co-lead by M.C. and J.O.-G.
-
 \end{document}

### 2021-08-19 doc/pnas/supplement/PNAS-template-supplement.tex
pnas submission

\documentclass[9pt,twoside,lineno]{pnas-new}
% Use the lineno option to display guide line numbers if required.

\templatetype{pnassupportinginfo}

\title{Fine-scale Data Reveals Neighborhood Inequalities In Effects of Heat on Expressed Mood}
\author{Matthew Cooper, Jeremiah Osborne-Gowey, Zheng Liu, Portia Adade Williams, Jie Liu, Aaron Schwartz, Patrick Baylis} 
\correspondingauthor{Matthew Cooper \\E-mail: mcooper@hsph.harvard.edu}

\begin{document}

%% Comment out or remove this line before generating final copy for submission; this will also remove the warning re: "Consecutive odd pages found".
%\instructionspage

\maketitle

%% Adds the main heading for the SI text. Comment out this line if you do not have any supporting information text.
\SItext

\section*{Alternative Sentiment Algorithms to Estimate Mood}

\subsection*{Hedonometer}

In addition to the Valence Aware Dictionary for Sentiment Reasoning (VADER) method \cite{gilbert_vader_2014}, we also used the hedonometer method \cite{Dodds2011Dec}.  This metric yielded very similar findings to those presented in the main paper with the VADER method.

The Hedonometer is a corpus-based technique to get mood scores from multi-lingual texts. The core steps are (1) building human evaluations of the happiness of a set of individual words, and (2) using an algorithm for scaling up from individual words to texts. For the English corpus, Dodds and others \cite{Dodds2011Dec} collect 10,222 unique words and used crowd-sourcing platform Amazon Mechanical Turk to get human evaluation of happiness degree for each word in an integer scale from 1 to 9, representing a sad to happy spectrum. Score 5 represents neutral words. Some illustrative example of words are:

\[h_{avg} (\text{laughter}) = 8.50 \]
\[h_{avg} (\text{the}) = 4.98\]
\[h_{avg} (\text{hate}) = 2.34\]

The score for single text will be the mean happiness score for all words in the text.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/hedono-wbgt.png}
  \caption{Effects of temperature on mood, measured with the hedonometer method.}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/hedono-wbgt-income.png}
  \caption{Effects of temperature on mood and moderated by income, measured with the hedonometer method.}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/hedono-wbgt-race_q.png}
  \caption{Effects of temperature on mood and moderated by race, measured with the hedonometer method.}
\end{figure}

\newpage
\subsection*{AFINN}

AFINN is another wordlist used in mood analysis and designed for microblogs like twitter \cite{nielsen2011new}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/afinn-wbgt.png}
  \caption{Effects of temperature on mood, measured with the AFINN method.}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/afinn-wbgt-income.png}
  \caption{Effects of temperature on mood and moderated by income, measured with the AFINN method.}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/afinn-wbgt-race_q.png}
  \caption{Effects of temperature on mood and moderated by race, measured with the AFINN method.}
\end{figure}

\newpage
\section*{Alternative Interaction Variables}
For our main analysis, we treated income as a continuous (log-transformed) variable and race as a categorical variable.  We ran similar analyses with income as a categorical variable and race as a continuous variable (the percent of a census block's population that was non-white/minority).

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/wbgt-income_q.png}
  \caption{Effect of temperature on mood and moderated by income, with income as a categorical variable.}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/wbgt-race.png}
  \caption{Effect of temperature on mood and moderated by race, with race as a continuous variable.}
\end{figure}

\newpage
\section*{Effects of Precipitation and Sunshine}
While our main variable of interest was temperature as measured by Wet Bulb Globe Temperature, we also interacted our variables of interest with a binary variable for whether it was raining or not, as well as a continous variable for the amount of shortwave radiation in ($W/m^2$), or sunshine.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/raining-income.png}
  \caption{Effect of rain on mood and moderated by income.}
  \label{fig:timeseries}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/raining-race_q.png}
  \caption{Effect of rain on mood and moderated by race.}
  \label{fig:timeseries}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/srad-income.png}
  \caption{Effect of sunshine on mood and moderated by income.}
  \label{fig:timeseries}
\end{figure}

\begin{figure}[H]
  \centering
  \includegraphics[width=0.6\linewidth]{../../../res/srad-race_q.png}
  \caption{Effect of sunshine on mood and moderated by race.}
  \label{fig:timeseries}
\end{figure}

\newpage
\section*{Temperature Effects by Combined Statistical Area}
To explore whether cultural or environmental baselines may affect the relationship between heat and mood, we examine the impact of temperature on mood across Combined Statistical Areas (CSAs) and Metropolitan Statistical Areas (MSAs) containing over 1 million people (hereafter: city regions). We find that higher temperatures are associated with declining mood in 83.6\% of city regions (see Fig. \ref{fig:map}) and the effect is significant in 35.3\% of these. Furthermore, we find no significant association between higher temperatures and increased mood in any of the cities.

For this analysis by city, we run models with a similar form to Model 2, including the same fixed effects, although we fit a linear effect for $f_t()$.  Because we are not accounting for non-linearities in $f_t()$, we exclude tweets with a WBGT of less than 5\textdegree C, as our initial analysis showed that this is the temperature at which mood peaks.

\begin{figure}[H]
\centering
  \includegraphics[width=\linewidth]{../../../res/map_wbgt.png}
    \caption{The change in mood for a 25-degree increase in WBGT for city regions with a population of over 1 million.  For cities shown with a dark outline, the effect is significant at a 95\% confidence interval.}
  \label{fig:map}
\end{figure}

\newpage
\section*{Interaction between Race and Income}
Given that both low-income and majority black neighborhoods are more vulnerable to heat waves, we model the combined effects of race and income together to explore the intersection of these two types of vulnerability. Specifically, we model how the relationship between temperature and mood is moderated by the interaction between a neighborhood's per-capita income and the percentage of a neighborhood's inhabitants that are black. Additionally, we compare AICs for models with no term for vulnerability, a race term for vulnerability, an income term for vulnerability, and both race and income terms for vulnerability.

\begin{minipage}{.65\textwidth}
\centering
\includegraphics[width=\textwidth]{../../../res/wbgt-income-race.png}
\captionof{figure}{Relationship between WBGT and mood from text in tweets, moderated by the interaction of race and income.}
\label{fig:race-income}
\end{minipage}\hfill
\begin{minipage}{.35\textwidth}
\centering
\begin{tabular}{| l | l |}
\hline
Terms & AIC \\
\hline
Neither & 277,270,561 \\
Income & 277,182,239 \\
Race & 277,143,579 \\
Both & 277,108,676 \\
\hline
\end{tabular}
\captionof{table}{Akaike information criterion (AIC) of models with interaction terms for race, income, both race and income, as well as neither race nor income. The model with the best fit by AIC includes both race and income interaction terms.}
\label{tab:aics}
\end{minipage}

Graphing the combined effects of race and income, we find that race has a larger impact on vulnerability than income (see Fig. \ref{fig:race-income}). A Black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants. Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and mood. Using AIC as an indicator of how different terms affect model quality (see Tab. \ref{tab:aics}), we find that a model including both terms performs the best, indicating that both race and income matter for the impact of heat on mood. However, including only a race term improves the model more than including only an income term, further supporting the conclusion that race is more important than income for explaining the vulnerability of mood to heat.

\newpage
\section*{Weather Terms}
For all of our analysis, we excluded tweets that contained the following weather-related terms: aerovane, air, airstream, altocumulus, altostratus, anemometer, anemometers, anticyclone, anticyclones, arctic, arid, aridity, atmosphere, atmospheric, autumn, autumnal, balmy, baroclinic, barometer, barometers, barometric, blizzard, blizzards, blustering, blustery, blustery, breeze, breezes, breezy, brisk, calm, celsius, chill, chilled, chillier, chilliest, chilly, chinook, cirrocumulus, cirrostratus, cirrus, climate, climates, cloud, cloudburst, cloudbursts, cloudier, cloudiest, clouds, cloudy, cold, colder, coldest, condensation, contrail, contrails, cool, cooled, cooling, cools, cumulonimbus, cumulus, cyclone, cyclones, damp, damp, damper, damper, dampest, dampest, degree, degrees, deluge, dew, dews, dewy, doppler, downburst, downbursts, downdraft, downdrafts, downpour, downpours, dried, drier, dries, driest, drizzle, drizzled, drizzles, drizzly, drought, droughts, dry, dryline, fall, farenheit, flood, flooded, flooding, floods, flurries, flurry, fog, fogbow, fogbows, fogged, fogging, foggy, fogs, forecast, forecasted, forecasting, forecasts, freeze, freezes, freezing, frigid, frost, frostier, frostiest, frosts, frosty, froze, frozen, gale, gales, galoshes, gust, gusting, gusts, gusty, haboob, haboobs, hail, hailed, hailing, hails, haze, hazes, hazy, heat, heated, heating, heats, hoarfrost, hot, hotter, hottest, humid, humidity, hurricane, hurricanes, ice, iced, ices, icing, icy, inclement, landspout, landspouts, lightning, lightnings, macroburst, macrobursts, maelstrom, mercury, meteorologic, meteorologist, meteorologists, meteorology, microburst, microbursts, microclimate, microclimates, millibar, millibars, mist, misted, mists, misty, moist, moisture, monsoon, monsoons, mugginess, muggy, nexrad, nippy, NOAA, nor'easter, nor'easters, noreaster, noreasters, overcast, ozone, parched, parching, pollen, precipitate, precipitated, precipitates, precipitating, precipitation, psychrometer, radar, rain, rainboots, rainbow, rainbows, raincoat, raincoats, rained, rainfall, rainier, rainiest, raining, rains, rainy, sandstorm, sandstorms, scorcher, scorching, searing, shower, showering, showers, skiff, sleet, slicker, slickers, slush, slushy, smog, smoggier, smoggiest, smoggy, snow, snowed, snowier, snowiest, snowing, snowmageddon, snowpocalypse, snows, snowy, spring, sprinkle, sprinkles, sprinkling, squall, squalls, squally, storm, stormed, stormier, stormiest, storming, storms, stormy, stratocumulus, stratus, subtropical, summer, summery, sun, sunnier, sunniest, sunny, temperate, temperature, tempest, thaw, thawed, thawing, thaws, thermometer, thunder, thundered, thundering, thunders, thunderstorm, thunderstorms, tornadic, tornado, tornadoes, tropical, troposphere, tsunami, turbulent, twister, twisters, typhoon, typhoons, umbrella, umbrellas, vane, warm, warmed, warming, warms, warmth, waterspout, waterspouts, weather, wet, wetter, wettest, wind, windchill, windchills, windier, windiest, windspeed, windy, winter, wintery, and wintry.

\bibliography{TextDataClimateShocks}

\end{document}

### 2021-08-19 doc/temp-sentiment.tex
pnas submission

@@
 \author{}
-% \author[1,*]{Matthew Cooper}
-% \author[2]{Jeremiah Osborne-Gowey}
-% \author[3]{Zheng Liu}
-% \author[4]{Portia Adade Williams}
-% \author[5]{Jie Liu}
-% \author[6]{Aaron Schwartz}
-% \author[7]{Patrick Baylis} 
-% 
-% \affil[1]{T.H. Chan School of Public Health, Harvard University, U.S.A.}
-% \affil[2]{Environmental Studies Program, University of Colorado Boulder, U.S.A.}
-% \affil[3]{Department of Geographical Sciences, University of Maryland College Park, U.S.A.}
-% \affil[4]{CSIR-Science and Technology Policy Research Institute, Ghana}
-% \affil[5]{School of Business, East China University of Science and Technology, China} 
-% \affil[6]{Department of Ecology \& Evolutionary Biology, University of Colorado Boulder, U.S.A.}
-% \affil[7]{Vancouver School of Economics, University of British Columbia, Canada}
-% \affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
+\author[1,*]{Matthew Cooper}
+\author[2]{Jeremiah Osborne-Gowey}
+\author[3]{Zheng Liu}
+\author[4]{Portia Adade Williams}
+\author[5]{Jie Liu}
+\author[6]{Aaron Schwartz}
+\author[7]{Patrick Baylis}
+
+\affil[1]{T.H. Chan School of Public Health, Harvard University, U.S.A.}
+\affil[2]{Environmental Studies Program, University of Colorado Boulder, U.S.A.}
+\affil[3]{Department of Geographical Sciences, University of Maryland College Park, U.S.A.}
+\affil[4]{CSIR-Science and Technology Policy Research Institute, Ghana}
+\affil[5]{School of Business, East China University of Science and Technology, China} 
+\affil[6]{Department of Ecology \& Evolutionary Biology, University of Colorado Boulder, U.S.A.}
+\affil[7]{Vancouver School of Economics, University of British Columbia, Canada}
+\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
 
@@
 
-\section*{Acknowledgements}
+% \section*{Acknowledgements}
 %Official SESYNC wording
-This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
+% This work was supported by the National Socio-Environmental Synthesis Center (SESYNC) under funding received from the National Science Foundation DBI-1639145.
 Additionally, computational resources for this project were provided for the Microsoft Azure cloud through Microsoft's AI for Earth program.
 
-This data came from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
+% This data came from Twitter via the University of Vermont’s (UVM) agreement with Twitter to access its streaming API - colloquially referred to as the Decahose.
 The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and this work complies with all the terms of service for Twitter and UVM.
@@
 
-\noindent This paper was the result of a SESYNC graduate pursuits project co-lead by M.C. and J.O.-G.
+% \noindent This paper was the result of a SESYNC graduate pursuits project co-lead by M.C. and J.O.-G.
 

### 2021-08-19 viz/urban_poverty_tab.R

@@ -1,0 +1,73 @@
+library(sf)
+library(tidyverse)
+
+setwd('~/tweets/')
+
+#Because it is 5-year average, use the middle year of the 5.  So 2011 is 2007-2011, so most accurate to 2009
+
+year <- 2012
+data_year <- ifelse(year < 2017, year + 2, 2018)
+
+sp <- read_sf(paste0('wealth/nhgis0002_shape/US_blck_grp_', data_year, '.shp'))
+
+sp_pts <- st_centroid(sp)
+
+urb_rur <- read_sf('wealth/tl_2017_us_uac10.shp')
+
+st_pts_pj <- st_transform(sp_pts, st_crs(urb_rur))
+
+cts <- st_within(st_pts_pj, urb_rur)
+
+sp$urban <- sapply(cts, length)
+
+dat <- read.csv(list.files(path='wealth/nhgis0003_csv/',
+                           full.names=T,
+                           pattern=paste0(data_year, '5_', data_year, '.*grp_E.csv')))
+
+
+names(dat)[38:94] <- c("TOTAL_POPULATION", "RACE_TOT", "RACE_NHISP_TOTXX", "RACE_WHITE", 
+                       "RACE_BLACK", "RACE_INDIAN", "RACE_ASIAN", "RACE_OTHER1", "RACE_OTHER2",
+                       "RACE_OTHER3", "RACE_OTHER4", "RACE_OTHER5", "RACE_HISP", 
+                       "RACE_HISPXX1", "RACE_HISPXX2", "RACE_HISPXX3", "RACE_HISPXX4", 
+                       "RACE_HISPXX5", "RACE_HISPXX6", "RACE_HISPXX7", "RACE_HISPXX8", 
+                       "RACE_HISPXX9", "TRAN_TOT", "TRAN_CAR1", "TRAN_CAR2", "TRAN_CAR3", 
+                       "TRAN_CAR4", "TRAN_CAR5", "TRAN_CAR6", "TRAN_CAR7", "TRAN_CAR8", 
+                       "TRAN_PUB1", "TRAN_PUB2", "TRAN_PUB3", "TRAN_PUB4", "TRAN_PUB5", 
+                       "TRAN_PUB6", "TRAN_CAR9", "TRAN_EXP1", "TRAN_EXP2", "TRAN_EXP3", 
[…]

### 2022-02-16 doc/temp-sentiment-supplement.tex
EHP submission

@@
 ]{biblatex}
+\usepackage{lineno}
+
+\linenumbers
 
@@
 \fancyhf{}
-\lfoot{Supplemental Materials for \textit{Fine-scale Data Reveals Neighborhood Inequalities Effects of Heat on Expressed Mood}}
+\lfoot{Supplemental Materials for \textit{Neighborhood Income and Race Moderate the Effect of Heat on Expressed Mood}}
 \rfoot{\thepage}
@@
 \begin{center}
-\section*{Supplemental Materials for \textit{Fine-scale Twitter Big Data Reveals Neighborhood Inequalities in Mental Health Effects of High Temperatures}}
+\section*{Supplemental Materials for \textit{Neighborhood Income and Race Moderate the Effect of Heat on Expressed Mood}}
 \end{center}
@@
 
+\newpage
 \section{Alternative Interaction Variables}
@@
 
+\newpage
 \section{Effects of Precipitation and Sunshine}
-While our main variable of interest was temperature as measured by Wet Bulb Globe Temperature, we also interacted our variables of interest with a binary variable for whether it was raining or not, as well as a continous variable for the amount of shortwave radiation in ($W/m^2$), or sunshine.
+While our main variable of interest was temperature as measured by Wet Bulb Globe Temperature, we also interacted our variables of interest with a binary variable for whether it was raining or not, as well as a continuous variable for the amount of shortwave radiation in ($W/m^2$), or sunshine.
 
@@
 
+\newpage
 \section*{Temperature Effects by Combined Statistical Area}
@@
 
+\newpage
+\section*{Interaction between Race and Income}
+Given that both low-income and majority black neighborhoods are more vulnerable to heat waves, we model the combined effects of race and income together to explore the intersection of these two types of vulnerability.
+Specifically, we model how the relationship between temperature and mood is moderated by the interaction between a neighborhood's per-capita income and the percentage of a neighborhood's inhabitants that are black.
+Additionally, we compare AICs for models with no term for vulnerability, a race term for vulnerability, an income term for vulnerability, and both race and income terms for vulnerability.
+
+\begin{minipage}{.65\textwidth}
+\centering
+\includegraphics[width=\textwidth]{../res/wbgt-income-race.png}
+\captionof{figure}{Relationship between WBGT and mood from text in tweets, moderated by the interaction of race and income.}
+\label{fig:race-income}
+\end{minipage}\hfill
+\begin{minipage}{.35\textwidth}
+\centering
+\begin{tabular}{| l | l |}
+\hline
+Terms & AIC \\
+\hline
+Neither & 277,270,561 \\
+Income & 277,182,239 \\
+Race & 277,143,579 \\
+Both & 277,108,676 \\
+\hline
+\end{tabular}
+\captionof{table}{Akaike information criterion (AIC) of models with interaction terms for race, income, both race and income, as well as neither race nor income.
+The model with the best fit by AIC includes both race and income interaction terms.}
+\label{tab:aics}
+\end{minipage}
+
+Graphing the combined effects of race and income, we find that race has a larger impact on vulnerability than income (see Fig.
+\ref{fig:race-income}).
+A Black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants.
+Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and mood.
+Using AIC as an indicator of how different terms affect model quality (see Tab.
+\ref{tab:aics}), we find that a model including both terms performs the best, indicating that both race and income matter for the impact of heat on mood.
+However, including only a race term improves the model more than including only an income term, further supporting the conclusion that race is more important than income for explaining the vulnerability of mood to heat.
+
+\newpage
 \section{Weather Terms}

### 2022-02-16 doc/tmp/temp-sentiment.tex
EHP submission

--- 2021-06-30 doc/temp-sentiment.tex
@@
-\documentclass[fleqn,10pt]{wlscirep}
+\documentclass[titlepage]{article}
 \usepackage[utf8]{inputenc}
@@
 \usepackage{xcolor}
-% \usepackage[
-% style=numeric,
-% citestyle=numeric-comp,
-% backend=biber,
-% doi=true,
-% natbib=true,
-% sorting=none
-% ]{biblatex}
+\usepackage[
+  style=alphabetic,
+  citestyle=authoryear,
+  backend=biber,
+  bibstyle=authortitle,
+  doi=true,
+  natbib=true,
+  sorting=nyt
+]{biblatex}
 \usepackage{scalerel,xparse}
+\usepackage{lineno}
+
+\DeclareFieldFormat{labelalphawidth}{#1}
+
+\bibliographystyle{plain}
+
+\linenumbers
+
+%\addbibresource{TextDataClimateShocks.bib}
 
@@
 
+\begin{document}
+
 %https://www.overleaf.com/project/5f918edd1afc380001feca0d
-\title{Fine-scale Data Reveals Neighborhood Inequalities In Effects of Heat on Expressed Mood}
+\title{The Effects of Neighborhood Income and Race In Moderating the Effect of Heat on Expressed Mood in the United States}
 
@@
 
-\affil[1]{T.H. Chan School of Public Health, Harvard University}
-\affil[2]{Environmental Studies Program, University of Colorado Boulder}
-\affil[3]{Department of Geographical Sciences, University of Maryland College Park}
-\affil[4]{CSIR-Science and Technology Policy Research Institute, Ghana}
-\affil[5]{School of Business, East China University of Science and Technology}
-\affil[6]{University of Colorado Boulder}
-\affil[7]{University of British Columbia}
-\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu}
+\affil[1]{T.H. Chan School of Public Health, Harvard University, Brookline, MA, U.S.A.}
+\affil[2]{Environmental Studies Program, University of Colorado Boulder, Boulder, CO, U.S.A.}
+\affil[3]{Department of Geographical Sciences, University of Maryland College Park, College Park, MD, U.S.A.}
+\affil[4]{CSIR-Science and Technology Policy Research Institute, Accra, Ghana}
+\affil[5]{School of Business, East China University of Science and Technology, Shanghai, China} 
+\affil[6]{Department of Ecology \& Evolutionary Biology, University of Colorado Boulder, Boulder, CO, U.S.A.}
+\affil[7]{Vancouver School of Economics, University of British Columbia, Vancouver, BC, Canada}
+\affil[*]{Corresponding Author: mcooper@hsph.harvard.edu, 677 Huntington Ave, Boston, MA 02115}
+
+\date{\parbox{\linewidth}{\centering The authors declare they have no actual or potential competing financial interests.}}
+
+\maketitle
 
 \begin{abstract}
-%Limit 150 words:
 
-%Actually there are lots of examples of articles with > 150 words, but I think we should keep it concise
-Higher temperatures associated with climate change are expected to have major impacts on human mental health, and previous work has found strong associations between heat and mental well-being.
+  \textbf{Background} Previous studies have found strong linkages between heat and many indicators of mental health.
+However, these studies have not found any heterogeneities in vulnerability based on income or other factors, in spite of the fact that there are strong theoretical reasons to expect the poor and other marginalized groups to be more vulnerable to higher temperates.
+
+  \textbf{Objectives} We tested the hypothesis that the impacts of heat on mental health are heterogeneous based on neighborhood characteristics.
+
+  \textbf{Methods} We used the expressed mood in 242 million geolocated tweets, paired with census block data on neighborhood characteristics and hourly weather data to test our hypothesis with a segmented regression.
+
+  \textbf{Results} We found that the effect of heat on expressed mood is much higher in low-income and majority Black neighborhoods.
+Additionally, we found that the effects of temperature on mood are greatest in the early morning.
+
+  \textbf{Discussion} Previous analysis using city-level data have been unable to find heterogeneities in vulnerability to heat, with consequences for our understanding of vulnerability and adaptation.
+We present the first evidence of neighborhood heterogeneity in the vulnerability of mental health to heat.
+Additionally, examining the timing of impacts of heat on mood suggests that sleep quality plays a role in the effects of heat on mood.
+
+%Higher temperatures associated with climate change are expected to have major impacts on human mental health, and previous work has found strong associations between heat and mental well-being.
 However, these studies typically use data reported at the city or county level, and, to date, have found no difference in vulnerability based on income or race.
@@
 \end{abstract}
-\begin{document}
-
-\raggedbottom
-\maketitle
-\thispagestyle{empty}
 
@@
 Climate change is impacting many aspects of human well-being.
-A large body of research has explored the ongoing impacts of climate change on agricultural outcomes, economic growth, and physical health \cite{pachauri2014climate}.
-However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \cite{Berry2018Apr}.
-While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \cite{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \cite{Berry2018Apr, Collins2011Jul}.
-Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \cite{Schwartz2017Aug}, as well as linkages between higher temperatures and a variety of mental health outcomes \cite{baylis_weather_2018, Mullins2019Dec, Obradovich2018Oct}.
-Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \cite{allen2019technical}, more research is needed into the impacts of increased heat on mental health, especially for identifying the most vulnerable communities.
+A large body of research has explored the ongoing impacts of climate change on agricultural outcomes, economic growth, and physical health \citep{pachauri2014climate}.
+However, academic and medical researchers have highlighted the relative paucity of empirical studies on the impacts of climate change on mental health \citep{Berry2018Apr, Berrang-Ford2021}.
+While mental health is under-studied, it accounts for up 13\% of the global burden of disease, representing a major share of total human suffering \citep{Collins2011Jul}, leading to calls for more research into potential linkages between climate change and mental health \citep{Berry2018Apr, Collins2011Jul}.
+Researchers have begun to answer these calls, finding associations between climate disasters and post-traumatic stress \citep{Schwartz2017Aug}, as well as linkages between higher temperatures and a variety of mental health outcomes \citep{baylis_weather_2018, Mullins2019Dec, Obradovich2018Oct}.
+Nevertheless, with average global temperatures projected to increase by 1.5\textdegree C within a decade \citep{allen2019technical}, more research is needed into the impacts of increased heat on mental health, especially for identifying the most vulnerable communities.
 
 %Theoretical: frameworks and pathways
-Researchers have proposed a number pathways by which temperature can affect mental health \cite{Berry2018Apr, Palinkas2020Apr}.
-Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \cite{Lohmus2018Jul}.
-Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \cite{Berry2007}.
-Additionally, exposure to higher temperatures can affect productivity and income \cite{Burke2015Nov}, leading to second-order impacts on mental health outcomes.
+Researchers have proposed a number pathways by which temperature can affect mental health \citep{Berry2018Apr, Palinkas2020Apr}.
+Work on biological mechanisms emphasizes the mental health effects from dehydration and other consequences of maintaining stable body temperature in high heat \citep{Lohmus2018Jul}.
+Other studies point to linkages between increased temperatures, overall physical health and increases in injuries which can in turn affect mental health \citep{Berry2007}.
+Additionally, exposure to higher temperatures can affect productivity and income \citep{Burke2015Nov}, leading to second-order impacts on mental health outcomes.
 
@@
 Drawing on these frameworks, results from an emerging literature are beginning to document linkages between higher temperatures worsened mental health outcomes.
-Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \cite{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \cite{Qi2014Dec, Likhvar2011Jan}.
-Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \cite{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \cite{Lee2007Jan, Sung2013Feb}.
-Higher temperatures are also linked to increased mortality among people with pre-existing mental health disorders \cite{Hansen2008Oct}.
-Finally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \cite{Obradovich2017May, Mullins2019Dec}, although these studies have relied on retrospective survey data and have not used data from across the diurnal cycle.
+Multiple studies find strong evidence that higher temperatures are associated with increases in suicides in the United States \citep{Burke2018Aug, Mullins2019Dec, Dixon2007May} with other studies finding a similar relationship in locations around the world \citep{Qi2014Dec, Likhvar2011Jan}.
+Higher temperatures are also correlated with increased hospitalizations related to general mental health issues \citep{Obradovich2018Oct, Mullins2019Dec} and incidents related specific conditions like bipolar disorder and schizophrenia \citep{Lee2007Jan, Sung2013Feb}.
+Higher temperatures are also linked to increased mortality among people with pre-existing mental health disorders \citep{Hansen2008Oct}.
+Finally, recent studies find nighttime temperatures affect sleep quality, with consequences for mental health \citep{Obradovich2017May, Mullins2019Dec}, although these studies have relied on retrospective survey data and have not used data from across the diurnal cycle.
 
 The effects of heat exposure on mental health are by now relatively well established, yet results from research conducted at national scales indicates surprisingly little heterogeneity in impacts, with consequences for our understanding of groups that are most vulnerable as well as potential adaptations to future heat stress.
-For example, a large study of temperature and suicide in the USA and Mexico found "no significant difference in suicide response to temperature between rich and poor municipalities or counties" \cite{Burke2018Aug}.
-Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \cite{Mullins2019Dec}.
+For example, a large study of temperature and suicide in the USA and Mexico found "no significant difference in suicide response to temperature between rich and poor municipalities or counties" \citep{Burke2018Aug}.
+Similarly, a national study found no effect of income as modifier of the effect of temperature on suicides, emergency department visits, or self-reported mental health status across the USA \citep{Mullins2019Dec}.
 A major challenge for these studies, however, is their reliance on data that is typically aggregated to the municipality or county level.
@@
 These findings of uniform vulnerability to heat are surprising given that minority groups and low-income people are more likely to be exposed to undesirable temperatures and less able to mitigate the effects of higher heat.
-For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \cite{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \cite{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \cite{Samuelson2020Jun}.
+For example, people from poorer socio-economic backgrounds are more likely to work outside rather than indoors \citep{Gubernot2014Oct}, more likely to rely on public transportation, bicycling, or walking instead of commuting in their own air-conditioned car \citep{Karner2015Dec}, and more likely to live in housing that is less insulated with poorer quality air-conditioning \citep{Samuelson2020Jun}.
 Many of these factors that expose low-income people to heat may be exacerbated and compounded for racial minorities.
-For example, Black Americans face discrimination in employment \cite{Kang2016, Penner2008} and housing access \cite{Desmond2015, Akbar2019}, forcing them into jobs and homes that may expose them to more heat.
-Additionally, policymakers have historically been less likely to take measures to address environmental hazards in Black neighborhoods \cite{Banzhaf2019}.
-Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \cite{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \cite{ferre2019hurricane, Gruebner2015Jun}.
+For example, Black Americans face discrimination in employment \citep{Kang2016, Penner2008} and housing access \citep{Desmond2015, Akbar2019}, forcing them into jobs and homes that may expose them to more heat.
+Additionally, policymakers have historically been less likely to take measures to address environmental hazards in Black neighborhoods \citep{Banzhaf2019}.
+Some studies find large differences in vulnerability between neighborhoods for a variety of impacts of heat on physical and mental health \citep{Belanger2015Mar, Uejio2011Mar}, as well as for vulnerability to other types of climate shocks like hurricanes \citep{ferre2019hurricane, Gruebner2015Jun}.
 Thus, there are theoretical reasons to expect heterogeneity in the impact of heat on mental health outcomes for different populations, particularly where socioeconomic, racial and other social inequities intersect, although there is little evidence of this to date.
@@
 While public health data on mental health outcomes is not available at the neighborhood scale, spontaneous, in situ data from the social media platform Twitter can provide an important, inexpensive and timely proxy for mental health at extremely precise spatial and temporal scales.
-Many studies use the mood expressed in social media posts as an indicator of mental health \cite{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
-For example, studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder (PTSD) in individuals even before their formal diagnosis \cite{Reece2017Oct}.
-Similar methods applied to posts on Facebook can predict the onset of depression \cite{Eichstaedt2018Oct}.
-In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \cite{Kolliakou2020Feb}.
-Across the USA, more positive tweets are associated with a variety of metrics of human well-being \cite{Mitchell2013May}, and twitter has provided insight into population-level patterns of happiness over the course of the day \cite{Dodds2011}.
-Finally, the mood expressed in tweets has been strongly associated with local weather conditions \cite{baylis_weather_2018} and been used as evidence to support the temperature-suicide relationship \cite{Burke2018Aug}.
+Many studies use the mood expressed in social media posts as an indicator of mental health and well-being \citep{Edo-Osagie2020Jul, Sinnenberg2016Dec}.
+For example, studies have used Twitter data to identify the onset of Post Traumatic Stress Disorder (PTSD) in individuals even before their formal diagnosis \citep{Reece2017Oct}.
+Similar methods applied to posts on Facebook can predict the onset of depression \citep{Eichstaedt2018Oct}.
+In London, day-to-day changes in mental health indicators derived from the text of tweets were associated with changes in mental health crisis episodes \citep{Kolliakou2020Feb}.
+Across the USA, more positive tweets are associated with a variety of metrics of human well-being \citep{Mitchell2013May}, and twitter has provided insight into population-level patterns of happiness over the course of the day \citep{Dodds2011}.
+Finally, the mood expressed in tweets has been strongly associated with local weather conditions and air quality \citep{baylis_weather_2018, Zheng2019} and been used as evidence to support the temperature-suicide relationship \citep{Burke2018Aug}.
 Thus, posts on Twitter provide an proxy for mental health at fine enough spatial scales to examine neighborhood effects on vulnerability to higher temperatures.
 To explore how neighborhood conditions affect the relationship between heat and mental heath, we draw on a data set of 243 million geolocated tweets from across the continental United States covering the period 2009-2019, each paired with data on local weather and neighborhood conditions.
-
-\section*{Effect of Heat on Sentiment}
-We examine the relationship between the mood expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \cite{budd2008wet}.
-We find higher temperatures are associated with worsened mood (Fig.
-\ref{fig:wbgt}) after controlling for a variety of fixed effects across all tweets.
-Mood is highest at 5\textdegree C WBGT (associated in our data set with a median dry bulb temperature on 12\textdegree C/54\textdegree F) with the largest declines between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
-Mood also declines with colder temperatures (below 5\textdegree C WBGT), although only slightly.
-Dis-aggregating our analysis by metropolitan area and using alternative methods to estimate expressed mood all yield similar results (see Supplement).
-
-\begin{figure}[H]
- \centering
- \includegraphics[width=0.5\linewidth]{../res/wbgt.png}
- \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and expressed mood in tweets. As temperatures increase above 5\textdegree C WBGT, mood rapidly declines.}
- \label{fig:wbgt}
-\end{figure}
-
-We find that mean neighborhood income strongly moderates the relationship between temperature and mood (Fig.
-\ref{fig:hetero}), with large differences in mood between the poorest (5th percentile) and wealthiest (95th percentile) neighborhoods.
-As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), mood increases in the wealthiest neighborhoods but decreases in the median and lowest income neighborhoods.
-The wealthiest neighborhoods do not see decreases in mood until temperatures exceed 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point mood decreases relatively evenly regardless of neighborhood income percentile.
-
-We also explored how neighborhood racial characteristics affect the relationship between temperature and mood, and we find the effects of heat are felt disproportionately in neighborhoods that are majority Black (Fig.
-\ref{fig:hetero}).
-Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the mood of tweets in majority Black neighborhoods decrease four times as much as the mood of people in other neighborhoods.
-Additionally, at mild to warm temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower mood than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
-
-\begin{figure}[H]
-\centering
- \includegraphics[width=\linewidth]{../res/wbgt_combined.png}
- \caption{Effect of changes in WBGT on expressed mood as moderated by neighborhood income (a) and race (b).}
-\label{fig:hetero}
-\end{figure}
-
-Finally, given that both low-income and majority black neighborhoods are more vulnerable to heat waves, we model the combined effects of race and income together to explore the intersection of these two types of vulnerability.
-Specifically, we model how the relationship between temperature and mood is moderated by the interaction between a neighborhood's per-capita income and the percentage of a neighborhood's inhabitants that are black.
-Additionally, we compare AICs for models with no term for vulnerability, a race term for vulnerability, an income term for vulnerability, and both race and income terms for vulnerability.
-
-\begin{minipage}{.65\textwidth}
-\centering
-\includegraphics[width=\textwidth]{../res/wbgt-income-race.png}
-\captionof{figure}{Relationship between WBGT and mood from text in tweets, moderated by the interaction of race and income.}
-\label{fig:race-income}
-\end{minipage}\hfill
-\begin{minipage}{.35\textwidth}
-\centering
-\begin{tabular}{| l | l |}
-\hline
-Terms & AIC \\
-\hline
-Neither & 277,270,561 \\
-Income & 277,182,239 \\
-Race & 277,143,579 \\
-Both & 277,108,676 \\
-\hline
-\end{tabular}
-\captionof{table}{Akaike information criterion (AIC) of models with interaction terms for race, income, both race and income, as well as neither race nor income.
-The model with the best fit by AIC includes both race and income interaction terms.}
-\label{tab:aics}
-\end{minipage}
-
-Graphing the combined effects of race and income, we find that race has a larger impact on vulnerability than income (see Fig.
-\ref{fig:race-income}).
-A Black neighborhood at any income level will be much more impacted by heat than a poor neighborhood with few black inhabitants.
-Additionally, neighborhoods that are largely black are also more affected by cold temperatures, although there is no clear effect of income moderating the relationship between cold temperatures and mood.
-Using AIC as an indicator of how different terms affect model quality (see Tab.
-\ref{tab:aics}), we find that a model including both terms performs the best, indicating that both race and income matter for the impact of heat on mood.
-However, including only a race term improves the model more than including only an income term, further supporting the conclusion that race is more important than income for explaining the vulnerability of mood to heat.
-
-\section*{Comparison With Other Events}
-We compared the expressed mood from exposure to temperatures to two other events associated with impacts on mood and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact of a major natural disaster (see Fig.
-\ref{fig:compare}).
-For this comparison, we define heat impact as the changes in mood associated with temperatures increasing from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F); we define weekly changes as the difference between the mean Saturday mood and the mean Monday mood, calculated across the entire data set; and we define the impact of Hurricane Sandy as the change in mood from the week before the hurricane made landfall to the week after, calculated for only impacted counties.
-Both of these events are associated with mental health impacts: the decrease in expressed mood on Twitter from Saturdays to Mondays mirrors mirrors the weekly dynamics of suicide rates \cite{CDC2021}, while Hurricane Sandy was associated with mental health effects on victims including increased anxiety, PTSD, and depression \cite{Schwartz2017Aug, Lieberman-Cribbin2017}.
-
-\begin{figure}[H]
- \centering
- \includegraphics[width=0.66\linewidth]{../res/comparison_plot.png}
- \caption{Decline in mood as temperature changes from mild (5\textdegree C WBGT) high (25\textdegree C WBGT) across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.}
- \label{fig:compare}
-\end{figure}
-
-We find that the decline in mood in wealthier and majority white neighborhoods from exposure to increasing temperatures is less than the average weekly change in mood from Saturday to Monday (see Fig.
-\ref{fig:compare}).
-Conversely, the effects of exposure to increasing temperatures in poorer and majority Black neighborhoods are much larger than the average weekly change in mood.
-Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on mood from experiencing a major hurricane.
-
-\section*{Effect of Heat by Time of Day}
-In addition to providing data with high spatial resolution, Twitter data also comes with very high temporal resolution as each tweet is time-stamped.
-Thus, we were able to examine the impact of heat on mood over the course of the day (Fig.
-\ref{fig:ts-wbgt}).
-We found that heat is associated with improved mood for a brief period in the late morning, and then has a negative effect on mood during the latter half of the day, with a consistent effect from noon until 9pm.
-This effect weakens in the early evening through midnight, and then increases substantially throughout the night.
-Heat has the greatest effect on expressed mood at 6am, an effect over twice as large as during the day.
-
-\begin{figure}[H]
- \centering
- \includegraphics[width=0.5\linewidth]{../res/ts_heat.png}
- \caption{Effects of rising temperatures on mood by hour of the day, with a 95\% confidence interval.
-The value shown is the predicted change in VADER score for a 10\textdegree C WBGT increase in temperatures.}
- \label{fig:ts-wbgt}
-\end{figure}
-
-\section*{Discussion}
-% Compare weekly changes in mood to weekly changes in suicides
-We found that the change from an optimum temperature to a higher temperature is associated with an overall decrease in mood similar in magnitude to the degree of change in mood from Monday to Saturday, a weekly change associated with an increase in suicides of 23\% over the course of the study period \cite{CDC2021}, and that the same level of heat in a poor or Black neighborhood is associated with a change in mood nearly matching the impact of a major Hurricane, an event associated with suicides and PTSD \cite{Schwartz2017Aug, Lieberman-Cribbin2017}.
-Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of expressed mood and mental health at specific time scales and following certain events.
-This suggests that the changes in expressed mood we observe at neighborhood scales is indicative of other mental health outcomes such as suicides and hospitalizations.
-
-% Discuss the NOVELTY: quote Burke and Mullins, refute their interpretations, people have said XX, but we have shown ~XX.
-% Variations in vulnerability also indicates ADAPTATION
-Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income.
-Such analyses have led to conclusions that the mental health effects of climate change will be uniform, least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
-For example, Mullins et al.
-conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \cite{Mullins2019Dec}.
-Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income.
-This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the vulnerable and marginalized.
-More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health may be addressable.
-
-Examining the role of race in vulnerability, we find that majority Black non-Hispanic neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
-Surprisingly, we did not find an effect for people with Hispanic ethnicity, given that Hispanic peoples can be marginalized in employment and housing.
-These findings may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans.
-It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
-Additionally, there are many other marginalized groups in the United States, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
-Examining the effects of both race and income, we find that race is more important for vulnerability than income.
-This probably due in part to the fact that Black Americans have historically been forced through redlining into neighborhoods with fewer trees that are much more impacted by heat \cite{Hoffman2020, Locke2021}.
-This suggests that racism and patterns of housing discrimination contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups, and that reducing existing discrimination in housing and employment is an important adaptation strategy.
-
-% Connect to bigger literature on vulnerability
-This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts.
-Additionally, it shows how measuring outcomes at fine spatial scales necessary for understanding how factors like race and income matter for vulnerability \cite{Tong2021}.
-Even within a relatively wealthy nation, we find that high temperatures affect certain populations to the same degree as a major hurricane.
-Thus, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \cite{Raymond2020May}, heatwaves are already under-counted \cite{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \cite{Biardeau2020Jan}.
-
-In addition to exploring heterogeneities in vulnerability, using twitter data lets us explore how the effect of heat varies by time of date and infer possible important pathways in the heat-mental health relationship.
-Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \cite{Obradovich2017May, Mullins2019Dec}, and our temporal analysis largely supports this hypothesis.
-We found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
-This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
-
-% Caveats
-% Mood is a fuzzy indicator but we have BIG DATA
-While these findings are robust to different model specifications and metrics of mood, there are important considerations in this analysis.
-One caveat associated with this data is that we were only able to locate the tweets within the census block from which the tweet was sent, and in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are found that census block.
-For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
-These issue may in fact lead us to under-estimate the true effect of heat on mental health for poor and Black people.
-Another issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \cite{Pew2020Sep}.
-Nevertheless, we found a large volume of tweets across all neighborhood types.
-
-% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
-While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how unevenly these impacts will be distributed.
-People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \cite{bullard2012wrong}.
-While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they will not also be highly unevenly distributed.
-Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect.
-By using fine-scale Twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States.
-These findings have significant implications for urban planning, climate and environmental justice, and mental health.
 
@@
 We used data from 243 million geo-located English-language tweets from the continental United States from the years 2009 to 2019.
+The data were collected by the University of Vermont through a special agreement with Twitter, and represent a random sample from the population of geolocated tweets.
 The Twitter data consists of publicly posted messages, or tweets, that are short status messages users post to the platform.
@@
 Expressed mood is a widely used metric to assess mental health and well-being from posts on social media platforms.
-We use the VADER metric to assess the mood of Twitter posts as it was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
-The VADER (Valence Aware Dictionary for Sentiment Reasoning) corpus \cite{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to moods expressed on social media and includes several features, as:
+We use the VADER metric to assess the mood of Twitter posts as it was specifically designed for microblogs like Twitter \citep{hutto2014vader}.
+The VADER (Valence Aware Dictionary for Sentiment Reasoning) corpus \citep{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to moods expressed on social media and includes several features, as:
 
@@
 We used data on local weather conditions from the North American Land Data Assimilation System (NLDAS), a gridded product developed by several collaborative institutions, including NOAA, NASA, Princeton University, and the University of Washington.
-This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution, and integrates a large quantity of observation-based and modeled data \cite{xia_continental-scale_2012}.
+This dataset is available at an hourly temporal resolution, and at 1/8th decimal degree spatial resolution, and integrates a large quantity of observation-based and modeled data \citep{xia_continental-scale_2012}.
 We extracted temperature, specific humidity, air pressure, total precipitation, shortwave radiation, and wind speed for the exact hour and location of each tweet.
@@
 The WBGT can be estimated given normal temperature, relative humidity, solar radiation, and wind speed.
-Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing \cite{budd2008wet}.
-Metrics like WBGT that account for the effects of humidity and other factors on heat stress have been associated with diminished economic output \cite{rao2020projections}, increased crime \cite{hu2017impact}, increased mortality \cite{chien2016spatiotemporal, armstrong2019role}, and worsened mental health outcomes \cite{vida2012relationship, ding2016importance}.
+Because evaporative cooling is how humans cool themselves through perspiration, this temperature better indicates the heat stress that people are experiencing \citep{budd2008wet}.
+Metrics like WBGT that account for the effects of humidity and other factors on heat stress have been associated with diminished economic output \citep{rao2020projections}, increased crime \citep{hu2017impact}, increased mortality \citep{chien2016spatiotemporal, armstrong2019role}, and worsened mental health outcomes \citep{vida2012relationship, ding2016importance}.
 
 We derived relative humidity using methods described by Bolton et al.
-\cite{bolton_computation_1980} which use temperature, specific humidity, and pressure.
+\citep{bolton_computation_1980} which use temperature, specific humidity, and pressure.
 We then calculated the WBGT using the formula described by Heo et al.
-\cite{heo2019comparison}.
+\citep{heo2019comparison}.
 Throughout the paper we give the median dry bulb temperature observed at each given wet bulb temp degree across our data set.
@@
 Because the most recent available ACS was from 2014-2018, all tweets from year years greater than 2016 were matched to this dataset.
-We used data from the Integrated Public Use Microdata Series (IPUMS USA) National Historical Geographic Information System (NHGIS) service provided by the University of Minnesota \cite{ruggles2018ipums}.
+We used data from the Integrated Public Use Microdata Series (IPUMS USA) National Historical Geographic Information System (NHGIS) service provided by the University of Minnesota \citep{ruggles2018ipums}.
 
@@
 
-In addition to our models exploring how income and race moderate the effect of heat on mood individually, we fit an additional model with an interaction terms for both race and income.
-Due to the complexity of a model with segmented terms interacting with two different variables, using a categorical variable for race would have yielded a model matrix too large for the largest cloud virtual machine available to us.
-Thus, we used a continuous metric for both income as well as race (percent black).
-This model extends our main model to the following form, interacting income and race with a temperature segmented spline as well as precipitation in sunshine:
-
-\begin{equation}
- y = \beta_0 + f_t(t) + f_{m_{i}t}(m_{i} t) + f_{m_{r}t}(m_{r} t) +  \beta_p p + \beta_{m_{r} p} m_{r} p + \beta_{m_{i} p} m_{i} p + \beta_s s + \beta_{m_{i} s} m_{i} s +  \beta_{m_{r} s} m_{r} s + 
- \Phi + \epsilon
- \label{mod:race-inc}
-\end{equation}
-
-In this new model, $m_{i}$ is average income in the census block where the tweet originated, while $m_{r}$ is the racial composition of the census block.
-
 Finally, we fit a model for how heat affects mood by time of day.
@@
 
+\section*{Results}
+
+\subsection*{Effect of Heat on Sentiment}
+We examine the relationship between the mood expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \citep{budd2008wet}.
+We find higher temperatures are associated with worsened mood (Fig.
+\ref{fig:wbgt}) after controlling for a variety of fixed effects across all tweets.
+Mood is highest at 5\textdegree C WBGT (associated in our data set with a median dry bulb temperature on 12\textdegree C/54\textdegree F) with the largest declines between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
+Mood also declines with colder temperatures (below 5\textdegree C WBGT), although only slightly.
+Dis-aggregating our analysis by metropolitan area and using alternative methods to estimate expressed mood all yield similar results (see Supplement).
+
+\begin{figure}[H]
+ \centering
+ \includegraphics[width=\linewidth]{Xfigure1.pdf}
+ \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and expressed mood in tweets.
+As temperatures increase above 5\textdegree C WBGT, mood rapidly declines.
+Shaded area shows the 95\% confidence interval estimated with bootstrapping.}
+ \label{fig:wbgt}
+\end{figure}
+
+We find that mean neighborhood income strongly moderates the relationship between temperature and mood (Fig.
+\ref{fig:hetero}), with large differences in mood between the poorest (5th percentile) and wealthiest (95th percentile) neighborhoods.
+As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), mood increases in the wealthiest neighborhoods but decreases in the median and lowest income neighborhoods.
+The wealthiest neighborhoods do not see decreases in mood until temperatures exceed 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point mood decreases relatively evenly regardless of neighborhood income percentile.
+
+We also explored how neighborhood racial characteristics affect the relationship between temperature and mood, and we find the effects of heat are felt disproportionately in neighborhoods that are majority Black (Fig.
+\ref{fig:hetero}).
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the mood of tweets in majority Black neighborhoods decrease four times as much as the mood of people in other neighborhoods.
+Additionally, at mild to warm temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower mood than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
+
+\begin{figure}[H]
+  %\makebox[\textwidth][c]{\includegraphics[width=1.4\textwidth]{Xfigure2.pdf}}
+  \includegraphics[width=\textwidth]{Xfigure2.pdf}
+  \caption{Effect of changes in WBGT on expressed mood as moderated by neighborhood income (a) and race (b). Shaded area shows the 95\% confidence interval estimated with bootstrapping.}
+\label{fig:hetero}
+\end{figure}
+
+\subsection*{Comparison With Other Events}
+We compared the expressed mood from exposure to temperatures to two other events associated with impacts on mood and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact of a major natural disaster (see Fig.
+\ref{fig:compare}).
+For this comparison, we define heat impact as the changes in mood associated with temperatures increasing from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F); we define weekly changes as the difference between the mean Saturday mood and the mean Monday mood, calculated across the entire data set; and we define the impact of Hurricane Sandy as the change in mood from the week before the hurricane made landfall to the week after, calculated for only impacted counties.
+Both of these events are associated with mental health impacts: the decrease in expressed mood on Twitter from Saturdays to Mondays mirrors mirrors the weekly dynamics of suicide rates \citep{CDC2021}, while Hurricane Sandy was associated with mental health effects on victims including increased anxiety, PTSD, and depression \citep{Schwartz2017Aug, Lieberman-Cribbin2017}.
+
+\begin{figure}[H]
+ \centering
+ \includegraphics[width=\linewidth]{Xfigure3.pdf}
+ \caption{Decline in mood as temperature changes from mild (5\textdegree C WBGT) high (25\textdegree C WBGT) across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.
+Error bars show the 95\% confidence interval estimated with bootstrapping.}
+ \label{fig:compare}
+\end{figure}
+
+We find that the decline in mood in wealthier and majority white neighborhoods from exposure to increasing temperatures is less than the average weekly change in mood from Saturday to Monday (see Fig.
+\ref{fig:compare}).
+Conversely, the effects of exposure to increasing temperatures in poorer and majority Black neighborhoods are much larger than the average weekly change in mood.
+Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on mood from experiencing a major hurricane.
+
+\subsection*{Effect of Heat by Time of Day}
+In addition to providing data with high spatial resolution, Twitter data also comes with very high temporal resolution as each tweet is time-stamped.
+Thus, we were able to examine the impact of heat on mood over the course of the day (Fig.
+\ref{fig:ts-wbgt}).
+We found that heat is associated with improved mood for a brief period in the late morning, and then has a negative effect on mood during the latter half of the day, with a consistent effect from noon until 9pm.
+This effect weakens in the early evening through midnight, and then increases substantially throughout the night.
+Heat has the greatest effect on expressed mood at 6am, an effect over twice as large as during the day.
+
+\begin{figure}[H]
+ \centering
+ \includegraphics[width=\linewidth]{Xfigure4.pdf}
+ \caption{Effects of rising temperatures on mood by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted change in VADER score for a 10\textdegree C WBGT increase in temperatures.
+Shaded area shows the 95\% confidence interval estimated with bootstrapping.}
+ \label{fig:ts-wbgt}
+\end{figure}
+
+\section*{Discussion}
+% Compare weekly changes in mood to weekly changes in suicides
+We found that the change from an optimum temperature to a higher temperature is associated with an overall decrease in mood similar in magnitude to the degree of change in mood from Monday to Saturday, a weekly change associated with an increase in suicides of 23\% over the course of the study period \citep{CDC2021}, and that the same level of heat in a poor or Black neighborhood is associated with a change in mood nearly matching the impact of a major Hurricane, an event associated with suicides and PTSD \citep{Schwartz2017Aug, Lieberman-Cribbin2017}.
+Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of expressed mood and mental health at specific time scales and following certain events.
+This suggests that the changes in expressed mood we observe at neighborhood scales is indicative of other mental health outcomes such as suicides and hospitalizations.
+
+% Discuss the NOVELTY: quote Burke and Mullins, refute their interpretations, people have said XX, but we have shown ~XX.
+% Variations in vulnerability also indicates ADAPTATION
+Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income.
+Such analyses have led to conclusions that the mental health effects of climate change will be uniform, least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
+For example, Mullins et al.
+conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \citep{Mullins2019Dec}.
+Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income.
+This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the vulnerable and marginalized.
+More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health may be addressable.
+
+Examining the role of race in vulnerability, we find that majority Black non-Hispanic neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
+Surprisingly, we did not find an effect for people with Hispanic ethnicity, given that Hispanic peoples can be marginalized in employment and housing.
+These findings may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans.
+It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
+Additionally, there are many other marginalized groups in the United States, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
+Examining the effects of both race and income, we find that race is more important for vulnerability than income.
+This probably due in part to the fact that Black Americans have historically been forced through redlining into neighborhoods with fewer trees that are much more impacted by heat \citep{Hoffman2020, Locke2021}.
+This suggests that racism and patterns of housing discrimination contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups, and that reducing existing discrimination in housing and employment is an important adaptation strategy.
+
+% Connect to bigger literature on vulnerability
+This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts.
+Additionally, it shows how measuring outcomes at fine spatial scales necessary for understanding how factors like race and income matter for vulnerability \citep{Tong2021}.
+Even within a relatively wealthy nation, we find that high temperatures affect certain populations to the same degree as a major hurricane.
+Thus, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \citep{Raymond2020May}, heatwaves are already under-counted \citep{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \citep{Biardeau2020Jan}.
+
+In addition to exploring heterogeneities in vulnerability, using twitter data lets us explore how the effect of heat varies by time of date and infer possible important pathways in the heat-mental health relationship.
+Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \citep{Obradovich2017May, Mullins2019Dec}, and our temporal analysis largely supports this hypothesis.
+We found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
+This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
+
+% Caveats
+% Mood is a fuzzy indicator but we have BIG DATA
+While these findings are robust to different model specifications and metrics of mood, there are important considerations in this analysis.
+One caveat associated with this data is that we were only able to locate the tweets within the census block from which the tweet was sent, and in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are found that census block.
+For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
+These issue may in fact lead us to under-estimate the true effect of heat on mental health for poor and Black people.
+Another issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \citep{Pew2020Sep}.
+Nevertheless, we found a large volume of tweets across all neighborhood types.
+
+% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
+While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how unevenly these impacts will be distributed.
+People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \citep{bullard2012wrong}.
+While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they will not also be highly unevenly distributed.
+Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect.
+By using fine-scale Twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States.
+These findings have significant implications for urban planning, climate and environmental justice, and mental health.
+
 \section*{Acknowledgements}
@@
 The UVM special agreement with Twitter allows for access to this data for research and analysis purposes and this work complies with all the terms of service for Twitter and UVM.
-
-%\printbibliography
-\bibliography{TextDataClimateShocks}
 
@@
 
+%\printbibliography
+\input{temp-sentiment.bbl}
 \end{document}

### 2022-02-16 doc/author-response-to-reviewers.tex

Associate editor comments:

Both reviewers found this manuscript important and interesting. Both offer some critiques on modeling choices that must be addressed in a revision. Please also pay close attention to reviewers comments about areas where statements may not fully reflect current literature and where the manuscript overstates its novelty.

Reviewer comments:

Reviewer 1: First, I would like to commend the authors for addressing a vital topic in their research. Research into how the sentiment-related effects of the weather differ by demographic status is very important as it relates to potential adaptation and resilience strategies. The authors have put together a very nice manuscript, and it was a pleasure to read.

That said, I think this is a solid manuscript that isn't quite ready for publication and needs some minor and some major tweaks. I have a number of comments, questions, and concerns that should be addressed prior to publication. I think once these concerns are addressed this manuscript will be worthy of publication.

Comments:

Comments follow in order of appearance in the manuscript.

1. The abstract is a bit incorrect. The authors cite a number of studies that, themselves, have sections where those studies examine heterogeneous impacts as a function of demographic vulnerability - even individually reported income. There's no need to over-sell their novelty on this point. The authors should mention/detail the heterogeneous impacts already well-documented in the climate/mental health literature, the prior effect size differences, alongside whether those differences were significant or not. Simply stating 'heterogeneous tests haven't been done or haven't seen anything' is incorrect.
% #Wording: Over-selling "first evidence". Change the last sentence of abstract.

2. "Expressed mood" or "mood" represents a sizable conceptual leap from 'sentiment' to affective states. If the authors want to argue that their construct measures affective states, they will need to justify this measurement leap a bit more strongly. To my knowledge the link between 'sentiment' and 'mood' is still only weakly validated, at best.
%#Major/#Minor: Leap or link between mood and sentiment is weak. Need additional literature review.

3. That "previous analysis [sic] using city-level data have been unable to find heterogeneities" isn't correct. The authors need to tone down the strength of this claim and better incorporate prior heterogeneous findings in the social impacts literature into their introduction and discussion. E.g. from Obradovich et al., 2018:

"To examine if poorer respondents are more sensitive to temperature, we stratify our sample along income quartiles and estimate regressions for the lowest and highest quartiles (SI Appendix, Table S8). Fig. 2A shows that the negative effect of temperatures greater than 30 °C on the probability of mental health issues is larger for low-income respondents (coefficient: 2.309, P = 0.002, n = 438,518). This is 1.6 times the effect observed among the highest-income adults in the sample (coefficient: 1.458, P = 0.005, n = 509,608).

Fig. 2B shows that the negative effect of temperatures greater than 30 °C on the probability of mental health difficulties is largest for women (coefficient: 1.41, P = 0.005, n = 1,211,220). This is 1.6 times the effect observed among men in our sample (coefficient: 0.879, P = 0.031, n = 750,523). Combining these insights, the effect observed in the subsample of low-income women (coefficient: 2.742, P = 0.002, n = 300,570) is approximately two times the magnitude of the effect observed in the high-income men in the sample (coefficient: 1.373, p = 0.03, n = 235,098)."
%#Wording: Change the tone for previous analysis 
4. "We present the first evidence…" can also be toned down.
%#Wording: Change the saying for overall novelty

5. There are a number of typos throughout the manuscript. Another proof read could be useful.
%#Minor: Check the typos

6. We've found the same sleep effects within actigraph-based data across the diurnal cycle, as well: https://arxiv.org/abs/2011.07161 (we also find quite substantial country-level heterogeneity across income in the effects of temperature on sleep).
%#Minor: Need additional literature review
7. "research conducted at national scales indicates surprisingly little heterogeneity in impacts" simply isn't consistent with my understanding of the literature. Some literature doesn't find significant differences, others find sizable heterogeneity.
%#Minor / #Wording: Need to review and summarise literature again
8. You cite, in part, Mullins and White's use of individual BRFSS data to justify the statement that a major challenge is that the data are aggregated. The survey measures used by Mullins and White and Obradovich et al. contain more precisely measured demographic covariates than the ecological inference approach you employ here…
% #Dispute: The aggregation level is the difference.

10. Zheng et al. 2020 examine by-city heterogeneous effects of income as well as gender on the temperature-sentiment relationship. Given the extensive literature on gender and heterogeneous mental health, the authors might do well to discuss this dimension of vulnerability, as well (though I understand their Twitter data does not easily enable investigation in this setting; Weibo data enabled more precise measurement). https://www.sciencedirect.com/science/article/pii/S2590332220302529#fig3

Further, Mullins and White found sizable differences by gender, see their Table A12.
% #Minor: Need additional literature review
11. The whole introduction is framed around a set of findings that is not nicely consistent with the literature as I understand it. There's no need to motivate this paper as the first paper to investigate heterogeneous effects in climate/mental health. The effects documented are interesting and important in their own right.
% #Wording: Change the tone of "first paper"

12. That many studies use sentiment as a measure of mood does not mean the measurement leap is well-validated. I wish it were better validated, but am not convinced that it yet is and have been convinced otherwise by some expert psychometricians.
% #Major: Change the usage of "mood" to "sentiment" See 2.
13. Jumbling up sentiment into the various Twitter-mental health papers is coarse. There are quite a large number of differences between those studies that try to derive features predictive of mental health outcomes and the use of sentiment as a direct proxy for those outcomes. This should be made more clear.
% #Major: Reorg the part of literature of Twitter-mental health
14. How did the authors define 'weather-related words'? It would be good to state this in the manuscript.
% #Minor: Add weather-related words into Appendix?

15. "Expressed mood is a widely used metric to assess mental health and well-being from posts on social media platforms." You are speaking of sentiment here, not 'expressed mood'. If you want to argue that one is a valid measure for the other, you will need to make a more compelling argument on that front. That others make the leap, by itself, isn't particularly compelling to me.
% #Major: Change the usage of "mood" to "sentiment". See 2 and 12.
16. One of the biggest conceptual/measurement leaps that the authors ask the reader to make in this manuscript is that neighborhood characteristics map to the individuals who post geolocated tweets. This is a very large conceptual leap. We know that a) Twitter users are not representative of the general public and b) those who geolocate their tweets are not representative of Twitter users. To map census characteristics onto these users - in the primary rationale for this entire paper - requires convincing the reader that such census characteristics do an adequate job serving as a proxy for individual geo-tweet users' characteristics. The authors should conduct more extensive testing and validation to justify this measurement leap. Noting it merely as a limitation in the discussion isn't sufficient.
% #Major: Do test to show the representative of geo-located Twitter is good enough to be mapped by census characteristics. May need literature review on twitter related papers

17. The cold side of the functional form presented by the authors in Fig. 1 does not map well to any prior result on this topic in the literature that I am aware of, which is concerning. Every other published paper that I'm aware of on this topic, using much the same data, finds a decline in sentiment on the cold side of the distribution (Baylis et al 2018, Baylis 2020, Wang et al 2020, etc.). (The global functional form also had a cold-side effect in this study: https://www.thelancet.com/journals/lancet/article/PIIS0140-6736(21)01787-6/fulltext)
% #Minor: Comparing the cold side difference from other paper.

18. Looking into the methods that might produce this different functional form, I see that the authors do not include space-by-time fixed effects that are standard in the climate econometrics literature. Why is this? The authors' primary functional form - and model - should match that of the literature in which they are attempting to investigate heteregeneous effects. Space-by-time fixed effects are an important way to rule out confounding variation that varies across spatial units (counties?) over time. I strongly suggest the authors modify their empirical approach accordingly.

Understanding the heterogeneous effects that might exist on the well-documented cold side of the distribution is just as important as understanding how these effects operate on the hot side of the distribution, if the goal is to understand the net-impact of warming across heterogeneous demographics.
% #Major: Rerun regression model with space-time fixed effect
19. Why do the authors not extend model (3) to the cold side of the temperature distribution? It is just as important to understand time-of-day effects on that side of the distribution, too (even within the relatively flat effects they observe on average, likely due to omission of fixed effects important in this type of modeling, there could easily be within-day effects that average out, right?). I'm fine if they model that side as a linear function of temperature, too, but not that they exclude it entirely.
% #Major: Rerun regression model for cold side
20. Income is a continuous variable and the goal of this study is to document heterogeneous effects. The authors choose two income categories to compare to one another, excluding 90\% of the income distribution. What happens across the rest of the income distribution? Are the heterogeneous effects mostly linear as they change across the distribution? Are their nonlinearities? This is critical to investigate, and the authors have the data to do so.
% #Major: Income group breakdown
21. Similarly, the authors' racial categories are continuous in measurement, but they for some reason choose to dichotomize them. This isn't really appropriate if the primary goal of this manuscript is to understand the nature of heterogeneous effects. The authors should present effects across the entire distribution of vulnerability, or at least show us how their results persist across more flexible and continuous definitions of their vulnerability measures.
% Zheng: I didn't get why racial categories are continuous
22. The by-time-of day analysis draws strange inference. We know a) that the effect of temperature on mood is curved and nonlinear and b) that it appears to be accelerating in hotter temperatures. That warming in early morning produces positive effects and in the afternoon produces negative effects could simply be an artifact of the absolute-location on the temperature distribution held by those time periods, on average. I would, instead, like to know: For any given time of day, how does the effect of temperature - across the distribution of all temperatures that occurred in that hour over the sample - compare to any other time of day. The authors could choose times of day with decent support in the data to compare, running a regression for each hour of the day and plotting them on the same axes.
% #Major: Change the display of Time of Day breakdown
23. "Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income." is of questionable accuracy.
% #Minor: Need additional literature review.
Overall I think the import of this study is high (no need to oversell it!) and the execution is solid. With the addressing of the above concerns I think it will be a useful addition to the literature.

Reviewer 2: The manuscript tackles an important climate change health impact of mental well-being and according to vulnerability of race and income using an innovative way of extrapolating Twitter data regarding expressions of mood with temperatures and socioeconomic data according to census blocks.
The manuscript is well written and is addressing neglected areas of climate change health impacts. There are some important findings regarding how increased temperatures are associated with lower mood as expressed by a large number of people through 242 million tweets across the United States, and especially among Black Americans. The authors carefully describe the methods and modelling and include covariates of time of day, day of the week, and the day in the year, rain, sunshine, as well as income and ethnicity.
There are the following issues that weaken the manuscript and its potential interpretations:
1. Without inclusion of air conditioning status/prevalence within each of the census blocks, or if not available, at least within the areas that such data is available as secondary analyses, the findings of the association between mood and heat can be easily explained by the lack of air conditioning rather than the other factors of race, income, time of the data, day of the week, etc.
% #Major  #Improvement #If-Applicable: introduce air conditioning with the modeling
2. There is a discussion to make the case that mood as expressed in Twitter equals mental health and mental illness (like lines 345-348 in the discussion), but this is a bit of stretch. For example, Reece et al uses an algorithm model to predict PTSD but cautions that this is just a proof of concept and not to be used before further research and refinement. The VADER sentiment analyses authors do not claim their scoring has anything to do with mental illnesses.
% #Minor #Wording: Relation between mood and mental health. See Reviewer1:2,12,15
3. The use of tweets from 2014 to be correlated to socioeconomic and racial distribution of the census blocks from the American Community Service data during 2012-2016 but then tweets after 2016 correlated with American Community Service data during 2014-2018 was not consistent and there is a possibility of changes in these populations or how they tweet their mood given twitter is a relatively recent and growing phenomena that became more common after 2014.
% #Major  #Improvement: Add discussion between using two ACS datasets
4. There is an assumption that impacts of heat on mood is immediate, while in general the longer the heat wave the more likely it will impact mental health, so a lag of several days or limiting the analyses to heat waves of several days would have been more appropriate.
% #Major #Improvement: Consider the time lag of heat wave effect
5. Similarly, heat during the night is more important than heat and high temperature during the day. This was evident from the fact that those who tweet at 6am are more likely to have a bad mood. Carrying out the analyses only for high temperatures during the night compared to any time of the day would shed more light on this matter.
% #Major #Improvement: time of day breakdown; See Reviewer1:22
6. Although in the supplement data there is a graph showing the different metropolitan cities across the nation all had negative mood tweets during higher temperature, it is a well-known fact that there is wide variety and range of tolerance to heat waves and higher temperature in places like Arizona, Texas, or New Mexico compared to the Northeast or Midwest. In our work we even found within the same cities those living in the coastal area exposed to the sea breeze are less tolerant to heat waves than those inlands who are more used to higher average temperatures. This matter can be addressed by carrying out the analyses with different cutoff/range of dry bulb for each region in the US.
% #Major #Improvement: cities trend groupby dry bulb
7. Similar to point 6, it is not clear why those living in high socioeconomic status area have better mood with higher temperature than those with low income status but then at higher temperature their mood declines. If those with high socioeconomic status have air conditioning then they would not be impacted by higher temperature. If both groups do not have air conditioning then one would expect that those with high income are less tolerant than the low socioeconomic status group.
% Zheng: I didn't see why wealthy people would be totally tolerant to high temperature.
8. I presume the tweets do not differentiate individuals who carry out multiple tweets, but this was not addressed. If someone is a very active tweeter and engaged with many on twitter, they will likely bias the results. If this data is available (tweets per individual) maybe do a weighting for such tweets from them same source, and if the data is not available then this should be addressed in the discussion.
% #Major #If-Applicable: Unique users problem
9. Given the nature of this data from twitter and the many assumptions in attributing how personal expressions of a mood can be related to heat at an ecological census block level, I would not agree with the concluding remarks and the certainty that heat waves will impact the mental health of low-income populations. It is a bit of a leap. This is especially true given past studies have not shown this differentiation, and the above mentioned issues with the data lacking air conditioning rates, and the counterintuitive higher tolerance to heat among the higher income populations, all weaken interpretations of the results from this study.

% #Major: Limitation of using Twitter weaken the conclusion

Requests from EHP

Note to authors: EHP conducts content editing during peer review as a service to authors and an internal quality assurance measure. Please verify that the revised version of your manuscript addresses the following comments before submitting your revision (also attached). In addition, please see attached reporting guidelines for EHP research articles. Doing so may avoid the need for additional revisions that could delay final decisions and publication.

- Line 40, Add qualifier, “first, to our knowledge” or similar.
- Line 153: Is there any knowledge of the % of this sample of tweets from the total number of geolocated tweets in this area/time frame? If so, please specify.
- Was there any linkage to individual users, or were the tweets sample and analysis limited to an analysis unit of a tweet (for example, was this a random sample of tweets or was there sampling/clustering by user)? Please clarify.
- Does this include only public tweets/accounts, or all tweets/accounts? Please specify.
- Line 175: Is there an upper or lower bound to the VADER score? If so, please include the bounds here.
- Line 176-178. Provide a brief description of the Hedonmeter and AFINN methods. Please also include some discussion of why the VADER method was used for the primary analysis over the other available methods (is one shown to be more valid than the other, is one more widely used, etc?).
- Please confirm if the analysis accounted for region, as response to temperature can vary widely due to acclimatization. For example 90 degrees in Seattle can elicit a different response than 90 degrees in Los Angeles. This concern was also brought up by reviewer 2.
- Clarify in method how the census block majority was calculated.
- Justify why you collapse race groups (line 215-216, line 227-230). Add to discussion, the limitations of this classification and it’s use as a proxy for unmeasured consequences of racism. Please see EHP guidelines on the use and reporting of race/ethnicity in research articles: https://ehp.niehs.nih.gov/authors/research-articles.
- Add to discussion, how air conditioner availability by region may impact results. Consider a secondary analysis including % population with air conditioning (there is public data on this available). This issue was also brought up by reviewer 2.
- Add to discussion, are there other secular trends over time that might impact results?
- Line 251: Typo, “were” should be “where”. Please also correct other typos throughout.
- Please describe all secondary analyses in the methods section, and refer to results of these analyses in the results section. Line 299 and 324.
- Figure 1: Unclear what “tweet count” on the y-axis refers to. Are there 2 sets of data presented on the same graph? This is not clear and there is no unit note about the tweet count. Please clarify on the figure and in the figure caption.
- Please refer to specific material in the supplement rather than “see supplement”. Such as (Figure S1.), or (Supplement, Title of section in the supplement)
- Figure 1, 2: Include in figure caption description of y and x-axes, make sure it is clear two sets of information are presented on the same graph.
- Figure 2: Change so both panel a and b use the same y axis
- Include definition of “poorest” and “wealthiest” neighborhoods in methods section.
- In fig 3 you use poor/rich, in the text you use wealthiest/poorest – please use consistent wording.
- Add to methods percentile cut points for income for analysis.
- Where there any census blocks with missing ACS data? If so, how was that handled in the analysis?
- Please include in the results, the coverage statistics of where tweets were located (% census blocks or census tracts in the US). Consider adding a jittered map of locations. Were certain areas or regions in the US more represented than others?
- Please present the number/proportion of total tweets by the sub-populations you examine.
- Add to methods mention of the comparison to natural disasters. Methods for all results presented in results should be described in the methods section. How was Hurricane Sandy chosen for comparison? How were “impacted counties” defined? What was the size for the tweets included in the analysis for Hurricane Sandy.
- Add to methods a description of the Saturday to Monday comparison.
- Add to methods, comparison with CDC WONDER data
- Line 367-369 – what is this based on? Include a citation.
- Line 418 – include as standard citation, not as a hyperlink
- Please add to discussion what the change in VADER score might represent (how do you interpret a .5 change, a .05 change?).
- Please include tables with corresponding results for figures in the supplement for transparency.
- Figure S7 – Include \$ associated with presented tertiles.
- In supplemental material, the temp effects by CBSA needs to be mentioned in the main text. Same with interaction of race and income.
- Please ensure all material in the supplement is specifically called-out in the main text. For example,
call out weather terms supplement material in methods.
- EHP submission format: Please provide main text figure captions on a new page after any tables (which should be on a new page after the references).
- Revise all current findings, interpretations, and conclusions to be past tense rather than present
- Note that EHP is required by law to conform to 508 standards for accessibility, which provides guidelines to ensure that the content is accessible in some form to all readers, including those with color vision disabilities. In general, do not identify elements of a figure by color alone, as individuals with color vision disabilities will not be able to discern that color. Instead, use contrast, texture, or symbols. Color may be used to enhance figures, but to ensure that figures are accessible to readers with impaired color vision, color should not be used as the only means of conveying information unless absolutely necessary. Use different symbols, shading/textures, and line patterns instead of (or in addition to) color to distinguish among different data points, or use contrasting colors that can be easily distinguished from each other when the figure is printed in black and white. Alternatively, select a “colorblind palette” if available in the software package being used to generate the figure. If colors are used to identify data, include a color key in the graphical portion of the figure: do not refer to colors by name (i.e., red, blue) in the figure title or caption. For example: Figure 2.
- Revise figure captions: Tables and figures do not have to be 100\% stand-alone, but they should include enough information to be easily interpreted by a casual reader who may be “skimming” the manuscript to decide whether it is worth a more detailed assessment. This applies to tables and figures provided in the separate Supplemental Material file, as well as tables and figures in the main text. If you are in doubt about the amount of information to include, please err on the side of excess information.
o Provide brief information about the study population (including location, overall size, and time period), usually in the table or figure title.
o Tables or figures reporting model-based estimates should include brief information regarding the methods used to derive the estimates, including the type of model, covariates, missing data methods (e.g., imputation), and methods to account for values o Report numbers of observations used to derive each individual estimate reported in the table or figure, after accounting for missing data and subgrouping. As appropriate, indicate numbers of observations by subgroup. (These can be provided in the supplemental material)

### 2022-02-16 doc/temp-sentiment.tex
Updates from Overleaf

--- 2021-06-30 doc/temp-sentiment.tex
@@
 \usepackage{xcolor}
-% \usepackage[
-% style=numeric,
-% citestyle=numeric-comp,
-% backend=biber,
-% doi=true,
-% natbib=true,
-% sorting=none
-% ]{biblatex}
+\usepackage[
+  style=alphabetic,
+  citestyle=authoryear,
+  backend=biber,
+  bibstyle=authortitle,
+  doi=true,
+  natbib=true,
+  sorting=nyt
+]{biblatex}
 \usepackage{scalerel,xparse}
+\usepackage{lineno}
+
+\DeclareFieldFormat{labelalphawidth}{#1}
+
+\linenumbers
+
+\addbibresource{TextDataClimateShocks.bib}
 
@@
 We used data from 243 million geo-located English-language tweets from the continental United States from the years 2009 to 2019.
+The data were collected by the University of Vermont through a special agreement with Twitter, and represent a random sample from the population of geolocated tweets.
 The Twitter data consists of publicly posted messages, or tweets, that are short status messages users post to the platform.
 We only considered users’ original content, thus did not include retweets in our analysis.
-Additionally, we excluded all tweets that contained weather-related terms, to ensure that the mood expressed in the tweets was reflective of a user's mental state, and not a commentary on the weather.
+Additionally, we excluded all tweets that contained weather-related terms, to ensure that the mood expressed in the tweets was reflective of a user's mental state, and not a commentary on the weather, yielding a final dataset of 236 million tweets used in our analysis.
 
 Expressed mood is a widely used metric to assess mental health and well-being from posts on social media platforms.
-We use the VADER metric to assess the mood of Twitter posts as it was specifically designed for microblogs like Twitter \cite{hutto2014vader}.
-The VADER (Valence Aware Dictionary for Sentiment Reasoning) corpus \cite{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to moods expressed on social media and includes several features, as:
+We use the VADER metric to assess the mood of Twitter posts as it was specifically designed for microblogs like Twitter \citep{hutto2014vader}.
+The VADER (Valence Aware Dictionary for Sentiment Reasoning) corpus \citep{gilbert_vader_2014} is a lexical, rule-based sentiment analysis tool that is specifically attuned to moods expressed on social media and includes several features, as:
 
@@
 
+\section*{Results}
+
+\subsection*{Effect of Heat on Sentiment}
+We examine the relationship between the mood expressed in tweets and the prevailing Wet Bulb Globe Temperature (WBGT), a temperature metric that accounts for dry-bulb temperature, humidity, wind speed, and solar radiation to more accurately describe the effects of heat on the human body \citep{budd2008wet}.
+We find higher temperatures are associated with worsened mood (Fig.
+\ref{fig:wbgt}) after controlling for a variety of fixed effects across all tweets.
+Mood is highest at 5\textdegree C WBGT (associated in our data set with a median dry bulb temperature on 12\textdegree C/54\textdegree F) with the largest declines between 20\textdegree -25\textdegree C WBGT (dry bulb 29\textdegree-36\textdegree C/84\textdegree-97\textdegree F).
+Mood also declines with colder temperatures (below 5\textdegree C WBGT), although only slightly.
+Dis-aggregating our analysis by metropolitan area and using alternative methods to estimate expressed mood all yield similar results (see Supplement).
+
+\begin{figure}[H]
+ \centering
+% \includegraphics[width=\linewidth]{Xfigure1.pdf}
+ \caption{Relationship between Wet Bulb Globe Temperature (WBGT) and expressed mood in tweets.
+As temperatures increase above 5\textdegree C WBGT, mood rapidly declines.
+Shaded area shows the 95\% confidence interval estimated with bootstrapping.}
+ \label{fig:wbgt}
+\end{figure}
+
+We find that mean neighborhood income strongly moderates the relationship between temperature and mood (Fig.
+\ref{fig:hetero}), with large differences in mood between the poorest (5th percentile) and wealthiest (95th percentile) neighborhoods.
+As temperatures increase to a modest 20\textdegree C WBGT (29\textdegree C/84\textdegree F), mood increases in the wealthiest neighborhoods but decreases in the median and lowest income neighborhoods.
+The wealthiest neighborhoods do not see decreases in mood until temperatures exceed 20\textdegree C WBGT (29\textdegree C/84\textdegree F), at which point mood decreases relatively evenly regardless of neighborhood income percentile.
+
+We also explored how neighborhood racial characteristics affect the relationship between temperature and mood, and we find the effects of heat are felt disproportionately in neighborhoods that are majority Black (Fig.
+\ref{fig:hetero}).
+Relative to an optimum temperature of 5\textdegree C WBGT (12\textdegree C/54\textdegree F), as temperatures increase to 30\textdegree C WBGT (42\textdegree C/108\textdegree F), the mood of tweets in majority Black neighborhoods decrease four times as much as the mood of people in other neighborhoods.
+Additionally, at mild to warm temperatures of 10\textdegree C WBGT (19\textdegree C/66\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F), people in majority Hispanic neighborhoods have slightly lower mood than people in majority white or other neighborhoods, although this gap narrows at higher temperatures.
+
+\begin{figure}[H]
+  %\makebox[\textwidth][c]{\includegraphics[width=1.4\textwidth]{Xfigure2.pdf}}
+%  \includegraphics[width=\textwidth]{Xfigure2.pdf}
+  \caption{Effect of changes in WBGT on expressed mood as moderated by neighborhood income (a) and race (b). Shaded area shows the 95\% confidence interval estimated with bootstrapping.}
+\label{fig:hetero}
+\end{figure}
+
+\subsection*{Comparison With Other Events}
+We compared the expressed mood from exposure to temperatures to two other events associated with impacts on mood and other mental health outcomes: the weekly change from Saturday to Monday, as well as the impact of a major natural disaster (see Fig.
+\ref{fig:compare}).
+For this comparison, we define heat impact as the changes in mood associated with temperatures increasing from 5\textdegree C WBGT (12\textdegree C/54\textdegree F) to 25\textdegree C WBGT (36\textdegree C/97\textdegree F); we define weekly changes as the difference between the mean Saturday mood and the mean Monday mood, calculated across the entire data set; and we define the impact of Hurricane Sandy as the change in mood from the week before the hurricane made landfall to the week after, calculated for only impacted counties.
+Both of these events are associated with mental health impacts: the decrease in expressed mood on Twitter from Saturdays to Mondays mirrors mirrors the weekly dynamics of suicide rates \citep{CDC2021}, while Hurricane Sandy was associated with mental health effects on victims including increased anxiety, PTSD, and depression \citep{Schwartz2017Aug, Lieberman-Cribbin2017}.
+
+\begin{figure}[H]
+ \centering
+ %\includegraphics[width=\linewidth]{Xfigure3.pdf}
+ \caption{Decline in mood as temperature changes from mild (5\textdegree C WBGT) high (25\textdegree C WBGT) across different neighborhood types, compared with the impacts of the change from Saturday to Monday, as well as the change associated with Hurricane Sandy.
+Error bars show the 95\% confidence interval estimated with bootstrapping.}
+ \label{fig:compare}
+\end{figure}
+
+We find that the decline in mood in wealthier and majority white neighborhoods from exposure to increasing temperatures is less than the average weekly change in mood from Saturday to Monday (see Fig.
+\ref{fig:compare}).
+Conversely, the effects of exposure to increasing temperatures in poorer and majority Black neighborhoods are much larger than the average weekly change in mood.
+Additionally, the impacts of exposure to these changes in temperature for the more vulnerable neighborhoods are close in severity to the impact on mood from experiencing a major hurricane.
+
+\subsection*{Effect of Heat by Time of Day}
+In addition to providing data with high spatial resolution, Twitter data also comes with very high temporal resolution as each tweet is time-stamped.
+Thus, we were able to examine the impact of heat on mood over the course of the day (Fig.
+\ref{fig:ts-wbgt}).
+We found that heat is associated with improved mood for a brief period in the late morning, and then has a negative effect on mood during the latter half of the day, with a consistent effect from noon until 9pm.
+This effect weakens in the early evening through midnight, and then increases substantially throughout the night.
+Heat has the greatest effect on expressed mood at 6am, an effect over twice as large as during the day.
+
+\begin{figure}[H]
+ \centering
+% \includegraphics[width=\linewidth]{figure4.pdf}
+ \caption{Effects of rising temperatures on mood by hour of the day, with a 95\% confidence interval.
+The value shown is the predicted change in VADER score for a 10\textdegree C WBGT increase in temperatures.
+Shaded area shows the 95\% confidence interval estimated with bootstrapping.}
+ \label{fig:ts-wbgt}
+\end{figure}
+
+\section*{Discussion}
+% Compare weekly changes in mood to weekly changes in suicides
+We found that the change from an optimum temperature to a higher temperature is associated with an overall decrease in mood similar in magnitude to the degree of change in mood from Monday to Saturday, a weekly change associated with an increase in suicides of 23\% over the course of the study period \citep{CDC2021}, and that the same level of heat in a poor or Black neighborhood is associated with a change in mood nearly matching the impact of a major Hurricane, an event associated with suicides and PTSD \citep{Schwartz2017Aug, Lieberman-Cribbin2017}.
+Thus, while research on linkages between expressed mood and other mental health outcomes like suicides and hospitalizations is still nascent, there are clear similarities in patters of expressed mood and mental health at specific time scales and following certain events.
+This suggests that the changes in expressed mood we observe at neighborhood scales is indicative of other mental health outcomes such as suicides and hospitalizations.
+
+% Discuss the NOVELTY: quote Burke and Mullins, refute their interpretations, people have said XX, but we have shown ~XX.
+% Variations in vulnerability also indicates ADAPTATION
+Previous work conducted with county-scale data found no heterogeneity in mental health vulnerability by income.
+Such analyses have led to conclusions that the mental health effects of climate change will be uniform, least in developed countries, and that adaptation with technologies like air-conditioning will not mitigate the mental health effects of climate change.
+For example, Mullins et al.
+conclude that ``individuals have not been able to successfully reduce the negative effects of higher temperatures on mental health" \citep{Mullins2019Dec}.
+Contrary to these analyses, we found large heterogeneity in vulnerability by neighborhood racial composition and per-capita income.
+This suggests that the mental health effects of climate change will not be uniform and, like other impacts, will fall disproportionately on the vulnerable and marginalized.
+More positively, it does suggest that adaptation is possible, and, for communities with infrastructure and working conditions similar to those of higher-income Americans, the effects of high temperatures on mental health may be addressable.
+
+Examining the role of race in vulnerability, we find that majority Black non-Hispanic neighborhoods are much more affected by higher temperatures than neighborhoods with a non-Black majority.
+Surprisingly, we did not find an effect for people with Hispanic ethnicity, given that Hispanic peoples can be marginalized in employment and housing.
+These findings may be because the ethnic category of Hispanic is more broad and encompasses many more groups with more diverse histories and living conditions than Black Americans.
+It may also be due to that fact that more marginalized and heat-vulnerable Hispanic people were more likely to tweet in the Spanish language, which we did not include our analysis.
+Additionally, there are many other marginalized groups in the United States, particularly indigenous people, that we did not have sufficient data to examine with respect to heat and mental health.
+Examining the effects of both race and income, we find that race is more important for vulnerability than income.
+This probably due in part to the fact that Black Americans have historically been forced through redlining into neighborhoods with fewer trees that are much more impacted by heat \citep{Hoffman2020, Locke2021}.
+This suggests that racism and patterns of housing discrimination contribute to making Black Americans more vulnerable to heat, even at similar levels of income as other groups, and that reducing existing discrimination in housing and employment is an important adaptation strategy.
+
+% Connect to bigger literature on vulnerability
+This study shows the importance of examining heterogeneities in vulnerability in analyses of climate change impacts.
+Additionally, it shows how measuring outcomes at fine spatial scales necessary for understanding how factors like race and income matter for vulnerability \citep{Tong2021}.
+Even within a relatively wealthy nation, we find that high temperatures affect certain populations to the same degree as a major hurricane.
+Thus, it is likely that heat waves have a much greater mental health impact throughout the developing world, where temperatures are expected to be much higher \citep{Raymond2020May}, heatwaves are already under-counted \citep{Harrington2020Sep}, and adaptive technologies like air conditioning are more scarce \citep{Biardeau2020Jan}.
+
+In addition to exploring heterogeneities in vulnerability, using twitter data lets us explore how the effect of heat varies by time of date and infer possible important pathways in the heat-mental health relationship.
+Recent research has suggested that the impacts of heat on sleep quality may play a large role in the observed mental health effects of heat \citep{Obradovich2017May, Mullins2019Dec}, and our temporal analysis largely supports this hypothesis.
+We found much stronger effects of heat on expressed sentiment in the early morning, adding weight to the sleep quality pathway linking heat and mental health that other authors have found.
+This suggests that efforts to improve mental health during heat waves may have the largest impact by focusing on providing electricity and air conditioning at night.
+
+% Caveats
+% Mood is a fuzzy indicator but we have BIG DATA
+While these findings are robust to different model specifications and metrics of mood, there are important considerations in this analysis.
+One caveat associated with this data is that we were only able to locate the tweets within the census block from which the tweet was sent, and in some cases the income level of a census block may be only weakly indicative of the wealth of the people who are found that census block.
+For example, some public spaces are estimated to have very low income levels even though people from a variety of income levels may occupy those spaces throughout the day.
+These issue may in fact lead us to under-estimate the true effect of heat on mental health for poor and Black people.
+Another issue with using Twitter data is that, while Twitter is used by more than one in five Americans and at similar rates across racial groups, Twitter users may not be representative of the general population, as they are typically younger, wealthier, and more educated \citep{Pew2020Sep}.
+Nevertheless, we found a large volume of tweets across all neighborhood types.
+
+% NCC doesnt have a "conclusion" section, but this is intended as an overall summary paragraph
+While climate change will have widespread and severe impacts on human well-being, our work emphasizes just how unevenly these impacts will be distributed.
+People with more money, access to aid and infrastructure, and who belong to ethnic groups in power are less affected by climate shocks and natural disasters and more able to adapt \citep{bullard2012wrong}.
+While the physical health impacts of these shocks are more visible and easy to measure, the mental health impacts of climate change are also causing severe human suffering and there is no reason to believe they will not also be highly unevenly distributed.
+Thus, there are strong theoretical priors behind the hypothesis that low-income and marginalized people are more vulnerable to the mental health impacts of higher temperatures, even though previous work at coarse spatial scales had not found such an effect.
+By using fine-scale Twitter data, we show that there are indeed stark differences in the mental health impacts of heat among neighborhoods in the United States.
+These findings have significant implications for urban planning, climate and environmental justice, and mental health.
+
 \section*{Acknowledgements}
@@
 
-%\printbibliography
-\bibliography{TextDataClimateShocks}
-
 \section*{Author Contributions}
@@
 and J.O.-G.
-analyzed data; and M.C., J.O.-G., P.A.W., and J.L.
+analyzed data; and M.C., J.O.-G., P.A.W., J.L.
+and Z.L.
 wrote the paper.
@@
 
+\printbibliography
 \end{document}
